### 2014-08-12 proposal/basic-prop.tex
Added basic research proposal

\documentclass[a4paper]{article}
\usepackage[cm]{fullpage}

\title{Honours 2014: Research Proposal}
\author{Tianyu Pu}

\date{Semester 1, 2014}

\begin{document}

\maketitle

\section{Title}
The research question is still being refined, but the title will be something
along the lines of ``Using data mining techniques to predict the length of stay
or mortality outcome of patients admitted to a major trauma centre".

\section{Supervisor}
Irena Koprinska, \texttt{irena.koprinska@sydney.edu.au} \\
School of Information Technologies, 4E

\section{Abstract}
The aim of the study is to derive a model using various artificial intelligence
and data mining techniques to predict the length of stay and mortality outcomes
of patients that are admitted to a major trauma centre. This model will be
internally validated using data from the Royal Prince Alfred Hospital. A mobile
application will be developed to assist planning and decision-making for
doctors in the trauma centre. There will also be some room to investigate
whether such a model can be generalised or used by other medical departments to
assist in their resource allocation and treatment preparation.

\section{Desk No.}
4E14

\bibliographystyle{plain}
\bibliography{references}
\nocite{*}

\end{document}

### 2014-08-12 thesis/introduction/introduction.tex
Added thesis skeleton with lit review fleshed out

\chapter{Introduction}

The introduction goes in here.

### 2014-08-12 thesis/litreview/litreview.tex
Added thesis skeleton with lit review fleshed out

\chapter{Literature Review} \label{chap:litreview}

\section{Introduction}
Hospitals often have limited funding, and therefore managing and utilising
their resources efficiently has always been a priority\citep{Walczak2003}.
One key measurement of hospital activity, health care utilisation, and hospital
cost is the patient length of stay\citep{Omachonu2004,Ng2006}. The length of
stay has been used in a wide variety of medical domains, such as burns
\citep{Yang2010}, intensive
care\citep{Tu1993,Harper2005,Perez2006,Dybowski1996},
psychiatry\citep{Lowell1997}, and acute pancreatitis\citep{Pofahl1998}, in
order to assist hospitals in allocating beds and other patient management
resources as efficiently as possible. However, given the many different
conditions that hospitals manage, the length of stay of a patient is
not easy to predict\citep{Walczak2003}, especially upon a patient's admission.
Many methods have been explored in the literature, from the realm of statistics
to techniques in data mining, and they have been explored for many different
medical domains.

One of the more well-understood and widely used statistical methods for
predicting the length of stay is \textit{logistic regression}\citep{Tu1996}.
Logistic regression models are straightforward to construct and use via the
help of statistical software packages such as
SAS\footnote{SAS Institute, Cary, NC, USA}, and
also make the contribution of each independent variable to the final
prediction explicit. Thus they is easy to interpret when compared to more
``black box'' approaches such as artificial neural networks\citep{Adams2012}.
A wide variety of other statistical approaches have been proposed, notably:
various scoring systems from univariate analysis\citep{Adams2012,Lavoie2005},
linear regression\citep{Yang2010}, exponential models\citep{Clark2007},
survival analysis\citep{Vasilakis2005}, and Markov
models\citep{Perez2006,Jain1989,Kapadia2000}.

Out of the development of the various statistically-motivated models, data
mining techniques, particularly artificial neural networks,
began to become more widely adopted following a number of
early applications to length of stay prediction in intensive
care\citep{Tu1993,Dybowski1996}, in addition to disease diagnosis and mortality
prediction\citep{Silva2006}. Artificial neural networks are able to detect
complex and non-linear relationships between input and output variables and can
be developed using many different techniques. Much work has been done in using
neural networks to predict patient length of stay in various medical conditions
such as acute pancreatitis\citep{Walczak2003}, gastroenteritis\citep{Ng2006},
surgical intensive care\citep{Buchman1994,Tu1993} and
psychiatry\citep{Lowell1997}. Other data mining techniques have also been
applied, such as support vector machines and tree-based
methods\citep{Harper2005}.

Given the benefits to the hospital of accurate length of stay prediction,
this research aims to build upon improving length of stay prediction for trauma
patients by utilising state-of-the-art data mining techniques. Being able to
predict the length of stay for trauma patients is especially crucial because of
the scarcity of patient care resource in this area\citep{Yang2010}.
Additionally,
despite the wealth of research into length of stay prediction in various
medical domains, there has been little work done in generalising models from
one domain to another, or to predict a combination of outcomes (not only
length of stay, but mortality, or rehabilitation requirement). Being able to
do this would mean that hospital staff are better equipped to handle patient
cases and prepare for possible critical outcomes well in advance, resulting
in time and resources saved for both hospital and patient.

This literature review is structured as follows. We will first describe the
various statistical modelling techniques for predicting length of stay that
have been used in the literature, followed by a review of the data mining
techniques and how they compare in terms of accurately being able to predict
patient length of stay. The review will conclude with a summary and a
discussion on the gaps in the literature and how this work intends to address
those gaps.

\section{Statistical methods}
\subsection{Regression}
As we stated earlier, logistic regression is a commonly-used modelling
technique for prediction in the medical domain\citep{Tu1996}, and especially
in the trauma data analysis\citep{Gabbe2005}. A series of
predictor (independent) variables (or features) determine the probability of a
particular outcome by an equation of the form
\begin{equation}
\mathrm{log}\left(\frac{p}{1-p}\right) = \beta_0 + \beta_1X_1 + \beta_2X_2 + \dots + \beta_nX_n
\end{equation}
where $p$ is the probability of the desired outcome, $\beta_i$ are coefficients
associated with each variable, and $X_i$ are the values of the independent
variables\citep{Hosmer1989}. A stepwise selection algorithm is often used to
choose which variables to include in the model out of all available variables:
different algorithms produce different models which are then compared against
one another to find the best model. Usually, a standard maximum-likelihood
procedure is followed in order to obtain values for the $\beta_i$'s, and models
can be tested using measures such as the area under the receiver operating
characteristic (ROC) curve (AUC) and the Hosmer-Lemeshow statistic.

This is the method that Gabbe et al. used in their study to identify factors
that contributed to trauma patient length of stay and use these factors to
subsequently predict length of stay\citep{Gabbe2005}. The prediction problem
they proposed was essentially a binary classification problem, to which
logistic regression is well suited\citep{Tu1996}: will length of stay be
\textit{less than 5 days}, or \textit{equal to 5 days}? They used a
\textit{backward} stepwise selection algorithm, whereby all available variables
were initially included in the model and then progressively eliminated if their
likelihood ratio (LR) test $P$-value was $>0.15$. Features could only be
re-included if $P<0.05$. For length of stay
prediction, this left them with 9 features out of the 15 that they originally
included. To validate their model, they used a data set collected from an
independent source which had slightly different statistical characteristics for
the various features and outcomes. Their length of stay prediction model was
found to have 95\% sensitivity and 24\% specificity, with an area under the ROC
curve (AUC) of 0.62. This indicates that the model is not able to discriminate
length of stay cases of less than 5 days and equal to 5 days (i.e., the two
desired outcomes) very well, and only does
slightly better than simply random guessing (which has AUC 0.50).

In a similar study, Dinh et al. sought to predict whether or not trauma
patients could be predicted to stay in hospital \textit{less than 2 days} or
\textit{greater than 2 days}\citep{Dinh2013a}.
They also used a logistic regression model and
selected their features using a stepwise selection algorithm with the same
$P$-value threshold (0.05). Even though we cannot compare their result to that
of Gabbe et al. above because the models were tested on completely different
data sets, it is interesting to note that the model obtained by Dinh et al.
had an AUC of 0.83 after bootstrap cross-validation\footnote{A technique of
sampling a single data set with replacement: e.g. given a set of 100 records,
we can create a bootstrap sample that also contains 100 records which contains
records from the original set drawn at random with replacement. This means that
some records may appear in the bootstrap sample more than once and some may not
appear at all. In the work of Dinh et al., they performed this sampling 500
times (thereby creating 500 data sets) to evaluate the performance of their
model}. This indicates that the model had much better discrimination between
the two values for the outcome. Again, while we cannot directly compare, the
better performance could be due to the availability of more relevant features
to the outcome: Gabbe et al. included very few features specific to trauma
such as place of injury and mechanism of injury\citep{Gabbe2005} whereas Dinh
et al. also had variables for certain kinds of injuries (such as if it was
penetrating), specific body regions that were affected, and whether operations
were required. Given the nature of the length of stay, it would seem that Dinh
et al. derived a model that had included features with more explanatory power
on the outcome, thereby obtaining better results. This idea is consistent with
Buchman et al., who used logistic regression to predict the probability of a
patient in a surgical intensive care unit staying greater than a
week\citep{Buchman1994}.

While logistic regression is simple to use and its results straightforward
to interpret, the above approaches have described a coarse two-class prediction
problem, namely that of predicting the length of stay less than or greater than
a certain threshold. A number of \textit{linear} regression models have been
explored in the literature that examine the task of predicting length of stay
as a continuous quantity.

The key difference between linear and logistic regression is the nature of the
dependent variable: in linear regression it is treated as continuous (i.e., it
can take on any value), whereas in logistic regression it is a dichotomy of
only two values. The works discussed above only considered predicting the
length of stay as two possible outcomes, less than or greater than a particular
threshold of time. We have already stressed the importance of accurate length
of stay prediction for hospitals, and therefore it makes sense that attempting
to predict length of stay as a continous variable could give us more accurate
and useful insight and facilitate better resource allocation.

Yang et al.\citep{Yang2010} used a linear regression model as the baseline for
performance benchmarking when trying to predict the length of stay of burn
patients. Interestingly, unlike the logistic regression models discussed above,
they did not perform any feature selection when constructing their models and
found that this gave better performance against models constructed using
correlation-based feature selection\footnote{Abbreviated CFS, correlation-based
feature selection is a technique of choosing which independent variables to
include in a model based on the hypothesis that good variable subsets should
be highly correlated with the dependent variable but not with each
other\citep{Hall2000}.}. Unlike other studies into length of stay prediction,
they also looked at predicting the length of stay at various stages of the
clinical process, rather than simply at admission. By making predictions
at the admission, acute, and post-treatment stages, they were able to decrease
their predictions' mean absolute error (MAE) from 9.532 days at admission, to
6.314 days at post-treatment, where MAE is defined as
\begin{equation}
MAE = \frac{\sum_{i=1}^n |\mathrm{PredictedLOS}_i-\mathrm{ActualLOS}_i|}{n}
\end{equation}
and PredictedLOS$_i$ is the predicted length of stay for the $i$-th instance,
ActualLOS$_i$ is the actual length of stay for the $i$-th instance, and $n$
is the total number of instances. Their approach was very much appropriate to
their medical domain (burn patients) as they tended to have longer lengths of
stay (which is more suitable for linear regression rather than logistic
regression) and the extended period of stay means that there will be enough
data from several phases of the clinical process for such patients. It is
difficult to apply such a phase method to our problem of trauma length of
stay prediction, however, as there is no standard clinical process due to
the varying nature of trauma causes and the lengths of stay are typically not
as long.

\subsection{Markov models}
There have been a number of studies attempting to model the length of stay
using discrete-time Markov processes. Kapadia et al. constructed a transition
probability matrix using 4 states for a pediatric intensive care unit (ICU)
application\citep{Kapadia2000}:
\begin{center}
\begin{tabular}{l|llll}
& Low & Medium & High & Exit \\
\hline
Low & 0.309 & 0.223 & 0.024 & 0.444 \\
Medium & 0.137 & 0.513 & 0.096 & 0.254 \\
High & 0.086 & 0.364 & 0.379 & 0.171 \\
Exit & 0 & 0 & 0 & 1 \\
\end{tabular}
\end{center}
From the matrix, we can say that the probability of transitioning from state
`Medium' to state `Low' is 0.137. Note that each row sums to 1, and that `Exit'
is an absorbing state: once entered, one can only re-enter the `Exit' state
with 100\% certainty. This represents patients leaving the ICU. How did they
obtain these probabilities? From their data set, they classified each patient
with an appropriate state (Low, Medium, High) corresponding to the severity of
their illness based on a particular mortality score (PRISM), and did this daily
over a few months; this gave them patient records that were simply strings of
L, M and H. To calculate the probability of transitioning from `Medium' to
`Low', they counted all occurrences of `Medium' to `Low' transitions and
divided this by the number of all transitions out of `Medium', and repeated
this for every transition. Using the theory of Markov processes, they are able
to calculate the average time taken before a person enters the `Exit' state,
which is precisely the length of stay. They found that the Markov model fitted
the data well and produced predictions that were very close to the observed
results in the test set\citep{Kapadia2000}, indicating that using such models
justifies further investigation.

Perez et al. also used a Markov model to estimate the length of stay of
patients in a number of different destinations after one day of ICU admission
\citep{Perez2006}. They created six models for different body regions and their
system had eight states. Using these models, they were able to obtain a
detailed breakdown prediction of how long each patient would be expected to
stay in a particular state given their injury. Their models had varying levels
of fit to the data, suggesting that some systems did not satisfy the Markov
property. Using such models gives practitioners a greater level of insight
into how patients are expected to progress through their particular situation
and allows them to prepare accordingly, rather than simply providing a number
for the length of stay. Clearly, there is an opportunity here to investigate
the types of medical problems that Markov modelling works well for, as well as
the problems that do not work so well.

\subsection{Other statistical techniques}
A number of other techniques have been used in length of stay prediction, which
we will describe briefly.
Yau et al.\citep{Yau2003} proposed a \textit{mixture regression model} in
modelling
length of stay to overcome the problem of correlated observations in hospital
data, which allowed them to predict neonatal length of stay across heterogenous
sub-groups (short-stay and long-stay). The idea is that these two sub-groups
have different length of stay patterns, and the length of stay of any patient
is just the probability of them being in one of these two groups (as well as
taking into account the underlying distribution of the group itself). Applying
this method to a set of neonatal length of stay data, they were able to
identify which input factors contributed to which component (short or long
stay) and derive useful insight for hospitals.

Clark et al. used a similar idea in their study of mortality and length of stay
prediction\citep{Clark2002}, by employing \textit{piecewise exponential models}
to model different probabilities of transition over distinct periods of time.
They obtained predictions for the length of stay and the probability of
survival that were fairly close to the actual outcome, for different age groups
and trauma types. These approaches have the advantage of being able to fit the
data more closely, but are more complex to estimate and use.

\section{Data mining methods}
Data mining is the analysis of (often large) observational data sets to find
unsuspected relationships and to summarise the data in novel ways that are both
understandable and useful to the data owner\citep{Hand2001}. Compared to
traditional statistical, it is a relatively recent field that emerged in the
1990s, and has
given researchers a variety of new tools to analyse and understand biomedical
datasets\citep{Bellazzi2008,Yoo2012}. Bellazzi \& Zupan\citep{Bellazzi2008}
have written a
comprehensive review paper of data mining techniques in clinical prediction
settings, and we refer the reader to their article for a more detailed
treatment of the techniques and their applications. Here we will describe the
major methods in the context of their application to predicting hospital length
of stay.

\subsection{Artificial neural networks}
Artificial neural networks have been widely used in clinical prediction
following their use in diagnosing acute myocardial infarction and various
applications in radiology\citep{Baxt1995}.
One of the earliest applications of neural networks to
length of stay prediction was by Tu \& Guerriere, who used a neural network
with the following structure\citep{Tu1993} to predict length of stay in
intensive care following cardiac surgery:
\begin{figure}[h]
\centering
\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

There are 15 nodes (or ``neurons'') in the input layer, one for each of the
input features (predictor variables). For each example in the training set,
the input variable values are presented to the corresponding node in the input
layer, which are then propagated into the hidden layer and subsequently the
output layer. At the hidden layer, each of the 12 nodes sums the weighted
value of the 15 inputs (weighted value = input value $\times$ connection
weight), applies a non-linear activation function (usually the sigmoid function
$\frac{1}{1+e^{-t}}$), and sends this value as its output to the next layer
(in this case, the output layer). The single output node sums the 12 values
it receives from the nodes in the hidden layer and applies the same activation
function to produce a value between 0 and 1 that indicates the class.

The neural network is commonly ``trained'' by the backpropagation algorithm.
Once the network predicts a certain value for the given input record, this
predicted value is compared to the actual (or known) value, an error is
calculated and the connection weights are adjusted in order to minimise the
error quantity, which was the root mean square error in Tu \& Guerriere study.
The training is complete when the error has converged to a minimum value.

Using this technique,
Tu \& Guerriere wanted to predict whether a patient was going to stay
\textit{less than 2 days} or \textit{more than 2 days} following cardiac
surgery. After testing their trained neural network on the test data, they
found that it had an AUC of 0.6960, indicating a reasonable amount of
discriminating ability between the two outcomes. Even though we cannot compare
their result to any of the studies conducted on length of stay prediction
using logistic regression techniques, there are a number of well-known reasons
why neural networks are advantageous to modelling complex clinical situations:
\begin{enumerate}
\item Neural networks are able to model arbitrarily complex relations between
independent and dependent variables\citep{Tu1996,Buchman1994}, especially if
the relation is of a non-linear nature
\item Neural networks are able to handle models where the features (independent
variables) are not completely independent of each other
\item Neural networks can be ``trained'' in different ways, which changes their
predictive power -- this has been studied by Walczak et al.\citep{Walczak2003},
discussed in the next section
\end{enumerate}

However, there were a few difficulties in using the neural network as Tu \&
Guerriere discovered, which are problems often cited in the use of such
networks. These include a high sensitivity to the starting parameters (such as
network architecture\footnote{For example, how many hidden layers to have, and
how many nodes in each hidden layer. The difficulty of finding the right
architecture is pointed out by Tu \& Guerriere themselves -- in their paper
they state, ``The optimal structure of the neural network model was determined
by `trial and error' ''.}
and learning rate\footnote{A constant that is set by the
network developer that governs how much the connection weights should adjust at
each iteration of the training algorithm.}), as well as the difficulty in
inducing sensible rules from the model (even by domain
experts)\citep{Bellazzi2008}. Additionally, they are also prone to overfitting
and require substantially more computational power to train when compared to
the standard logistic regression technique, and do not always perform
better\citep{Harper2005}. A number of studies have attempted to address some of
these issues in length of stay prediction, building on from Tu \& Guerriere's
work, and these will be described in the next section.

\subsection{Extensions to the artificial neural network model}
Buchman et al. published a comparison of the performance of logistic regression
and neural networks on predicting whether or not patients would stay greater
than 7 days in a surgical intensive care unit\citep{Buchman1994}. Instead of
using one type of training algorithm as Tu \& Guerriere did, they obtained
results from three different neural network algorithms: backpropagation (the
standard against which other training methods are compared), probabilistic, and
generalised regression. Probabilistic neural networks are trained by a single
pass over the training data and use Bayes strategies and Parzen windows to
estimate a probability density function for prediction. Generalised regression
neural networks allow the definition of arbitrarily complex regression surfaces
which allows for greater modelling power and is able to produce continuous
values instead of simply a binary classification\citep{Sprecht1991}. They found
that all three neural networks performed noticeably better than the logistic
regression model, but that they all achieved fairly similar results (by the ROC
test). However, the AUC for the four models were not given.

Dybowski et al. decided to address the problem of neural network architecture
by applying a genetic algorithm approach to selecting the best
structure\citep{Dybowski1996} --
an improvement from the ``trial and error'' method Tu \& Guerriere used. It is
interesting to briefly discuss their method even though they were interested
in predicting the outcome of critically ill patients in intensive care, rather
than the length of stay which is the focus of our work. The genetic algorithm
in choosing an optimal neural network structure is as
follows\citep{Dybowski1996}:
\begin{enumerate}
\item A large population of random neural networks is created, each with
a random number of nodes in three hidden layers.
\item Each neural network optimises
its connection weights with respect to the same data set using
backpropagation.
\item Each neural network is expressed as a string which represents the
substructure of the network -- this is the ``chromosome''.
\item The ``chromosomes'' are copied to form a new population whereby the
expected relative frequency of a chromosome is proportional to how well
its neural network performed -- this is the ``survival of the fittest''
aspect.
\item The replica chromosomes undergo crossover so that better-performing
components can be combined. Mutations are also present, through the random
modification of a chromosome.
\item The process is repeated until a best-performing neural network forms.
\end{enumerate}
Through this Dybowski et al. were able to show that successive ``generations''
of neural networks had greater AUC measures, indicating better predictive
ability. Additionally, after seven generations, they obtained a neural network
that had an AUC of 0.863 compared to 0.753 for an equivalent logistic
regression model.

Walczak et al. also attempted to address the problem of arbitrary neural
network structure in their study to predict length of stay in two hospital
settings, acute pancreatitis and pediatric trauma\citep{Walczak2003}. Instead
of a genetic algorithm approach, they generated and tested every possible
neural network with one or two hidden layers, the quantity of nodes in each
layer varying from $2N$ to $5N$ where $N$ is the number of nodes in the
preceding layer, with a step size of 5. To address overfitting they also
discarded any networks generated this way whose performance dropped after
an initial increase. This comprehensive approach gave them over 100
backpropagation neural networks to evaluate for each of the two medical
domains. 
Rowan et al. followed a similar
approach to mitigate the effect of neural network structure, but instead of
selecting a single optimal network, they used ensembles of neural networks
to improve predictive power\citep{Rowan2007}. Many different combinations of
trained neural networks were tested in order to find the combination with the
greatest classification power according to the AUC measure.

Additionally, Walczak et al. also examined not only backpropagation, which
is the training algorithm used by most of the applications, but also radial
basis function and fuzzy ARTMAP. Using the mean absolute difference, they
found that a single hidden layer backpropagation neural network performed
better than all other networks in pediatric trauma, and was able to predict
the length of stay to within 1 day for 52\% of all cases. They also obtained
promising results using the fuzzy ARTMAP method, which warrants further
investigation as this training method has not been applied to modelling
length of stay in other medical domains.

One concern that has been raised with the use of neural networks in prediction
is that they are usually trained once and are not subsequently re-trained with
new data to fine-tune their connections. This has been worked on by Lowell et
al. in predicting psychiatric length of stay\citep{Lowell1997}, and also by
Ng et al. who proposed an incremental learning approach for neural networks
and applied it to gastroenteritis length of stay data\citep{Ng2006}. Lowell et
al. trained an initial model in an earlier paper to predict the length of stay
of psychiatric patients, and then subsequently deployed this model to three
separate sites. Using the additional site-specific data to re-train the initial
neural network resulted in increased prediction accuracy of 3-15\% with a less
than 10\% change in the training data\citep{Lowell1997}. Ng et al. took a more
theoretical approach and proposed a mixture of experts neural network with an
incremental learning algorithm based on expectation-maximisation. They wanted
to investigate how the neural network trained all at once (batch-mode)
would perform against the same network trained all at once and then
incrementally with new records as they were obtained. This allowed for an
on-line prediction system which takes into account new data, which had not
been investigated very much in the area of length of stay prediction. They
applied their proposed technique to a set of gastroenteritis data, and found
that the incremental learning method had a lower mean absolute difference (MAD)
on the test set (indicating that predictions were generally closer to the
actual value), and had a greater proportion of predictions with MAD less than
or equal to 1 day (meaning that more of the incremental learning neural
network's predictions were accurate to within a day), compared to the standard
batch-mode learning method\citep{Ng2006}.

\subsection{Other data mining techniques}
As discussed, artificial neural networks have been used extensively in
predicting patient length of stay. However, a few other data mining techniques
have also been investigated, namely tree-based methods and support vector
machines.

In a review published in 2005, Harper compared the performance of four
classification algorithms on four different medical data sets, trying to
model five different scenarios\citep{Harper2005}. Two of these outcomes were
relevant to our work: one was predicting the length of stay in the intensive
care unit, and another was predicting the length of stay in a particular
hospital ward. The classification techniques he used were discriminant
analysis, regression, CART (classification and regression trees),
and artificial neural networks.

CART has been extensively used in medical decision making, and at its heart
involves an algorithm that is used to split the data set into sub-groups of
increasing similarity. At each node in the tree, the algorithm examines the
independent variables and splits the data set on the variable that produces
a node with the smallest variance (i.e., the greatest similarity). This is
continued until defined stopping rules are met. One of the major reasons that
such tree-based methods are more popular than neural networks in the
healthcare domain is that the rules it ``learns'' are very simple to extract
from the resulting tree, unlike the ``black box'' that neural networks commonly
are. Harper found that CART performed almost as well as neural networks in
predicting intensive care length of stay, and much better in predicting
hospital length of stay in terms of predicting accuracy (the only evaluation
statistic reported). He also found that CART was superior to regression, a
finding that Yang et al. also reported when predicting the length of stay of
burn patients\citep{Yang2010}.

Yang et al. investigated the performance of linear regression, M5 (another
tree-based classification method), and support vector machines in predicting
length of stay for burn patients. The difference between their tree method,
M5, with that of Harper (CART), is that the nodes contain linear regression
functions instead of the usual discrete (two-class) outputs that are used to
successively split the data set. Additionally, they also use support vector
machines (SVMs), perhaps the most powerful classification algorithm in terms of
predictive accuracy\citep{Bellazzi2008}. SVMs work on the principle that when
the data is represented in a high enough dimensional space, a linear hyperplane
can be found where the margin between two different records is as large as
possible. In figure \ref{svm} from Yoo et al.\citep{Yoo2012}, hyperplane 2
is the better plane because its margin is as large as possible. 
\begin{figure}[h]
\label{svm}
\centering
\includegraphics[scale=1.05]{images/a2-svm}
\end{figure}

However, because SVMs are typically
used for binary classification, Yang et al. adopted the technique of SVM
regression in order to predict a continuous value for the length of stay. The
idea is that the desired function would have the same margin-maximising
properties of the SVM, while being able to approximate the mapping between the
input domain and the real numbers\citep{Yang2010}. Evaluating the performance
of linear regression, M5 and SVM regression using the mean absolute error
(MAE) described earlier, they found that predictions from SVM regression
produced consistently lower MAEs across all three clinical stages when compared
to both the regression and tree-based methods. Regression was the poorest
performing predictor, which is consistent with many of the findings we have
discussed. The paucity of studies using tree-based methods and SVM techniques
to predict length of stay warrants more investigation, especially since the
SVM has high predictive power which would be useful to medical professionals.

It is important to note that length of stay prediction is most often a
\textit{supervised} problem, i.e. the desired outcome is known in the training
data, and we train our model to be able to produce the correct output based on
this data. Once we present this trained model with data with which the outcome
is not known, we expect that it will give us a reasonable prediction. This is
what all of the techniques described above fall under, and this is expected
given the nature of the problem. However, we should note that Huang et al. 
used an unsupervised technique, $K$-nearest neighbour clustering (KNN), in 
predicting the length of stay for clinical
treatment processes\citep{Huang2013}. Their approach assumes that similar
patients have a similar length of
stay. They define a patient trace to be a sequence of clinical events that are
performed on a patient, and define a similarity measure that follows the
principle of temporal edit distance\footnote{The \textit{edit distance}
between two states (or traces in this case) is the total ``cost'' of edit
operations (insertions, deletions, substitutions) required to change from
one state to another. The costs of each operation are defined based on the
application.}. Using this similarity measure, they find the $K$ most similar
patient traces to a given trace of unknown length of stay,
and find a similarity-weighted average of the lengths of stay
of the $K$ traces to produce a prediction for the unknown trace. They point
out some difficulty in determining the most suitable number for $K$, as well
as in measuring the performance of the algorithms due to the temporal nature
of the patient traces (very different error amounts depending on at which
time points measurements are taken at), but generally found that their
approach performed better than the widely used techniques of regression and
neural networks. This is a novel approach and should be investigated in further
studies for length of stay prediction.

\section{Conclusion}
Despite the wide range of statistical and data mining techniques that have been
used to predict the hospital length of stay, we have raised a few avenues for
further work throughout the literature review. Firstly, there has been no work
done to predict the length of stay of trauma patients using data mining
techniques, which is something worth investigating as trauma is a complex
medical condition. We have pointed out that there is also value in
investigating a number of the best data mining techniques, i.e. using more
than just neural networks. There is also value in investigating the performance
of such a model on data sets from different locations, since most studies are
only validated on one or two data sets. Additionally, length of stay prediction
has largely been restricted to specific medical domains, and therefore it is
worthwhile to examine if there are factors present in other domains which allow
for a more general model of the length of stay.

### 2014-08-12 proposal/a2-prop.tex
Added the more fleshed out proposal

\documentclass[a4paper]{article}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{natbib}

\title{INFO5993 IT Research Methods \\
    Assignment 2: Outline of Research Approach}
\author{Tianyu Pu \\
    Supervisor: Irena Koprinska}

\date{\today}

\begin{document}

\maketitle

\section{Research Contribution}
\subsection{Background}
Hospitals are always subject to limitations in funding, and as a result,
it is critical that they effectiely utilise the resources at their
disposal.
A key measure of hospital activity and utilisation is the length of stay
of patients, usually expressed in days. To this end, much research has been
conducted in the past few decades to find accurate and reliable models to
assist in predicting the length of stay of patients, so as to allow hospitals
to allocate their resources and plan their treatment processes more
efficiently. This is especially important in settings where there is a
prominent scarcity of a resource, such as patient beds\citep{Tu1993}.

Due to the numerous functions and units of hospitals, and the absence of a
standard in medical data collection, many models have been devised in the
literature applying statistical or data mining techniques (or both) to
length of stay prediction for various medical domains, such as burn injuries
\citep{Yang2010} and psychiatry\citep{Lowell1997}. Among these, logistic
regression remains a commonly-used and well-understood technique
\citep{Gabbe2005}\citep{Tu1996}. Artificial neural networks are also gaining
attention in the medical domain through their use in predicting various
outcomes of diseases such as breast cancer\citep{Bellazzi2008}. However,
many works have restricted themselves to looking at a few statistical or data
mining methods in a very specific domain.

\subsection{Intended Contribution}
As part of my research, I intend to:
\begin{enumerate}
\item apply a range of data mining techniques (artificial neural
networks, support vector machines, decision trees, na\"{i}ve Bayesian
classifier, Bayesian networks) to predict the length of stay of patients that
are admitted to a trauma centre. This builds upon the work of Dinh et al.
\citep{Dinh2013a}, who constructed a logistic regression model;
\item develop a new, accurate, and most importantly reliable method for
predicting the length of stay of trauma patients, based on the advantages and
drawbracks of the investigated data mining methods
\item use data mining techniques to discover what patient factors contribute
to their transfer to rehabilitation after trauma, and to predict the
probability of requiring rehabilitation upon admission;
\item generalise my findings in the trauma domain into other hospital functions
to see if there are over-arching, universal, or common factors that contribute
to how long patients stay in hospital; and
\item develop a mobile application that is easy-to-use and reliable that will
incorporate the proposed methods to assist medical professionals in resource
planning.
\end{enumerate}

The value of the work will be multi-faceted: accurately predicting how long
a patient will stay when they are first admitted will allow hospital staff to
better plan their resources (such as beds, medication, and the staff
themselves) to efficiently use the funding that they receive; additionally,
knowing about common warning signs across all medical domains that indicate
longer stays means that staff can plan for long-term structural improvements
in the hospital (for example, if a particular ward receives patients that
characteristically require longer-term care, then they could look at adding
more beds or other decisions to improve the delivery of care to that ward).

\section{Evaluation}
In order to gauge how the data mining models perform, I will compare them to
the logistic regression model that Dinh et al.\citep{Dinh2013a} came up with.
In that study, the area under the receiver operating characteristic curve was
used to measure the performance of the model, and I will also use this
measurement to evaluate the models that I construct. Additionally, I will
also use a number of other metrics (such as sensitivity, specificity,
classification accuracy, and mean absolute difference\citep{Walczak2003}) to
evaluate and compare the performance of the various techniques.

It is also important to keep the training and testing data separate as part
of the evaluation process. To this end, I will use ten-fold cross-validation
and the average of the runs will be reported. In order for this to be
comparable to Dinh et al.\citep{Dinh2013a}, who used bootstrap
cross-validation, I will re-run their method and use ten-fold cross-validation
to evaluate the performance of their model. These techniques ensure that we
obtain an accurate record of the performance of the constructed models, and
allows us to make comparisons and draw conclusions and insights from them.

\bibliographystyle{abbrv}
\bibliography{../hons/references}

\end{document}

### 2014-08-12 poster/myposter.tex
Added poster template and skeleton files

\documentclass[portrait]{usydposter}
\usepackage{LI}
\usepackage{xspace}

\newcommand{\acronym}[1]{\textsc{#1}\xspace}
\newcommand{\cf}[1]{\mbox{$\it{#1}$}}
\newcommand{\todo}[1]{{\color{red} #1}}

\newcommand{\ngram}{n-gram\xspace}
\newcommand{\ngrams}{{\ngram}s\xspace}
\newcommand{\candc}{C\&C\xspace}
\newcommand{\ccgbank}{CCGBank\xspace}
\newcommand{\ccg}{\acronym{ccg}}
\newcommand{\cky}{\acronym{cky}}
\newcommand{\nlp}{\acronym{nlp}}
\newcommand{\np}{\acronym{np}}
\newcommand{\pos}{\acronym{pos}}
\newcommand{\wsj}{\acronym{wsj}}

\flushbottom

\title{CCG Parsing with One Structure per n-gram}
\author{Tim Dawborn}

\begin{document}
\makeheader

\begin{multicols}{3}

% =============================================================================
\section{Heading RGB 206, 17, 38 -- Arial 28}
\noindent This is some text here...

\begin{itemize}
  \item Parsing is the process of deriving the grammatical structure for a given sentence
  \item Natural language parsing is currently slow, with the best algorithms running in $O(n^3)$ with respect to the length of the sentence
  \item There is an inherent \emph{redundancy} in natural languages where certain common phrases (or \ngrams) appear frequently in text, each time with the same grammatical structure
  \item We explore the idea of exploiting this redundancy through the construction and \emph{memoisation} of the parse structures for these frequent \ngrams
\end{itemize}

% =============================================================================
\section{Background}
\begin{itemize}
  \item \cky is a $O(n^3)$ parsing algorithm which uses a chart data structure, illustrated in Figure~1. A chart is a hierarchical triangular structure  which stores the parse trees for all substrings of the sentence being parsed
\end{itemize}

\begin{figure}
  \pgfimage[width=0.7\columnwidth]{images/chart}
  \caption{An illustration of the chart data structure used in parsing algorithms such as \cky}
\end{figure}

\begin{itemize}
  \item Combinatory Categorial Grammar (\ccg) \cite{steedman:00} is a \emph{lexicalised grammar formalism}
  \item Each word in a sentence is assigned a composite object that reflects its role within the sentence
  \item The following is an example derivation using \ccg, where the lexical categories for each token are shown
\end{itemize}

\begin{center}
  \small
  \deriv{4}{
    \rm I & \rm like & \rm the & \rm cat \\
    \uline{1} & \uline{1} & \uline{1} & \uline{1} \\
    \mc{1}{\cf{NP}} & \mc{1}{\cf{(S[dcl]\bs NP)/NP}} & \mc{1}{\cf{NP[nb]/N}} & \mc{1}{\cf{N}} \\
    & & \fapply{2} \\
    & & \mc{2}{\cf{NP[nb]}} \\
    & \bapply{3}\\
    & \mc{3}{\cf{S[dcl]\bs NP}}\\
    \bapply{4}\\
    \mc{4}{\cf{S[dcl]}}\\
  }
\end{center}

% =============================================================================
\section{Problems with Pizza}
\emph{Syntactic ambiguity} is a problem when parsing, and is a particular problem for our one structure per \ngram problem. Consider the sentence

\begin{center}
  \small
  \deriv{5}{
    \rm I & \rm ate & \rm pizza & \rm with & \rm anchovies \\
    \uline{1} & \uline{1} & \uline{1} & \uline{1} & \uline{1} \\
    \mc{1}{\cf{NP}} & \mc{1}{\cf{(S\bs NP)/NP}} & \mc{1}{\cf{NP}} & \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{1}{\cf{NP}} \\
    & & & \fapply{2} \\
    & & & \mc{2}{\cf{NP\bs NP}} \\
    & & \bapply{3}\\
    & & \mc{3}{\cf{NP}}\\
    & \fapply{4} \\
    & \mc{4}{\cf{S\bs NP}} \\
    \bapply{5}\\
    \mc{5}{\cf{S}}\\
  }
\end{center}

and compare it to the sentence

\begin{center}
  \small
  \deriv{5}{
    \rm I & \rm ate & \rm pizza & \rm with & \rm cutlery \\
    \uline{1} & \uline{1} & \uline{1} & \uline{1} & \uline{1} \\
    \mc{1}{\cf{NP}} & \mc{1}{\cf{(S\bs NP)/NP}} & \mc{1}{\cf{NP}} & \mc{1}{\cf{((S\bs NP)\bs(S\bs NP))/NP}} & \mc{1}{\cf{NP}} \\
    & \fapply{2} & \fapply{2} \\
    & \mc{2}{\cf{S\bs NP}} & \mc{2}{\cf{(S\bs NP)\bs(S\bs NP)}} \\
    & \bapply{4} \\
    & \mc{4}{\cf{S\bs NP}} \\
    \bapply{5}\\
    \mc{5}{\cf{S}}\\
  }
\end{center}

These two sentences differ by only one word, yet they have two completely different derivations due to the different roles the word \textrm{with} plays in each sentence.

The word \textrm{with} thus is syntactically ambiguous. Without knowing the surrounding context, it is unknown which grammatical role the word is undertaking.

% =============================================================================
\section{Non-constituent-forming \ngrams}
Memoisation of \ngrams which do not form constituents can be achieved through the use of {\ccg}'s \emph{composition} and \emph{type raising} combinators.

\begin{center}
  \small
  \deriv{3}{
    \rm of & \rm the & \rm potato \\
    \uline{1} & \uline{1} & \uline{1} \\
    \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{1}{\cf{NP/N}} & \mc{1}{\cf{N}} \\
    & \fapply{2}\\
    & \mc{2}{\cf{NP}} \\
    \fapply{3}\\
    \mc{3}{\cf{NP\bs NP}}\\
  }
\end{center}

\textrm{The} is joined to \textrm{potato} before \textrm{of} can be joined with \textrm{the}. Instead, we could construct the following derivation and insert it into the pre-constructed database:

\begin{center}
  \small
  \deriv{2}{
    \rm of & \rm the \\
    \uline{1} & \uline{1} \\
    \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{1}{\cf{NP/N}} \\
    \fcomp{2} \\
    \mc{2}{\cf{(NP\bs NP)/N}} \\
  }
\end{center}

Here we use \ccg forwards composition to combine \textrm{of} and \textrm{the} into a constituent-forming analysis. This chart structure could be reused to construct a span of the original phrase in the following manner:

\begin{center}
  \small
  \deriv{3}{
    \rm of & \rm the & \rm potato \\
    \uline{1} & \uline{1} & \uline{1} \\
    \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{1}{\cf{NP/N}} & \mc{1}{\cf{N}} \\
    \fcomp{2} & \\
    \mc{2}{\cf{(NP\bs NP)/N}} & \\
    \fapply{3}\\
    \mc{3}{\cf{NP\bs NP}}\\
  }
\end{center}

One situation where this technique does not work is is with \emph{prepositional phrase attachment}. The correct \ccg derivation for the phrase \textrm{on the king of England} is

\begin{center}
  \small
  \deriv{6}{
    \rm X & \rm on & \rm the & \rm king & \rm of & \rm England \\
    \uline{1} & \uline{1} & \uline{2} & \uline{2}  \\
    \mc{1}{\cf{NP}} & \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{2}{\cf{NP[nb]}} & \mc{2}{\cf{NP\bs NP}} \\
    & & \bapply{4}\\
    & & \mc{4}{\cf{NP}}\\
    & \fapply{5} \\
    & \mc{5}{\cf{NP\bs NP}} \\
    \bapply{6} \\
    \mc{6}{\cf{NP}} \\
  }
\end{center}

If we were to use in this example the same forward composed derivation of the bigram \textrm{of the} as described earlier for the bigram \textrm{on the}, the wrong analysis would be constructed.

\begin{center}
  \small
  \deriv{6}{
    \rm X & \rm on & \rm the & \rm king & \rm of & \rm England \\
    \uline{1} & \uline{2} & \uline{1} & \uline{2}  \\
    \mc{1}{\cf{NP}} & \mc{2}{\cf{(NP\bs NP)/N}} & \mc{1}{\cf{N}} & \mc{2}{\cf{NP\bs NP}} \\
    & \fapply{3} & & \\
    & \mc{3}{\cf{NP\bs NP}} & & \\
    \bapply{4} & & \\
    \mc{4}{\cf{NP}} & & \\
    \bapply{6} \\
    \mc{6}{\cf{NP}} \\
  }
\end{center}

% =============================================================================
\section{Implementation}
The \candc parser \cite{clark-curran:07} is a state of the art \ccg parser which uses \cky. The \candc parser was used to memoise constituent-forming \ngrams into databases (Figure~2). When parsing sentences in the future, the memoised databases can be used to insert pre-constructed parses straight into the chart (Figure~3).

\begin{figure}
  \pgfimage[width=0.5\columnwidth]{images/prebuilt1}
  \caption{When creating the trigram database, a trigram is added to the database if it forms a constituent in the chart}
  \label{fig:prebuilt1}
\end{figure}

\begin{figure}
  \pgfimage[width=0.7\columnwidth]{images/prebuilt2}
  \caption{Illustration of using the \ngram databases. The trigram {\tt B~C~D} is loaded from the pre-constructed database, and blocks the corresponding cells}
\end{figure}

% =============================================================================
\section{Results}
Empirical results for the memoisation of constituent-forming \ngrams can be seen in the table below.

\begin{itemize}
  \item Different grammar models were used over varying sized \ngrams, with the  parsing time, F-score, and coverage figures being recorded
  \item No statistically significant change was observed in the F-scores or parsing times
  \item These results backup the theoretical hypothesis that persisting only constituent-forming \ngrams would not be enough to get the one structure per \ngram idea to work in practice
\end{itemize}

\begin{table}
  \small
  \begin{tabular}{cc||c|cc}
    \hline
    Model & & Baseline & 2-gram & 3-gram \\
    \hline \hline
           &  Time  & 82.82   & 82.83   & 82.57  \\
    \wsj   &   LF   & 85.26   & 85.26   & 85.26  \\
    derivs &   UF   & 92.01   & 92.01   & 92.01  \\
           &   Cov  & 98.64   & 98.64   & 98.64  \\
%%%    \hline
%%%           &  Time  & 84.15   & 83.83   & 83.87  \\
%%%    \wsj   &   LF   & 87.40   & 87.40   & 87.40  \\
%%%    hybrid &   UF   & 93.10   & 93.10   & 93.10  \\
%%%           &   Cov  & 98.64   & 98.64   & 98.64  \\
%%%    \hline
%%%           &  Time  & 84.10   & 83.96   & 84.26  \\
%%%    \np    &   LF   & 83.60   & 83.60   & 83.60  \\
%%%    derivs &   UF   & 90.48   & 90.48   & 90.48  \\
%%%           &   Cov  & 98.54   & 98.54   & 98.54  \\
%%%    \hline
%%%           &  Time  & 84.82   & 84.88   & 85.32  \\
%%%    \np    &   LF   & 85.88   & 85.88   & 85.88  \\
%%%    hybrid &   UF   & 91.63   & 91.63   & 91.63  \\
%%%           &   Cov  & 98.54   & 98.54   & 98.54  \\
    \hline
  \end{tabular}
  \caption{The speed versus performance trade-off for varying sized \ngrams on \ccgbank 02 to 21}
  \label{tab:results}
\end{table}

%%%Experiments were also conducted using only non-constituent-forming \ngrams over the same datasets. However the outcome of these experiments were the same; no statistically significant change was observed.

% =============================================================================
\section{Conclusion}
Thorough analysis of this one structure per \ngram idea using \ccg as a grammar formalism combined with a set of empirical results has shown that this one structure per \ngram idea does not work in the general when using token based \ngrams.

% =============================================================================
\section{Future work}
\begin{itemize}
  \item Recent work has been undertaken to explore the one structure per \ngram idea using \ngrams constructed of supertags instead of tokens
  \item The initial results for this new direction are positive, showing show a 10\% parser speedup being achieved with minimal loss of F-score
\end{itemize}

% =============================================================================
\references
\bibliography{references}

\end{multicols}
\end{document}

### 2014-08-28 code/weka/run_weka.sh
Updated weka script to take arbitrary config

@@ -1,0 +1,57 @@
+#!/bin/bash
+
+# A script to automate the running of Weka on a given dataset.
+#
+# Given an ARFF file, train and test different classifiers on it and write the
+# results to a file (whose filename contains the current timestamp) for closer
+# examination after the run.
+#
+# The classifiers to be tested are loaded in a separate file, whose name is
+# specified as the second filename on the command line.
+
+# Weka configuration
+weka_cp=~/Downloads/weka-3-7-11/weka.jar
+common_options="-t $1" # options common to all weka classifiers
+jvm_heapsize=1024m
+
+# Output file configuration
+prefix=weka_result
+suffix=`date +%Y%m%d%H%M%S`
+separator="########################################"
+outdir=../../data/results/
+outfile=$outdir$prefix$suffix
+
+# if we didn't specify an input file name to this script, exit
+if [ -z $1 ]; then
+  echo "Please specify an input ARFF file. Exiting."
+  exit 1
+fi
+
+# if we didn't specify a config file
+if [ -z $2 ]; then
+  echo "Please specify a list of weka classifiers. Exiting."
+  exit 2
+fi
+
+# read all lines of weka config
+declare -a clfs
+readarray -t clfs < $2
+
+cat $2 >> $outfile
+
+# run all classifiers from a configuration file
+num_lines=${#clfs[@]}
+i=0
+while [ $i -lt $num_lines ]
+do
+  echo "Training classifier $i: ${clfs[i]}..."
+  java -Xmx$jvm_heapsize -cp $weka_cp ${clfs[i]} $common_options >> $outfile
+  for j in {0..1}
+  do
+    echo $separator >> $outfile
+  done
+  i=$[$i+1]
+done
+
+echo "Complete."
+exit 0

### 2014-08-30 presentation/example.tex
Added template for hons presentation

\documentclass{beamer}
\usepackage{graphicx}
\usepackage{lmodern}

% Load the usyd template
\input{usyd-beamer/usyd-beamer.tex}

\begin{document}

\usydtitleframe{Title}{Subtitle}{Author information}

\usydtitleframe
{Title goes here, this one is heaps long and wraps around}
{
	\vspace{3mm} % we want a little bit more room...
	Subtitle goes here, this is pretty long too hey? But not quite as long...
} {
	\textbf{Firstname Lastname} \hfill Student \\
	SID: 123 456 789 \\
	\vspace{3mm}
	\textbf{Firstname Lastname} \hfill Supervisor \\
	%\vspace{-12mm} % You might need to play here to get it positioned nicely...
}

\section{First Section}
\subsection{Subsection}
\begin{frame}{A sample slide}

A displayed formula:

\[ \int_{-\infty}^\infty e^{-x^2} \, dx = \sqrt{\pi} \]

An itemized list:

\begin{itemize}
  \item itemized item 1
  \item itemized item 2
  \item itemized item 3
\end{itemize}

\begin{theorem}
  In a right triangle, the square of hypotenuse equals
  the sum of squares of two other sides.
\end{theorem}

\end{frame}

\end{document}

### 2014-08-31 code/models/util.py
util removes samples that contain missing data

@@ -32,9 +32,10 @@ def strip_ws_normalise_headings(dir_path, infile_path, outfile_path):
                     row = [field.translate(table) for field in row]
                     headings = row
                 # strip whitespace in each item of the row
-                row = map(str.strip, row)
-                # write to file
-                csv_writer.writerow(list(row))
+                row = list(map(str.strip, row))
+                # write to file if all fields are available
+                if all(row):
+                    csv_writer.writerow(row)
 
 def separate_data(has_headings, *data_path):
     """Returns three lists, one with n samples and m features from the original

### 2014-08-31 code/weka/run_weka.sh
Updated result file output in run weka script

@@ -9,18 +9,6 @@
 # The classifiers to be tested are loaded in a separate file, whose name is
 # specified as the second filename on the command line.
 
-# Weka configuration
-weka_cp=~/Downloads/weka-3-7-11/weka.jar
-common_options="-t $1" # options common to all weka classifiers
-jvm_heapsize=1024m
-
-# Output file configuration
-prefix=weka_result
-suffix=`date +%Y%m%d%H%M%S`
-separator="########################################"
-outdir=../../data/results/
-outfile=$outdir$prefix$suffix
-
 # if we didn't specify an input file name to this script, exit
 if [ -z $1 ]; then
   echo "Please specify an input ARFF file. Exiting."
@@ -33,10 +21,24 @@ if [ -z $2 ]; then
   exit 2
 fi
 
+# Weka configuration
+weka_cp=~/Downloads/weka-3-7-11/weka.jar
+common_options="-t $1" # options common to all weka classifiers
+jvm_heapsize=1024m
+
+# Output file configuration
+prefix=weka_result
+timestamp=`date +%Y%m%d%H%M%S`
+suffix=$2 # indicates which config was used
+separator="########################################"
+outdir=../../data/results/
+outfile=$outdir$prefix$timestamp$suffix
+
 # read all lines of weka config
 declare -a clfs
 readarray -t clfs < $2
 
+echo $1 >> $outfile
 cat $2 >> $outfile
 
 # run all classifiers from a configuration file

### 2014-09-07 thesis/litreview/litreview.tex
Updated all cite commands in lit review

@@
 Hospitals often have limited funding, and therefore managing and utilising
-their resources efficiently has always been a priority\citep{Walczak2003}.
+their resources efficiently has always been a priority \cite{Walczak2003}.
 One key measurement of hospital activity, health care utilisation, and hospital
-cost is the patient length of stay\citep{Omachonu2004,Ng2006}. The length of
+cost is the patient length of stay \citep{Omachonu2004,Ng2006}. The length of
 stay has been used in a wide variety of medical domains, such as burns
  \citep{Yang2010}, intensive
-care\citep{Tu1993,Harper2005,Perez2006,Dybowski1996},
-psychiatry\citep{Lowell1997}, and acute pancreatitis\citep{Pofahl1998}, in
+care \citep{Tu1993,Harper2005,Perez2006,Dybowski1996},
+psychiatry \citep{Lowell1997}, and acute pancreatitis \citep{Pofahl1998}, in
 order to assist hospitals in allocating beds and other patient management
@@
 conditions that hospitals manage, the length of stay of a patient is
-not easy to predict\citep{Walczak2003}, especially upon a patient's admission.
+not easy to predict \citep{Walczak2003}, especially upon a patient's admission.
 Many methods have been explored in the literature, from the realm of statistics
@@
 One of the more well-understood and widely used statistical methods for
-predicting the length of stay is \textit{logistic regression}\citep{Tu1996}.
+predicting the length of stay is \textit{logistic regression} \citep{Tu1996}.
 Logistic regression models are straightforward to construct and use via the
@@
 prediction explicit. Thus they is easy to interpret when compared to more
-``black box'' approaches such as artificial neural networks\citep{Adams2012}.
+``black box'' approaches such as artificial neural networks \citep{Adams2012}.
 A wide variety of other statistical approaches have been proposed, notably:
-various scoring systems from univariate analysis\citep{Adams2012,Lavoie2005},
-linear regression\citep{Yang2010}, exponential models\citep{Clark2007},
-survival analysis\citep{Vasilakis2005}, and Markov
-models\citep{Perez2006,Jain1989,Kapadia2000}.
+various scoring systems from univariate analysis \citep{Adams2012,Lavoie2005},
+linear regression \citep{Yang2010}, exponential models \citep{Clark2007},
+survival analysis \citep{Vasilakis2005}, and Markov
+models \citep{Perez2006,Jain1989,Kapadia2000}.
 
@@
 early applications to length of stay prediction in intensive
-care\citep{Tu1993,Dybowski1996}, in addition to disease diagnosis and mortality
-prediction\citep{Silva2006}. Artificial neural networks are able to detect
+care \citep{Tu1993,Dybowski1996}, in addition to disease diagnosis and mortality
+prediction \citep{Silva2006}. Artificial neural networks are able to detect
 complex and non-linear relationships between input and output variables and can
@@
 neural networks to predict patient length of stay in various medical conditions
-such as acute pancreatitis\citep{Walczak2003}, gastroenteritis\citep{Ng2006},
-surgical intensive care\citep{Buchman1994,Tu1993} and
-psychiatry\citep{Lowell1997}. Other data mining techniques have also been
+such as acute pancreatitis \citep{Walczak2003}, gastroenteritis \citep{Ng2006},
+surgical intensive care \citep{Buchman1994,Tu1993} and
+psychiatry \citep{Lowell1997}. Other data mining techniques have also been
 applied, such as support vector machines and tree-based
-methods\citep{Harper2005}.
+methods \citep{Harper2005}.
 
@@
 predict the length of stay for trauma patients is especially crucial because of
-the scarcity of patient care resource in this area\citep{Yang2010}.
+the scarcity of patient care resource in this area \citep{Yang2010}.
 Additionally,
@@
 As we stated earlier, logistic regression is a commonly-used modelling
-technique for prediction in the medical domain\citep{Tu1996}, and especially
-in the trauma data analysis\citep{Gabbe2005}. A series of
+technique for prediction in the medical domain \citep{Tu1996}, and especially
+in the trauma data analysis \citep{Gabbe2005}. A series of
 predictor (independent) variables (or features) determine the probability of a
@@
 associated with each variable, and $X_i$ are the values of the independent
-variables\citep{Hosmer1989}. A stepwise selection algorithm is often used to
+variables \citep{Hosmer1989}. A stepwise selection algorithm is often used to
 choose which variables to include in the model out of all available variables:
@@
 that contributed to trauma patient length of stay and use these factors to
-subsequently predict length of stay\citep{Gabbe2005}. The prediction problem
+subsequently predict length of stay \citep{Gabbe2005}. The prediction problem
 they proposed was essentially a binary classification problem, to which
-logistic regression is well suited\citep{Tu1996}: will length of stay be
+logistic regression is well suited \citep{Tu1996}: will length of stay be
 \textit{less than 5 days}, or \textit{equal to 5 days}? They used a
@@
 patients could be predicted to stay in hospital \textit{less than 2 days} or
-\textit{greater than 2 days}\citep{Dinh2013a}.
+\textit{greater than 2 days} \citep{Dinh2013a}.
 They also used a logistic regression model and
@@
 to the outcome: Gabbe et al. included very few features specific to trauma
-such as place of injury and mechanism of injury\citep{Gabbe2005} whereas Dinh
+such as place of injury and mechanism of injury \citep{Gabbe2005} whereas Dinh
 et al. also had variables for certain kinds of injuries (such as if it was
@@
 patient in a surgical intensive care unit staying greater than a
-week\citep{Buchman1994}.
+week \citep{Buchman1994}.
 
@@
 
-Yang et al.\citep{Yang2010} used a linear regression model as the baseline for
+Yang et al. \citep{Yang2010} used a linear regression model as the baseline for
 performance benchmarking when trying to predict the length of stay of burn
@@
 be highly correlated with the dependent variable but not with each
-other\citep{Hall2000}.}. Unlike other studies into length of stay prediction,
+other \citep{Hall2000}.}. Unlike other studies into length of stay prediction,
 they also looked at predicting the length of stay at various stages of the
@@
 probability matrix using 4 states for a pediatric intensive care unit (ICU)
-application\citep{Kapadia2000}:
+application \citep{Kapadia2000}:
 \begin{center}
@@
 the data well and produced predictions that were very close to the observed
-results in the test set\citep{Kapadia2000}, indicating that using such models
+results in the test set \citep{Kapadia2000}, indicating that using such models
 justifies further investigation.
@@
 we will describe briefly.
-Yau et al.\citep{Yau2003} proposed a \textit{mixture regression model} in
+Yau et al. \citep{Yau2003} proposed a \textit{mixture regression model} in
 modelling
@@
 Clark et al. used a similar idea in their study of mortality and length of stay
-prediction\citep{Clark2002}, by employing \textit{piecewise exponential models}
+prediction \citep{Clark2002}, by employing \textit{piecewise exponential models}
 to model different probabilities of transition over distinct periods of time.
@@
 unsuspected relationships and to summarise the data in novel ways that are both
-understandable and useful to the data owner\citep{Hand2001}. Compared to
+understandable and useful to the data owner \citep{Hand2001}. Compared to
 traditional statistical, it is a relatively recent field that emerged in the
@@
 given researchers a variety of new tools to analyse and understand biomedical
-datasets\citep{Bellazzi2008,Yoo2012}. Bellazzi \& Zupan\citep{Bellazzi2008}
+datasets \citep{Bellazzi2008,Yoo2012}. Bellazzi \& Zupan \citep{Bellazzi2008}
 have written a
@@
 following their use in diagnosing acute myocardial infarction and various
-applications in radiology\citep{Baxt1995}.
+applications in radiology \citep{Baxt1995}.
 One of the earliest applications of neural networks to
 length of stay prediction was by Tu \& Guerriere, who used a neural network
-with the following structure\citep{Tu1993} to predict length of stay in
+with the following structure \citep{Tu1993} to predict length of stay in
 intensive care following cardiac surgery:
 \begin{figure}[h]
+\caption{}
 \centering
@@
 \item Neural networks are able to model arbitrarily complex relations between
-independent and dependent variables\citep{Tu1996,Buchman1994}, especially if
+independent and dependent variables \citep{Tu1996,Buchman1994}, especially if
 the relation is of a non-linear nature
@@
 \item Neural networks can be ``trained'' in different ways, which changes their
-predictive power -- this has been studied by Walczak et al.\citep{Walczak2003},
+predictive power -- this has been studied by Walczak et al. \citep{Walczak2003},
 discussed in the next section
@@
 inducing sensible rules from the model (even by domain
-experts)\citep{Bellazzi2008}. Additionally, they are also prone to overfitting
+experts) \citep{Bellazzi2008}. Additionally, they are also prone to overfitting
 and require substantially more computational power to train when compared to
 the standard logistic regression technique, and do not always perform
-better\citep{Harper2005}. A number of studies have attempted to address some of
+better \citep{Harper2005}. A number of studies have attempted to address some of
 these issues in length of stay prediction, building on from Tu \& Guerriere's
@@
 and neural networks on predicting whether or not patients would stay greater
-than 7 days in a surgical intensive care unit\citep{Buchman1994}. Instead of
+than 7 days in a surgical intensive care unit \citep{Buchman1994}. Instead of
 using one type of training algorithm as Tu \& Guerriere did, they obtained
@@
 which allows for greater modelling power and is able to produce continuous
-values instead of simply a binary classification\citep{Sprecht1991}. They found
+values instead of simply a binary classification \citep{Sprecht1991}. They found
 that all three neural networks performed noticeably better than the logistic
@@
 by applying a genetic algorithm approach to selecting the best
-structure\citep{Dybowski1996} --
+structure \citep{Dybowski1996} --
 an improvement from the ``trial and error'' method Tu \& Guerriere used. It is
@@
 in choosing an optimal neural network structure is as
-follows\citep{Dybowski1996}:
+follows \citep{Dybowski1996}:
 \begin{enumerate}
@@
 network structure in their study to predict length of stay in two hospital
-settings, acute pancreatitis and pediatric trauma\citep{Walczak2003}. Instead
+settings, acute pancreatitis and pediatric trauma \citep{Walczak2003}. Instead
 of a genetic algorithm approach, they generated and tested every possible
@@
 selecting a single optimal network, they used ensembles of neural networks
-to improve predictive power\citep{Rowan2007}. Many different combinations of
+to improve predictive power \citep{Rowan2007}. Many different combinations of
 trained neural networks were tested in order to find the combination with the
@@
 new data to fine-tune their connections. This has been worked on by Lowell et
-al. in predicting psychiatric length of stay\citep{Lowell1997}, and also by
+al. in predicting psychiatric length of stay \citep{Lowell1997}, and also by
 Ng et al. who proposed an incremental learning approach for neural networks
-and applied it to gastroenteritis length of stay data\citep{Ng2006}. Lowell et
+and applied it to gastroenteritis length of stay data \citep{Ng2006}. Lowell et
 al. trained an initial model in an earlier paper to predict the length of stay
@@
 neural network resulted in increased prediction accuracy of 3-15\% with a less
-than 10\% change in the training data\citep{Lowell1997}. Ng et al. took a more
+than 10\% change in the training data \citep{Lowell1997}. Ng et al. took a more
 theoretical approach and proposed a mixture of experts neural network with an
@@
 network's predictions were accurate to within a day), compared to the standard
-batch-mode learning method\citep{Ng2006}.
+batch-mode learning method \citep{Ng2006}.
 
@@
 classification algorithms on four different medical data sets, trying to
-model five different scenarios\citep{Harper2005}. Two of these outcomes were
+model five different scenarios \citep{Harper2005}. Two of these outcomes were
 relevant to our work: one was predicting the length of stay in the intensive
@@
 finding that Yang et al. also reported when predicting the length of stay of
-burn patients\citep{Yang2010}.
+burn patients \citep{Yang2010}.
 
@@
 machines (SVMs), perhaps the most powerful classification algorithm in terms of
-predictive accuracy\citep{Bellazzi2008}. SVMs work on the principle that when
+predictive accuracy \citep{Bellazzi2008}. SVMs work on the principle that when
 the data is represented in a high enough dimensional space, a linear hyperplane
 can be found where the margin between two different records is as large as
-possible. In figure \ref{svm} from Yoo et al.\citep{Yoo2012}, hyperplane 2
+possible. In figure \ref{svm} from Yoo et al. \citep{Yoo2012}, hyperplane 2
 is the better plane because its margin is as large as possible. 
@@
 \label{svm}
+\caption{}
 \centering
@@
 properties of the SVM, while being able to approximate the mapping between the
-input domain and the real numbers\citep{Yang2010}. Evaluating the performance
+input domain and the real numbers \citep{Yang2010}. Evaluating the performance
 of linear regression, M5 and SVM regression using the mean absolute error
@@
 predicting the length of stay for clinical
-treatment processes\citep{Huang2013}. Their approach assumes that similar
+treatment processes \citep{Huang2013}. Their approach assumes that similar
 patients have a similar length of

### 2014-09-20 code/weka/run_weka.sh
Run WEKA script now runs however many times the user specifies #10

@@ -10,17 +10,31 @@
 # specified as the second filename on the command line. If they require using
 # one classifier in another, make sure <OPTS> is specified at the correct place
 # to include the training file commands in the right place.
+#
+# An optional third argument to the script is to specify the number of runs
+# of the classifiers. This must be a positive integer, and defaults to 1.
 
-# if we didn't specify an input file name to this script, exit
+# If we didn't specify an input file name to this script, exit
 if [ -z $1 ]; then
   echo "Please specify an input ARFF file. Exiting."
   exit 1
 fi
 
-# if we didn't specify a config file
+# If we didn't specify a config file
 if [ -z $2 ]; then
   echo "Please specify a list of weka classifiers. Exiting."
   exit 2
+fi
+
+# Check if the number of runs has been specified
+if [ -z $3 ]; then
+  echo "Number of runs not specified; defaulting to 1."
+  RUNS=1
+else
+  if [ $3 -gt 0 ]; then
+    echo "Number of runs specified: $3."
+    RUNS=$3
+  fi
 fi
 
 # Weka configuration
@@ -28,31 +42,38 @@ WEKA_CP=~/Downloads/weka-3-7-11/weka.jar
 COMMON_OPTIONS="-t $1" # options common to all weka classifiers
 JVM_HEAPSIZE=1024m
 
-# Output file configuration
+# Output file base configuration
 PREFIX=weka_result
-TIMESTAMP=`date +%Y%m%d%H%M%S`
 CONFIG=$2 # indicates which config was used
 OUTDIR=../../data/results/
-OUTFILE=$OUTDIR$PREFIX$TIMESTAMP$CONFIG
 
 # Read all lines of weka config into an array
 declare -a CLFS
 readarray -t CLFS < $2
 
-echo $1 >> $OUTFILE
-cat $2 >> $OUTFILE
-
-# Run all classifiers from a configuration file, substituting the filename
-# in the correct place to overcome problems with supplying arguments to
-# weka classifiers via the command line.
-# sed is run twice to remove a possible double occurrence of $COMMON_OPTIONS
-NUM_LINES=${#CLFS[@]}
-LINENUM=0
-while [ $LINENUM -lt $NUM_LINES ]
[…]

### 2014-09-20 thesis/introduction/introduction.tex
Added structure of introduction
Started work on background of introduction #11

@@
 
-The introduction goes in here.
+\section{Background}
+Suppose that we have a (perhaps large) data set. It could consist of
+demographic data taken from a census, details of transactions in a shopping
+centre, or simply a collection of text documents. The term used to describe
+the methods of analysis used to find unknown relationships and summarise
+information in these data sets is \textit{data mining}. Broadly speaking,
+data mining can be broken down into five areas (or \textit{tasks}) depending
+on the aims of the person analysing the data:
+
+\begin{enumerate}
+\item Exploratory data analysis (EDA), which consists of simply exploring some
+given data set without any particular goal or ideas of what to look for.
+Methods in EDA usually involve interaction and visualisation, but this can
+become difficult if there are too many data points;
+\item Descriptive modelling, as its name suggests, seeks to describe the data
+or the processes that generate it. The major areas here are clustering (finding
+partitions of the data) and density estimation (finding the probability
+distribution of the data);
+\item Predictive modelling aims to create models that will allow the value of
+some desired quantity or metric to be predicted based on some known or observed
+values in the data set. The two major sub-areas of predictive modelling are
+\textit{classification} and \textit{regression}, distinguished only by the
+nature of the quantity we want to predict. If we want to predict whether
+something belongs in a category, such as the disease of a patient, then we
+are dealing with classification; otherwise, if we wish to predict a numeric
+quantity, such as the value of the stock market at some future data, then
+it is a regression problem;
+\item Pattern and rule discovery, which covers anomaly detection (finding data
+points which do not fit an expected trend, like detecting credit card fraud)
+and association rule learning (discovering relationships between various items
+in transaction databases, such as the food items purchased frequently together
+in a supermarket);
+\item Content retrieval, where the user wishes to find patterns in a data set
+that are similar to the one he or she specifies. Search engines are a familiar
+application of this area of data mining.
+\end{enumerate}
+
+\subsection{Problem Statement}
+
+\subsection{Previous Work}
+
+\section{Contributions}
+
+\section{Results and Evaluation}

### 2014-09-21 thesis/introduction/introduction.tex
Finished background and problem statement #11

@@
 
+Data mining overlaps substantially with machine learning, a sub-field of
+artificial intelligence (AI) dedicated to the study of systems that can learn
+from data rather than executing explicitly programmed instructions. The
+techniques and issues discussed in this thesis are equally relevant in both
+areas, so we will not attempt to define a clear distinction here.
+
+In this thesis we will focus on the task of predictive modelling. Specifically,
+our problem will be a classification one. We will introduce the necessary
+terminology in the next section as we describe the problem.
+
 \subsection{Problem Statement}
+Consider a data set of patients admitted to the trauma ward of a hospital.
+This data set looks like a very large table: there are many columns, each one
+describing some aspect of a patient (for example, age); there are also many
+rows, each corresponding to a particular patient that was admitted. We will
+refer to the columns as \textit{features} or \textit{attributes}, so-named
+because they describe some particular aspect of the rows, which we will call
+\textit{samples}. Concretely, as an example, our data set could contain
+two features (age and gender) for five patients: as a table, this means that
+we have two columns, one for age and another for gender, with five rows
+containing the age and gender data for each patient. Attributes can be numeric
+(such as age) and potentially take on an infinite number of different values,
+or categorical (such as gender) and can only have one of a limited, pre-defined
+list of values.
+
+In a general predictive modelling problem, we wish to predict a value for some
+desired quantity given some values for the attributes. In our problem, we wish
+to predict the length of stay (LOS) for a patient admitted to a hospital trauma
+ward. To do this, we use our data set of trauma patients with \textit{known}
+values for the LOS and apply various predictive modelling techniques to deduce
+an effective model that will be able to predict the LOS of patients in advance.
+Note that the LOS is simply another attribute of our data set; however, because
+it is the quantity that we are interested in predicting, it is also called the
+\textit{target} or \textit{outcome} variable.
+
+Like the other attributes, the target or outcome variable can also be numeric
+or categorical. This determines the sub-task of predictive modelling that we
+will subsequently be dealing with: regression if the target variable is
+numeric, and classification if it is categorical. The various values that the
+categorical target variable can take on are also called \textit{classes}.
+Our problem is one of classification, divided into two sub-problems:
+
+\begin{enumerate}
+\item The first problem is a \textit{binary} classification task in which we
+predict one of two possible outcomes (or classes) for the patient: LOS less
+than 2 days, or greater than 2 days. Given this, what model performs best?
+\item The second problem is a \textit{multi-class} classification in which we
+define several possible outcomes with different ranges for the LOS and predict
+which range the patient will fall into. An example of possible outcomes can be
+1-3 days, 4-10 days, and 11 or more days, giving us three classes for the
+target variable. The optimal division of LOS into ranges will be investigated
+as part of the problem.
 

### 2014-09-22 thesis/thanks/thanks.tex
Added acknowledgements

\chapter*{Acknowledgements}

I would like to thank my supervisor, Dr Irena Koprinska, for her unwavering
support throughout the year. Irena, thank you for always being willing to
listen, and for raising points that inspired me to pursue new directions in
my research.

I would also like to thank Dr Michael Dinh from the Royal Prince Alfred
Hospital, without whom the research in this thesis would not have been
possible. Thank you,
Michael, for providing access to the data we needed for this study, and for
patiently answering all my questions about the data set and your previous
experiments.

Thank you to my close friends Daniel Buckmaster, Stephen Caraher and Christine
Au Yeung for the selfless encouragement, help and sanity checks that have
made this year-long journey possible.

And finally, but no less significant -- to my parents Caixiu and Ning, and
my sister Lily -- thank you for tirelessly supporting me throughout this year
and in everything that I do.

### 2014-09-22 thesis/conclusion/conclusion.tex
Updated heading of conclusion to 'conclusions'

\chapter{Conclusions} \label{chap:conclusion}

The conclusion goes here.

### 2014-09-22 thesis/introduction/introduction.tex
Finished a preliminary version of the introduction #11

@@
 the methods of analysis used to find unknown relationships and summarise
-information in these data sets is \textit{data mining}. Broadly speaking,
+information in these data sets is \textit{data mining} \citep{Hand2001}.
+Broadly speaking,
 data mining can be broken down into five areas (or \textit{tasks}) depending
@@
 
-In this thesis we will focus on the task of predictive modelling. Specifically,
+In this thesis we will focus on the task of predictive modelling, also
+commonly known as \textit{supervised learning}. Specifically,
 our problem will be a classification one. We will introduce the necessary
@@
 
-\subsection{Problem Statement}
+\section{Problem Statement}
 Consider a data set of patients admitted to the trauma ward of a hospital.
@@
 as part of the problem.
+\end{enumerate}
 
-\subsection{Previous Work}
+The LOS is an important measure of resource utilisation in hospitals around
+the world \citep{Ng2006},
+and many studies have been carried out to investigate whether the
+LOS of a patient in various medical domains -- after surgery, burns, or
+intensive care treatment to name a few -- can be predicted accurately.
+Regardless of the
+specifics of the problem, accurate LOS prediction results in better planning
+and management of limited resources such as staff, equipment and medical
+supplies. This is particularly critical as hospitals often face the problem
+of limited funding \citep{Walczak2003}.
+As we will see in the next chapter, there has been limited
+application of a wide range of data mining techniques to predict the LOS in
+various medical domains, including trauma. Our work will aim to fill this
+gap by providing insights on the application of data mining techniques to
+LOS prediction.
 
 \section{Contributions}
+The work presented in this thesis aims to address several gaps in the current
+research into LOS prediction, by building upon the work of Dinh et. al.
+\citep{Dinh2013a}.
+We apply and assess a range of techniques from different data mining paradigms
+to a data set of trauma patients, and investigate whether formulating the LOS
+prediction problem differently yields different results. The insights gained
+from applying various data mining techniques will lead to a novel method of
+predicting the LOS which will yield reliable and accurate results and support
+the decision-making of medical professionals.
 
-\section{Results and Evaluation}
+\section{Thesis Outline}
+This thesis is structured as follows. Chapter \ref{chap:litreview} describes
+the major methods in predictive modelling, interweaved with an analysis of
+past work on LOS prediction. We explore in detail the major algorithms,
+summarise the current state of research, and identify where our work fits
+within this area.
+
+Chapter \ref{chap:experiments} describes the experimental setup, detailing
+each step in the process from data pre-processing to model creation and
+finally to performance evaluation.
+
+After the methods and setup are described, the results are presented and
+discussed in Chapter \ref{chap:results}. We conclude in Chapter
+\ref{chap:conclusion}, with indications for future work in this area.

### 2014-09-22 thesis/experiments/experiments.tex
Started writing up experimental setup

\chapter{Experimental Setup} \label{chap:experiments}

In this chapter we detail the steps carried out as part of our investigation.
We first outline the nature of the data set, describe how it was cleaned and
pre-processed before it was used to train various classifiers. The methods of
evaluation are then presented, as these are critical to assessing the
performance of the models created.

\section{Trauma patient data set}

\subsection{Collection}
The data set we used consisted of trauma registry data from the trauma centre
at the Royal Prince Alfred Hospital, a major trauma centre in New South Wales,
Australia. It covers all adult (age 15 and over) inhospital admissions to the
trauma centre from 2007--2011.

All patients are first admitted to the trauma ward until discharged
or transferred to an appropriate unit within the hospital.

\subsection{Characteristics}

\section{Data pre-processing}

\subsection{Cleaning} % missing values, irrelevant features

\subsection{Normalisation}

\subsection{Feature selection}

\section{Model derivation}

\subsection{Single classifier}

\subsubsection{Parameter tuning}

\subsection{Ensemble classifiers}

\section{Performance evaluation}

\subsection{Metrics}

\subsection{Cross-validation}

\subsection{Statistical tests}

### 2014-09-24 thesis/results/results.tex
Updated results title

\chapter{Results and Discussion} \label{chap:results}

The results go here.

### 2014-09-24 thesis/experiments/experiments.tex
Wrote data preparation section of experimental setup

@@
 \subsection{Collection}
-The data set we used consisted of trauma registry data from the trauma centre
+The data set used consisted of trauma registry data from the trauma centre
 at the Royal Prince Alfred Hospital, a major trauma centre in New South Wales,
-Australia. It covers all adult (age 15 and over) inhospital admissions to the
+Australia. It covered all adult (age 15 and over) inpatient admissions to the
 trauma centre from 2007--2011.
 
-All patients are first admitted to the trauma ward until discharged
-or transferred to an appropriate unit within the hospital.
+All patients were first admitted to the trauma ward until discharged
+or transferred to an appropriate unit within the hospital. A single trained
+data manager recorded a variety of attributes about the admitted patient,
+such as age, gender, blood pressure, mechanism of injury and body regions
+that were injured. We received this data in a Microsoft Excel spreadsheet.
 
 \subsection{Characteristics}
+There were 2546 patient records in the data set we received, comprising of 74
+features, one of which was the target variable \textbf{LOS48}. LOS48 was a
+binary variable -- that is, it could only take two values, 0 or 1 -- with 1
+indicating that the patient stayed two days or less, and 0 indicating a stay
+of greater than 2 days.
 
-\section{Data pre-processing}
+\todo{Perhaps insert some more information about the data itself, such
+as the various attributes and their data types? Not sure if that should go
+here, in the results, or in an appendix -- it is relevant but can disturb
+the flow of the main text.}
 
-\subsection{Cleaning} % missing values, irrelevant features
+\section{Data preparation}
+In any data mining task, adequately preparing the data ensures the best
+chance of success: this includes deciding how to deal with missing and
+inaccurate values, as well as appropriately transforming the data
+\citep{Witten2005} and selecting or constructing the most suitable features
+for the task \citep{Kotsiantis2006}.
+In this section we describe how we addressed each of
+the major issues and the reasons for our choice.
 
-\subsection{Normalisation}
+\subsection{Cleaning and normalisation} % missing values, irrelevant features
+Since the data was initially in an Excel spreadsheet, we explored the data
+by hand using Microsoft Excel's filter function on each column and
+consolidated many categorical values simply
+by correcting spelling and punctuation. For example:
+\begin{itemize}
+\item In the \textbf{sex} attribute, there were values of `Female' and
+`Femal' which were consolidated into `Female'.
+\item In the \textbf{mechanism} attribute, which was textual and indicated
+the category of how the patient was injured, there were such categories as
+`Other Vehi' and `Other Vehicle' as well as `Pedal Cycl' and `Pedal Cyclist'.
+Such categories were combined into `Other Vehicle' and `Pedal Cyclist'
+respectively.
+\item Some categories in \textbf{mechanism} were also divided into whether or
+not the mechanism of injury was self-inflicted, such as `Shooting' and
+`Shooting-selfinflicted'. When we asked the domain expert, Michael Dinh, his
+suggestion was to combine them into one category: in this case, simply
+`Shooting'; this was done for all the categories in this attribute.
+\item \textbf{disp from ED} indicated the section of the hospital that the
+patient was discharged to. This was also a textual categorical attribute.
+There were redundancies like `G. ICU' and `G.ICU' and even `GICU', which we
+all combined into one category, `GICU'.
+\end{itemize}
+
+The data was then exported into CSV (comma-separated value) format, so that
+we could continue to process it programmatically rather than manually. Many
+fields in the data set contained leading or trailing whitespace, which was
+removed with a few lines of Python code (see appendix \todo{[add reference
+to appendix]}). All attribute names were converted to lowercase. In the case
+where attribute names consisted of several words, the spaces were replaced
+with underscores (\_) to ensure that the full attribute name was read.
+\todo{Include a list of these converted attribute names? In the appendix?}
+
+In order for our results to be comparable to that of Dinh
+et. al. \citep{Dinh2013a}, we removed all instances with missing values,
+leaving 2517 instances with which to use to create our prediction models.
 
 \subsection{Feature selection}
+After data preparation, selecting a subset of features or attributes to use
+in a model is also important, and helps to reduce noise and redundant
+attributes from becoming part of the final model. Experiments have found that
+the performance of various classification methods such as decision trees
+\textit{deteriorates} in the presence of useless attributes.
+Our data set contains 73
+non-target attributes, which is a fairly high number. Many of these attributes
+are derived from others in the set, such as \textbf{age} and \textbf{age65}.
 
-\section{Model derivation}
+We began the process of feature selection by removing attributes that were
+not available upon admission, as these tended to be highly related to the
+LOS48 outcome and were a consequence, not a predictor, of LOS. These attributes
+were: iculos, outcome, disp from ED, disp from hosp, died, rehab and los.
+The id attribute was also removed because it simply numbered each record and
+indicated nothing about the patient themself. This left us with 67 attributes.
+
+The work of actually selecting the attributes out of the 67 available was
+carried out in two ways: automated using feature selection methods, and
+manually using advice from a domain expert.
+
+\subsubsection{Automated feature selection methods}
+There are a number of methods that are commonly used to automate the selection
+of a subset of features for the final classifier. The methods we used were:
+\begin{itemize}
+\item \textit{Correlation-based feature selection} \citep{Hall2000},
+which evaluates the `merit' of a
+subset of features by considering their degree of correlation with the target
+variable compared to their degree of correlation amongst each other. High
+correlation between variables indicates redundancy. Under
+this scheme, a set of features that are highly correlated with the target
+while having low correlation with other features in the set is ideal.
+\item Selection based on the
+\textit{Pearson product-moment correlation coefficient}, a measure of the
+linear correlation or dependence between two variables.\todo{add some more info}
+\item \textit{Information gain} \todo{explain a bit about this}
+\item Feature selection based on the \textit{OneR classifier}
+\citep{Holte1993}, which uses one
+attribute as the predictor for the target. This attribute is typically the one
+that has the lowest prediction error. Feature selection using this scheme
+involves first ranking attributes based on their error, with the lowest error
+attributes as the most desirable. We then selected the top 20 attributes to
+use in the building of the model.
+\end{itemize}
+
+We used the WEKA \citep{Hall2009} implementation of the above methods, after
+converting the data into the Attribute-Relation File Format
+(ARFF)\footnote{\url{http://weka.wikispaces.com/ARFF+\%28developer+version\%29}}
+most suitable for the program.
+\todo{give more details, ie specifying binary/numeric
+attributes, maybe specify in appendix?}
+
+\subsubsection{Expert selection}
+We also wanted to compare the feature selection of an expert in the domain
+of the data set, namely trauma, with the models obtained from the features
+selected using the automated methods described above. This resulted in two
+feature subsets. The first was obtained simply by presenting the entire list
+of 73 features to the domain expert and asking him to select the features he
+thought were most indicative of the target variable from experience. The
+second set was obtained from the work of Dinh et. al. \citep{Dinh2013a} that
+we are building on, giving us a baseline to compare our results to.
+
+\section{Model construction}
 
@@
 
-\section{Performance evaluation}
+\section{Performance evaluation} % don't forget cross-validation
 
@@
 
-\subsection{Cross-validation}
-
 \subsection{Statistical tests}

### 2014-09-25 thesis/experiments/experiments.tex
Wrote more of the experimental setup

@@
 
-We used the WEKA \citep{Hall2009} implementation of the above methods, after
-converting the data into the Attribute-Relation File Format
+We used the WEKA 3.7.11 \citep{Hall2009} implementation of the above
+methods, after converting the data into the Attribute-Relation File Format
 (ARFF)\footnote{\url{http://weka.wikispaces.com/ARFF+\%28developer+version\%29}}
@@
 
+Altogether, the various feature selection methods resulted in a total of
+7 data sets (4 automated, 2 expert selected, and 1 with all features) which
+we used in the model construction phase.
+
 \section{Model construction}
+Here we present the various classification schemes we used in order to find
+the best model for the trauma LOS data.
 
 \subsection{Single classifier}
+First we sought to compare the suitability and performance of various
+classifiers by themselves (later we looked at \textit{ensembles} of
+classifiers, when more than one is involved in making a prediction).
+These are the
+classifiers studied: \todo{expand a bit more here, maybe list the
+configuration parameters}
+\begin{itemize}
+\item Na\"{i}ve Bayes
+\item Decision tree (C4.5)
+\item Logistic regression
+\item Support vector machine
+\item $K$-nearest neighbour
+\item Feed-forward neural network (multilayer perceptron)
+\end{itemize}
+
+Each classifier was trained using ten-fold cross-validation for each of the
+7 data sets obtained from the various feature selection methods described
+above. This was repeated 30 times for each classifier and the reasons for
+this, as well as the metrics used to assess the performance of these
+classifiers, are described in Section \ref{sec:perfeval} of this chapter.
+
+In addition to the feature selection methods, we also used the WEKA 3.7.11
+\citep{Hall2009} implementation of the above classifiers for our experiments.
+Training was automated using a Bash shell script (whose source can be found
+in the appendix) that programmatically read in a list of classifiers to
+train and repeated the procedure any specified number of times, writing the
+results to a specified directory for later reference.
 
 \subsubsection{Parameter tuning}
+It is difficult in data mining to apply an algorithm out-of-the-box and have
+it perform optimally on a given data set \citep{Witten2005}. Therefore, time
+must be spent tweaking and tuning the parameters of a classifier in order to
+improve classification performance. This was the next step we took. We
+considered a range of values for various parameters of each classification
+scheme, and tried some of these new configurations by hand and some by using
+WEKA's cross-validation parameter selection class. The configurations tested,
+as well as how they were executed, are listed in this table:
+\todo{insert table!}
 
@@
 \section{Performance evaluation} % don't forget cross-validation
+\label{sec:perfeval}
 

### 2014-09-26 poster/myposter.tex
Updated poster template to the latest style #14

@@
-\documentclass[portrait]{usydposter}
-\usepackage{LI}
+\documentclass[landscape]{usydposter}
 \usepackage{xspace}
@@
 
-\title{CCG Parsing with One Structure per n-gram}
-\author{Tim Dawborn}
+\title{Improved Prediction of Hospital Length of Stay for Severe Injury}
+\author{Tianyu Pu \texttt{tianyu.pu@sydney.edu.au}}
 
 \begin{document}
+
 \makeheader
@@
 % =============================================================================
-\section{Heading RGB 206, 17, 38 -- Arial 28}
+\section{Heading RGB 0, 152, 219 -- Arial 28}
+\subsection{Subheading black -- Arial 24}
 \noindent This is some text here...
@@
 \section{Background}
-\begin{itemize}
-  \item \cky is a $O(n^3)$ parsing algorithm which uses a chart data structure, illustrated in Figure~1.
-A chart is a hierarchical triangular structure  which stores the parse trees for all substrings of the sentence being parsed
-\end{itemize}
-
-\begin{figure}
-  \pgfimage[width=0.7\columnwidth]{images/chart}
-  \caption{An illustration of the chart data structure used in parsing algorithms such as \cky}
-\end{figure}
-
-\begin{itemize}
-  \item Combinatory Categorial Grammar (\ccg) \cite{steedman:00} is a \emph{lexicalised grammar formalism}
-  \item Each word in a sentence is assigned a composite object that reflects its role within the sentence
-  \item The following is an example derivation using \ccg, where the lexical categories for each token are shown
-\end{itemize}
-
-\begin{center}
-  \small
-  \deriv{4}{
-    \rm I & \rm like & \rm the & \rm cat \\
-    \uline{1} & \uline{1} & \uline{1} & \uline{1} \\
-    \mc{1}{\cf{NP}} & \mc{1}{\cf{(S[dcl]\bs NP)/NP}} & \mc{1}{\cf{NP[nb]/N}} & \mc{1}{\cf{N}} \\
-    & & \fapply{2} \\
-    & & \mc{2}{\cf{NP[nb]}} \\
-    & \bapply{3}\\
-    & \mc{3}{\cf{S[dcl]\bs NP}}\\
-    \bapply{4}\\
-    \mc{4}{\cf{S[dcl]}}\\
-  }
-\end{center}
 
@@
 \section{Problems with Pizza}
-\emph{Syntactic ambiguity} is a problem when parsing, and is a particular problem for our one structure per \ngram problem. Consider the sentence
-
-\begin{center}
-  \small
-  \deriv{5}{
-    \rm I & \rm ate & \rm pizza & \rm with & \rm anchovies \\
-    \uline{1} & \uline{1} & \uline{1} & \uline{1} & \uline{1} \\
-    \mc{1}{\cf{NP}} & \mc{1}{\cf{(S\bs NP)/NP}} & \mc{1}{\cf{NP}} & \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{1}{\cf{NP}} \\
-    & & & \fapply{2} \\
-    & & & \mc{2}{\cf{NP\bs NP}} \\
-    & & \bapply{3}\\
-    & & \mc{3}{\cf{NP}}\\
-    & \fapply{4} \\
-    & \mc{4}{\cf{S\bs NP}} \\
-    \bapply{5}\\
-    \mc{5}{\cf{S}}\\
-  }
-\end{center}
-
-and compare it to the sentence
-
-\begin{center}
-  \small
-  \deriv{5}{
-    \rm I & \rm ate & \rm pizza & \rm with & \rm cutlery \\
-    \uline{1} & \uline{1} & \uline{1} & \uline{1} & \uline{1} \\
-    \mc{1}{\cf{NP}} & \mc{1}{\cf{(S\bs NP)/NP}} & \mc{1}{\cf{NP}} & \mc{1}{\cf{((S\bs NP)\bs(S\bs NP))/NP}} & \mc{1}{\cf{NP}} \\
-    & \fapply{2} & \fapply{2} \\
-    & \mc{2}{\cf{S\bs NP}} & \mc{2}{\cf{(S\bs NP)\bs(S\bs NP)}} \\
-    & \bapply{4} \\
-    & \mc{4}{\cf{S\bs NP}} \\
-    \bapply{5}\\
-    \mc{5}{\cf{S}}\\
-  }
-\end{center}
-
-These two sentences differ by only one word, yet they have two completely different derivations due to the different roles the word \textrm{with} plays in each sentence.
-
-The word \textrm{with} thus is syntactically ambiguous. Without knowing the surrounding context, it is unknown which grammatical role the word is undertaking.
 
@@
 
-\begin{center}
-  \small
-  \deriv{3}{
-    \rm of & \rm the & \rm potato \\
-    \uline{1} & \uline{1} & \uline{1} \\
-    \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{1}{\cf{NP/N}} & \mc{1}{\cf{N}} \\
-    & \fapply{2}\\
-    & \mc{2}{\cf{NP}} \\
-    \fapply{3}\\
-    \mc{3}{\cf{NP\bs NP}}\\
-  }
-\end{center}
-
 \textrm{The} is joined to \textrm{potato} before \textrm{of} can be joined with \textrm{the}. Instead, we could construct the following derivation and insert it into the pre-constructed database:
-
-\begin{center}
-  \small
-  \deriv{2}{
-    \rm of & \rm the \\
-    \uline{1} & \uline{1} \\
-    \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{1}{\cf{NP/N}} \\
-    \fcomp{2} \\
-    \mc{2}{\cf{(NP\bs NP)/N}} \\
-  }
-\end{center}
 
@@
 
-\begin{center}
-  \small
-  \deriv{3}{
-    \rm of & \rm the & \rm potato \\
-    \uline{1} & \uline{1} & \uline{1} \\
-    \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{1}{\cf{NP/N}} & \mc{1}{\cf{N}} \\
-    \fcomp{2} & \\
-    \mc{2}{\cf{(NP\bs NP)/N}} & \\
-    \fapply{3}\\
-    \mc{3}{\cf{NP\bs NP}}\\
-  }
-\end{center}
-
 One situation where this technique does not work is is with \emph{prepositional phrase attachment}. The correct \ccg derivation for the phrase \textrm{on the king of England} is
-
-\begin{center}
-  \small
-  \deriv{6}{
-    \rm X & \rm on & \rm the & \rm king & \rm of & \rm England \\
-    \uline{1} & \uline{1} & \uline{2} & \uline{2}  \\
-    \mc{1}{\cf{NP}} & \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{2}{\cf{NP[nb]}} & \mc{2}{\cf{NP\bs NP}} \\
-    & & \bapply{4}\\
-    & & \mc{4}{\cf{NP}}\\
-    & \fapply{5} \\
-    & \mc{5}{\cf{NP\bs NP}} \\
-    \bapply{6} \\
-    \mc{6}{\cf{NP}} \\
-  }
-\end{center}
 
@@
 
-\begin{center}
-  \small
-  \deriv{6}{
-    \rm X & \rm on & \rm the & \rm king & \rm of & \rm England \\
-    \uline{1} & \uline{2} & \uline{1} & \uline{2}  \\
-    \mc{1}{\cf{NP}} & \mc{2}{\cf{(NP\bs NP)/N}} & \mc{1}{\cf{N}} & \mc{2}{\cf{NP\bs NP}} \\
-    & \fapply{3} & & \\
-    & \mc{3}{\cf{NP\bs NP}} & & \\
-    \bapply{4} & & \\
-    \mc{4}{\cf{NP}} & & \\
-    \bapply{6} \\
-    \mc{6}{\cf{NP}} \\
-  }
-\end{center}
-
 % =============================================================================
 \section{Implementation}
-The \candc parser \cite{clark-curran:07} is a state of the art \ccg parser which uses \cky.
-The \candc parser was used to memoise constituent-forming \ngrams into databases (Figure~2).
-When parsing sentences in the future, the memoised databases can be used to insert pre-constructed parses straight into the chart (Figure~3).
-
-\begin{figure}
-  \pgfimage[width=0.5\columnwidth]{images/prebuilt1}
-  \caption{When creating the trigram database, a trigram is added to the database if it forms a constituent in the chart}
-  \label{fig:prebuilt1}
-\end{figure}
-
-\begin{figure}
-  \pgfimage[width=0.7\columnwidth]{images/prebuilt2}
-  \caption{Illustration of using the \ngram databases. The trigram {\tt B~C~D} is loaded from the pre-constructed database, and blocks the corresponding cells}
-\end{figure}
 
@@
 \end{itemize}
-
-\begin{table}
-  \small
-  \begin{tabular}{cc||c|cc}
-    \hline
-    Model & & Baseline & 2-gram & 3-gram \\
-    \hline \hline
-           &  Time  & 82.82   & 82.83   & 82.57  \\
-    \wsj   &   LF   & 85.26   & 85.26   & 85.26  \\
-    derivs &   UF   & 92.01   & 92.01   & 92.01  \\
-           &   Cov  & 98.64   & 98.64   & 98.64  \\
-%%%    \hline
-%%%           &  Time  & 84.15   & 83.83   & 83.87  \\
-%%%    \wsj   &   LF   & 87.40   & 87.40   & 87.40  \\
-%%%    hybrid &   UF   & 93.10   & 93.10   & 93.10  \\
-%%%           &   Cov  & 98.64   & 98.64   & 98.64  \\
-%%%    \hline
-%%%           &  Time  & 84.10   & 83.96   & 84.26  \\
-%%%    \np    &   LF   & 83.60   & 83.60   & 83.60  \\
-%%%    derivs &   UF   & 90.48   & 90.48   & 90.48  \\
-%%%           &   Cov  & 98.54   & 98.54   & 98.54  \\
-%%%    \hline
-%%%           &  Time  & 84.82   & 84.88   & 85.32  \\
-%%%    \np    &   LF   & 85.88   & 85.88   & 85.88  \\
-%%%    hybrid &   UF   & 91.63   & 91.63   & 91.63  \\
-%%%           &   Cov  & 98.54   & 98.54   & 98.54  \\
-    \hline
-  \end{tabular}
-  \caption{The speed versus performance trade-off for varying sized \ngrams on \ccgbank 02 to 21}
-  \label{tab:results}
-\end{table}
-
-%%%Experiments were also conducted using only non-constituent-forming \ngrams over the same datasets.
-However the outcome of these experiments were the same; no statistically significant change was observed.
 
@@
 \references
-\bibliography{references}
+\bibliography{../thesis/references}
 

### 2014-09-26 poster/myposter.tex
Updated structure of poster and tested citing references

@@
-\documentclass[landscape]{usydposter}
+\documentclass[portrait]{usydposter}
 \usepackage{xspace}
@@
 % =============================================================================
-\section{Heading RGB 0, 152, 219 -- Arial 28}
-\subsection{Subheading black -- Arial 24}
+\section{Aims}
 \noindent This is some text here...
-
-\begin{itemize}
-  \item Parsing is the process of deriving the grammatical structure for a given sentence
-  \item Natural language parsing is currently slow, with the best algorithms running in $O(n^3)$ with respect to the length of the sentence
-  \item There is an inherent \emph{redundancy} in natural languages where certain common phrases (or \ngrams) appear frequently in text, each time with the same grammatical structure
-  \item We explore the idea of exploiting this redundancy through the construction and \emph{memoisation} of the parse structures for these frequent \ngrams
-\end{itemize}
+\cite{Dinh2013a}
 
 % =============================================================================
-\section{Background}
+\section{Introduction}
 
 % =============================================================================
-\section{Problems with Pizza}
-
-% =============================================================================
-\section{Non-constituent-forming \ngrams}
-Memoisation of \ngrams which do not form constituents can be achieved through the use of {\ccg}'s \emph{composition} and \emph{type raising} combinators.
-
-\textrm{The} is joined to \textrm{potato} before \textrm{of} can be joined with \textrm{the}. Instead, we could construct the following derivation and insert it into the pre-constructed database:
-
-Here we use \ccg forwards composition to combine \textrm{of} and \textrm{the} into a constituent-forming analysis.
-This chart structure could be reused to construct a span of the original phrase in the following manner:
-
-One situation where this technique does not work is is with \emph{prepositional phrase attachment}. The correct \ccg derivation for the phrase \textrm{on the king of England} is
-
-If we were to use in this example the same forward composed derivation of the bigram \textrm{of the} as described earlier for the bigram \textrm{on the}, the wrong analysis would be constructed.
-
-% =============================================================================
-\section{Implementation}
+\section{Method}
 
@@
 \section{Results}
-Empirical results for the memoisation of constituent-forming \ngrams can be seen in the table below.
-
-\begin{itemize}
-  \item Different grammar models were used over varying sized \ngrams, with the  parsing time, F-score, and coverage figures being recorded
-  \item No statistically significant change was observed in the F-scores or parsing times
-  \item These results backup the theoretical hypothesis that persisting only constituent-forming \ngrams would not be enough to get the one structure per \ngram idea to work in practice
-\end{itemize}
 
 % =============================================================================
-\section{Conclusion}
-Thorough analysis of this one structure per \ngram idea using \ccg as a grammar formalism combined with a set of empirical results has shown that this one structure per \ngram idea does not work in the general when using token based \ngrams.
+\section{Discussion}
 
 % =============================================================================
-\section{Future work}
-\begin{itemize}
-  \item Recent work has been undertaken to explore the one structure per \ngram idea using \ngrams constructed of supertags instead of tokens
-  \item The initial results for this new direction are positive, showing show a 10\% parser speedup being achieved with minimal loss of F-score
-\end{itemize}
+\section{Conclusions and Future work}
 

### 2014-09-26 thesis/experiments/experiments.tex
Finished basic writeup of experimental setup #12

@@
 \subsection{Ensemble classifiers}
+In order to improve upon the performance of individual classifiers, we also
+tested the performance of ensembles of classifiers -- that is, combining the
+predictions of several different classifiers, or several of the same
+classifier. Specifically, the ensemble methods used were:
+\begin{itemize}
+  \item Random Forests
+  \item Bagging
+  \item Stacked generalisation (stacking)
+  \item Boosting (AdaBoostM1)
+\end{itemize}
+
+Again, we were able to create predictive models using the implementations of
+these ensemble methods in WEKA 3.7.11. Since ensemble classifiers are not
+actual classifiers but a way to train and combine the predictive power of more
+than one classifier, we list the combinations of classifiers used in our
+ensemble methods:
+\todo{insert table!}
 
@@
 \label{sec:perfeval}
+Here we describe the methods used in order to assess the performance of the
+models constructed through single classifiers and ensemble methods described
+above.
 
 \subsection{Metrics}
+There are many available and commonly used metrics that we can use when
+evaluating and comparing the performance of prediction models. The ones that
+we use are: \todo{maybe put these in a table}
+\begin{itemize}
+  \item Accuracy
+  \item Specificity
+  \item Sensitivity
+  \item Area under the receiver-operating characteristic (ROC) curve, or AUC
+\end{itemize}
+
+Since we only have one data set, we use ten-fold stratified cross-validation
+to obtain values for the above metrics. We run the ten-fold cross-validation
+40 times for each classifier (single or ensemble) for each of the 7 data sets
+produced from feature selection. 40 runs was necessary in order for us to
+compute confidence intervals for each of the metrics, which makes use of the
+Central Limit Theorem from statistics.
+
+\subsubsection{Confidence intervals}
+In order to compare our results with that of Dinh et. al. \citep{Dinh2013a},
+we are especially interested in the mean cross-validated AUC of our
+classifiers as well as
+the upper and lower bounds of the 95\% confidence interval associated with this
+mean. The intervals are constructed using this argument:
+
+For each classifier, let the population of all of its ten-fold cross-validated
+AUC scores be some random variable $X$, which follows a distribution whose
+parameters (such as mean and standard deviation) we do not know.
+The Central Limit Theorem states that given a random
+sample of $n$ items from $X$, namely $X_1,X_2,\dots,X_n$, the sample mean
+$\bar{x}$
+is approximately normally distributed when $n$ is at least roughly around 30
+and the $X_i$'s are independent and identically distributed.
+Consequently, we can construct a $(1-\alpha)$\% confidence interval for
+$\bar{x}$ -- where $\alpha$ is commonly called the
+\textit{significance level} -- as follows:
+\begin{equation}
+  \left[\bar{x} - z_{1-\alpha/2}\dfrac{s}{\sqrt{n}},
+    \bar{x} + z_{1-\alpha/2}\dfrac{s}{\sqrt{n}}\right]
+\end{equation}
+where $\bar{x}$ is the sample mean, $\alpha$ is the significance level
+(5\% or $0.05$ in our case), $s$ is the sample standard deviation, $n$
+is the number of observations in our sample, and $z_{1-\alpha/2}$ is the value
+of the standard normal variable $Z \sim \mathcal{N}(0,1)$ with cumulative
+probability $1-\alpha/2$. Since we have taken 40 independent samples of the
+ten-fold cross-validated AUC, the Central Limit Theorem holds and we can
+compute the mean and standard deviation of our 40 samples in order to
+construct a 95\% confidence interval for the mean AUC of each classifier.
 
 \subsection{Statistical tests}
+To decide whether or not the difference in values of each metric between
+classifiers was statistically significant, paired $t$-tests were conducted for
+a subset of the results: \todo{elaborate}
+\begin{itemize}
+  \item Some feature set vs other feature set
+  \item Single classifier vs other single classifier
+  \item Ensemble vs single
+  \item Our best method vs Dinh2013's method
+\end{itemize}

### 2014-09-27 poster/myposter.tex
Started writing up poster #14

@@
 \title{Improved Prediction of Hospital Length of Stay for Severe Injury}
-\author{Tianyu Pu \texttt{tianyu.pu@sydney.edu.au}}
+\author{Tianyu Pu \xspace \texttt{tianyu.pu@sydney.edu.au}}
 
@@
 % =============================================================================
-\section{Aims}
-\noindent This is some text here...
-\cite{Dinh2013a}
+\section{Problem}
+\noindent There are limited beds in hospital trauma wards, and yet there is a
+constant demand for these beds by the inflow of severely injured patients. Many
+patients occupy these beds when they could be better suited in another
+specialised ward. If we could accurately predict how long an injured patient
+would need to be hospitalised, we could make a more informed decision to place
+them in another ward, freeing up these beds for the urgently injured who need
+them most.
 
 % =============================================================================
-\section{Introduction}
+\section{Motivation}
+\noindent Accurate prediction of the hospital length of stay in various medical
+domains has been extensively studied, as it is a key indicator of the level of
+resource utilisation in hospitals and directly impacts the use of their limited
+funding \cite{Walczak2003}.
 
 % =============================================================================
-\section{Method}
+\section{Contribution}
+\noindent
+
+% =============================================================================
+\section{Dataset}
+
+% =============================================================================
+\section{Approach}
+
+% =============================================================================
+\section{Evaluation Metrics}
 
@@
 % =============================================================================
-\section{Discussion}
+\section{Conclusions and Future Work}
 
 % =============================================================================
-\section{Conclusions and Future work}
 
-% =============================================================================
 \references

### 2014-09-27 code/weka/run_weka.sh
Updated run weka script with more helpful error messages

@@ -14,15 +14,24 @@
 # An optional third argument to the script is to specify the number of runs
 # of the classifiers. This must be a positive integer, and defaults to 1.
 
+# Define a function that prints out usage information for various missing
+# argument cases
+usage_msg ()
+{
+  echo "Usage: ./run_weka.sh <input arff file> <filename with list of weka commands> [<number of runs>]"
+}
+
 # If we didn't specify an input file name to this script, exit
 if [ -z $1 ]; then
   echo "Please specify an input ARFF file. Exiting."
+  usage_msg
   exit 1
 fi
 
 # If we didn't specify a config file
 if [ -z $2 ]; then
   echo "Please specify a list of weka classifiers. Exiting."
+  usage_msg
   exit 2
 fi
 

### 2014-09-27 code/weka/run_weka.sh
Rewrote control logic of run weka script
Outputs results from all runs into single result file
Groups output from each classifier
Suppresses training data statistics
Uses a `random' number seed per run

@@ -48,41 +48,42 @@ fi
 
 # Weka configuration
 WEKA_CP=~/Downloads/weka-3-7-11/weka.jar
-COMMON_OPTIONS="-t $1" # options common to all weka classifiers
+# options common to all weka classifiers; -v suppresses training data output
+COMMON_OPTIONS="-v -t $1"
 JVM_HEAPSIZE=1024m
 
 # Output file base configuration
 PREFIX=weka_result
 CONFIG=$2 # indicates which config was used
 OUTDIR=../../data/results/
+TIMESTAMP=`date +%Y%m%d%H%M%S`
+OUTFILE=$OUTDIR$PREFIX$TIMESTAMP${CONFIG}_$RUNS # use a single file for all runs
+
+# Record what training file and classifiers (and their options) we used and
+# how many runs were performed
+echo $RUNS >> $OUTFILE
+echo $1 >> $OUTFILE
+cat $2 >> $OUTFILE
 
 # Read all lines of weka config into an array
 declare -a CLFS
 readarray -t CLFS < $2
 
-# Run the classifiers $RUNS times
-for RUN in $(seq 1 $RUNS)
+# Run all classifiers from a configuration file, substituting the filename
+# in the correct place to overcome problems with supplying arguments to
+# weka classifiers via the command line.
+# sed is run twice to remove a possible double occurrence of $COMMON_OPTIONS
+NUM_LINES=${#CLFS[@]}
+LINENUM=0
+while [ $LINENUM -lt $NUM_LINES ]
 do
-  echo "Executing run $RUN..."
-  # Create the output file based on the current timestamp of the run
-  TIMESTAMP=`date +%Y%m%d%H%M%S`
-  OUTFILE=$OUTDIR$PREFIX$TIMESTAMP$CONFIG
-  # Record what data file and classifiers we are using in the results file
-  # for later reference
-  echo $1 >> $OUTFILE
-  cat $2 >> $OUTFILE
-  # Run all classifiers from a configuration file, substituting the filename
-  # in the correct place to overcome problems with supplying arguments to
-  # weka classifiers via the command line.
[…]

### 2014-09-28 code/weka/run_weka.sh
Extensive rewrite of run weka script
* Command line arguments use flags and can appear in any order
* Added the ability to indicate if ROC curve data is desired
* If ROC curve data is desired, then files are automatically created
* Sanity checking for input files

@@ -13,61 +13,115 @@
 #
 # An optional third argument to the script is to specify the number of runs
 # of the classifiers. This must be a positive integer, and defaults to 1.
+#
+# Another optional fourth argument to the script is to specify whether or not
+# to save ROC curve data for the model. Specify the -r flag to do so.
 
 # Define a function that prints out usage information for various missing
 # argument cases
 usage_msg ()
 {
-  echo "Usage: ./run_weka.sh <input arff file> <filename with list of weka commands> [<number of runs>]"
+  echo "Usage: ./run_weka.sh -i <input arff file> -w <filename with list of weka commands> [-n <number of runs>] [-r]"
 }
 
-# If we didn't specify an input file name to this script, exit
-if [ -z $1 ]; then
-  echo "Please specify an input ARFF file. Exiting."
+# Initialise some variables
+TRAINING_FILE=
+WEKACONFIG=
+OUTPUT_ROC=false
+
+# Process command line arguments
+while [[ $# > 0 ]]; do
+  key="$1"
+  shift
+  case $key in
+    -i)
+      TRAINING_FILE="$1"
+      shift
+      ;;
+    -w)
+      WEKACONFIG="$1"
+      shift
+      ;;
+    -n)
+      RUNS="$1"
+      shift
+      ;;
+    -r)
+      OUTPUT_ROC=true
+      ;;
+    *)
+      echo "Unknown option: $key"
+      usage_msg
+      ;;
+  esac
+done
+
+echo "TRAINING_FILE: $TRAINING_FILE"
+echo "WEKACONFIG: $WEKACONFIG"
+echo "RUNS: $RUNS"
+echo "OUTPUT_ROC: $OUTPUT_ROC"
+
+# Sanity check for the input files
+if [ -z "$TRAINING_FILE" ] || [ -z "$WEKACONFIG" ]; then
+  echo "Input ARFF file or WEKA config not specified. Exiting."
   usage_msg
   exit 1
 fi
-
-# If we didn't specify a config file
-if [ -z $2 ]; then
-  echo "Please specify a list of weka classifiers. Exiting."
-  usage_msg
+if [ ! -e "$TRAINING_FILE" ]; then
[…]

### 2014-09-28 code/weka/summarise_results.sh
Wrote a script to summarise various metrics from results #15
Shows average # correct, TP/FP rate, precision, recall, and AUC

@@ -1,0 +1,89 @@
+#!/usr/bin/env bash
+
+usage_msg () {
+  echo "Usage: ./summarise_results.sh -n <num groups> -x <num runs> -f <input file>"
+}
+
+while [[ $# > 0 ]]; do
+  key=$1
+  shift
+  case $key in
+    -n)
+      NUMGROUPS="$1"
+      shift
+      ;;
+    -x)
+      NUMRUNS="$1"
+      shift
+      ;;
+    -f)
+      RESULTFILE="$1"
+      shift
+      ;;
+    *)
+      echo "Unknown option: $key"
+      ;;
+  esac
+done
+
+if [ -z $NUMGROUPS ] || [ -z $NUMRUNS ] || [ -z $RESULTFILE ]; then
+  echo "Please specify all arguments."
+  usage_msg
+  exit 1
+fi
+if [ ! -e $RESULTFILE ]; then
+  echo "Input file does not exist."
+  exit 2
+fi
+
+NUMBER=^[1-9][0-9]*
+if [[ ! $NUMGROUPS =~ $NUMBER ]] || [[ ! $NUMRUNS =~ $NUMBER ]]; then
+  echo "Number of groups or runs is not a valid number."
+  exit 3
+fi
+
+echo "NUMGROUPS: $NUMGROUPS"
+echo "NUMRUNS: $NUMRUNS"
+echo "RESULTFILE: $RESULTFILE"
+echo "TRAINING FILE: $(head $RESULTFILE -n 2 | tail -n 1)"
+
+build_cmd () {
+  search_string=$1
+  fields_to_retrieve=$2
+  awk_main=$3
+  cat $RESULTFILE | grep -E "$search_string" | tail -n+$(($group*$NUMRUNS+1)) | \
+    head -n $NUMRUNS | awk "$fields_to_retrieve" | awk -v runs=$NUMRUNS "$awk_main"
+}
+
+echo "avg_correct"
+for group in $(seq 0 $(($NUMGROUPS-1)))
+do
+  build_cmd "^Correctly" '{print $4}' 'BEGIN {numcorrect = 0} {numcorrect += $0} END {print numcorrect/runs}'
+done
+
+echo "avg_tp avg_fp avg_prec avg_rec avg_auc"
+for group in $(seq 0 $(($NUMGROUPS-1)))
+do
+  build_cmd "^[ ]{15,}0.*0$" '{print $1, $2, $3, $4, $7}' '
[…]

### 2014-09-29 code/weka/compute_ci.py
Updated summarise results script, wrote a script to calculate CIs #15
summarise_results.sh outputs numbers from the result files for compute_ci.py
to process and calculate confidence intervals.

summarise_results.sh can also output averages if given a specific option

@@ -1,0 +1,66 @@
+#!/usr/bin/env python3
+
+import re
+import sys
+
+# A pattern for matching numbers with at least 1 digit and optional decimals.
+number = re.compile(r'[0-9]+(([.]?[0-9]+)|([0-9]*))')
+
+# Compute a normal 95% CI for a sample.
+def compute_ci(sample):
+    z95 = 1.95996
+    sample = list(map(float, sample))
+    n = len(sample)
+    mean = sum(sample) / n
+
+    stdev = 0
+    for x in sample:
+        stdev += (x-mean)**2
+    stdev *= 1.0 / (n-1)
+    stdev **= 0.5
+
+    uncertainty = z95 * stdev / n**0.5
+
+    return (mean, uncertainty)
+
+# These are assumed to be given from the summarise_results.sh script.
+ngroups = int(sys.stdin.readline().strip())
+nruns = int(sys.stdin.readline().strip())
+
+lines = [line.strip() for line in sys.stdin.readlines()]
+
+# Get the headings for each of the various statistics that we are summarising.
+# Keep track of the actual headings, as well as the indices that they appear.
+# This is important for output and partitioning.
+headings = []
+heading_index = 0
+indices = []
+while heading_index < len(lines):
+    indices.append(heading_index)
+    headings.append(lines[heading_index])
+    heading_index += ngroups * nruns + 1
+
+# For each heading, get the output values and compute a mean and 95% CI for
+# them, grouped by the number of distinct classifiers.
+for i in range(len(headings)):
+    heading = headings[i]
+    print(heading)
+    # Find the start and end points of all the data belonging to this heading.
+    start_index = indices[i] + 1
+    if i + 1 >= len(headings):
+        partition = lines[start_index:]
+    else:
+        end_index = indices[i+1]
+        partition = lines[start_index:end_index]
+    # Split each line of the partition up, in case there are several metrics
+    # on one line.
+    partition = [line.split() for line in partition]
+    # Split the partition into its corresponding groups and summarise the
+    # numbers for each group.
+    for j in range(0, len(partition), nruns):
+        sample = partition[j:j+nruns]
+        results = []
+        for col in zip(*sample):
+            m, u = compute_ci(col)
+            results.append('{0:.3f}({1:.5f})'.format(m, u))
+        print(' '.join(results))

### 2014-09-29 code/weka/summarise_results.sh

@@ -1,9 +1,19 @@
 #!/usr/bin/env bash
 
+# A script to get certain parts of result files and either summarise them
+# or output them for summarising by compute_ci.py.
+
 usage_msg () {
-  echo "Usage: ./summarise_results.sh -n <num groups> -x <num runs> -f <input file>"
+  echo "Usage: ./summarise_results.sh -n <num groups> -x <num runs> -f <input file> [-v] [-a]"
+  echo "If -a is specified, then -v is turned on as well as it is assumed the output won't
+    be piped into compute_ci.py"
 }
 
+# Default behaviour
+VERBOSE=false
+FIND_AVGS=false
+
+# Process command line arguments
 while [[ $# > 0 ]]; do
   key=$1
   shift
@@ -20,12 +30,19 @@ while [[ $# > 0 ]]; do
       RESULTFILE="$1"
       shift
       ;;
+    -v)
+      VERBOSE=true
+      ;;
+    -a)
+      FIND_AVGS=true
+      ;;
     *)
       echo "Unknown option: $key"
       ;;
   esac
 done
 
+# Check that the inputs make sense
 if [ -z $NUMGROUPS ] || [ -z $NUMRUNS ] || [ -z $RESULTFILE ]; then
   echo "Please specify all arguments."
   usage_msg
@@ -42,19 +59,45 @@ if [[ ! $NUMGROUPS =~ $NUMBER ]] || [[ ! $NUMRUNS =~ $NUMBER ]]; then
   exit 3
 fi
 
-echo "NUMGROUPS: $NUMGROUPS"
-echo "NUMRUNS: $NUMRUNS"
-echo "RESULTFILE: $RESULTFILE"
-echo "TRAINING FILE: $(head $RESULTFILE -n 2 | tail -n 1)"
+# Special output for verbose mode
+verbose () {
+  echo "NUMGROUPS: $NUMGROUPS"
+  echo "NUMRUNS: $NUMRUNS"
+  echo "RESULTFILE: $RESULTFILE"
+  echo "TRAINING FILE: $(head $RESULTFILE -n 2 | tail -n 1)"
+}
 
+# If -a is set, assume manual mode and set verbose on
+if [ $FIND_AVGS = true ]; then
+  VERBOSE=true
+fi
+
+# If -v set, then print verbose output, otherwise print terse output for
+# compute_ci.py
+if [ $VERBOSE = true ]; then
+  verbose
+else
+  echo "$NUMGROUPS"
[…]

### 2014-09-29 code/weka/run_weka.sh
Fixed terrible bug in run weka script with replacing common options

@@ -126,7 +126,7 @@ readarray -t CLFS < "$WEKACONFIG"
 # Run all classifiers from a configuration file, substituting the filename
 # in the correct place to overcome problems with supplying arguments to
 # weka classifiers via the command line.
-# sed is run twice to remove a possible double occurrence of $COMMON_OPTIONS
+# sed is run twice to remove a possible double occurrence of $OPTIONS
 NUM_LINES=${#CLFS[@]}
 LINENUM=0
 while [ $LINENUM -lt $NUM_LINES ]; do
@@ -138,7 +138,8 @@ while [ $LINENUM -lt $NUM_LINES ]; do
       ROC_FILE="${LINENUM}_$(echo ${CLFS[LINENUM]} | awk '{print $1}')_$run.arff"
       ROC_OPTION="-threshold-file $ROC_FOLDER/$ROC_FILE"
     fi
-    eval $(echo "java -Xmx$JVM_HEAPSIZE -cp $WEKA_CP ${CLFS[LINENUM]} $ROC_OPTION -s $SEED $COMMON_OPTIONS >> $RESULTFILE" | sed "s:<OPTS>:$COMMON_OPTIONS:g" | sed "s:$COMMON_OPTIONS::2g")
+    OPTIONS="$ROC_OPTION -s $SEED $COMMON_OPTIONS"
+    eval $(echo "java -Xmx$JVM_HEAPSIZE -cp $WEKA_CP ${CLFS[LINENUM]} $OPTIONS >> $RESULTFILE" | sed "s:<OPTS>:$OPTIONS:g" | sed "s:$OPTIONS::2g")
   done
   LINENUM=$[$LINENUM+1]
 done

### 2014-10-12 poster/myposter.tex
Lots of poster writing. #14

--- 2014-09-26 poster/myposter.tex
@@
 \usepackage{xspace}
+\usepackage{graphicx}
 
@@
 
+\newcommand{\los}{\acronym{los}}
+\newcommand{\roc}{\acronym{roc}}
+\newcommand{\auc}{\acronym{auc}}
+\newcommand{\nn}{\acronym{nn}}
+
+\DeclareGraphicsExtensions{.png}
+
 \flushbottom
@@
 % =============================================================================
-\section{Aims}
-\noindent This is some text here...
-\cite{Dinh2013a}
+\section{Problem}
+\noindent There are limited beds in hospital trauma wards, and yet there is a
+constant demand for these beds by the inflow of severely injured patients. Many
+patients are initially allocated to these beds when they could be better
+treated in another specialised ward.
+\begin{center}
+\includegraphics{problem-diag}
+\end{center}
+
+\noindent If we could accurately classify patients with hospital length of
+stay (\los) of 2 days or less versus those who require longer stays, we could
+make a more informed decision whether or not to place them in another ward
+when they are admitted, rather than having to waste time and resources
+transferring them to another ward later. This also frees up the high-demand
+trauma beds for the critically injured who need them most.
+\\
 
 % =============================================================================
-\section{Introduction}
+\section{Motivation}
+\noindent Accurate prediction of the \los in various medical
+domains (such as burns and acute pancreatitis) has been extensively studied,
+as it is a key indicator of the level of
+resource utilisation in hospitals and directly impacts the use of their limited
+funding \cite{Walczak2003}.
+\\
+
+\noindent However, the current literature focuses extensively
+on logistic regression and generalised linear models, with some applications
+of techniques from machine learning such as artificial neural networks.
+There has been no work done in systematically
+applying a range of machine learning techniques to \los prediction, and
+specifically not to the trauma medical domain.
+\\
+
+\noindent Additionally, there has been little work done in feature
+transformation and selection when \los prediction models are created.
+\\
 
 % =============================================================================
-\section{Method}
+\section{Contribution}
+\noindent Building upon the work of trauma \los prediction in \cite{Dinh2013a},
+we systematically investigate feature transformation and selection
+techniques in the construction of a \los prediction model for trauma patients.
+We also apply and evaluate a comprehensive range of
+classification algorithms on data from the trauma domain as well as from a
+general hospital setting.
+\\
+
+\noindent Additionally, we
+propose a new nearest-neighbour (\nn) algorithm, ranked \nn, which takes
+into account the
+predictive relevance of features when computing the distance to the nearest
+neighbours.
+\\
+
+% =============================================================================
+\section{Datasets}
+\noindent Our study was conducted on two datasets: one with 2546 records from
+the Trauma Services Centre at the Royal Prince Alfred Hospital in Sydney,
+consisting of
+users admitted to the centre between 2007--11; the other from the Hospital das
+Foras Armadas in Portugal with 17546 records collected from 2000--13.
+\\
+
+\noindent
+The dataset from Sydney consists solely of patients admitted to Trauma
+Services, whereas the dataset
+from Portugal is a collection of hospital-wide inpatient admissions spanning
+a range of diagnoses.
+\\
+
+% =============================================================================
+\section{Approach}
+\includegraphics{approach}
+
+\subsection{Ranked Nearest Neighbour}
+\noindent \textit{Idea: the contribution of a feature to the distance between
+two records should be proportional to its level of `relevance' or predictive
+power with respect to the outcome variable, \los.} This is how it works:
+\begin{center}
+\includegraphics{ranked-nn}
+\end{center}
+
+% =============================================================================
+\section{Evaluation Metrics}
+\noindent \textit{Area under the receiver-operating characteristic curve (\auc)},
+the extent to which the classification algorithm is able to distinguish between
+patients that require a stay of less than 2 days or those who require longer.
+Randomly picking between two outcomes has \auc 0.5, and the maximum is 1,
+indicating perfect discriminating ability.
+\\
+
+\noindent Our \textit{baseline} is the logistic regression model derived in
+\cite{Dinh2013a}, as we aim to improve upon the results they achieved for
+trauma patients.
+\\
+
+\noindent All \auc figures are obtained from
+\textit{10-fold stratified cross-validation}.
+\\
 
@@
 \section{Results}
+\noindent
 
 % =============================================================================
-\section{Discussion}
+\section{Conclusions and Future Work}
+\noindent
 
 % =============================================================================
-\section{Conclusions and Future work}
+\section{Acknowledgements}
+\noindent
+Many thanks to Dr Michael Dinh, co-director of Royal Prince Alfred
+Hospital Trauma Services, for providing the trauma data and medical domain
+guidance; and thanks also to Associate Professor Paulo Cortez, from the
+Department of Information Systems at the University of
+Minho in Portugal, for providing an extensive general \los dataset.
+\\
 
 % =============================================================================
+
 \references

### 2014-10-13 poster/myposter.tex
Finished poster writeup #14

--- 2014-09-26 poster/myposter.tex
@@
 
+\newcommand{\los}{\acronym{los}}
+\newcommand{\roc}{\acronym{roc}}
+\newcommand{\auc}{\acronym{auc}}
+\newcommand{\nn}{\acronym{nn}}
+\newcommand{\cfs}{\acronym{cfs}}
+
+\DeclareGraphicsExtensions{.png}
+
 \flushbottom
@@
 % =============================================================================
-\section{Aims}
-\noindent This is some text here...
-\cite{Dinh2013a}
+\section{Problem}
+\noindent There are limited beds in hospital trauma wards, and yet there is a
+constant demand for these beds by the inflow of severely injured patients. Many
+patients are initially allocated to these beds when they could be better
+treated in another specialised ward.
+\begin{center}
+\includegraphics[scale=0.75]{problem-diag}
+\end{center}
+\noindent If we could accurately classify patients with hospital length of
+stay (\los) of 2 days or less versus those who require longer stays, we could
+make a more informed decision whether or not to place them in another ward
+when they are admitted, rather than wasting time and resources
+transferring them to another ward later.
 
 % =============================================================================
-\section{Introduction}
+\section{Motivation}
+\noindent Accurate prediction of the \los in various medical
+domains (such as burns) has been extensively studied
+as it is a key indicator of how hospitals use their limited resources.
+The current literature focuses extensively
+on generalised linear models, with some applications
+of techniques from machine learning such as neural networks.
+There has been no work done in systematically
+applying a range of machine learning techniques to \los prediction, and
+specifically not to the trauma medical domain.
+Additionally, there has been little work done in feature
+transformation and selection when \los prediction models are created.
 
 % =============================================================================
-\section{Method}
+\section{Contribution}
+\noindent We systematically investigate feature transformation and selection
+techniques in the construction of a \los prediction model for trauma patients.
+We also apply and evaluate a comprehensive range of
+classification algorithms on data from the trauma domain as well as from a
+general hospital setting.
+
+In addition, we
+propose a new nearest-neighbour (\nn) algorithm, ranked \nn, which takes
+into account the
+predictive relevance of features when computing the distance to the nearest
+neighbours.
+
+% =============================================================================
+\section{Datasets}
+\noindent Our study was conducted on two datasets: one with 2546 records from
+the Trauma Services Centre at the Royal Prince Alfred Hospital in Sydney,
+consisting of
+trauma patients admitted to the centre between 2007--11;
+the other from the Hospital das
+Foras Armadas in Portugal with 17546 records collected from 2000--13 and
+covering a wide range of medical diagnoses.
+
+% =============================================================================
+\section{Approach}
+\subsection{Feature preprocessing and investigation of learning algorithms}
+\begin{center}
+\includegraphics[scale=0.75]{approach}
+\end{center}
+
+\subsection{Ranked nearest neighbour}
+\noindent \textit{Idea: the contribution of a feature to the distance between
+two records should be proportional to its level of `relevance' or predictive
+power with respect to the outcome variable, \los.} This is how it works:
+\begin{center}
+\includegraphics[scale=0.75]{ranked-nn}
+\end{center}
+
+% =============================================================================
+\section{Evaluation Metrics}
+\noindent \textit{Area under the receiver-operating characteristic curve (\auc)},
+the extent to which the classification algorithm is able to distinguish between
+patients that require a stay of less than 2 days or those who require longer.
+Randomly picking between two outcomes has \auc 0.5; the maximum is 1,
+indicating perfect discriminating ability.
+
+Our \textit{baseline} is the logistic regression model derived in
+\cite{Dinh2013a}, as we aim to improve upon the results they achieved for
+trauma patients.
+All \auc figures are obtained from
+\textit{10-fold stratified cross-validation}.
 
@@
 \section{Results}
+\subsection{Trauma dataset}
+Discretisation significantly improved the performance of all classifiers,
+regardless of feature selection method. Feature selection usually improved
+performance but depended on the method, with \cfs the most effective. Ranked
+\nn is not affected by discretisation and does not require feature selection,
+and improves upon baseline \auc by 1-2\% (0.82-0.83)
+while being faster to train than more sophisticated methods.
+The best classifiers were logistic regression
+and SVM with discretisation with mean \auc 0.84 (baseline 0.81).
+\includegraphics[scale=0.53]{trauma}
+
+\subsection{General dataset}
+All classification algorithms performed very well regardless of discretisation
+or feature selection as the original features were very good at predicting
+\los. Again, performing discretisation makes the method of feature selection
+less important, as it stabilises performance overall. The best results were
+mean \auc 0.994 with logistic regression and SVM regardless of discretisation;
+ranked \nn had mean \auc 0.992 with less training time.
+\includegraphics[scale=0.53]{portugal}
 
 % =============================================================================
-\section{Discussion}
+\section{Conclusions and Future Work}
+\noindent Careful application of discretisation and feature selection can
+improve the predictive power of classification algorithms. We proposed a
+general method to improve the \los predictions for a trauma and general
+hospital domain, and showed overall improvements in \auc from 2-4\% from the
+baseline. This general method will be a useful starting point for those
+investigating \los prediction.
+
+Furthermore, ranked \nn performs 1-2\% better than the baseline, is
+moderately fast to train, and is not strongly affected by discretisation and
+irrelevant features.
+
+There is
+room to further evaluate this method in other medical areas, as well as
+implementation of such a model as a decision support tool for physicians.
+Additionally, improving upon the ranked \nn algorithm to account for different
+methods of weighting features and calculating ranks can also be investigated.
 
 % =============================================================================
-\section{Conclusions and Future work}
+\section{Acknowledgements}
+\noindent
+Many thanks to Dr Michael Dinh, co-director of Royal Prince Alfred
+Hospital Trauma Services, for providing the trauma data and medical domain
+guidance; and thanks also to Associate Professor Paulo Cortez, from the
+Department of Information Systems at the University of
+Minho in Portugal, for providing an extensive general \los dataset.
 
 % =============================================================================
+
 \references

### 2014-10-15 vagrant/setup-python.sh
setup-python.sh: download pydot with the expected filename for unpacking

@@ -35,7 +35,8 @@ mkdir $temp_dir
 cd $temp_dir
 
 echo "=== DOWNLOADING PACKAGES FOR SCIKIT-LEARN, NUMPY, SCIPY AND DEPENDENCIES"
-wget $sklearn_url $numpy_url $scipy_url $nose_url $pyparsing_url $pydot_url
+wget $sklearn_url $numpy_url $scipy_url $nose_url $pyparsing_url
+wget -O $pydot_name.zip $pydot_url
 
 # Extract, build and install tars
 for package in $numpy_name $scipy_name $sklearn_name $nose_name

### 2014-10-15 vagrant/provision.sh
provision.sh: add return checks to source commands, fail more gracefully

@@ -13,13 +13,13 @@ echo "=== INSTALLING OPENJDK 7 JRE"
 apt-get install -y openjdk-7-jre > /dev/null
 
 # TeX Live
-source setup-texlive.sh
+source setup-texlive.sh || exit 1
 
 # Python 3.4 and scipy, numpy, and scikit-learn
-source setup-python.sh
+source setup-python.sh || exit 1
 
 # Go to the vagrant folder in the project root folder and run user provision scripts
-su -c "source user-provision.sh" vagrant
+su -c "source user-provision.sh" vagrant || exit 1
 
 # State that we are done
 echo "=== SETUP OF VIRTUAL MACHINE COMPLETE"

### 2014-10-15 code/weka/RankedDistance.java
Added ranked distance function for k-NN classifier

@@ -1,0 +1,287 @@
+/**
+ * 
+ */
+package weka.core;
+
+import weka.attributeSelection.ASEvaluation;
+import weka.attributeSelection.ASSearch;
+import weka.attributeSelection.CorrelationAttributeEval;
+import weka.attributeSelection.InfoGainAttributeEval;
+import weka.attributeSelection.Ranker;
+import weka.core.Instance;
+import weka.core.Instances;
+import weka.core.NormalizableDistance;
+import weka.core.RevisionUtils;
+import weka.core.TechnicalInformation;
+import weka.core.TechnicalInformationHandler;
+import weka.core.TechnicalInformation.Field;
+import weka.core.TechnicalInformation.Type;
+import weka.core.neighboursearch.PerformanceStats;
+
+/**
+ * @author tian
+ *
+ */
+public class RankedDistance
+extends NormalizableDistance
+implements TechnicalInformationHandler {
+
+	private static final long serialVersionUID = 1068606253458807903L;
+
+	/** Holds the ranked attributes */
+	private int[] m_RankedAttributes;
+
+	/** Holds the "merit" of the ranked attributes. */
+	private double[] m_RankedMerits;
+	
+	/** Holds the weighting of the ranked attributes used by this distance function */
+	private double[] m_RankWeights;
+
+	/** The actual ranker that performs attribute ranking */
+	private ASSearch m_AttributeRanker;
+
+	/** The class that evaluates the goodness of the feature set. */
+	protected ASEvaluation m_AttributeEvaluator;
+	
+	/** How much to update a weight if we find it important. */
+	protected static double m_WeightUpdateFactor = 0.001;
+
+	/**
+	 * 
+	 */
+	public RankedDistance() {
+		super();
+	}
+
+	/**
+	 * @param data
+	 * @throws Exception 
+	 */
+	public RankedDistance(Instances data) {
+		super(data);
+	}
+
+	/**
+	 * initializes the ranges and the attributes being used.
+	 */
+	protected void initialize() {
+		super.initialize();
+
+		m_AttributeRanker = new Ranker();
+		m_AttributeEvaluator = new CorrelationAttributeEval();
+		m_RankedAttributes = new int[m_Data.numAttributes()];
+		m_RankedMerits = new double[m_Data.numAttributes()];
+		m_RankWeights = new double[m_Data.numAttributes()];
+		
+		try {
+			m_AttributeEvaluator.buildEvaluator(m_Data);
[…]

### 2014-10-16 code/weka/RankedDistance.java
Updated ranked distance k-NN class

@@ -29,22 +29,22 @@ implements TechnicalInformationHandler {
 	private static final long serialVersionUID = 1068606253458807903L;
 
 	/** Holds the ranked attributes */
-	private int[] m_RankedAttributes;
+	protected int[] m_RankedAttributes;
 
 	/** Holds the "merit" of the ranked attributes. */
-	private double[] m_RankedMerits;
+	protected double[] m_RankedMerits;
 	
 	/** Holds the weighting of the ranked attributes used by this distance function */
-	private double[] m_RankWeights;
+	protected double[] m_RankWeights;
 
 	/** The actual ranker that performs attribute ranking */
-	private ASSearch m_AttributeRanker;
+	protected ASSearch m_AttributeRanker;
 
 	/** The class that evaluates the goodness of the feature set. */
 	protected ASEvaluation m_AttributeEvaluator;
 	
 	/** How much to update a weight if we find it important. */
-	protected static double m_WeightUpdateFactor = 0.001;
+	protected static double m_WeightUpdateFactor = 0.0001;
 
 	/**
 	 * 
@@ -84,26 +84,19 @@ implements TechnicalInformationHandler {
 				double attributeMerit = tempRanked[i][1];
 				m_RankedAttributes[i] = attributeIndex;
 				m_RankedMerits[attributeIndex] = attributeMerit;
-				m_RankWeights[attributeIndex] = 1.0 / Math.pow(i+1, 2.0/3.0);
+				m_RankWeights[attributeIndex] = decayFunction(i);
 			}
 			
 		} catch (Exception e) {
 			e.printStackTrace();
 		}
-/*
-		int rankedAttributeIndex;
-
-		for (int i=0; i<m_RankWeights.length; i++) {
-			if (i < m_RankedAttributes.length) {
-				rankedAttributeIndex = m_RankedAttributes[i]; // the index of a ranked attribute
-				// if m_RankedAttributes = [13, 23 ...], then m_RankWeights = [...0...1] at the 13th and 23rd indices
-				m_RankWeights[rankedAttributeIndex] = 1.0 / Math.pow(i+1, 2.0/3.0);
-				//m_RankWeights[rankedAttributeIndex] = m_RankedMerits[rankedAttributeIndex];
-			} else {
-				m_RankWeights[i] = 0;
-			}
-		}
-*/
+	}
+	
+	/**
+	 * Given an attribute index, describes the decay of the ranking as it decreases.
+	 */
+	protected double decayFunction(int index) {
[…]

### 2014-10-16 code/weka/SimilarityDistance.java
Added another distance function related to similarity

@@ -1,0 +1,222 @@
+package weka.core;
+
+import weka.core.TechnicalInformation.Field;
+import weka.core.TechnicalInformation.Type;
+import weka.core.neighboursearch.PerformanceStats;
+
+public class SimilarityDistance
+extends RankedDistance
+implements TechnicalInformationHandler {
+
+	/**
+	 * 
+	 */
+	private static final long serialVersionUID = 4926634536990124857L;
+
+	public SimilarityDistance() {
+		super();
+	}
+
+	public SimilarityDistance(Instances data) {
+		super(data);
+	}
+
+	public void initialize() {
+		super.initialize();
+		
+		// normalise the rank weights 
+		double sum = 0;
+		for (int i=0; i<m_RankedMerits.length; i++) {
+			sum += m_RankedMerits[i];
+		}
+		for (int i=0; i<m_RankWeights.length; i++) {
+			m_RankWeights[i] = m_RankWeights[i] * m_RankedMerits[i] / sum;
+		}
+	}
+	
+	@Override
+	public String getRevision() {
+		// TODO Auto-generated method stub
+		return null;
+	}
+
+	/* (non-Javadoc)
+	 * @see weka.core.NormalizableDistance#globalInfo()
+	 */
+	@Override
+	public String globalInfo() {
+		return 
+				"Implementing Euclidean distance (or similarity) function.\n\n"
+				+ "One object defines not one distance but the data model in which "
+				+ "the distances between objects of that data model can be computed.\n\n"
+				+ "Attention: For efficiency reasons the use of consistency checks "
+				+ "(like are the data models of the two instances exactly the same), "
+				+ "is low.\n\n"
+				+ "For more information, see:\n\n"
+				+ getTechnicalInformation().toString();
+	}
+
+	/**
+	 * Calculates the distance between two instances. Offers speed up (if the
+	 * distance function class in use supports it) in nearest neighbour search by
+	 * taking into account the cutOff or maximum distance. Depending on the
+	 * distance function class, post processing of the distances by
+	 * postProcessDistances(double []) may be required if this function is used.
+	 * 
+	 * @param first the first instance
+	 * @param second the second instance
+	 * @param cutOffValue If the distance being calculated becomes larger than
[…]

### 2014-10-16 code/weka/run_weka.sh
Added option to use a varying random number seed #18 in run weka script

@@ -16,18 +16,23 @@
 #
 # Another optional fourth argument to the script is to specify whether or not
 # to save ROC curve data for the model. Specify the -r flag to do so.
+#
+# Another optional argument is to specify whether or not to use a random
+# number seed that varies. The default is to use the same seed for all runs,
+# namely 1.
 
 # Define a function that prints out usage information for various missing
 # argument cases
 usage_msg ()
 {
-  echo "Usage: ./run_weka.sh -i <input arff file> -w <filename with list of weka commands> [-n <number of runs>] [-r]"
+  echo "Usage: ./run_weka.sh -i <input arff file> -w <filename with list of weka commands> [-n <number of runs>] [-r] [-s]"
 }
 
 # Initialise some variables
 TRAINING_FILE=
 WEKACONFIG=
 OUTPUT_ROC=false
+USE_RANDOM=false
 
 # Process command line arguments
 while [[ $# > 0 ]]; do
@@ -49,6 +54,9 @@ while [[ $# > 0 ]]; do
     -r)
       OUTPUT_ROC=true
       ;;
+    -s)
+      USE_RANDOM=true
+      ;;
     *)
       echo "Unknown option: $key"
       usage_msg
@@ -60,6 +68,7 @@ echo "TRAINING_FILE: $TRAINING_FILE"
 echo "WEKACONFIG: $WEKACONFIG"
 echo "RUNS: $RUNS"
 echo "OUTPUT_ROC: $OUTPUT_ROC"
+echo "USE_RANDOM: $USE_RANDOM"
 
 # Sanity check for the input files
 if [ -z "$TRAINING_FILE" ] || [ -z "$WEKACONFIG" ]; then
@@ -132,7 +141,11 @@ LINENUM=0
 while [ $LINENUM -lt $NUM_LINES ]; do
   for run in $(seq 1 $RUNS)
   do
-    SEED=$RANDOM
+    if [ $USE_RANDOM = true ]; then
+      SEED=$RANDOM
+    else
+      SEED=1
+    fi
     echo "Training classifier $[$LINENUM+1] run #$run with seed $SEED: ${CLFS[LINENUM]}"
     if [ $OUTPUT_ROC ]; then
       ROC_FILE="${LINENUM}_$(echo ${CLFS[LINENUM]} | awk '{print $1}')_$run.arff"

### 2014-10-16 code/weka/run_weka.sh
Fixed output_roc flag check in run weka script

@@ -147,7 +147,7 @@ while [ $LINENUM -lt $NUM_LINES ]; do
       SEED=1
     fi
     echo "Training classifier $[$LINENUM+1] run #$run with seed $SEED: ${CLFS[LINENUM]}"
-    if [ $OUTPUT_ROC ]; then
+    if [ $OUTPUT_ROC = true ]; then
       ROC_FILE="${LINENUM}_$(echo ${CLFS[LINENUM]} | awk '{print $1}')_$run.arff"
       ROC_OPTION="-threshold-file $ROC_FOLDER/$ROC_FILE"
     fi

### 2014-10-18 thesis/method/classification.tex
Restructure of experimental method section #12

--- 2014-09-25 thesis/experiments/experiments.tex
@@
-\chapter{Experimental Setup} \label{chap:experiments}
-
-In this chapter we detail the steps carried out as part of our investigation.
-We first outline the nature of the data set, describe how it was cleaned and
-pre-processed before it was used to train various classifiers. The methods of
-evaluation are then presented, as these are critical to assessing the
-performance of the models created.
-
-\section{Trauma patient data set}
-
-\subsection{Collection}
-The data set used consisted of trauma registry data from the trauma centre
-at the Royal Prince Alfred Hospital, a major trauma centre in New South Wales,
-Australia. It covered all adult (age 15 and over) inpatient admissions to the
-trauma centre from 2007--2011.
-
-All patients were first admitted to the trauma ward until discharged
-or transferred to an appropriate unit within the hospital. A single trained
-data manager recorded a variety of attributes about the admitted patient,
-such as age, gender, blood pressure, mechanism of injury and body regions
-that were injured. We received this data in a Microsoft Excel spreadsheet.
-
-\subsection{Characteristics}
-There were 2546 patient records in the data set we received, comprising of 74
-features, one of which was the target variable \textbf{LOS48}. LOS48 was a
-binary variable -- that is, it could only take two values, 0 or 1 -- with 1
-indicating that the patient stayed two days or less, and 0 indicating a stay
-of greater than 2 days.
-
-\todo{Perhaps insert some more information about the data itself, such
-as the various attributes and their data types? Not sure if that should go
-here, in the results, or in an appendix -- it is relevant but can disturb
-the flow of the main text.}
-
-\section{Data preparation}
-In any data mining task, adequately preparing the data ensures the best
-chance of success: this includes deciding how to deal with missing and
-inaccurate values, as well as appropriately transforming the data
-\citep{Witten2005} and selecting or constructing the most suitable features
-for the task \citep{Kotsiantis2006}.
-In this section we describe how we addressed each of
-the major issues and the reasons for our choice.
-
-\subsection{Cleaning and normalisation} % missing values, irrelevant features
-Since the data was initially in an Excel spreadsheet, we explored the data
-by hand using Microsoft Excel's filter function on each column and
-consolidated many categorical values simply
-by correcting spelling and punctuation. For example:
-\begin{itemize}
-\item In the \textbf{sex} attribute, there were values of `Female' and
-`Femal' which were consolidated into `Female'.
-\item In the \textbf{mechanism} attribute, which was textual and indicated
-the category of how the patient was injured, there were such categories as
-`Other Vehi' and `Other Vehicle' as well as `Pedal Cycl' and `Pedal Cyclist'.
-Such categories were combined into `Other Vehicle' and `Pedal Cyclist'
-respectively.
-\item Some categories in \textbf{mechanism} were also divided into whether or
-not the mechanism of injury was self-inflicted, such as `Shooting' and
-`Shooting-selfinflicted'. When we asked the domain expert, Michael Dinh, his
-suggestion was to combine them into one category: in this case, simply
-`Shooting'; this was done for all the categories in this attribute.
-\item \textbf{disp from ED} indicated the section of the hospital that the
-patient was discharged to. This was also a textual categorical attribute.
-There were redundancies like `G. ICU' and `G.ICU' and even `GICU', which we
-all combined into one category, `GICU'.
-\end{itemize}
-
-The data was then exported into CSV (comma-separated value) format, so that
-we could continue to process it programmatically rather than manually. Many
-fields in the data set contained leading or trailing whitespace, which was
-removed with a few lines of Python code (see appendix \todo{[add reference
-to appendix]}). All attribute names were converted to lowercase. In the case
-where attribute names consisted of several words, the spaces were replaced
-with underscores (\_) to ensure that the full attribute name was read.
-\todo{Include a list of these converted attribute names? In the appendix?}
-
-In order for our results to be comparable to that of Dinh
-et. al. \citep{Dinh2013a}, we removed all instances with missing values,
-leaving 2517 instances with which to use to create our prediction models.
-
-\subsection{Feature selection}
-After data preparation, selecting a subset of features or attributes to use
-in a model is also important, and helps to reduce noise and redundant
-attributes from becoming part of the final model. Experiments have found that
-the performance of various classification methods such as decision trees
-\textit{deteriorates} in the presence of useless attributes.
-Our data set contains 73
-non-target attributes, which is a fairly high number. Many of these attributes
-are derived from others in the set, such as \textbf{age} and \textbf{age65}.
-
-We began the process of feature selection by removing attributes that were
-not available upon admission, as these tended to be highly related to the
-LOS48 outcome and were a consequence, not a predictor, of LOS. These attributes
-were: iculos, outcome, disp from ED, disp from hosp, died, rehab and los.
-The id attribute was also removed because it simply numbered each record and
-indicated nothing about the patient themself. This left us with 67 attributes.
-
-The work of actually selecting the attributes out of the 67 available was
-carried out in two ways: automated using feature selection methods, and
-manually using advice from a domain expert.
-
-\subsubsection{Automated feature selection methods}
-There are a number of methods that are commonly used to automate the selection
-of a subset of features for the final classifier. The methods we used were:
-\begin{itemize}
-\item \textit{Correlation-based feature selection} \citep{Hall2000},
-which evaluates the `merit' of a
-subset of features by considering their degree of correlation with the target
-variable compared to their degree of correlation amongst each other. High
-correlation between variables indicates redundancy. Under
-this scheme, a set of features that are highly correlated with the target
-while having low correlation with other features in the set is ideal.
-\item Selection based on the
-\textit{Pearson product-moment correlation coefficient}, a measure of the
-linear correlation or dependence between two variables.\todo{add some more info}
-\item \textit{Information gain} \todo{explain a bit about this}
-\item Feature selection based on the \textit{OneR classifier}
-\citep{Holte1993}, which uses one
-attribute as the predictor for the target. This attribute is typically the one
-that has the lowest prediction error. Feature selection using this scheme
-involves first ranking attributes based on their error, with the lowest error
-attributes as the most desirable. We then selected the top 20 attributes to
-use in the building of the model.
-\end{itemize}
-
-We used the WEKA 3.7.11 \citep{Hall2009} implementation of the above
-methods, after converting the data into the Attribute-Relation File Format
-(ARFF)\footnote{\url{http://weka.wikispaces.com/ARFF+\%28developer+version\%29}}
-most suitable for the program.
-\todo{give more details, ie specifying binary/numeric
-attributes, maybe specify in appendix?}
-
-\subsubsection{Expert selection}
-We also wanted to compare the feature selection of an expert in the domain
-of the data set, namely trauma, with the models obtained from the features
-selected using the automated methods described above. This resulted in two
-feature subsets. The first was obtained simply by presenting the entire list
-of 73 features to the domain expert and asking him to select the features he
-thought were most indicative of the target variable from experience. The
-second set was obtained from the work of Dinh et. al. \citep{Dinh2013a} that
-we are building on, giving us a baseline to compare our results to.
-
-Altogether, the various feature selection methods resulted in a total of
-7 data sets (4 automated, 2 expert selected, and 1 with all features) which
-we used in the model construction phase.
+\chapter{Classification and Supervised Learning Algorithms}
+  \label{chap:classification}
 
@@
 \subsection{Ensemble classifiers}
+In order to improve upon the performance of individual classifiers, we also
+tested the performance of ensembles of classifiers -- that is, combining the
+predictions of several different classifiers, or several of the same
+classifier. Specifically, the ensemble methods used were:
+\begin{itemize}
+  \item Random Forests
+  \item Bagging
+  \item Stacked generalisation (stacking)
+  \item Boosting (AdaBoostM1)
+\end{itemize}
 
-\section{Performance evaluation} % don't forget cross-validation
-\label{sec:perfeval}
-
-\subsection{Metrics}
-
-\subsection{Statistical tests}
+Again, we were able to create predictive models using the implementations of
+these ensemble methods in WEKA 3.7.11. Since ensemble classifiers are not
+actual classifiers but a way to train and combine the predictive power of more
+than one classifier, we list the combinations of classifiers used in our
+ensemble methods:
+\todo{insert table!}

### 2014-10-18 thesis/method/evaluation.tex
Restructure of experimental method section #12

--- 2014-09-26 thesis/experiments/experiments.tex
@@
-\chapter{Experimental Setup} \label{chap:experiments}
+\chapter{Evaluation} \label{chap:evaluation}
 
-In this chapter we detail the steps carried out as part of our investigation.
-We first outline the nature of the data set, describe how it was cleaned and
-pre-processed before it was used to train various classifiers. The methods of
-evaluation are then presented, as these are critical to assessing the
-performance of the models created.
-
-\section{Trauma patient data set}
-
+\section{Data Sets}
 \subsection{Collection}
@@
 the flow of the main text.}
-
-\section{Data preparation}
-In any data mining task, adequately preparing the data ensures the best
-chance of success: this includes deciding how to deal with missing and
-inaccurate values, as well as appropriately transforming the data
-\citep{Witten2005} and selecting or constructing the most suitable features
-for the task \citep{Kotsiantis2006}.
-In this section we describe how we addressed each of
-the major issues and the reasons for our choice.
-
-\subsection{Cleaning and normalisation} % missing values, irrelevant features
-Since the data was initially in an Excel spreadsheet, we explored the data
-by hand using Microsoft Excel's filter function on each column and
-consolidated many categorical values simply
-by correcting spelling and punctuation. For example:
-\begin{itemize}
-  \item In the \textbf{sex} attribute, there were values of `Female' and
-  `Femal' which were consolidated into `Female'.
-  \item In the \textbf{mechanism} attribute, which was textual and indicated
-  the category of how the patient was injured, there were such categories as
-  `Other Vehi' and `Other Vehicle' as well as `Pedal Cycl' and `Pedal Cyclist'.
-  Such categories were combined into `Other Vehicle' and `Pedal Cyclist'
-  respectively.
-  \item Some categories in \textbf{mechanism} were also divided into whether or
-  not the mechanism of injury was self-inflicted, such as `Shooting' and
-  `Shooting-selfinflicted'. When we asked the domain expert, Michael Dinh, his
-  suggestion was to combine them into one category: in this case, simply
-  `Shooting'; this was done for all the categories in this attribute.
-  \item \textbf{disp from ED} indicated the section of the hospital that the
-  patient was discharged to. This was also a textual categorical attribute.
-  There were redundancies like `G. ICU' and `G.ICU' and even `GICU', which we
-  all combined into one category, `GICU'.
-\end{itemize}
-
-The data was then exported into CSV (comma-separated value) format, so that
-we could continue to process it programmatically rather than manually. Many
-fields in the data set contained leading or trailing whitespace, which was
-removed with a few lines of Python code (see appendix \todo{[add reference
-to appendix]}). All attribute names were converted to lowercase. In the case
-where attribute names consisted of several words, the spaces were replaced
-with underscores (\_) to ensure that the full attribute name was read.
-\todo{Include a list of these converted attribute names? In the appendix?}
-
-In order for our results to be comparable to that of Dinh
-et. al. \citep{Dinh2013a}, we removed all instances with missing values,
-leaving 2517 instances with which to use to create our prediction models.
-
-\subsection{Feature selection}
-After data preparation, selecting a subset of features or attributes to use
-in a model is also important, and helps to reduce noise and redundant
-attributes from becoming part of the final model. Experiments have found that
-the performance of various classification methods such as decision trees
-\textit{deteriorates} in the presence of useless attributes.
-Our data set contains 73
-non-target attributes, which is a fairly high number. Many of these attributes
-are derived from others in the set, such as \textbf{age} and \textbf{age65}.
-
-We began the process of feature selection by removing attributes that were
-not available upon admission, as these tended to be highly related to the
-LOS48 outcome and were a consequence, not a predictor, of LOS. These attributes
-were: iculos, outcome, disp from ED, disp from hosp, died, rehab and los.
-The id attribute was also removed because it simply numbered each record and
-indicated nothing about the patient themself. This left us with 67 attributes.
-
-The work of actually selecting the attributes out of the 67 available was
-carried out in two ways: automated using feature selection methods, and
-manually using advice from a domain expert.
-
-\subsubsection{Automated feature selection methods}
-There are a number of methods that are commonly used to automate the selection
-of a subset of features for the final classifier. The methods we used were:
-\begin{itemize}
-  \item \textit{Correlation-based feature selection} \citep{Hall2000},
-  which evaluates the `merit' of a
-  subset of features by considering their degree of correlation with the target
-  variable compared to their degree of correlation amongst each other. High
-  correlation between variables indicates redundancy. Under
-  this scheme, a set of features that are highly correlated with the target
-  while having low correlation with other features in the set is ideal.
-  \item Selection based on the
-  \textit{Pearson product-moment correlation coefficient}, a measure of the
-  linear correlation or dependence between two variables.\todo{add some more info}
-  \item \textit{Information gain} \todo{explain a bit about this}
-  \item Feature selection based on the \textit{OneR classifier}
-  \citep{Holte1993}, which uses one
-  attribute as the predictor for the target. This attribute is typically the one
-  that has the lowest prediction error. Feature selection using this scheme
-  involves first ranking attributes based on their error, with the lowest error
-  attributes as the most desirable. We then selected the top 20 attributes to
-  use in the building of the model.
-\end{itemize}
-
-We used the WEKA 3.7.11 \citep{Hall2009} implementation of the above
-methods, after converting the data into the Attribute-Relation File Format
-(ARFF)\footnote{\url{http://weka.wikispaces.com/ARFF+\%28developer+version\%29}}
-most suitable for the program.
-\todo{give more details, ie specifying binary/numeric
-attributes, maybe specify in appendix?}
-
-\subsubsection{Expert selection}
-We also wanted to compare the feature selection of an expert in the domain
-of the data set, namely trauma, with the models obtained from the features
-selected using the automated methods described above. This resulted in two
-feature subsets. The first was obtained simply by presenting the entire list
-of 73 features to the domain expert and asking him to select the features he
-thought were most indicative of the target variable from experience. The
-second set was obtained from the work of Dinh et. al. \citep{Dinh2013a} that
-we are building on, giving us a baseline to compare our results to.
-
-Altogether, the various feature selection methods resulted in a total of
-7 data sets (4 automated, 2 expert selected, and 1 with all features) which
-we used in the model construction phase.
-
-\section{Model construction}
-Here we present the various classification schemes we used in order to find
-the best model for the trauma LOS data.
-
-\subsection{Single classifier}
-First we sought to compare the suitability and performance of various
-classifiers by themselves (later we looked at \textit{ensembles} of
-classifiers, when more than one is involved in making a prediction).
-These are the
-classifiers studied: \todo{expand a bit more here, maybe list the
-configuration parameters}
-\begin{itemize}
-  \item Na\"{i}ve Bayes
-  \item Decision tree (C4.5)
-  \item Logistic regression
-  \item Support vector machine
-  \item $K$-nearest neighbour
-  \item Feed-forward neural network (multilayer perceptron)
-\end{itemize}
-
-Each classifier was trained using ten-fold cross-validation for each of the
-7 data sets obtained from the various feature selection methods described
-above. This was repeated 30 times for each classifier and the reasons for
-this, as well as the metrics used to assess the performance of these
-classifiers, are described in Section \ref{sec:perfeval} of this chapter.
-
-In addition to the feature selection methods, we also used the WEKA 3.7.11
-\citep{Hall2009} implementation of the above classifiers for our experiments.
-Training was automated using a Bash shell script (whose source can be found
-in the appendix) that programmatically read in a list of classifiers to
-train and repeated the procedure any specified number of times, writing the
-results to a specified directory for later reference.
-
-\subsubsection{Parameter tuning}
-It is difficult in data mining to apply an algorithm out-of-the-box and have
-it perform optimally on a given data set \citep{Witten2005}. Therefore, time
-must be spent tweaking and tuning the parameters of a classifier in order to
-improve classification performance. This was the next step we took. We
-considered a range of values for various parameters of each classification
-scheme, and tried some of these new configurations by hand and some by using
-WEKA's cross-validation parameter selection class. The configurations tested,
-as well as how they were executed, are listed in this table:
-\todo{insert table!}
-
-\subsection{Ensemble classifiers}
-In order to improve upon the performance of individual classifiers, we also
-tested the performance of ensembles of classifiers -- that is, combining the
-predictions of several different classifiers, or several of the same
-classifier. Specifically, the ensemble methods used were:
-\begin{itemize}
-  \item Random Forests
-  \item Bagging
-  \item Stacked generalisation (stacking)
-  \item Boosting (AdaBoostM1)
-\end{itemize}
-
-Again, we were able to create predictive models using the implementations of
-these ensemble methods in WEKA 3.7.11. Since ensemble classifiers are not
-actual classifiers but a way to train and combine the predictive power of more
-than one classifier, we list the combinations of classifiers used in our
-ensemble methods:
-\todo{insert table!}
 

### 2014-10-18 thesis/method/features.tex
Restructure of experimental method section #12

--- 2014-09-24 thesis/experiments/experiments.tex
@@
-\chapter{Experimental Setup} \label{chap:experiments}
-
-In this chapter we detail the steps carried out as part of our investigation.
-We first outline the nature of the data set, describe how it was cleaned and
-pre-processed before it was used to train various classifiers. The methods of
-evaluation are then presented, as these are critical to assessing the
-performance of the models created.
-
-\section{Trauma patient data set}
-
-\subsection{Collection}
-The data set used consisted of trauma registry data from the trauma centre
-at the Royal Prince Alfred Hospital, a major trauma centre in New South Wales,
-Australia. It covered all adult (age 15 and over) inpatient admissions to the
-trauma centre from 2007--2011.
-
-All patients were first admitted to the trauma ward until discharged
-or transferred to an appropriate unit within the hospital. A single trained
-data manager recorded a variety of attributes about the admitted patient,
-such as age, gender, blood pressure, mechanism of injury and body regions
-that were injured. We received this data in a Microsoft Excel spreadsheet.
-
-\subsection{Characteristics}
-There were 2546 patient records in the data set we received, comprising of 74
-features, one of which was the target variable \textbf{LOS48}. LOS48 was a
-binary variable -- that is, it could only take two values, 0 or 1 -- with 1
-indicating that the patient stayed two days or less, and 0 indicating a stay
-of greater than 2 days.
-
-\todo{Perhaps insert some more information about the data itself, such
-as the various attributes and their data types? Not sure if that should go
-here, in the results, or in an appendix -- it is relevant but can disturb
-the flow of the main text.}
+\chapter{Feature Preprocessing and Selection} \label{chap:features}
 
@@
 
+All numeric attributes were normalised into the range $[-1,+1]$
+
 \subsection{Feature selection}
@@
 
-\section{Model construction}
-
-\subsection{Single classifier}
-
-\subsubsection{Parameter tuning}
-
-\subsection{Ensemble classifiers}
-
-\section{Performance evaluation} % don't forget cross-validation
-
-\subsection{Metrics}
-
-\subsection{Statistical tests}
+Altogether, the various feature selection methods resulted in a total of
+7 data sets (4 automated, 2 expert selected, and 1 with all features) which
+we used in the model construction phase.

### 2014-10-19 code/weka/run_weka.sh
Fixes input file bug in run weka script #20

@@ -107,12 +107,14 @@ JVM_HEAPSIZE=1024m
 
 # Output file base configuration
 PREFIX=weka_result
-CONFIG="$WEKACONFIG" # indicates which config was used
+CONFIG=`basename $WEKACONFIG` # indicates which config was used
 OUT_FOLDER=../../data/results/
 TIMESTAMP=`date +%Y%m%d%H%M%S`
 OUTFILE=$PREFIX$TIMESTAMP${CONFIG}_$RUNS # use a single file naming scheme for all runs
 ROC_FOLDER=${OUT_FOLDER}roc/$OUTFILE # sub folder for roc threshold outputs
 RESULTFILE=$OUT_FOLDER$OUTFILE
+
+echo "RESULTFILE: $RESULTFILE"
 
 # Create the output folders if they don't exist
 if [ ! -d "$OUT_FOLDER" ]; then

### 2014-10-19 thesis/method/classification.tex
Wrote introduction to classification and moved stuff to evaluation #12

@@
-\chapter{Classification and Supervised Learning Algorithms}
+\chapter{LOS Classification and Supervised Learning}
   \label{chap:classification}
 
-\section{Model construction}
-Here we present the various classification schemes we used in order to find
-the best model for the trauma LOS data.
+In order to use a supervised learning algorithm to identify objects with
+a particular classification, we first need \textit{feature vectors} to
+describe these objects. In our case, each patient is converted into one
+feature vector, where each \textit{feature} in this vector tells us
+something about the
+patient that can help us decide which LOS category he or she should be
+placed in. The features comprising each feature vector must be the same,
+but with different values. By comparing feature values between different
+vectors, we are able to decide which \textit{class} or category a patient
+belongs in. For example, if we wanted to classify the weather on a given
+day as ``good'' or ``bad'', we could use two features -- temperature and
+the amount of rain -- which are likely to be key indicators. Classifying
+several days simply involves creating a feature vector that corresponds
+to each day, with differing values for the temperature and the amount of
+rain, and using these values to decide whether or not a day had good or
+bad weather.
 
-\subsection{Single classifier}
-First we sought to compare the suitability and performance of various
-classifiers by themselves (later we looked at \textit{ensembles} of
-classifiers, when more than one is involved in making a prediction).
-These are the
-classifiers studied: \todo{expand a bit more here, maybe list the
-configuration parameters}
-\begin{itemize}
-  \item Na\"{i}ve Bayes
-  \item Decision tree (C4.5)
-  \item Logistic regression
-  \item Support vector machine
-  \item $K$-nearest neighbour
-  \item Feed-forward neural network (multilayer perceptron)
-\end{itemize}
+Since we will be using supervised learning techniques, we need
+\textit{training examples}, which consist of a set of feature vectors
+that are already labelled with the class that it belongs in. These
+training examples are then used by a classifier to ``learn'' the
+relationship linking the feature vectors to their class labels. In this
+chapter we describe each of the learning algorithms we used to model the
+relationship between the feature vectors and the LOS.
 
-Each classifier was trained using ten-fold cross-validation for each of the
-7 data sets obtained from the various feature selection methods described
-above. This was repeated 30 times for each classifier and the reasons for
-this, as well as the metrics used to assess the performance of these
-classifiers, are described in Section \ref{sec:perfeval} of this chapter.
+\section{ZeroR and OneR}
 
-In addition to the feature selection methods, we also used the WEKA 3.7.11
-\citep{Hall2009} implementation of the above classifiers for our experiments.
-Training was automated using a Bash shell script (whose source can be found
-in the appendix) that programmatically read in a list of classifiers to
-train and repeated the procedure any specified number of times, writing the
-results to a specified directory for later reference.
+\section{Na\"{i}ve Bayes}
 
-\subsubsection{Parameter tuning}
-It is difficult in data mining to apply an algorithm out-of-the-box and have
-it perform optimally on a given data set \citep{Witten2005}. Therefore, time
-must be spent tweaking and tuning the parameters of a classifier in order to
-improve classification performance. This was the next step we took. We
-considered a range of values for various parameters of each classification
-scheme, and tried some of these new configurations by hand and some by using
-WEKA's cross-validation parameter selection class. The configurations tested,
-as well as how they were executed, are listed in this table:
-\todo{insert table!}
+\section{Decision Tree (C4.5)}
 
-\subsection{Ensemble classifiers}
-In order to improve upon the performance of individual classifiers, we also
-tested the performance of ensembles of classifiers -- that is, combining the
-predictions of several different classifiers, or several of the same
-classifier. Specifically, the ensemble methods used were:
-\begin{itemize}
-  \item Random Forests
-  \item Bagging
-  \item Stacked generalisation (stacking)
-  \item Boosting (AdaBoostM1)
-\end{itemize}
+\section{Logistic Regression}
 
-Again, we were able to create predictive models using the implementations of
-these ensemble methods in WEKA 3.7.11. Since ensemble classifiers are not
-actual classifiers but a way to train and combine the predictive power of more
-than one classifier, we list the combinations of classifiers used in our
-ensemble methods:
-\todo{insert table!}
+\section{Support Vector Machine (SVM)}
+
+\section{$k$-Nearest Neighbour}
+\subsection{Basic Algorithm}
+
+\subsection{Extensions}
+\subsubsection{K*: an Entropy-Based Distance Function}
+
+\subsubsection{Ranked Distance}
+
+\section{Feed-forward Neural Network}

### 2014-10-19 thesis/method/evaluation.tex
Wrote introduction to classification and moved stuff to evaluation #12

@@
 above.
+
+\subsection{Single classifier}
+First we sought to compare the suitability and performance of various
+classifiers by themselves (later we looked at \textit{ensembles} of
+classifiers, when more than one is involved in making a prediction).
+These are the
+classifiers studied: \todo{expand a bit more here, maybe list the
+configuration parameters}
+\begin{itemize}
+  \item Na\"{i}ve Bayes
+  \item Decision tree (C4.5)
+  \item Logistic regression
+  \item Support vector machine
+  \item $K$-nearest neighbour
+  \item Feed-forward neural network (multilayer perceptron)
+\end{itemize}
+
+Each classifier was trained using ten-fold cross-validation for each of the
+7 data sets obtained from the various feature selection methods described
+above. This was repeated 30 times for each classifier and the reasons for
+this, as well as the metrics used to assess the performance of these
+classifiers, are described in Section \ref{sec:perfeval} of this chapter.
+
+In addition to the feature selection methods, we also used the WEKA 3.7.11
+\citep{Hall2009} implementation of the above classifiers for our experiments.
+Training was automated using a Bash shell script (whose source can be found
+in the appendix) that programmatically read in a list of classifiers to
+train and repeated the procedure any specified number of times, writing the
+results to a specified directory for later reference.
+
+\subsubsection{Parameter tuning}
+It is difficult in data mining to apply an algorithm out-of-the-box and have
+it perform optimally on a given data set \citep{Witten2005}. Therefore, time
+must be spent tweaking and tuning the parameters of a classifier in order to
+improve classification performance. This was the next step we took. We
+considered a range of values for various parameters of each classification
+scheme, and tried some of these new configurations by hand and some by using
+WEKA's cross-validation parameter selection class. The configurations tested,
+as well as how they were executed, are listed in this table:
+\todo{insert table!}
+
+\subsection{Ensemble classifiers}
+In order to improve upon the performance of individual classifiers, we also
+tested the performance of ensembles of classifiers -- that is, combining the
+predictions of several different classifiers, or several of the same
+classifier. Specifically, the ensemble methods used were:
+\begin{itemize}
+  \item Random Forests
+  \item Bagging
+  \item Stacked generalisation (stacking)
+  \item Boosting (AdaBoostM1)
+\end{itemize}
+
+Again, we were able to create predictive models using the implementations of
+these ensemble methods in WEKA 3.7.11. Since ensemble classifiers are not
+actual classifiers but a way to train and combine the predictive power of more
+than one classifier, we list the combinations of classifiers used in our
+ensemble methods:
+\todo{insert table!}
 

### 2014-10-20 thesis/method/classification.tex
Wrote about 0R and 1R and created acronyms for the classifiers #12

@@
 training examples are then used by a classifier to ``learn'' the
-relationship linking the feature vectors to their class labels. In this
+relationship linking the feature vectors to their class labels. Once we have
+``trained'' the classifier to learn this relationship with the pre-labelled
+data, it will then be able to predict the class of any feature vector that we
+give it.
+
+In this
 chapter we describe each of the learning algorithms we used to model the
@@
 
-\section{ZeroR and OneR}
+\section{0R and 1R}
+Short for ``0-rule'' and ``1-rule'' respectively, 0R and 1R are
+very simple classifiers. 0R uses no features to predict the class, and
+1R only uses one.
 
-\section{Na\"{i}ve Bayes}
+Given a data set, the 0R classification algorithm will find the class that
+occurs the most often, and predict this class for all new examples. For LOS
+classification, if one class in a data set was ``LOS $\leq$ 2 days'' and
+the other was
+``LOS $>$ 2 days'', and feature vectors (which represent patients in this
+scenario) were more commonly labelled ``LOS $\leq$ 2 days'', then 0R will
+predict all unseen patients with ``LOS $\leq$ 2 days''.
+The prediction decision of 0R is very easy to understand,
+and its results also provide an apt baseline for evaluating the more
+sophisticated classifiers used in this work.
 
-\section{Decision Tree (C4.5)}
+1R works by considering each feature individually.
+For each of the values of the
+feature, 1R finds the class with the highest frequency and creates a rule
+from this observation. If the feature is numeric (that is, its values are
+continuous rather than discrete), then 1R will partition the spectrum of
+values into discrete buckets. \todo{describe buckets more}
+The error rate, expressed as the ratio of correct predictions to the total
+number of examples in the data set,
+is calculated for each rule that is generated
+in this way. The rule with the lowest error rate is then chosen as the 1-rule
+for this classifier,
+and if there is a tie, it is broken randomly. As with 0R, the prediction
+decision of 1R is transparent and it provides another baseline upon which
+more sophisticated classifiers can be evaluated.
 
-\section{Logistic Regression}
+Concretely, given a feature
+vector of the form \textit{(age, gender, has\_cancer)},
+1R will split the ages of
+all patients into discrete categories
+(such as \textit{0-20, 21-40, 41-65, 66+}). It
+will then check the class labels for all training examples in the data set
+that have \textit{age} in the 0-20 category
+(as that is the first value for the partitioned \textit{age} feature) and
+note the most common class, proceeding similarly for the rest of the
+categories of \textit{age}. Then it will consider \textit{gender}, which has
+two possible values $\{$\textit{male, female}$\}$, and finally \textit{has\_cancer}
+with two values $\{$\textit{true, false}$\}$. This results in 8 rules,
+corresponding to three features.
+An example of a rule is that \textit{age 0-20} most commonly
+has ``LOS $\leq$ 2 days''. For each of these rules, 1R will attempt to
+use only that rule to predict the LOS class for each example in the data set,
+noting down the error rate. 1R then chooses the feature that produces rules
+with the lowest total error rate. \todo{maybe add a table}
+
+\section{Na\"{i}ve Bayes (NB)}
+Unlike 1R, which considers only one feature in the
+
+\section{C4.5 Decision Tree (DT)}
+
+\section{Logistic Regression (LR)}
 
@@
 
-\section{$k$-Nearest Neighbour}
+\section{$k$-Nearest Neighbour (kNN)}
 \subsection{Basic Algorithm}
@@
 
-\section{Feed-forward Neural Network}
+\section{Feed-Forward Neural Network}

### 2014-10-20 thesis/method/classification.tex
Wrote up Naive Bayes, logistic regression, and some k-NN

@@
 continuous rather than discrete), then 1R will partition the spectrum of
-values into discrete buckets. \todo{describe buckets more}
+values into discrete buckets. This is known as \textit{discretisation}.
 The error rate, expressed as the ratio of correct predictions to the total
@@
 that have \textit{age} in the 0-20 category
-(as that is the first value for the partitioned \textit{age} feature) and
+(as that is the first value for the discretised \textit{age} feature) and
 note the most common class, proceeding similarly for the rest of the
@@
 noting down the error rate. 1R then chooses the feature that produces rules
-with the lowest total error rate. \todo{maybe add a table}
+with the lowest total error rate.
 
 \section{Na\"{i}ve Bayes (NB)}
-Unlike 1R, which considers only one feature in the
+Unlike 1R, which considers only one feature in the feature vector, Na\"{i}ve
+Bayes (NB) classifiers allows all features to contribute equally to the
+classification decision. Such classifiers apply Bayes' theorem with the
+(unrealistic) assumption of independence between features, hence the
+description \textit{na\"{i}ve}. This means that we
+calculate the probability that a particular feature value belongs in a class
+without considering the effect of other features on this probability. This
+probability can be computed by
+counting the frequency with which the value is associated with a class in the
+training data. After these probabilities are computed, classifying a feature
+vector simply involves multiplying the probabilities of each feature belonging
+to each possible class. The predicted class for that feature vector is then the
+class with the highest probability.
+
+Speaking mathematically, Bayes' theorem states that:
+\begin{equation*}
+\mathbb{P}(Class_i|F_1,F_2,\ldots,F_n) =
+  \dfrac{\mathbb{P}(F_1,F_2,\ldots,F_n|Class_i)\mathbb{P}(Class_i)}{\mathbb{P}(F_1,F_2,\ldots,F_n)}
+\end{equation*}
+Since we assume that the features are independent, we can simplify the above as:
+\begin{equation*}
+\mathbb{P}(Class_i|F_1,F_2,\ldots,F_n) =
+\dfrac{\mathbb{P}(F_1|Class_i)\mathbb{P}(F_2|Class_i)\ldots \mathbb{P}(F_n|Class_i)\mathbb{P}(Class_i)}{\mathbb{P}(F_1,F_2,\ldots,F_n)}
+\end{equation*}
+This gives us the probability that the feature vector $(F_1,F_2,\ldots,F_n)$
+belongs in the $i$-th class. Classifying the feature vector involves
+calculating this probability for each for each class and choosing the class
+with the highest probability.
+
+Since we compute the probability of each feature value belonging in a
+particular class by counting the frequency of each value's occurrence with
+respect to the class, it is sometimes possible that a feature value has zero
+probability of being in a certain class. This would make the numerator of the
+above equation zero. A feature vector that could otherwise have been highly
+likely to belong in class $i$ is now counted as having zero probability of
+belonging in the class. In order to ensure that each probability is non-zero,
+a small constant is added to all probabilities in a process called
+\textit{Laplace smoothing}, which is commonly used for NB classifiers.
+
+Like 1R, numeric features also need to be handled separately from nominal ones.
+The typical approach is to assume that the values of a numeric feature follow a
+normal (or Gaussian) distribution with mean and standard deviation calculated
+from the values that the feature takes with respect to each class.
+This allows us to use the probability
+density function of the normal distribution to calculate the probabilty of the
+feature belonging to a particular class:
+\begin{equation*}
+\mathbb{P}(F_i=x_i|Class_i) =
+  \dfrac{1}{\sqrt{2\pi}\sigma}e^\frac{(x_i-\mu)^2}{2\sigma^2}
+\end{equation*}
+
+Despite the na\"{i}ve independence assumption of the NB classifier, it often
+works surprisingly well in practice and is relatively fast to train. We use it
+in our work both as a simple classifier in its own right, and also as a more
+sophisticated baseline to 0R and 1R.
 
@@
 \section{Logistic Regression (LR)}
+Logistic regression is a regression technique used for when the dependent
+variable is dichotomous or two-valued, and thus it is a suitable candidate to
+use for our LOS classification problem.
+Previous work on trauma LOS prediction by
+Dinh et. al. \cite{Dinh2013a}, as well as many attempts at modelling the LOS
+in various medical domains, have all used logistic regression, often as a
+starting point.
+
+For a class variable with two possible values that occur with
+probability $p$ and $1-p$ respectively, and a feature vector
+$(F_1,F_2,\ldots,F_n)$, the logistic regression model is:
+\begin{equation*}
+\mathrm{logit}(p) &=& \mathrm{log} \left(\dfrac{p}{1-p}\right) \\
+  &=& \beta_0 + \beta_1F_1 + \beta_2F_2 + \ldots + \beta_nF_n
+\end{equation*}
+This can be rearranged to:
+\begin{equation*}
+p &=& \mathbb{P}(Class=\mathrm{LOS }>2\mathrm{ days}|F_1,F_2,\ldots,F_n) \\
+  &=& \dfrac{e^{\beta_0 + \sum^n_{i=1}\beta_i F_i}}{1 + e^{\beta_0 + \sum^n_{i=1}\beta_i F_i}}
+\end{equation*}
+
+Training the logistic regression classifier, then, is a matter of finding the
+$\beta_i$s that will maximise the likelihood (conditional probability of the
+data given the values for $\beta_i$) of the given data set. This is done using
+numerical methods, where estimates of the $\beta_i$s are chosen, the likelihood
+computed, and the $\beta_i$ are updated, and so on: this process is repeated
+until parameter estimates converge or do not change more than a given threshold
+value.
 
@@
 
-\section{$k$-Nearest Neighbour (kNN)}
+\section{$k$-Nearest Neighbour ($k$-NN)}
+In this section we will describe the $k$-nearest neighbour ($k$-NN) algorithm
+and discuss some extensions which we use in our work. Additionally, we will
+describe our contribution, which is an extension of the distance metric for
+finding the nearest neighbours.
+
 \subsection{Basic Algorithm}
+The $k$-nearest neighbour classifier is an example of an
+\textit{instance-based} learning algorithm. This is because it does not, unlike
+the other classifiers described in this section, attempt to deduce or
+generalise a relationship between the features and the class: it simply stores
+all training examples and classifies an unseen example by finding the $k$
+``nearest'' examples (or \textit{instances}) and assigning the majority class
+to the new example. In the simplest form of the algorithm, $k$ is chosen to be
+1 and the Euclidean distance is used to measure the closeness of neighbours:
+\begin{equation*}
+\mathrm{Distance} = \sqrt{(x_1-y_1)^2 + (x_2-y_2)^2 + \ldots + (x_n-y_n)^2}
+\end{equation*}
+where $x_i,y_i$ are the values of the $i$-th feature for two different feature
+vectors (or instances).
+
+Like the Na\"{i}ve Bayes classifier, $k$-NN allows all features to be taken
+into account equally when classifying a new example. However, since feature
+values often span different ranges and units of measurement, a particularly
+large range of values for a feature would contribute much more to the distance
+than another feature that spans a lower set of values.
+To avoid this, it is important to \textit{normalise} the
+values of each numeric feature before using $k$-NN: for each feature $i$, we
+calculate the normalised value of the feature for each training example. This
+is done using the following relationship:
+\begin{equation*}
+a_i = \dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}
+\end{equation*}
+where $a_i$ is the normalised value, $v_i$ is the actual value in the data
+set, and max $v_i$ and min $v_i$ are taken over all training examples.
 
 \subsection{Extensions}
+There have been many extensions proposed to the $k$-NN algorithm described
+above.
+
 \subsubsection{K*: an Entropy-Based Distance Function}
@@
 
-\section{Feed-Forward Neural Network}
+\section{Multi-layer Perceptron (MLP)}

### 2014-10-21 code/weka/run_weka.sh
Made run weka script output more manageable by overwriting the last line

@@ -140,6 +140,9 @@ readarray -t CLFS < "$WEKACONFIG"
 # sed is run twice to remove a possible double occurrence of $OPTIONS
 NUM_LINES=${#CLFS[@]}
 LINENUM=0
+PREV_LINE=`tput cuu1`
+CLEAR_LINE=`tput el`
+echo "Begin training..."
 while [ $LINENUM -lt $NUM_LINES ]; do
   for run in $(seq 1 $RUNS)
   do
@@ -148,6 +151,8 @@ while [ $LINENUM -lt $NUM_LINES ]; do
     else
       SEED=1
     fi
+    # clear the previous line and write over it so that the output is more readable
+    echo -en "${PREV_LINE}\r${CLEAR_LINE}"
     echo "Training classifier $[$LINENUM+1] run #$run with seed $SEED: ${CLFS[LINENUM]}"
     if [ $OUTPUT_ROC = true ]; then
       ROC_FILE="${LINENUM}_$(echo ${CLFS[LINENUM]} | awk '{print $1}')_$run.arff"

### 2014-10-21 thesis/method/classification.tex
More writing about classification algorithms...

@@
-\chapter{LOS Classification and Supervised Learning}
+\chapter{LOS Classification Using Supervised Learning}
   \label{chap:classification}
@@
 decision of 1R is transparent and it provides another baseline upon which
-more sophisticated classifiers can be evaluated.
+more sophisticated classifiers can be evaluated. Studies have also found that
+1R performs quite well in terms of accuracy relative to other classifiers
+\cite{Holte1993}.
 
@@
 $(F_1,F_2,\ldots,F_n)$, the logistic regression model is:
-\begin{equation*}
-\mathrm{logit}(p) &=& \mathrm{log} \left(\dfrac{p}{1-p}\right) \\
-  &=& \beta_0 + \beta_1F_1 + \beta_2F_2 + \ldots + \beta_nF_n
-\end{equation*}
-This can be rearranged to:
-\begin{equation*}
-p &=& \mathbb{P}(Class=\mathrm{LOS }>2\mathrm{ days}|F_1,F_2,\ldots,F_n) \\
-  &=& \dfrac{e^{\beta_0 + \sum^n_{i=1}\beta_i F_i}}{1 + e^{\beta_0 + \sum^n_{i=1}\beta_i F_i}}
-\end{equation*}
+\begin{align*}
+\mathrm{logit}(p) &= \mathrm{log} \left(\dfrac{p}{1-p}\right) \\
+  &= \beta_0 + \beta_1F_1 + \beta_2F_2 + \ldots + \beta_nF_n
+\end{align*}
+For LOS classification, this can be rearranged to:
+\begin{align*}
+p &= \mathbb{P}(Class=\mathrm{LOS }\leq2\mathrm{ days}|F_1,F_2,\ldots,F_n) \\
+  &= \dfrac{e^{\beta_0 + \sum^n_{i=1}\beta_i F_i}}{1 + e^{\beta_0 + \sum^n_{i=1}\beta_i F_i}}
+\end{align*}
 
@@
 value.
+
+To classify a feature vector $x = (x_1,x_2,\ldots,x_n)$, we substitute the
+$x_i$'s into the above equation in place of the $F_i$'s, which gives us a
+probability. If this probability is above a particular threshold, then we
+classify the example as belonging to the class ``LOS $\leq$ 2 days'', otherwise
+it is classified as ``LOS $>$ 2 days''. Usually this threshold is 0.5.
 
@@
 There have been many extensions proposed to the $k$-NN algorithm described
-above.
+above. These usually seek to find a more suitable distance metric for the
+problem domain at hand, or to reduce the storage requirements of the original
+$k$-NN algorithm. We will focus on two extensions to the Euclidean distance:
+K* and our contribution, Ranked Distance.
 
-\subsubsection{K*: an Entropy-Based Distance Function}
+\subsubsection{K*: Instance-based Classifier with Entropic Distance Function}
 
 \subsubsection{Ranked Distance}
+\paragraph{Intuition}
+As we mentioned earlier, all features are considered equally in a $k$-NN
+classifier, which implies that all features are equally important in predicting
+the class. In practice, however, this is not always the case: to predict the
+LOS of a trauma patient, the severity of their injury should affect the final
+LOS more than whether or not they can speak English (which could be just two
+of many features that are recorded about a patient). Our contribution is thus
+a modified distance function to take the relative importance of attributes into
+account.
+
+\paragraph{Mathematical formulation}
+Given two instances $x = (x_1,x_2,\ldots,x_n)$ and $y = (y_1,y_2,\ldots,y_n)$,
+the distance between them is given by:
+\begin{equation*}
+\mathrm{Distance} = \sum_{i=1}^n w_i |x_i-y_i|
+\end{equation*}
+where the $w_i$s weight the contribution of the $i$-th feature to the overall
+distance that is computed by the $k$-NN algorithm. We assume that all feature
+values have been normalised, as described above.
+
+\paragraph{Assignment of weights}
+The weights $w_i$ can be tuned to match the particular problem that is being
+investigated. For our LOS classification problem, we will consider two ways
+of weighting features:
+\begin{enumerate}
+\item The weight of each feature is the magnitude of the correlation
+coefficient
+between it and the class, meaning that features that are more highly correlated
+with the class will contribute more to the distance. The correlation
+coefficient $r_i$ for feature $i$ with the class $c$ can be computed by:
+\begin{equation*}
+r_i = \dfrac{\sum_{i=1}^m (x_i-\bar{x})(c-\bar{c})}{\sqrt{(\sum_{i=1}^m x_i-\bar{x})(\sum_{i=1}^m c-\bar{c})}}
+\end{equation*}
+where $m$ is the number of training examples, $x_i$ is the value of feature
+$i$, and $\bar{x}$ and $\bar{c}$ are the arithmetic averages of the feature
+values and the class values respectively. 
+\item We compute the correlation coefficients as described in the point above.
+This gives us a ranking of the importance of each feature to the prediction of
+the class, with 1 being the highest rank (indicating the most importance) and
+$n$ being the lowest rank (indicating the least importance).
+Instead of using the values of the correlation directly to weight
+the distance contribution of each feature, we specify a function:
+\begin{equation*}
+f : \mathrm{Rank} \rightarrow \mathbb{R}, \mathrm{Rank} \in \{1,2,\ldots,n\}
+\end{equation*}
+that describes
+how the weights of the features vary with the rank. This function should
+inituitively be non-increasing and should produce lower values when the rank
+number is greater, meaning that less important features have a lower weight
+in the distance calculation.
+Note that choosing $f(\mathrm{Rank}) = k$ for any non-zero constant $k$
+results in all features contributing equally to the distance calculation.
+
+\noindent We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}}$ and
+$f(\mathrm{Rank}) = n-\mathrm{Rank}$.
+\end{enumerate}
 
 \section{Multi-layer Perceptron (MLP)}
+
+\section{Conclusion}

### 2014-10-22 thesis/method/classification.tex
Added outline of decision trees

@@
 \section{C4.5 Decision Tree (DT)}
+Decision trees (DT) can be specified recursively and are an example of a
+``divide-and-conquer'' approach to learning. First, we choose a feature to
+place at the root node and then create one branch for each possible value.
+This partitions the data set into subsets, one for every value that the
+feature takes. We repeat this for each branch, choosing different features
+each time and continuing to divide the data set into progressively smaller
+subsets, growing the tree.
+If all examples in a subset have the same class at any point during
+this process, we stop splitting that subset.
+
+The only thing left to specify
+now is how to decide which attribute to split on. Note that when splitting,
+we would like to create subsets that are as ``pure''
+-- that is, containing as many
+instances of the same class -- as possible, so that the splitting may
+terminate earlier. Since we would like our trees to be as small as possible to
+prevent overfitting and allow the best generalisation ability,
+we would therefore like to choose the feature to split on such that the
+created subsets were as pure as possible. The measure of purity used in
+learning decision trees is the \textit{information}, and is measured in
+\textit{bits}. At any node (or subset) of the tree, this measure represents
+the expected amount of information that would be needed to classify a new
+example, given that the example reached the node. The information at each
+node is calculated using the \textit{entropy} function:
+\begin{equation*}
+\mathrm{entropy}(p_1,p_2,\ldots,p_n) = -p_1\mathrm{log}p_1 -p_2\mathrm{log}p_2 \ldots -p_n\mathrm{log}p_n
+\end{equation*}
+Given $k$ classes, $c_1,\ldots,c_k$ is the number of instances in each
+class at a node, and $\sum_{i=1}^k c_i = c$ is the total number of instances
+at that node, the information of that node is then:
+\begin{equation*}
+\mathrm{info}(c_1,\ldots,c_k) = \mathrm{entropy}\left(\frac{c_1}{c},\frac{c_2}{c},\ldots,\frac{c_k}{c}\right)
+\end{equation*}
+
+To find the attribute to split on, we assume we are splitting on a particular
+attribute, and calculate the information at the node given the split. We do
+this for each attribute, and compute the \textit{information gain} from
+splitting on that attribute:
+\begin{equation*}
+\mathrm{gain}(feature) = \mathrm{info}(\text{child node}) - \mathrm{info}(\text{parent node})
+\end{equation*}
+The feature that is chosen is therefore the one with the highest information
+gain.
+
+Decision trees are fast to train and produce models that are straightforward
+for humans to interpret, the latter of which is particularly valuable in a
+clinical environment. We will use the C4.5 tree induction algorithm, which is
+considered the state-of-the-art and handles different types of feature values
+as well as missing values.
 

### 2014-10-22 thesis/method/classification.tex
Misc fixes to equations
Support vector machines!

@@
 $(F_1,F_2,\ldots,F_n)$, the logistic regression model is:
-\begin{align*}
+\begin{equation*}
+\begin{aligned}
 \mathrm{logit}(p) &= \mathrm{log} \left(\dfrac{p}{1-p}\right) \\
   &= \beta_0 + \beta_1F_1 + \beta_2F_2 + \ldots + \beta_nF_n
-\end{align*}
+\end{aligned}
+\end{equation*}
 For LOS classification, this can be rearranged to:
-\begin{align*}
+\begin{equation*}
+\begin{aligned}
 p &= \mathbb{P}(Class=\mathrm{LOS }\leq2\mathrm{ days}|F_1,F_2,\ldots,F_n) \\
   &= \dfrac{e^{\beta_0 + \sum^n_{i=1}\beta_i F_i}}{1 + e^{\beta_0 + \sum^n_{i=1}\beta_i F_i}}
-\end{align*}
+\end{aligned}
+\end{equation*}
 
@@
 \section{Support Vector Machine (SVM)}
+Support vector machines (SVMs) are an example of a supervised learning algorithm
+capable of classifying training examples into two classes. They are one of the
+best algorithms in terms of predictive accuracy \cite{Bellazzi2008}. Despite
+this, they have not been widely applied to predicting the LOS, which gives us
+an opportunity to evaluate their performance in our LOS classification problem.
+
+Given a set of $n$-feature vectors, we can represent each vector as a point in
+$n$-dimensional space, where each feature is a dimension. The SVM will try to
+find a \textit{hyperplane} that separates the two classes. Usually there are
+many such hyperplanes; the distinguishing aspect of an SVM is that it will
+find the hyperplane that maximises the distance between the two nearest points
+in each class. This distance is called the \textit{margin}, and the intuition
+is that maximising it will give us the greatest possible separation between the
+classes -- the classifier comes no closer to either class than it needs to.
+This gives the SVM the best generalising ability to classify new examples.
+
+Figure \ref{fig:svm} shows an example of a two-dimensional feature space, with
+two classes of points (black and white). We can see the hyperplane, in this
+case a line since we are only in two dimensions, separating the two classes
+with as much margin as possible. In
+this case, the points are linearly separable: all white points lie on one side
+of the hyperplane, and all black points lie on the other. In practice, this is
+not common, and often data cannot be separated cleanly by a hyperplane.
+Note that the points closest to the margin are those that are most likely to be
+ambiguous when the data cannot be cleanly separated. The SVM will progressively
+remove more and more of these ambiguous points from consideration when it
+cannot find a hyperplane until it finds a suitable one.
+
+\begin{figure}[h]
+\label{fig:svm}
+\caption{}
+\centering
+%\includegraphics[scale=0.95]{images/a2-nn}
+\end{figure}
+
+Instead of removing points from consideration by the SVM, we can also specify
+a \textit{kernel function}, which transforms the original feature space into
+some other non-linear space. Commonly, polynomial functions of degree 2 or
+greater are used, and theoretically they can approximate non-linear boundaries
+between classes to any arbitrary accuracy. However, this is usually not
+practically possible with data sets containing even a modest number of features
+(around 10), due to the computational complexity of estimating a large number
+of coefficients that are introduced by the transformation.
+
+SVMs can be specified mathematically as a type of optimisation problem known
+as \textit{constrained quadratic optimisation}. Given $m$ training vectors
+$\mathbf{x_i} \in \mathbb{R}^n, i=1,\ldots,m$ and their labels
+$y_i \in \{-1,+1\}, i=1,\ldots,m$, the optimisation problem can be written as:
+\begin{equation}
+\label{eqn:svm1}
+\begin{aligned}
+& \underset{\mathbf{w},b}{\mathrm{minimise}}
+  && ||\mathbf{w}|| + C\sum_{i=1,\ldots,m} s_i \\
+& \text{subject to}
+  && y_i(\mathbf{w}^T\phi(\mathbf{x_i})+b) \geq 1 \\
+& && s_i \geq 0, i = 1,\ldots,m
+\end{aligned}
+\end{equation}
+where $\phi(\mathbf{x})$ is the kernel function, and
+$\mathbf{w}\cdot\mathbf{x_i} + b = 0$ is the hyperplane that the SVM finds.
+The $s_i$'s allow the SVM to ignore some examples if it cannot find a suitable
+hyperplane, and $C$ (sometimes called the SVM's complexity constant) is a
+constant specified by the user that controls how much we want to balance
+finding the best possible margin with allowing misclassified examples.
+Note that for a linear kernel, $\phi(\mathbf{x}) = \mathbf{x}$. Training the
+SVM involves solving this optimisation problem and obtaining values for
+$\mathbf{w}$ and $b$.
+
+To classify an instance $\mathbf{x_{new}}$, we simply find the sign of
+$\mathbf{w}\cdot\mathbf{x_{new}}+b$. This will tell us which one of the
+two classes the new example belongs to.
 

### 2014-10-22 thesis/method/classification.tex
Outlined K* very briefly

@@
 \subsubsection{K*: Instance-based Classifier with Entropic Distance Function}
+Cleary and Trigg developed K*, an instance-based classifier that uses entropy
+to measure the distances between instances \cite{Cleary1995}. They consider
+the distance between two instances to be the complexity of transforming one
+into another. Define a sequence of transformations on an instance:
+\begin{equation*}
+\bar{t}(a) = t_n(t_{n-1}(\ldots t_1(a)\ldots)), \bar{t} \in T
+\end{equation*}
+where $\bar{t} = t_1,\ldots,t_n$. We can define a probability function $p$
+that gives the probabilities of these transformations occurring, and then we
+can define a probability function that gives the probabilities of all paths
+(transformations) from example $a$ to $b$:
+\begin{equation*}
+P^*(b|a) = \sum_{\bar{t}\in T:\bar{t}(a)=b} p(\bar{t})
+\end{equation*}
+Within this framework, we can define transformations for both real-valued and
+categorical features, which allows K* to work with all types of features.
+The K* function is then defined as:
+\begin{equation*}
+K^*(b|a) = -\mathrm{log}_2P^*(b|a)
+\end{equation*}
+
+To calculate the probability of an instance $a$ being in a class $C$, we sum
+the probabilities from $a$ to each instance of $C$, and repeat for each class.
+The predicted class will be the one with the highest probability:
+\begin{equation*}
+P^*(C|a) = \sum_{b\in C} P^*(b|a)
+\end{equation*}
+
+We use K* as a comparison to the basic NN algorithm and also as a comparison
+to $k$-NN with our distance function.
 

### 2014-10-22 thesis/method/classification.tex
Multi-layer perceptron and the conclusion of the chapter

@@
 \section{Multi-layer Perceptron (MLP)}
+Multi-layer perceptrons (MLPs) are a generalised form of the
+\textit{perceptron}, first
+studied extensively by Rosenblatt \todo{add citation entry}. The original
+motivation of the perceptron 
+was to develop a very simple artificial neural network to model the
+behaviour of neurons in the brain. The perceptron model for a two-class
+problem is a linear combination of the features: that is,
+$h(\mathbf{x}) = \sum w_i x_i$ where the $w_i$ are the weights of the model.
+There is an input node, called a \textit{neuron}, for each feature, and the
+weights are applied to the connections between these input neurons to the
+output.
+Usually, a constant term of 1 called the \textit{bias} is also added to the
+model for an additional trainable input. To classify an item, we substitute it
+into $h(\mathbf{x})$; if the value is above a particular threshold, we label it
+as class 1, else it belongs in class 2. For convenience, we can let this
+threshold be 0. Training the perceptron involves the following steps:
+\begin{enumerate}
+\item Start with random numbers for the weights $w_i$.
+\item For each training example $\mathbf{x_i}$: calculate $h(\mathbf{x_i})$.
+If it is correctly classified, do not update the weights.
+If it is incorrectly classified, add a multiple of the misclassified
+instance to the current weights using
+$\mathbf{w} = \mathbf{w} + \lambda \mathbf{x_i}$, where $\lambda$ is a small
+constant.
+\item Repeat the previous step until all training examples are correctly
+classified.
+\end{enumerate}
+If the data are linearly separable, the perceptron is guaranteed to find a
+hyperplane that separates the two classes. However, it is not able to do this
+when there is a non-linear decision boundary -- for that we will need the
+multi-layer perceptron.
+
+The \textit{multi-layer perceptron}, also known as a
+\textit{feed-forward neural network}, is a type of artificial neural network
+that generalises the perceptron model outlined above. Instead of only one layer
+of inputs, MLPs consist of more than one layer, with the outputs from previous
+layers serving as inputs to the next layer. At each layer, the inputs are
+combined linearly and then transformed nonlinearly for input into the next
+layer. For a network with one layer of transformations between the input and
+output layers -- that is, one ``hidden'' layer -- the mathematical description
+is:
+\begin{equation*}
+y = \sum_j w^{(2)}_j f_j \left(\sum_i w^{(1)}_i x_i\right)
+\end{equation*}
+The $w$ and the $f_j$ represent the linear combinations and the nonlinear
+transformations respectively, and $y$ is the predicted outcome.
+The nonlinearity of the transformations ensures
+that we can classify data which are not linearly separable. Figure
+\ref{fig:mlp} shows an MLP with one hidden layer.
+
+\begin{figure}[h]
+\label{fig:mlp}
+\caption{}
+\centering
+%\includegraphics[scale=0.95]{images/a2-nn}
+\end{figure}
+
+MLPs can be trained in a number of ways, the most common of which is
+\textit{backpropagation}. Suppose that we have the network above, with one
+input, hidden and output layer respectively. From the perceptron learning
+algorithm, we can update the weights of the hidden neurons leading to the
+output neuron, but we do not know what the outputs of these hidden neurons are.
+Backpropagation solves this problem by updating the weights of the various
+connections to the hidden units based on the strength of the unit's
+contribution to the final prediction, using a mathematical optimisation
+algorithm called \textit{gradient descent}. This requires taking derivatives,
+so the threshold function we used for the perceptron is usually replaced with
+the sigmoid function $f(x) = \frac{1}{1+e^{-x}}$. We also define an error
+function $E=\frac{1}{2}(\text{actual} - \text{predicted})^2$ -- the one most
+commonly used -- which is
+minimised in the gradient descent process. Updating the weights for each
+connection in the network now involves finding the value of the derivative,
+multiplying it by a small constant between 0 and 1 called the
+\textit{learning rate}, and subtracting this value from the current weight
+value, until the minimum for the error function is reached. Note that the
+error function is quadratic, so it is guaranteed to have a minimum. The
+updating of the weights is continued until the minimum is reached or if we
+are ``close enough'' to the minimum.
+
+It can be shown that MLPs with a single hidden layer are powerful enough to
+model any continuous function, provided there are enough neurons in that
+hidden layer \cite{Hand2001}. However, as they must iterate over the entire
+data set many times, and due to the sometimes slow convergence of gradient
+descent, MLPs are often slower to train than other learning algorithms we
+have described, such as decision trees. We use them in this work on LOS
+classification as we would like to see how they perform in modelling the
+complex relationships between various medical attributes.
 
 \section{Conclusion}
+In this chapter we have outlined all of the learning algorithms that we will
+be evaluating with respect to our classification problem, from the simple
+0R and 1R to Na\"{i}ve Bayes, decision trees, logistic regression,
+$k$-nearest neighbour (and extensions K* and Ranked Distance), to the more
+sophisticated support vector machines and multi-layer perceptrons. The reason
+for our inclusion of so many algorithms is two-fold. Firstly, many of
+these algorithms have not been extensively applied in various medical settings
+to the problem of LOS prediction. Using a wide range of algorithms from various
+paradigms will enable us to see what algorithms perform better than others.
+Secondly, evaluating several classifiers will allow us to make comparisons
+between different feature selection methods that are independent of the
+classifier and give us more insight into the nature of the data set and of the
+problem.

### 2014-10-23 thesis/method/features.tex
Updated feature preprocessing section

@@
 
-\section{Data preparation}
+In the previous chapter we began by describing each training example as a
+vector of feature values, and then described how a range of different
+learning algorithms used those feature vectors to learn a way to
+distinguish examples of one class from another. Ideally, all features given
+in a data set should be relevant and useful in not only describing the training
+example but also predicting the class.
+In practice, and particularly
+in medicine, a large number of attributes about a patient is usually recorded.
+This is detrimental for two reasons: the training time is significantly
+increased because the classifier must incorporate irrelevant information, and
+this irrelevant information tends to degrade the performance
+of even state-of-the-art classifiers such as C4.5 decision trees
+\cite{Witten2005}. Clearly, there is an advantage in considering only a subset
+of the most relevant features; to find them, we will use feature selection.
+
+It is also common for data to require preprocessing before a classifier can
+be trained upon it -- indeed, even before feature selection is performed.
+Otherwise, it will be hard to know if the results we see are reflective of
+the algorithms we use or simply a case of lucky data. Preprocessing involves
+cleaning and transforming the data, and is critical to the performance of
+learning algorithms that will be used to learn from it.
+
+In this chapter we describe the various preprocessing steps performed on
+the data, as well as the feature selection methods we evaluate in our work.
+
+\section{Preprocessing}
 In any data mining task, adequately preparing the data ensures the best
@@
 
-\subsection{Cleaning and normalisation} % missing values, irrelevant features
+\subsection{Cleaning} % missing values, irrelevant features
+The exact steps involved in \textit{cleaning} the data will vary between data
+sets, but the overall process is similar. Common issues to deal with include:
+\begin{itemize}
+\item Incorrect spelling or punctuation, perhaps as a result of mistakes in
+data entry
+\item Leading or trailing whitespace, and other garbage characters,
+perhaps as a result of data entry or
+the system that stores the data
+\item Missing values for various features or for the class label
+\item Values for features that do not make sense (such as a patient age of
+-1)
+\end{itemize}
+
 Since the data was initially in an Excel spreadsheet, we explored the data
@@
 with underscores (\_) to ensure that the full attribute name was read.
-\todo{Include a list of these converted attribute names? In the appendix?}
+For example, ``Any Cancer'' was converted into ``any\_cancer''.
 
@@
 
+\subsection{Normalisation}
 All numeric attributes were normalised into the range $[-1,+1]$
 
-\subsection{Feature selection}
+\subsection{Discretisation}
+
+\section{Feature Selection}
 After data preparation, selecting a subset of features or attributes to use
 in a model is also important, and helps to reduce noise and redundant
-attributes from becoming part of the final model. Experiments have found that
+attributes from becoming part of the final model. As we mentioned earlier,
+experiments have found that
 the performance of various classification methods such as decision trees
@@
 non-target attributes, which is a fairly high number. Many of these attributes
-are derived from others in the set, such as \textbf{age} and \textbf{age65}.
+are derived from others in the set: for example, whether or not a patient is
+over 65 years of age (the \textit{age65} feature) is directly dependent on the
+value of their \textit{age} attribute.
 
@@
 
-\subsubsection{Automated feature selection methods}
+\subsection{Automatic Methods}
 There are a number of methods that are commonly used to automate the selection
@@
   use in the building of the model.
+  \item Wrapper feature evaluation \todo{explain}
 \end{itemize}
@@
 
-\subsubsection{Expert selection}
+\subsection{Manual Selection}
 We also wanted to compare the feature selection of an expert in the domain

### 2014-10-24 thesis/method/classification.tex
Added reference to Rosenblatt (1962)

@@
 \begin{equation*}
-a_i = \dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}
+a_i = s\left(\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right) + t
 \end{equation*}
 where $a_i$ is the normalised value, $v_i$ is the actual value in the data
-set, and max $v_i$ and min $v_i$ are taken over all training examples.
+set, max $v_i$ and min $v_i$ are taken over all training examples, $s$ is a
+scaling factor $t$ indicates how much to translate the range by. If $s=1$
+and $t=0$, then the range of normalised values will be in $[0,1]$.
 
@@
 \textit{perceptron}, first
-studied extensively by Rosenblatt \todo{add citation entry}. The original
+studied extensively by Rosenblatt \cite{Rosenblatt1962}. The original
 motivation of the perceptron 

### 2014-10-24 thesis/method/preprocess.tex
Finished preprocessing section, restructure of overall methods section

\chapter{Feature Preprocessing} \label{chap:preprocess}

In the previous chapter we began by describing each training example as a
vector of feature values, and then described how a range of different
learning algorithms used those feature vectors to learn a way to
distinguish examples of one class from another. However, it is uncommon that
data can be used in a learning algorithm without first being preprocessed.
Preprocessing involves cleaning and transforming the data, and is critical
to the performance of learning algorithms that are applied to it. Without
preprocessing, it will be difficult to evaluate the true performance of
our classifiers on the data set and to draw meaningful conclusions from
comparisons between classifiers.

In this chapter give an overview of preprocessing and describe the steps we
took on our data.

\section{Cleaning} % missing values, irrelevant features
In any data mining task, adequately \textit{cleaning} the data ensures the best
chance of success: this includes deciding how to deal with missing and
inaccurate values, as well as appropriately transforming the data
\citep{Witten2005}.
The exact steps involved in data cleaning will vary between data
sets, but the overall process is similar. Common issues to deal with include:
\begin{itemize}
\item Incorrect spelling or punctuation, perhaps as a result of mistakes in
data entry
\item Leading or trailing whitespace, and other garbage characters,
perhaps as a result of data entry or
the system that stores the data
\item Missing values for various features or for the class label
\item Values for features that do not make sense (such as a patient age of
-1)
\end{itemize}
The following paragraphs address how we dealt with the issues that were
applicable to our data.

\paragraph{Incorrect spelling and punctuation.}
Since the data was initially in Excel spreadsheet format, we explored the data
by hand using Microsoft Excel's filter on each column. This allowed us to see
the range of possible values for each feature and easily detect anomalous
selling and punctuation:
\begin{itemize}
  \item In the \texttt{sex} attribute, there were values of `Female' and
  `Femal' which were consolidated into `Female'.
  \item In the \texttt{mechanism} attribute, which was textual and indicated
  the category of how the patient was injured, there were such categories as
  `Other Vehi' and `Other Vehicle' as well as `Pedal Cycl' and `Pedal Cyclist'.
  Such categories were combined into `Other Vehicle' and `Pedal Cyclist'
  respectively.
  \item Some categories in \texttt{mechanism} were also divided into whether or
  not the mechanism of injury was self-inflicted, such as `Shooting' and
  `Shooting-selfinflicted'. At the suggestion of a domain expert, we combined
  them into one category: in this case, simply
  `Shooting'; this was done for all the categories in this attribute.
  \item \texttt{disp from ED} indicated the section of the hospital that the
  patient was discharged to. This was also a textual categorical attribute.
  There were redundancies like `G. ICU' and `G.ICU' and even `GICU', which we
  all combined into one category, `GICU'.
\end{itemize}

\paragraph{Extraneous whitespace or other characters.}
After correcting for spelling and punctuation errors,
the data was then exported into CSV (comma-separated value) format, so that
we could continue to process it programmatically rather than manually without
having to use Excel. Many
fields in the data set contained leading or trailing whitespace, which was
removed with a few lines of Python code (see appendix \todo{[add reference
to appendix]}). All attribute names were converted to lowercase. In the case
where attribute names consisted of several words, the spaces were replaced
with underscores (\_) to ensure that the full attribute name was read.
For example, \texttt{Any Cancer} was converted into \texttt{any\_cancer}.

\paragraph{Missing values.}
There has been a varied treatment of missing values in the work of other
researchers, from discarding examples with missing values for any feature
to \textit{imputing} or ``filling in'' the missing values. We remove all
examples with missing feature values from consideration in all further
investigations, in line with the work of other research in LOS prediction
in the same domain \cite{Dinh2013a}.

\section{Normalisation}
Features represent different pieces of information about a training example.
Therefore, it is not surprising that not all features will have the same range
of values: a patient can have an \texttt{age} feature with a value of 20, but
their \texttt{heart rate} feature cannot possibly be 20. Learning algorithms
such as $k$-nearest neighbour (Chapter \ref{chap:classification})
are sensitive to the scales of feature values -- features with a larger range
of values will dominate a distance calculation. Therefore, we
\textit{normalise} all numeric (that is, continuous-valued)
feature values into the $[-1,1]$ range so that they are all expressed on the
same scale. Typically, values are normalised in the $[0,1]$ or the $[-1,1]$
ranges.
The formula to do this, also discussed in Chapter \ref{chap:classification},
is:
\begin{equation*}
a_i = 2\left[\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right] - 1
\end{equation*}
where we have used $s=2$ and $t=-1$ to transform values into the $[-1,1]$
range. This transformation was applied to each value of all numeric features,
and was performed after all cleaning steps.

\section{Discretisation}
Not all classification algorithms can handle numeric features, and some of the
ones that do require assumptions to be made about the underlying distribution
of the feature values that is not always satisfied -- recall the normality
assumption of the Na\"{i}ve Bayes classifier
from Chapter \ref{chap:classification}. This requires us
to \textit{discretise} the numeric features, which is a process of splitting
the continuous data into a small number of discrete, distinct ranges.
Additionally, there has been no work examining the use of discretisation to
improve the performance of LOS prediction models, so our work provides a
valuable guideline on what other researchers looking at a similar problem can
expect.

Two basic approaches to the discretisation problem exist: unsupervised and
supervised. \textit{Unsupervised} discretisation involves splitting each
feature without considering any class information, whereas \textit{supervised}
discretisation accounts for the class when discretising.
Unsupervised discretisation uses simple methods such as splitting the data into
fixed-width intervals -- each interval spans the same width of values,
but can contain different numbers of feature values depending on the
distribution of underlying data -- which
can often partition the continuous feature into unhelpful categories due to its
ignorance of class labels. Supervised discretisation, on the
other hand, is commonly performed using the entropy minimisation heuristic
described by Fayyad and Irani \cite{Fayyad1993}. Since we have labelled
data, we are able to use both approaches. However, we will focus on evaluating
the effect of supervised discretisation, as it is generally preferred to
unsupervised discretisation when class information about the training examples
is known. And the method proposed by Fayyad and Irani is one of the best known
general supervised discretisation techniques \cite{Witten2005}.

\subsection{Entropy-based Discretisation}
In Chapter \ref{chap:classification} we outlined the induction of a decision
tree on a data set, using the information gain as the criterion for deciding
the attribute to split on at each step. A similar approach is used in Fayyad
and Irani's entropy-based method of discretisation \cite{Fayyad1993}:
to discretise an attribute,
each possible split point is considered and the information gain from that
split is computed. The split point with the highest information gain is
therefore chosen. Once the first split is determined, the algorithm continues
recursively in the newly created upper and lower ranges.

The \textit{minimum description length (MDL)} principle is used to determine
the stopping criterion. The MDL value is calculated by adding the information
values from two cases: split or do not split. If we split, then we can encode
the split point in $\mathrm{log}_2(N-1)$ bits (where $N$ is the number of
instances), then the classes of those above and below it. If we do not split,
then each instance's label must be encoded. Mathematically, given $N$ examples
with entropy $E$, sub-intervals $k_1$ and $k_2$ with the entropy of examples
in each sub-interval denoted $E_1$ and $E_2$:
\begin{equation*}
\mathrm{MDL} = \dfrac{\mathrm{log}_2(N-1)}{N} + \dfrac{\mathrm{log}_2(3^k-2) - kE + k_1E_1 + k_2E_2}{N}
\end{equation*}
The first term is the information needed for the splitting point, and the
second is a correction necessary for encoding which classes belong to which
sub-interval. The decision rule is then to split if the information gain for
that split exceeds the MDL value: $\mathrm{gain} > \mathrm{MDL}$.

Since this requires exhaustive searching of the full range of feature values
to find potential split points, this method of supervised discretisation can
be time-consuming given large enough data sets. However, it has been proven
that any potential split that minimises the information value will never occur
between two examples of the same class. This means that we only need to
consider splits between regions of different classes, reducing the number of
possible points to consider.

\section{Conclusion}
In this chapter we discussed various issues encountered during the
preprocessing stage and how these were addressed. These steps are critical to
the success of our learning algorithms: without cleaned data, any subsequent
work is meaningless and difficult to draw valid conclusions from. By outlining
our steps, we hope to provide an example of how one can conduct the cleaning
process.

We then discussed normalisation, an important transformation of
continuous-valued data that ensures all features take on the same range of
values. This is important for ensuring that classifiers sensitive to the
relative range of values between features give valid outputs.

We concluded with outlining discretisation, in particular Fayyad and Irani's
entropy-based method. Such a technique has not been used in LOS prediction
and by including it in our work, we can evaluate its effectiveness in this
domain and let our results serve
as a baseline and reference for future work.

### 2014-10-24 thesis/method/selection.tex
Finished preprocessing section, restructure of overall methods section

--- 2014-10-18 thesis/method/features.tex
@@
-\chapter{Feature Preprocessing and Selection} \label{chap:features}
+\chapter{Feature Selection} \label{chap:selection}
 
-\section{Data preparation}
-In any data mining task, adequately preparing the data ensures the best
-chance of success: this includes deciding how to deal with missing and
-inaccurate values, as well as appropriately transforming the data
-\citep{Witten2005} and selecting or constructing the most suitable features
-for the task \citep{Kotsiantis2006}.
-In this section we describe how we addressed each of
-the major issues and the reasons for our choice.
+Recall that each feature tells
+us something about the training example. Ideally, it should also help us
+classify the example. However, in practice and particularly
+in medicine, a large number of attributes about a patient is usually recorded,
+not all of which are relevant in predicting the class.
+Having such a large number of features
+is detrimental for two reasons: the training time is significantly
+increased because the classifier must incorporate irrelevant information, and
+this irrelevant information tends to degrade the performance
+of even state-of-the-art classifiers such as C4.5 decision trees
+\cite{Witten2005}. Clearly, there is an advantage in considering only a subset
+of the most relevant features; to find them, we will use feature selection.
 
-\subsection{Cleaning and normalisation} % missing values, irrelevant features
-Since the data was initially in an Excel spreadsheet, we explored the data
-by hand using Microsoft Excel's filter function on each column and
-consolidated many categorical values simply
-by correcting spelling and punctuation. For example:
-\begin{itemize}
-  \item In the \textbf{sex} attribute, there were values of `Female' and
-  `Femal' which were consolidated into `Female'.
-  \item In the \textbf{mechanism} attribute, which was textual and indicated
-  the category of how the patient was injured, there were such categories as
-  `Other Vehi' and `Other Vehicle' as well as `Pedal Cycl' and `Pedal Cyclist'.
-  Such categories were combined into `Other Vehicle' and `Pedal Cyclist'
-  respectively.
-  \item Some categories in \textbf{mechanism} were also divided into whether or
-  not the mechanism of injury was self-inflicted, such as `Shooting' and
-  `Shooting-selfinflicted'. When we asked the domain expert, Michael Dinh, his
-  suggestion was to combine them into one category: in this case, simply
-  `Shooting'; this was done for all the categories in this attribute.
-  \item \textbf{disp from ED} indicated the section of the hospital that the
-  patient was discharged to. This was also a textual categorical attribute.
-  There were redundancies like `G. ICU' and `G.ICU' and even `GICU', which we
-  all combined into one category, `GICU'.
-\end{itemize}
-
-The data was then exported into CSV (comma-separated value) format, so that
-we could continue to process it programmatically rather than manually. Many
-fields in the data set contained leading or trailing whitespace, which was
-removed with a few lines of Python code (see appendix \todo{[add reference
-to appendix]}). All attribute names were converted to lowercase. In the case
-where attribute names consisted of several words, the spaces were replaced
-with underscores (\_) to ensure that the full attribute name was read.
-\todo{Include a list of these converted attribute names? In the appendix?}
-
-In order for our results to be comparable to that of Dinh
-et. al. \citep{Dinh2013a}, we removed all instances with missing values,
-leaving 2517 instances with which to use to create our prediction models.
-
-All numeric attributes were normalised into the range $[-1,+1]$
-
-\subsection{Feature selection}
+\section{Feature Selection}
 After data preparation, selecting a subset of features or attributes to use
-in a model is also important, and helps to reduce noise and redundant
-attributes from becoming part of the final model. Experiments have found that
+in a model is also important \citep{Kotsiantis2006},
+and helps to reduce noise and redundant
+attributes from becoming part of the final model. As we mentioned earlier,
+experiments have found that
 the performance of various classification methods such as decision trees

### 2014-10-25 thesis/method/classification.tex
Finished section on feature selection and added references to equations

@@
 \section{0R and 1R}
+\label{sec:zeror}
 Short for ``0-rule'' and ``1-rule'' respectively, 0R and 1R are
@@
 \section{C4.5 Decision Tree (DT)}
+\label{sec:c45}
 Decision trees (DT) can be specified recursively and are an example of a
@@
 node is calculated using the \textit{entropy} function:
-\begin{equation*}
+\begin{equation}
+\label{eqn:entropy}
 \mathrm{entropy}(p_1,p_2,\ldots,p_n) = -p_1\mathrm{log}p_1 -p_2\mathrm{log}p_2 \ldots -p_n\mathrm{log}p_n
-\end{equation*}
+\end{equation}
 Given $k$ classes, $c_1,\ldots,c_k$ is the number of instances in each
@@
 $y_i \in \{-1,+1\}, i=1,\ldots,m$, the optimisation problem can be written as:
-\begin{equation}
-\label{eqn:svm1}
+\begin{equation*}
 \begin{aligned}
@@
 \end{aligned}
-\end{equation}
+\end{equation*}
 where $\phi(\mathbf{x})$ is the kernel function, and

### 2014-10-25 thesis/method/selection.tex
Finished section on feature selection and added references to equations

@@
 Recall that each feature tells
-us something about the training example. Ideally, it should also help us
+us something about a training example. Ideally, it should also help us
 classify the example. However, in practice and particularly
@@
 is detrimental for two reasons: the training time is significantly
-increased because the classifier must incorporate irrelevant information, and
+increased because the classifier must incorporate more information than it
+needs to, and
 this irrelevant information tends to degrade the performance
@@
 \cite{Witten2005}. Clearly, there is an advantage in considering only a subset
-of the most relevant features; to find them, we will use feature selection.
+of the most relevant features; to find them, we will use what is called
+\textit{feature selection}.
 
-\section{Feature Selection}
-After data preparation, selecting a subset of features or attributes to use
-in a model is also important \citep{Kotsiantis2006},
-and helps to reduce noise and redundant
-attributes from becoming part of the final model. As we mentioned earlier,
-experiments have found that
-the performance of various classification methods such as decision trees
-\textit{deteriorates} in the presence of useless attributes.
-Our data set contains 73
-non-target attributes, which is a fairly high number. Many of these attributes
+Our data set contains 78 features aside from the class, which is a fairly high
+number. Many of these attributes
 are derived from others in the set: for example, whether or not a patient is
-over 65 years of age (the \textit{age65} feature) is directly dependent on the
-value of their \textit{age} attribute.
+over 65 years of age (the \texttt{age65} feature) is directly dependent on the
+value of their \texttt{age} attribute.
 
+\section{Preliminary Removal of Features}
 We began the process of feature selection by removing attributes that were
 not available upon admission, as these tended to be highly related to the
-LOS48 outcome and were a consequence, not a predictor, of LOS. These attributes
-were: iculos, outcome, disp from ED, disp from hosp, died, rehab and los.
-The id attribute was also removed because it simply numbered each record and
-indicated nothing about the patient themself. This left us with 67 attributes.
+\texttt{los48} outcome and were a consequence, not a predictor, of LOS. Had
+we not removed these features, they would be selected by the feature selection
+methods described below due to their high association with \texttt{los48}, but
+would actually tell us nothing about what causes a patient to require a
+hospital stay of more or less than 2 days.
+These attributes were: \texttt{iculos}, \texttt{outcome},
+\texttt{disp from ed}, \texttt{disp from hosp}, \texttt{died}, \texttt{rehab},
+\textit{los} and \texttt{eddc}.
+The \texttt{id} attribute was also removed because it
+simply numbered each record and
+indicated nothing about the patient. This left us with 69 attributes.
 
-The work of actually selecting the attributes out of the 67 available was
-carried out in two ways: automated using feature selection methods, and
-manually using advice from a domain expert.
+We then removed 10 other features which were directly derived from others in
+the data set: \texttt{gcs1}, \texttt{iss8}, \texttt{mechanism}, \texttt{male},
+\texttt{age65}, \texttt{bp1}, \texttt{pr1}, \texttt{rr1}, \texttt{fall} and
+\texttt{comorbidity}. At this stage there were 59 features -- still a large
+number.
 
-\subsection{Automatic Methods}
-There are a number of methods that are commonly used to automate the selection
-of a subset of features for the final classifier. The methods we used were:
-\begin{itemize}
-  \item \textit{Correlation-based feature selection} \citep{Hall2000},
-  which evaluates the `merit' of a
-  subset of features by considering their degree of correlation with the target
-  variable compared to their degree of correlation amongst each other. High
-  correlation between variables indicates redundancy. Under
-  this scheme, a set of features that are highly correlated with the target
-  while having low correlation with other features in the set is ideal.
-  \item Selection based on the
-  \textit{Pearson product-moment correlation coefficient}, a measure of the
-  linear correlation or dependence between two variables.\todo{add some more info}
-  \item \textit{Information gain} \todo{explain a bit about this}
-  \item Feature selection based on the \textit{OneR classifier}
-  \citep{Holte1993}, which uses one
-  attribute as the predictor for the target. This attribute is typically the one
-  that has the lowest prediction error. Feature selection using this scheme
-  involves first ranking attributes based on their error, with the lowest error
-  attributes as the most desirable. We then selected the top 20 attributes to
-  use in the building of the model.
-  \item Wrapper feature evaluation \todo{explain}
-\end{itemize}
+After eliminating redundant or inapplicable features as detailed above, we
+considered four \textit{automatic} (correlation-based
+feature selection, Pearson correlation, 1R and wrapper selection
+using C4.5) and two \textit{manual} feature selection methods (domain expert
+selection and the feature set used in existing work on trauma LOS prediction).
+These methods are described in the following sections.
 
-We used the WEKA 3.7.11 \citep{Hall2009} implementation of the above
-methods, after converting the data into the Attribute-Relation File Format
-(ARFF)\footnote{\url{http://weka.wikispaces.com/ARFF+\%28developer+version\%29}}
-most suitable for the program.
-\todo{give more details, ie specifying binary/numeric
-attributes, maybe specify in appendix?}
+\section{Automatic Methods}
+Automatic feature selection methods are useful for when we do not have a deep
+understanding of the underlying domain and what each attribute really means to
+the class. Even if we are well-versed in the domain, evaluating a few different
+methods of automatic feature selection can give us new insight into the
+learning problem at hand. Here we will describe a few of the methods we used.
 
-\subsection{Manual Selection}
-We also wanted to compare the feature selection of an expert in the domain
-of the data set, namely trauma, with the models obtained from the features
-selected using the automated methods described above. This resulted in two
-feature subsets. The first was obtained simply by presenting the entire list
-of 73 features to the domain expert and asking him to select the features he
-thought were most indicative of the target variable from experience. The
-second set was obtained from the work of Dinh et. al. \citep{Dinh2013a} that
-we are building on, giving us a baseline to compare our results to.
+\subsection{Correlation-based Feature Selection (CFS)}
+Correlation-based feature selection (CFS) is a method developed by Hall
+\cite{Hall2000} which looks for a subset of features that have high correlation
+with the class but low intercorrelation among the features. In this case,
+correlation between two features is measured using what is called the
+\textit{symmetric uncertainty}:
+\begin{equation*}
+U(A,B) = 2\dfrac{H(A)+H(B)-H(A,B)}{H(A)+H(B)}
+\end{equation*}
+where $A$ and $B$ are two nominal attributes and $H$ is the entropy function
+denoted by Equation \ref{eqn:entropy} in Chapter \ref{chap:classification}.
+The entropies are calculated from the probabilities of each feature value, and
+$H(A,B)$, the joint entropy of $A$ and $B$, is calculated from the
+probabilities of each combination of values of $A$ and $B$. Note that $U(A,B)$
+is always between 0 and 1.
 
-Altogether, the various feature selection methods resulted in a total of
-7 data sets (4 automated, 2 expert selected, and 1 with all features) which
-we used in the model construction phase.
+To evaluate the ``goodness'' of a subset of features with the class $C$, we
+compute:
+\begin{equation}
+\label{eqn:cfs}
+\dfrac{\sum_j U(A_j,C)}{\sqrt{\sum_i \sum_j U(A_i,A_j)}}
+\end{equation}
+where $i$ and $j$ range over the features in the subset being evaluated. The
+maximum value that Equation \ref{eqn:cfs} can obtain is 1 when all features
+correlate perfectly with the class and also with each other -- the numerator
+becomes $n$ and the denominator $\sqrt{n^2}$ if the subset has $n$ features.
+\todo{elaborate a bit more}
+
+\subsection{Pearson Correlation Coefficient}
+The Pearson correlation coefficient measures the degree of linear correlation
+between two variables. It is always in the range $[-1,1]$, where -1 is total
+negative correlation, 0 is no (linear) correlation, and 1 is total positive
+correlation. We compute the Pearson correlation between each feature and the
+class using:
+\begin{equation*}
+r_i = \dfrac{\sum_{i=1}^m (x_i-\bar{x})(c-\bar{c})}{\sqrt{(\sum_{i=1}^m x_i-\bar{x})(\sum_{i=1}^m c-\bar{c})}}
+\end{equation*}
+where $m$ is the number of training examples, $x_i$ is the value of feature
+$i$, and $\bar{x}$ and $\bar{c}$ are the arithmetic averages of the feature
+values and the class values respectively. (Recall that we also used this
+in Chapter \ref{chap:classification} to rank features for use in Ranked
+Distance nearest neighbours.)
+
+A higher value for $r_i$ indicates higher correlation with the class. We
+rank all features according to their $r_i$ from highest to lowest, and select
+some subset of these either by selecting a fixed number, or selecting all
+features that have a correlation above a specified threshold.
+
+\subsection{1R}
+Another way to evaluate the relevance of features is to use the idea behind the
+1R classifier described in Section \ref{sec:zeror}. Recall that the 1R
+classifier constructs simple 1-rules based on each feature and uses the feature
+that produces the lowest error rate (that is, the highest accuracy) on the data
+set to classify unseen examples. These accuracy rates can be calculated for
+each feature, which are then used to rank the features from highest to lowest
+accuracy. The worst accuracy for any feature is the 0R accuracy, which is the
+proportion of the majority class in the data set. This is because if 1R finds
+that the accuracy of the rule constructed using a feature is worse than 0R,
+then it will not use that rule and predict the majority class instead.
+
+We can also select a fixed number of features or all features above a certain
+1R accuracy threshold.
+
+\subsection{Wrapper-based Feature Selection}
+The feature selection methods described above (CFS, Pearson correlataion and
+1R) have been described as \textit{filter} approaches to attribute selection.
+Kohavi and John proposed another approach, the \textit{wrapper} approach:
+instead of performing feature selection independently of the learning
+algorithm using some measure of ``relevance'',
+the learning algorithm itself is used to evaluate the goodness of
+subsets of features \cite{Kohavi1997}. The difference between the filter
+and wrapper approaches is depicted in Figure \ref{fig:wrapper}.
+
+\begin{figure}[h]
+\label{fig:wrapper}
+\caption{}
+\centering
+%\includegraphics[scale=0.95]{images/a2-nn}
+\end{figure}
+
+A key advantage of the wrapper approach is that the selected feature sets are
+specifically tailored to a learning algorithm, which ensures the best
+performance of that algorithm given the data set. However, wrapper approaches
+can also be prohibitively slow if the data or feature sets are large and the
+chosen learning algorithm expensive to train (such as support vector machines
+and multi-layer perceptrons). We will evaluate the use of C4.5 decision trees
+(see Section \ref{sec:c45}) in selecting features using the wrapper approach.
+Using decision trees to select features is relatively fast as due to the
+divide and conquer nature of the learning algorithm, and the tendency of the
+trees to select fewer features for learning than other algorithms.
+
+\section{Manual Selection}
+All of the above automatic feature selection methods use some notion of
+``relevance'', such as symmetric uncertainty and linear correlation,
+in order to determine the best subset of features to select.
+Depending on the problem at hand, these definitions of relevance may or may
+not work well with the data. In the LOS prediction literature, studies
+have enlisted the help of medical domain experts to assist in feature
+selection, due to their experience and understanding of the underlying
+problem and the effect of the features on the outcome. It will therefore be
+worthwhile to compare the performance of classifiers using features selected
+automatically and manually.
+
+\subsection{Baseline Feature Set}
+The key previous work on trauma LOS prediction was by Dinh et. al.
+\cite{Dinh2013a}, who used a specific feature set from the data set that we
+use. We would like to use the features they selected as a baseline for
+evaluating the other feature selection methods we describe.
+
+\subsection{Domain Expert Selection}
+Consultation with a domain expert yielded another feature set that we could
+use to compare with the baseline and with the other automatic methods.
+
+\section{Conclusion}
+In this chapter we covered feature selection, which is concerned with reducing
+the number of attributes that a learning algorithm needs to consider when
+trying to learn a relationship between the training examples and their labels.
+We described how we removed redundant and meaningless features before the
+feature selection step, and discussed the automatic and manual feature
+selection methods we use in our work. The value in using several methods of
+each type of feature selection is two-fold: it allows us to not only compare
+the performance of the methods, but also to compare the performance of the
+different classifiers we use independently of the feature selection method.

### 2014-10-25 thesis/litreview/litreview.tex
Updated title of literature review
Update case of literature review headings

@@
-\chapter{Literature Review} \label{chap:litreview}
+\chapter{Predicting the LOS: A Review} \label{chap:litreview}
 
@@
 
-\section{Statistical methods}
+\section{Statistical Methods}
 \subsection{Regression}
@@
 
-\subsection{Markov models}
+\subsection{Markov Models}
 There have been a number of studies attempting to model the length of stay
@@
 
-\subsection{Other statistical techniques}
+\subsection{Other Statistical Techniques}
 A number of other techniques have been used in length of stay prediction, which
@@
 
-\section{Data mining methods}
+\section{Data Mining Methods}
 Data mining is the analysis of (often large) observational data sets to find
@@
 
-\subsection{Artificial neural networks}
+\subsection{Artificial Neural Networks}
 Artificial neural networks have been widely used in clinical prediction
@@
 
-\subsection{Extensions to the artificial neural network model}
+\subsection{Extensions to the Artificial Neural Network Model}
 Buchman et al. published a comparison of the performance of logistic regression
@@
 
-\subsection{Other data mining techniques}
+\subsection{Other Data Mining Techniques}
 As discussed, artificial neural networks have been used extensively in

### 2014-10-26 thesis/method/classification.tex
Added some section labels and updated some headings

@@
 \subsubsection{Ranked Distance}
+\label{sec:rankeddist}
 \paragraph{Intuition}
@@
 
-\section{Conclusion}
+\section{Summary}
 In this chapter we have outlined all of the learning algorithms that we will

### 2014-10-26 thesis/method/preprocess.tex
Added some section labels and updated some headings

@@
 
-\section{Conclusion}
+The data was discretised according to the above method using the supervised
+discretisation filter in WEKA. We keep both the original data set (without
+discretisation) and the discretised data set for comparison.
+
+\section{Summary}
 In this chapter we discussed various issues encountered during the

### 2014-10-26 thesis/method/selection.tex
Added some section labels and updated some headings

@@
 \section{Manual Selection}
+\label{sec:manual}
 All of the above automatic feature selection methods use some notion of

### 2014-10-26 thesis/method/evaluation.tex
Add evaluation section

--- 2014-10-18 thesis/method/evaluation.tex
@@
 
+The methods of evaluating the performance of classifiers in LOS prediction
+that are discussed in the literature encompass a wide range of medical
+domains and learning problems, which presents us with a challenge when we
+decide how to evaluate our work. Here we have chosen to rigorously evaluate
+our results (such as using stratified cross-validation when others may have
+used a single training-test split), while still using performance measures
+comparable to the literature and suitable for our particular learning problem.
+This chapter discusses the data sets, procedure and performance measures used
+in evaluating our work.
+
 \section{Data Sets}
-\subsection{Collection}
+In our work we obtained two independent sets of LOS data, one specifically for
+trauma and the other a general, hospital-wide set.
+
+\subsection{Trauma LOS Data}
+\subsubsection{Collection}
 The data set used consisted of trauma registry data from the trauma centre
@@
 
-\subsection{Characteristics}
-There were 2546 patient records in the data set we received, comprising of 74
-features, one of which was the target variable \textbf{LOS48}. LOS48 was a
-binary variable -- that is, it could only take two values, 0 or 1 -- with 1
-indicating that the patient stayed two days or less, and 0 indicating a stay
-of greater than 2 days.
+\subsubsection{Characteristics}
+There were 2546 patient records in the data set we received, comprising of 79
+features, one of which was the target variable \texttt{los48}.
+\texttt{los48} was a binary variable -- that is, it could only take two
+values, 0 or 1 -- with 1 indicating that the patient stayed two days or less,
+and 0 indicating a stay of greater than 2 days.
 
-\todo{Perhaps insert some more information about the data itself, such
-as the various attributes and their data types? Not sure if that should go
-here, in the results, or in an appendix -- it is relevant but can disturb
-the flow of the main text.}
+\subsection{General Hospital LOS Data}
+By collaborating with our colleagues from the University of Minho in Portugal,
+we were able to obtain LOS data for all inpatient hospitalisations from
+2000--2013 at the Hospital das Foras Armadas in Lisboa, Portugal. The data
+contained 26462 inpatient records spanning all medical departments within the
+hospital, comprising 15 features that were not specific to any medical domain.
 
-\section{Performance evaluation} % don't forget cross-validation
-\label{sec:perfeval}
-Here we describe the methods used in order to assess the performance of the
-models constructed through single classifiers and ensemble methods described
-above.
+\section{Evaluation Procedure}
+\subsection{Steps Involved}
+The steps that we took to evaluate the performance of the various classifiers
+and feature selection methods were:
+\begin{enumerate}
+\item The trauma LOS data was cleaned and preprocessed as described in Chapter
+\ref{chap:preprocess}, giving us two data sets: one with and one without
+discretisation. The general LOS data from Portugal was already cleaned when we
+received it; however, we discretised the continuous \texttt{los} variable to
+the same categories as for the trauma data (2 days or less, greater than 2
+days), normalised the numeric attribute \texttt{prev\_admissions}, and derived
+both a discretised and non-discretised data set. This resulted in 4 sets of
+data.
+\item For each data set, we evaluated the effect of the 4 automatic feature
+selection methods on each of the 8 classifiers described in Chapter
+\ref{chap:classification}, giving us 32 configurations for each data set, over
+4 sets. Evaluation was performed using \textit{ten-fold stratified
+cross-validation} (described below).
+\item For the two trauma LOS data sets, we additionally evaluated the two
+manual feature selection methods with all 8 classifiers. Recall from Section
+\ref{sec:manual} that one of the feature sets is from the key previous work
+on trauma LOS prediction that we use as a baseline, and the other feature set
+is a set of features included through consultation with a domain expert.
+\end{enumerate}
 
-\subsection{Metrics}
-There are many available and commonly used metrics that we can use when
-evaluating and comparing the performance of prediction models. The ones that
-we use are: \todo{maybe put these in a table}
+\subsection{Ten-fold Stratified Cross-Validation}
+Cross-validation is a method of evaluating the performance of a learning
+algorithm. The entire data set is partitioned into $k$ subsets, and the
+learning algorithm is trained on $k-1$ of these subsets and tested on one. This
+is repeated $k$ times, so that each of the subsets is used exactly once as a
+test set. After each iteration of training and testing, called a \textit{fold},
+various performance measures about the classifier are recorded. At the end of
+the $k$ folds, the $k$ performance measures obtained are averaged to arrive at
+a final measure of performance. \textit{Stratified} cross-validation ensures
+that the proportion of each class in each of the $k$ subsets is approximately
+the same as in the entire data set. This reduces the effect of random uneven
+class representation in the partitioned $k$ subsets during the cross-validation
+process.
+
+We use $k=10$ folds for our stratified cross-validation, which has been
+considered the standard method of measuring the performance of a learning
+algorithm on a given data set \cite{Witten2005}.
+
+\subsection{Technical Details}
+Here we specify the technical details of our work: namely, the programming
+languages and software used in our entire procedure.
+
+\subsubsection{Repository}
+All code relating to this work, including the \LaTeX code for this document,
+as well as results from our evaluations are available publicly online at
+\url{https://github.com/tianyupu/hons-thesis}. However, we were not able to
+obtain permission from neither the Royal Prince Alfred Hospital nor the
+Hospital das Foras Armadas to make the data publicly available due to the
+privacy of patient information. Interested readers should discuss with the
+author.
+
+\subsubsection{Software Used}
+For initial exploration and cleaning of data, we used Microsoft Excel 2007
+as the data was already in spreadsheet format. This was then exported to
+CSV. For the conversion to ARFF, and evaluation of the automatic
+feature selection methods and classification algorithms, we used the latest
+developer version of WEKA, 3.7.11 \cite{Hall2009}. We used WEKA exclusively
+through the command-line interface so that we could automate and configure
+our experiments easily using scripts which we wrote (see Section
+\ref{sec:code}).
+
+\subsubsection{Code Written}
+\label{sec:code}
+\todo{attach code in appendix?}
+Leading and trailing whitespace was removed and feature names transformed using
+code written in Python to manipulate the CSV. All runs of WEKA were executed by
+a program we wrote in Bash (a Unix shell) script. This program was able to run
+WEKA with the following features:
 \begin{itemize}
-  \item Accuracy
-  \item Specificity
-  \item Sensitivity
-  \item Area under the receiver-operating characteristic (ROC) curve, or AUC
+\item The ability to specify filenames corresponding to the input training data
+and the specification of classification or feature selection algorithms to run
+\item ROC curve data output can be optionally enabled and saved to disk after
+each cross-validation pass
+\item An experiment can be repeated an arbitrary number of times -- this is
+specified by the user upon invoking the script (default is one time)
+\item The ability to specify the inclusion of a random number seed -- this is
+necessary for repetitions of runs to generate different cross-validation splits
+\item Model information and performance measures are saved to disk after each
+cross-validation
 \end{itemize}
 
-Since we only have one data set, we use ten-fold stratified cross-validation
-to obtain values for the above metrics. We run the ten-fold cross-validation
-40 times for each classifier (single or ensemble) for each of the 7 data sets
-produced from feature selection. 40 runs was necessary in order for us to
-compute confidence intervals for each of the metrics, which makes use of the
-Central Limit Theorem from statistics.
+Because of the large numbers of results we obtained, we also wrote scripts in
+Bash and Python to process and summarise the performance measures we required.
+If we ran a set of classifiers 50 times in an experiment, the script would
+retrieve all 50 outputs of a performance measure (such as accuracy, described
+in the next section) for each classifier, and compute the mean and confidence
+interval (also discussed in the next section) in a single output that we could
+then save as CSV.
 
-\subsubsection{Confidence intervals}
+In order to implement our novel Ranked Distance function (Section
+\ref{sec:rankeddist}), we created our own class within the WEKA Java package
+hierarchy. This was implemented in Java and needed to implement the correct
+interfaces and override the right inherited methods in order for the $k$-NN
+classifier in WEKA to be able to use it as a distance function. This also
+involved updating certain properties of the \texttt{weka.jar} executable in
+order to use the new class, as well as building the entire WEKA source. The
+development environment we used was Eclipse \todo{insert version}, and the
+source was built using the 1.7 Java runtime. \todo{double check}
+
+\section{Performance Measures}
+There are many ways to evaluate the performance of learning algorithms. The
+two measures we will use are the \textit{accuracy} and the \textit{area under
+the receiver operating characteristic (ROC) curve}, commonly abbreviated AUC
+(for \textit{area under the curve}).
+
+\subsection{Accuracy}
+The accuracy is a common measure of performance for classification algorithms,
+and indicates the proportion of examples that were classified correctly from
+a given data set. Mathematically, assuming a two-class scenario, the expression
+is:
+\begin{equation*}
+\mathrm{Accuracy} = \dfrac{TP + TN}{TP + TN + FP + FN}
+\end{equation*}
+where:
+\begin{itemize}
+  \item $TP$ is the number of \textit{true positives} -- that is, the number of
+  examples belonging to class 1 that were classified as belonging in class 1;
+  \item $TN$ is the number of \textit{true negatives}, which is the number of
+  examples belonging to class 0 that were classified as belonging to class 0;
+  \item $FP$ is the number of \textit{false positives}, which is the number of
+  examples that should have been classified 0 but were classified as 1; and
+  \item $FN$ is the number of \textit{false negatives}, which is the number of
+  examples that should have been classified 1 but were classified as 0.
+\end{itemize}
+
+The accuracy can also be expressed more simply and intuitively as
+$\dfrac{\text{number of correct classifications}}{\text{total number of examples}}$
+but the terms introduced above will be necessary to understand the
+AUC measure below.
+
+We use accuracy as a performance measure as it is often used in previous work,
+which allows us to compare our work with others. Accuracy is also relatively
+straightforward to understand and interpret.
+
+\subsection{Area Under the Curve (AUC)}
+The area under the receiver operating characteristic (ROC) curve, often
+abbreviated AUC, is a measure of the ability of a binary classifier to
+distinguish between members of the two classes. Let the \textit{true
+positive rate} be $\dfrac{TP}{TP + FN}$ and the \textit{false positive rate}
+be $\dfrac{FP}{FP + TN}$. The ROC curve is a plot of the true positive rate
+on the vertical axis against the false positive rate on the horizontal axis.
+The area under this curve will lie between 0 and 1, with 1 indicating a
+100\% true positive rate and 0\% false positives (that is, everything was
+classified correctly). The AUC of randomly picking between the two classes
+with equal probability is 0.5, with a ROC curve that is a straight line
+dividing the graph diagonally in half. Figure \ref{fig:sampleroc} shows
+two ROC curves on the same graph:
+
+\begin{figure}[h]
+\label{fig:sampleroc}
+\caption{}
+\centering
+%\includegraphics[scale=0.95]{images/a2-nn}
+\end{figure}
+
+The AUC has not been commonly used in the literature to evaluate LOS
+classification problems, but we have included it in our work for a number of
+reasons. Firstly, it measures an aspect of classifiers that accuracy does not,
+and which is more important and useful in the domain: the ability to
+discriminate between instances of two classes. Recall from Chapter
+\ref{chap:classification} that 0R will always predict the majority class --
+this means that it will be evaluated positively on the basis of accuracy
+as long as there are sufficiently more examples of one class than the other.
+However, the AUC of the classifier will be \textit{less} than 0.5, because it
+is less able to distinguish between the two classes than random guessing is,
+due to the fact that it only ever predicts unseen examples to be in the
+majority class. Secondly, previous work on trauma LOS prediction by Dinh et al.
+also used the AUC as the key measure by which they evaluated their model
+\cite{Dinh2013a}, and we would like our work to be comparable to theirs.
+
+\subsection{Confidence Intervals}
 In order to compare our results with that of Dinh et. al. \citep{Dinh2013a},
@@
 \textit{significance level} -- as follows:
-\begin{equation}
+\begin{equation*}
   \left[\bar{x} - z_{1-\alpha/2}\dfrac{s}{\sqrt{n}},
     \bar{x} + z_{1-\alpha/2}\dfrac{s}{\sqrt{n}}\right]
-\end{equation}
+\end{equation*}
 where $\bar{x}$ is the sample mean, $\alpha$ is the significance level
@@
 
-\subsection{Statistical tests}
-To decide whether or not the difference in values of each metric between
-classifiers was statistically significant, paired $t$-tests were conducted for
-a subset of the results: \todo{elaborate}
-\begin{itemize}
-  \item Some feature set vs other feature set
-  \item Single classifier vs other single classifier
-  \item Ensemble vs single
-  \item Our best method vs Dinh2013's method
-\end{itemize}
+\subsection{Statistical Tests}
+To decide whether or not the difference in values of each measure between
+classifiers and feature selection methods was statistically significant,
+paired $t$-tests at a significance level of 0.05 were conducted.
+
+\section{Summary}

### 2014-10-26 thesis/method/preprocess.tex
Misc changes to preprocessing and selection

@@
 
+At the end of the cleaning process, the data was converted into the
+Attribute-Relation File Format (ARFF)
+\footnote{\url{http://www.cs.waikato.ac.nz/ml/weka/arff.html}} for use in
+WEKA, a suite of machine learning algorithms implemented in Java
+\cite{Hall2009}. Each feature was carefully checked to ensure that their
+type was specified correctly -- that is, the nominal attributes were
+distinguished from the numeric.
+
 \section{Normalisation}
@@
 where we have used $s=2$ and $t=-1$ to transform values into the $[-1,1]$
-range. This transformation was applied to each value of all numeric features,
-and was performed after all cleaning steps.
+range. This transformation was applied to each value of all numeric features
+using WEKA, and was performed after all cleaning steps described above.
 
@@
 and by including it in our work, we can evaluate its effectiveness in this
-domain and let our results serve
-as a baseline and reference for future work.
+domain and let our results serve as a baseline and reference for future work.

### 2014-10-26 thesis/method/selection.tex
Misc changes to preprocessing and selection
Change 'conclusion' heading in chapter to 'summary'

@@
 \texttt{disp from ed}, \texttt{disp from hosp}, \texttt{died}, \texttt{rehab},
-\textit{los} and \texttt{eddc}.
+\texttt{los} and \texttt{eddc}.
 The \texttt{id} attribute was also removed because it
@@
 
-\section{Conclusion}
+\section{Summary}
 In this chapter we covered feature selection, which is concerned with reducing

### 2014-10-26 thesis/method/evaluation.tex
Update evaluation section

@@
 
-\subsection{Statistical Tests}
+\subsection{Tests for Statistical Significance}
 To decide whether or not the difference in values of each measure between
@@
 paired $t$-tests at a significance level of 0.05 were conducted.
+\todo{elaborate}
 
 \section{Summary}
+In this section we described our method of evaluating the performance of each
+classifier and feature selection method that we use. We used accuracy and the
+area under the curve to measure two different aspects of our classifiers, and
+we obtained these figures using stratified ten-fold cross-validation. Paired
+$t$-tests were conducted to test for statistical significance.

### 2014-10-26 thesis/litreview/litreview.tex
Change 'conclusion' section heading of literature review to 'summary'

@@
 
-\section{Conclusion}
+\section{Summary}
 Despite the wide range of statistical and data mining techniques that have been

### 2014-10-27 thesis/method/selection.tex
Add outline of information gain feature selection

@@
 
-\subsection{Pearson Correlation Coefficient}
+\subsection{Pearson Correlation Coefficient (PCC)}
 The Pearson correlation coefficient measures the degree of linear correlation
@@
 features that have a correlation above a specified threshold.
+
+\subsection{Information Gain (IG)}
+Recall from our discussion of decision trees in in Section \ref{sec:c45} that
+the information gain (IG) is used to decide which attribute to split on at each
+step of the tree. We can also measure the IG of each feature as follows.
+If we have a class $C$ and a feature $F$, the entropy of the
+class before and after observing the feature is:
+\begin{equation*}
+\begin{aligned}
+& \mathrm{entropy}(C) = - \sum_{c \in C} p(c) \mathrm{log}_2 p(c) \\
+& \mathrm{entropy}(C|F) = - \sum_{f \in F} p(f) \sum_{c \in C} p(c|f) \mathrm{log}_2 p(c|f)
+\end{aligned}
+\end{equation*}
+The decrease in entropy reflects the gain in information provided by the
+feature and is given by:
+\begin{equation*}
+IG(C|F) = \mathrm{entropy}(C) - \mathrm{entropy}(C|F)
+\end{equation*}
+
+We can compute the IG for each feature and then select subsets of features that
+are greater than some threshold we specify.
 

### 2014-10-28 thesis/appendix/allresults.tex
Add appendices and results for trauma data

\chapter{Full Results Listing} \label{app:results}

\section{Trauma LOS}
\subsection{Performance Results}

\input{appendix/tr-all-summary.tex}

\input{appendix/tr-dinh-summary.tex}

\input{appendix/tr-expert-summary.tex}

\input{appendix/tr-cfs-summary.tex}

\input{appendix/tr-wrapper-summary.tex}

\input{appendix/tr-corr-threshold-disc.tex}
\input{appendix/tr-corr-threshold-nodisc.tex}

\input{appendix/tr-ig-threshold-summary-disc.tex}
\input{appendix/tr-ig-threshold-summary-nodisc.tex}

\input{appendix/tr-oner-threshold-summary-disc.tex}
\input{appendix/tr-oner-threshold-summary-nodisc.tex}

\subsection{Feature Selection Results}

\input{appendix/tr-nothreshold-features.tex}

\input{appendix/tr-corr-threshold-features.tex}

\input{appendix/tr-ig-threshold-features.tex}

\input{appendix/tr-oner-threshold-features.tex}

\section{General LOS}
\subsection{Performance Results}

\subsection{Feature Selection Results}

### 2014-10-28 thesis/appendix/paper.tex
Add appendices and results for trauma data

\chapter{Artificial Intelligence in Medicine -- Submitted Paper} \label{app:paper}

### 2014-10-28 thesis/appendix/tr-corr-threshold-features.tex
Add appendices and results for trauma data

\begin{table}[htbp]
\caption{}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{Pearson Correlation - Discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
0 & 57 & 73.08\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3, bodyregions, operation, lowerlimbnopelvis, neurosurgery, charlson, gcs, triss, bp, age, chest3, anypelvis, head3, headminor, mentalhealth, rr, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3, upper\_limbop, externalany, english \\ \hline
0.01 & 54 & 69.23\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3, bodyregions, operation, lowerlimbnopelvis, neurosurgery, charlson, gcs, triss, bp, age, chest3, anypelvis, head3, headminor, mentalhealth, rr, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3 \\ \hline
0.05 & 41 & 52.56\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3, bodyregions, operation, lowerlimbnopelvis, neurosurgery, charlson, gcs, triss, bp, age, chest3, anypelvis, head3, headminor, mentalhealth, rr, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, faceany, any\_heart, any\_cancer, face3, upperlimboponly \\ \hline
0.1 & 29 & 37.18\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3, bodyregions, operation, lowerlimbnopelvis, neurosurgery, charlson, gcs, triss, bp, age, chest3, anypelvis, head3, headminor, mentalhealth, rr, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface \\ \hline
0.125 & 20 & 25.64\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3, bodyregions, operation, lowerlimbnopelvis, neurosurgery, charlson, gcs, triss, bp, age, chest3, anypelvis, head3, headminor \\ \hline
0.15 & 12 & 15.38\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3, bodyregions, operation, lowerlimbnopelvis, neurosurgery, charlson \\ \hline
0.2 & 7 & 8.97\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3 \\ \hline
0.3 & 1 & 1.28\% & traumatier \\ \hline
0.5 & 0 & 0.00\% & - \\ \hline
0.75 & 0 & 0.00\% & - \\ \hline
\end{tabular}}
\end{table}

\begin{table}[htbp]
\caption{}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{11cm}|}
\hline
\multicolumn{ 4}{|c|}{Pearson Correlation - No discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
0 & 58 & 74.36\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss, charlson, age, gcs, operation, lowerlimbnopelvis, neurosurgery, chest3, anypelvis, headminor, head3, mentalhealth, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, rr, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, pr, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, bp, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3, upper\_limbop, externalany, english \\ \hline
0.01 & 55 & 70.51\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss, charlson, age, gcs, operation, lowerlimbnopelvis, neurosurgery, chest3, anypelvis, headminor, head3, mentalhealth, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, rr, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, pr, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, bp, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3 \\ \hline
0.05 & 41 & 52.56\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss, charlson, age, gcs, operation, lowerlimbnopelvis, neurosurgery, chest3, anypelvis, headminor, head3, mentalhealth, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, rr, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, pr, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, bp, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3 \\ \hline
0.1 & 27 & 34.62\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss, charlson, age, gcs, operation, lowerlimbnopelvis, neurosurgery, chest3, anypelvis, headminor, head3, mentalhealth, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, rr, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, pr, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, bp, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3 \\ \hline
0.125 & 19 & 24.36\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss, charlson, age, gcs, operation, lowerlimbnopelvis, neurosurgery, chest3, anypelvis, headminor, head3, mentalhealth, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, rr, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, pr, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, bp, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3 \\ \hline
0.15 & 15 & 19.23\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss, charlson, age, gcs, operation, lowerlimbnopelvis, neurosurgery \\ \hline
0.2 & 9 & 11.54\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss \\ \hline
0.3 & 2 & 2.56\% & iss, traumatier \\ \hline
0.5 & 0 & 0.00\% & - \\ \hline
0.75 & 0 & 0.00\% & - \\ \hline
\end{tabular}}
\end{table}

### 2014-10-28 thesis/appendix/tr-ig-threshold-features.tex
Add appendices and results for trauma data

\begin{table}[htbp]
\caption{}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{Information Gain - Discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
0 & 57 & 73.08\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, rr, mentalhealth, spine3, normalvitals, spineany, headany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, any\_cancer, any\_dm, any\_heart, faceany, face3, upperlimboponly, any\_kidney, thoracotomy, any\_liver, married, pelvic\_fracture, upperlimbany, ext3, alcohol, sex, neckany, upper3, transfer, neck3, upper\_limbop, externalany, english \\ \hline
0.001 & 47 & 60.26\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, rr, mentalhealth, spine3, normalvitals, spineany, headany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, any\_cancer, any\_dm, any\_heart, faceany, face3, upperlimboponly, any\_kidney, thoracotomy, any\_liver, married, pelvic\_fracture, upperlimbany \\ \hline
0.005 & 33 & 42.31\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, rr, mentalhealth, spine3, normalvitals, spineany, headany, headface, chestany, abdo3, penetrating \\ \hline
0.01 & 27 & 34.62\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, rr, mentalhealth, spine3, normalvitals \\ \hline
0.05 & 4 & 5.13\% & iss, triss, traumatier, icu \\ \hline
0.1 & 2 & 2.56\% & iss, triss \\ \hline
0.15 & 1 & 1.28\% & iss \\ \hline
0.2 & 0 & 0.00\% &  \\ \hline
0.3 & 0 & 0.00\% &  \\ \hline
0.5 & 0 & 0.00\% &  \\ \hline
\end{tabular}}
\end{table}

\begin{table}[htbp]
\caption{}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{Information Gain - No discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
0 & 57 & 73.08\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, rr, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, mentalhealth, spine3, normalvitals, spineany, headany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, any\_cancer, any\_dm, any\_heart, faceany, face3, upperlimboponly, any\_kidney, thoracotomy, any\_liver, married, pelvic\_fracture, upperlimbany, ext3, alcohol, sex, neckany, upper3, transfer, neck3, upper\_limbop, externalany, english \\ \hline
0.001 & 47 & 60.26\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, rr, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, mentalhealth, spine3, normalvitals, spineany, headany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, any\_cancer, any\_dm, any\_heart, faceany, face3, upperlimboponly, any\_kidney, thoracotomy, any\_liver, married, pelvic\_fracture, upperlimbany \\ \hline
0.005 & 33 & 42.31\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, rr, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, mentalhealth, spine3, normalvitals, spineany, headany, headface, chestany, abdo3, penetrating \\ \hline
0.01 & 27 & 34.62\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, rr, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, mentalhealth, spine3, normalvitals \\ \hline
0.05 & 5 & 6.41\% & iss, triss, traumatier, icu, bodyregions \\ \hline
0.1 & 2 & 2.56\% & iss, triss \\ \hline
0.15 & 1 & 1.28\% & iss \\ \hline
0.2 & 0 & 0.00\% &  \\ \hline
0.3 & 0 & 0.00\% &  \\ \hline
0.5 & 0 & 0.00\% &  \\ \hline
\end{tabular}}
\end{table}

### 2014-10-28 thesis/appendix/tr-nothreshold-features.tex
Add appendices and results for trauma data

\begin{table}[hbp]
\caption{}
\begin{tabular}{|l|r|r|p{8.5cm}|}
\hline
\multicolumn{ 4}{|c|}{Discretisation} \\ \hline
Method & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
CFS & 9 & 11.54\% & icu, iss, traumatier, bodyregions, lower3, lower\_limbop, triss, upperlimbonly, any\_cancer \\ \hline
C4.5 Wrapper & 11 & 14.10\% & age, bp, icu, iss, faceany, lowerlimbany, ext3, operation, upperlimboponly, lowerlimbnopelvis, any\_cancer \\ \hline
Baseline & 19 & 24.36\% & normalvitals, gcs1, iss8, age65, penetrating, headany, chestany, abdoany, spineany, upperlimbany, lowerlimbany, mentalhealth, operation, anymedical, mva, mba, pedestrian, fallsonly, cyclist \\ \hline
Expert Selection & 11 & 14.10\% & age, normalvitals, gcs1, icu1, iss, transfer, mechcode, upperlimbany, lowerlimbany, head3, charlson \\ \hline
\end{tabular}
\vspace{1cm}
\vfill
\begin{tabular}{|l|r|r|p{8.5cm}|}
\hline
\multicolumn{ 4}{|c|}{No discretisation} \\ \hline
Method & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
CFS & 9 & 11.54\% & icu, iss, traumatier, bodyregions, lower3, lower\_limbop, triss, upperlimbonly, any\_cancer \\ \hline
C4.5 Wrapper & 10 & 12.82\% & gcs, iss, abdo3, ext3, laparotomy, upper\_limbop, lower\_limbop, charlson, triss, any\_kidney \\ \hline
Baseline & 19 & 24.36\% & normalvitals, gcs1, iss8, age65, penetrating, headany, chestany, abdoany, spineany, upperlimbany, lowerlimbany, mentalhealth, operation, anymedical, mva, mba, pedestrian, fallsonly, cyclist \\ \hline
Expert Selection & 11 & 14.10\% & age, normalvitals, gcs1, icu1, iss, transfer, mechcode, upperlimbany, lowerlimbany, head3, charlson \\ \hline
\end{tabular}
\end{table}

### 2014-10-28 thesis/appendix/tr-oner-threshold-features.tex
Add appendices and results for trauma data

\begin{table}[htbp]
\caption{}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{1R Accuracy - Discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
66.905 & 59 & 75.64\% & iss, upperlimbonly, triss, arrivalmode, mechcode, ext3, chestany, headface, faceany, neckany, any\_stroketia, headany, bodyregions, externalany, lowerlimbany, spineany, upperlimbany, abdoany, traumateam, penetrating, face3, normalvitals, bp, sex, pr, rr, gcs, traumatier, transfer, icu, head3, chest3, neck3, headminor, operation, alcohol, mentalhealth, upperlimboponly, lowerlimbnopelvis, english, any\_heart, any\_cancer, any\_kidney, any\_liver, charlson, married, any\_dm, laparotomy, upper3, spine3, abdo3, lower3, thoracotomy, lower\_limbop, neurosurgery, upper\_limbop, pelvic\_fracture, anypelvis, age \\ \hline
67.905 & 4 & 5.13\% & iss, upperlimbonly, triss, arrivalmode \\ \hline
68.905 & 2 & 2.56\% & iss, upperlimbonly \\ \hline
69.905 & 2 & 2.56\% & iss, upperlimbonly \\ \hline
71.905 & 0 & 0.00\% &  \\ \hline
74.405 & 0 & 0.00\% &  \\ \hline
76.905 & 0 & 0.00\% &  \\ \hline
81.905 & 0 & 0.00\% &  \\ \hline
86.905 & 0 & 0.00\% &  \\ \hline
96.905 & 0 & 0.00\% &  \\ \hline
\end{tabular}}
\vspace{1cm}
\vfill
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{1R Accuracy - No discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
66.905 & 55 & 70.51\% & triss, iss, upperlimbonly, arrivalmode, mechcode, bodyregions, ext3, chestany, spineany, abdoany, head3, upperlimbany, lowerlimbany, headface, externalany, neckany, any\_stroketia, faceany, headany, gcs, normalvitals, sex, icu, transfer, traumatier, neck3, penetrating, traumateam, face3, chest3, any\_dm, headminor, operation, alcohol, mentalhealth, upperlimboponly, lowerlimbnopelvis, english, any\_heart, any\_cancer, any\_kidney, any\_liver, abdo3, charlson, married, thoracotomy, lower3, upper3, spine3, lower\_limbop, laparotomy, neurosurgery, anypelvis, upper\_limbop, pelvic\_fracture \\ \hline
67.905 & 4 & 5.13\% & triss, iss, upperlimbonly, arrivalmode \\ \hline
68.905 & 3 & 3.85\% & triss, iss, upperlimbonly \\ \hline
69.905 & 3 & 3.85\% & triss, iss, upperlimbonly \\ \hline
71.905 & 1 & 1.28\% & triss \\ \hline
74.405 & 0 & 0.00\% &  \\ \hline
76.905 & 0 & 0.00\% &  \\ \hline
81.905 & 0 & 0.00\% &  \\ \hline
86.905 & 0 & 0.00\% &  \\ \hline
96.905 & 0 & 0.00\% &  \\ \hline
\end{tabular}}
\end{table}

### 2014-10-28 code/weka/RankedDistance.java
Add functionality to dynamically specify degree of decay function

@@ -3,6 +3,9 @@
  */
 package weka.core;
 
+import java.util.Enumeration;
+import java.util.Vector;
+
 import weka.attributeSelection.ASEvaluation;
 import weka.attributeSelection.ASSearch;
 import weka.attributeSelection.CorrelationAttributeEval;
@@ -45,6 +48,9 @@ implements TechnicalInformationHandler {
 	
 	/** How much to update a weight if we find it important. */
 	protected static double m_WeightUpdateFactor = 0.0001;
+	
+	/** The degree of the decay function. */
+	protected static double m_DecayDegree = 1;
 
 	/**
 	 * 
@@ -61,6 +67,21 @@ implements TechnicalInformationHandler {
 		super(data);
 	}
 
+	/**
+	 * Sets the decay degree.
+	 */
+	protected void setDecayDegree(double decayDegree) {
+		m_DecayDegree = decayDegree;
+	}
+	
+	/**
+	 * Gets the decay degree.
+	 * @return the decay degree as a double
+	 */
+	protected double getDecayDegree() {
+		return m_DecayDegree;
+	}
+	
 	/**
 	 * initializes the ranges and the attributes being used.
 	 */
@@ -95,8 +116,8 @@ implements TechnicalInformationHandler {
 	/**
 	 * Given an attribute index, describes the decay of the ranking as it decreases.
 	 */
-	protected double decayFunction(int index) {
-		return 1.0 / Math.pow(index+1, 2.0/3.0);
+	protected static double decayFunction(int index) {
+		return 1.0 / Math.pow(index+1, m_DecayDegree);
 	}
 
 	/* (non-Javadoc)
@@ -121,6 +142,88 @@ implements TechnicalInformationHandler {
 				+ "is low.\n\n"
 				+ "For more information, see:\n\n"
 				+ getTechnicalInformation().toString();
+	}
+	/**
+	 * Returns an enumeration describing the available options.
+	 * 
+	 * @return an enumeration of all the available options.
+	 */
+	@Override
+	public Enumeration<Option> listOptions() {
+		Vector<Option> result = new Vector<Option>();
+
+		result.add(new Option("\tTurns off the normalization of attribute \n"
+				+ "\tvalues in distance calculation.", "D", 0, "-D"));
+
+		result.addElement(new Option(
+				"\tSpecifies list of columns to used in the calculation of the \n"
+						+ "\tdistance. 'first' and 'last' are valid indices.\n"
[…]

### 2014-10-29 thesis/appendix/allresults.tex
Update results tables for trauma and general data

@@
 
-\input{appendix/tr-corr-threshold-disc.tex}
-\input{appendix/tr-corr-threshold-nodisc.tex}
+\input{appendix/tr-corr-threshold-summary.tex}
 
-\input{appendix/tr-ig-threshold-summary-disc.tex}
-\input{appendix/tr-ig-threshold-summary-nodisc.tex}
+\input{appendix/tr-ig-threshold-summary.tex}
 
-\input{appendix/tr-oner-threshold-summary-disc.tex}
-\input{appendix/tr-oner-threshold-summary-nodisc.tex}
+\input{appendix/tr-oner-threshold-summary.tex}
 
+\clearpage
 \subsection{Feature Selection Results}
@@
 
+\clearpage
+\subsection{$k$-NN Results}
+
+\input{appendix/tr-knn-summary.tex}
+
+\input{appendix/tr-rankeddist-summary.tex}
+
+\clearpage
 \section{General LOS}
@@
 
+\input{appendix/pt-all-summary.tex}
+
+\input{appendix/pt-cfs-summary.tex}
+
+\input{appendix/pt-wrapper-summary.tex}
+
+\input{appendix/pt-corr-threshold-summary.tex}
+
+\input{appendix/pt-ig-threshold-summary.tex}
+
+\input{appendix/pt-oner-threshold-summary.tex}
+
+\clearpage
 \subsection{Feature Selection Results}
+
+\input{appendix/pt-nothreshold-features.tex}
+
+\input{appendix/pt-corr-threshold-features.tex}
+
+\input{appendix/pt-ig-threshold-features.tex}
+
+\input{appendix/pt-oner-threshold-features.tex}
+
+\clearpage
+\subsection{$k$-NN Results}
+
+\input{appendix/pt-knn-summary.tex}
+
+\input{appendix/pt-rankeddist-summary.tex}

### 2014-10-29 thesis/appendix/pt-corr-threshold-features.tex
Update results tables for trauma and general data

\begin{table}[htbp]
\caption{}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{Pearson Correlation - Discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
0 & 13 & 92.86\% & EpisodeType, MainDiagnosis, Education, InpatientService, OriginEpisodeType, MainProcedure, Sex, MedicalSpecialty, AdmissionHour, MaritalStatus, AdmissionDay, Age, AdmissionMonth \\ \hline
0.01 & 13 & 92.86\% & EpisodeType, MainDiagnosis, Education, InpatientService, OriginEpisodeType, MainProcedure, Sex, MedicalSpecialty, AdmissionHour, MaritalStatus, AdmissionDay, Age, AdmissionMonth \\ \hline
0.05 & 12 & 85.71\% & EpisodeType, MainDiagnosis, Education, InpatientService, OriginEpisodeType, MainProcedure, Sex, MedicalSpecialty, AdmissionHour, MaritalStatus, AdmissionDay, Age \\ \hline
0.1 & 10 & 71.43\% & EpisodeType, MainDiagnosis, Education, InpatientService, OriginEpisodeType, MainProcedure, Sex, MedicalSpecialty, AdmissionHour, MaritalStatus \\ \hline
0.125 & 7 & 50.00\% & EpisodeType, MainDiagnosis, Education, InpatientService, OriginEpisodeType, MainProcedure, Sex \\ \hline
0.15 & 6 & 42.86\% & EpisodeType, MainDiagnosis, Education, InpatientService, OriginEpisodeType, MainProcedure \\ \hline
0.2 & 4 & 28.57\% & EpisodeType, MainDiagnosis, Education, InpatientService \\ \hline
0.3 & 1 & 7.14\% & EpisodeType \\ \hline
0.5 & 1 & 7.14\% & EpisodeType \\ \hline
0.75 & 1 & 7.14\% & EpisodeType \\ \hline
\end{tabular}}
\end{table}

\begin{table}[htbp]
\caption{}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{11cm}|}
\hline
\multicolumn{ 4}{|c|}{Pearson Correlation - No discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
0 & 14 & 100.00\% & EpisodeType, MainDiagnosis, Education, InpatientService, OriginEpisodeType, MainProcedure, Sex, MedicalSpecialty, AdmissionHour, MaritalStatus, AdmissionDay, Age, prev\_adm, AdmissionMonth \\ \hline
0.01 & 14 & 100.00\% & EpisodeType, MainDiagnosis, Education, InpatientService, OriginEpisodeType, MainProcedure, Sex, MedicalSpecialty, AdmissionHour, MaritalStatus, AdmissionDay, Age, prev\_adm, AdmissionMonth \\ \hline
0.05 & 12 & 85.71\% & EpisodeType, MainDiagnosis, Education, InpatientService, OriginEpisodeType, MainProcedure, Sex, MedicalSpecialty, AdmissionHour, MaritalStatus, AdmissionDay, Age \\ \hline
0.1 & 10 & 71.43\% & EpisodeType, MainDiagnosis, Education, InpatientService, OriginEpisodeType, MainProcedure, Sex, MedicalSpecialty, AdmissionHour, MaritalStatus \\ \hline
0.125 & 7 & 50.00\% & EpisodeType, MainDiagnosis, Education, InpatientService, OriginEpisodeType, MainProcedure, Sex \\ \hline
0.15 & 6 & 42.86\% & EpisodeType, MainDiagnosis, Education, InpatientService, OriginEpisodeType, MainProcedure \\ \hline
0.2 & 4 & 28.57\% & EpisodeType, MainDiagnosis, Education, InpatientService \\ \hline
0.3 & 1 & 7.14\% & EpisodeType \\ \hline
0.5 & 1 & 7.14\% & EpisodeType \\ \hline
0.75 & 1 & 7.14\% & EpisodeType \\ \hline
\end{tabular}}
\end{table}

### 2014-10-29 thesis/appendix/pt-ig-threshold-features.tex
Update results tables for trauma and general data

\begin{table}[htbp]
\caption{}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{Information Gain - Discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
0 & 13 & 92.86\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty, MainProcedure, Education, AdmissionHour, AdmissionDay, MaritalStatus, Age, Sex, AdmissionMonth \\ \hline
0.001 & 13 & 92.86\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty, MainProcedure, Education, AdmissionHour, AdmissionDay, MaritalStatus, Age, Sex, AdmissionMonth \\ \hline
0.005 & 12 & 85.71\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty, MainProcedure, Education, AdmissionHour, AdmissionDay, MaritalStatus, Age, Sex \\ \hline
0.01 & 12 & 85.71\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty, MainProcedure, Education, AdmissionHour, AdmissionDay, MaritalStatus, Age, Sex \\ \hline
0.05 & 8 & 57.14\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty, MainProcedure, Education, AdmissionHour \\ \hline
0.1 & 7 & 50.00\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty, MainProcedure, Education \\ \hline
0.15 & 5 & 35.71\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty \\ \hline
0.2 & 3 & 21.43\% & EpisodeType, InpatientService, MainDiagnosis \\ \hline
0.3 & 2 & 14.29\% & EpisodeType, InpatientService \\ \hline
0.5 & 0 & 0.00\% &  \\ \hline
\end{tabular}}
\end{table}

\begin{table}[htbp]
\caption{}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{Information Gain - No discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
0 & 14 & 100.00\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty, MainProcedure, Education, AdmissionHour, AdmissionDay, MaritalStatus, Age, Sex, AdmissionMonth, prev\_adm \\ \hline
0.001 & 14 & 100.00\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty, MainProcedure, Education, AdmissionHour, AdmissionDay, MaritalStatus, Age, Sex, AdmissionMonth, prev\_adm \\ \hline
0.005 & 12 & 85.71\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty, MainProcedure, Education, AdmissionHour, AdmissionDay, MaritalStatus, Age, Sex \\ \hline
0.01 & 12 & 85.71\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty, MainProcedure, Education, AdmissionHour, AdmissionDay, MaritalStatus, Age, Sex \\ \hline
0.05 & 8 & 57.14\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty, MainProcedure, Education, AdmissionHour \\ \hline
0.1 & 7 & 50.00\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty, MainProcedure, Education \\ \hline
0.15 & 5 & 35.71\% & EpisodeType, InpatientService, MainDiagnosis, OriginEpisodeType, MedicalSpecialty \\ \hline
0.2 & 3 & 21.43\% & EpisodeType, InpatientService, MainDiagnosis \\ \hline
0.3 & 2 & 14.29\% & EpisodeType, InpatientService \\ \hline
0.5 & 0 & 0.00\% &  \\ \hline
\end{tabular}}
\end{table}

### 2014-10-29 thesis/appendix/pt-nothreshold-features.tex
Update results tables for trauma and general data

\begin{table}[h]
\caption{}
\begin{tabular}{|l|r|r|p{8.5cm}|}
\hline
\multicolumn{ 4}{|c|}{No Discretisation} \\ \hline
Method & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
CFS & 2 & 14.29\% & Episode Type, Main Diagnosis \\ \hline
C4.5 Wrapper & 4 & 28.57\% & Education, Episode Type, Inpatient Service, Main Diagnosis \\ \hline
\end{tabular}
\vspace{1cm}
\vfill
\begin{tabular}{|l|r|r|p{8.5cm}|}
\hline
\multicolumn{ 4}{|c|}{Discretisation} \\ \hline
Method & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
CFS & 2 & 14.29\% & Episode Type, Main Diagnosis \\ \hline
C4.5 Wrapper & 4 & 28.57\% & Education, Episode Type, Inpatient Service, Main Diagnosis \\ \hline
\end{tabular}
\end{table}

### 2014-10-29 thesis/appendix/pt-oner-threshold-features.tex
Update results tables for trauma and general data

\begin{table}[htbp]
\caption{}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{1R Accuracy - Discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
66.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
67.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
68.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
69.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
71.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
74.405 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
76.905 & 6 & 42.86\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure \\ \hline
81.905 & 3 & 21.43\% & EpisodeType, InpatientService, OriginEpisodeType \\ \hline
86.905 & 2 & 14.29\% & EpisodeType, InpatientService \\ \hline
96.905 & 0 & 0.00\% &  \\ \hline
\end{tabular}}
\end{table}

% nodisc is exactly the same

### 2014-10-29 thesis/appendix/pt-rankeddist-summary.tex
Update results tables for trauma and general data

\begin{sidewaystable}[htbp]
\vspace{15cm}
\caption{}
\medskip
\resizebox{\linewidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{|*{21}{l|}}
\hline
 & \multicolumn{20}{c|}{Ranked Distance $c$ - Discretisation} \\ \hline
k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
 & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} \\ \hline
1 & 92.77\% & 0.04\% & 96.45\% & 0.05\% & 94.72\% & 0.02\% & 96.86\% & 0.03\% & 94.45\% & 0.04\% & 96.71\% & 0.03\% & 94.62\% & 0.02\% & 96.71\% & 0.03\% & 94.38\% & 0.03\% & 96.51\% & 0.03\% \\ \hline
5 & 93.61\% & 0.06\% & 96.64\% & 0.03\% & 95.27\% & 0.03\% & 97.05\% & 0.04\% & 95.08\% & 0.03\% & 96.96\% & 0.02\% & 95.15\% & 0.03\% & 96.98\% & 0.04\% & 94.96\% & 0.03\% & 96.77\% & 0.02\% \\ \hline
10 & 93.76\% & 0.05\% & 96.92\% & 0.04\% & 95.38\% & 0.04\% & 97.25\% & 0.04\% & 95.04\% & 0.04\% & 97.18\% & 0.02\% & 95.10\% & 0.04\% & 97.17\% & 0.02\% & 94.97\% & 0.03\% & 97.02\% & 0.03\% \\ \hline
15 & 94.42\% & 0.04\% & 97.15\% & 0.03\% & 95.57\% & 0.02\% & 97.43\% & 0.02\% & 95.38\% & 0.04\% & 97.35\% & 0.02\% & 95.43\% & 0.03\% & 97.29\% & 0.02\% & 95.17\% & 0.02\% & 97.17\% & 0.02\% \\ \hline
20 & 95.69\% & 0.03\% & 97.33\% & 0.03\% & 96.30\% & 0.03\% & 97.56\% & 0.03\% & 96.14\% & 0.03\% & 97.47\% & 0.03\% & 96.21\% & 0.03\% & 97.41\% & 0.02\% & 95.97\% & 0.02\% & 97.27\% & 0.02\% \\ \hline
\end{tabular}}
\label{}
\vspace{0.5cm}
\vfill
\resizebox{\linewidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{|*{21}{l|}}
\hline
 & \multicolumn{20}{c|}{Ranked Distance $c$ - Discretisation} \\ \hline
k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
 & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} \\ \hline
1 & 0.904 & 0.0009 & 0.957 & 0.0008 & 0.963 & 0.0005 & 0.984 & 0.0003 & 0.970 & 0.0005 & 0.988 & 0.0004 & 0.973 & 0.0004 & 0.990 & 0.0003 & 0.974 & 0.0003 & 0.990 & 0.0002 \\ \hline
5 & 0.917 & 0.0009 & 0.960 & 0.0006 & 0.969 & 0.0007 & 0.985 & 0.0003 & 0.976 & 0.0003 & 0.989 & 0.0004 & 0.978 & 0.0005 & 0.991 & 0.0000 & 0.978 & 0.0002 & 0.991 & 0.0003 \\ \hline
10 & 0.920 & 0.0009 & 0.964 & 0.0008 & 0.969 & 0.0004 & 0.987 & 0.0003 & 0.976 & 0.0005 & 0.991 & 0.0004 & 0.977 & 0.0004 & 0.992 & 0.0002 & 0.979 & 0.0000 & 0.992 & 0.0003 \\ \hline
15 & 0.929 & 0.0009 & 0.966 & 0.0005 & 0.972 & 0.0004 & 0.988 & 0.0003 & 0.977 & 0.0002 & 0.991 & 0.0002 & 0.979 & 0.0004 & 0.992 & 0.0000 & 0.980 & 0.0004 & 0.993 & 0.0003 \\ \hline
20 & 0.947 & 0.0006 & 0.969 & 0.0004 & 0.979 & 0.0003 & 0.988 & 0.0004 & 0.984 & 0.0004 & 0.991 & 0.0003 & 0.985 & 0.0003 & 0.992 & 0.0004 & 0.986 & 0.0000 & 0.993 & 0.0002 \\ \hline
\end{tabular}}
\vspace{0.5cm}
\vfill
\resizebox{\linewidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{|*{21}{l|}}
\hline
 & \multicolumn{20}{c|}{Ranked Distance $c$ - No Discretisation} \\ \hline
k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
 & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} \\ \hline
1 & 92.89\% & 0.04\% & 96.48\% & 0.04\% & 94.81\% & 0.04\% & 97.07\% & 0.02\% & 94.52\% & 0.01\% & 96.83\% & 0.03\% & 94.69\% & 0.03\% & 96.83\% & 0.02\% & 94.41\% & 0.02\% & 96.61\% & 0.01\% \\ \hline
5 & 93.70\% & 0.05\% & 96.68\% & 0.02\% & 95.44\% & 0.04\% & 97.21\% & 0.03\% & 95.10\% & 0.04\% & 97.10\% & 0.03\% & 95.20\% & 0.04\% & 97.10\% & 0.02\% & 95.08\% & 0.04\% & 96.87\% & 0.03\% \\ \hline
10 & 93.93\% & 0.05\% & 96.94\% & 0.02\% & 95.56\% & 0.03\% & 97.42\% & 0.03\% & 95.30\% & 0.03\% & 97.26\% & 0.02\% & 95.35\% & 0.04\% & 97.27\% & 0.02\% & 95.16\% & 0.03\% & 97.10\% & 0.02\% \\ \hline
15 & 94.56\% & 0.04\% & 97.20\% & 0.03\% & 95.90\% & 0.03\% & 97.55\% & 0.02\% & 95.60\% & 0.02\% & 97.41\% & 0.03\% & 95.68\% & 0.02\% & 97.37\% & 0.02\% & 95.40\% & 0.02\% & 97.24\% & 0.01\% \\ \hline
20 & 95.75\% & 0.02\% & 97.38\% & 0.03\% & 96.52\% & 0.03\% & 97.63\% & 0.01\% & 96.30\% & 0.04\% & 97.52\% & 0.01\% & 96.36\% & 0.01\% & 97.49\% & 0.03\% & 96.13\% & 0.02\% & 97.32\% & 0.02\% \\ \hline
\end{tabular}}
\vspace{0.5cm}
\vfill
\resizebox{\linewidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{|*{21}{l|}}
\hline
 & \multicolumn{20}{c|}{Ranked Distance $c$ - No Discretisation} \\ \hline
k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
 & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} \\ \hline
1 & 0.905 & 0.0011 & 0.956 & 0.0007 & 0.965 & 0.0007 & 0.984 & 0.0003 & 0.973 & 0.0004 & 0.989 & 0.0003 & 0.975 & 0.0003 & 0.990 & 0.0003 & 0.977 & 0.0002 & 0.991 & 0.0003 \\ \hline
5 & 0.917 & 0.0012 & 0.958 & 0.0006 & 0.969 & 0.0006 & 0.986 & 0.0005 & 0.977 & 0.0004 & 0.990 & 0.0004 & 0.979 & 0.0004 & 0.991 & 0.0003 & 0.980 & 0.0003 & 0.992 & 0.0000 \\ \hline
10 & 0.922 & 0.0006 & 0.962 & 0.0006 & 0.971 & 0.0005 & 0.988 & 0.0002 & 0.978 & 0.0003 & 0.991 & 0.0003 & 0.980 & 0.0003 & 0.993 & 0.0003 & 0.981 & 0.0003 & 0.993 & 0.0003 \\ \hline
15 & 0.930 & 0.0011 & 0.966 & 0.0003 & 0.974 & 0.0004 & 0.988 & 0.0004 & 0.979 & 0.0004 & 0.991 & 0.0003 & 0.982 & 0.0003 & 0.993 & 0.0003 & 0.982 & 0.0003 & 0.993 & 0.0000 \\ \hline
20 & 0.946 & 0.0008 & 0.967 & 0.0007 & 0.979 & 0.0004 & 0.988 & 0.0003 & 0.986 & 0.0003 & 0.992 & 0.0003 & 0.987 & 0.0002 & 0.993 & 0.0003 & 0.987 & 0.0004 & 0.993 & 0.0002 \\ \hline
\end{tabular}}
\end{sidewaystable}

### 2014-10-29 thesis/appendix/tr-rankeddist-summary.tex
Update results tables for trauma and general data

\begin{sidewaystable}[htbp]
\vspace{15cm}
\caption{}
\medskip
\resizebox{\linewidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{|*{21}{l|}}
\hline
 & \multicolumn{20}{c|}{Ranked Distance $c$ - Discretisation} \\ \hline
k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
 & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} \\ \hline
1 & 68.16\% & 0.27\% & 70.59\% & 0.30\% & 72.28\% & 0.30\% & 75.05\% & 0.29\% & 74.28\% & 0.25\% & 76.14\% & 0.24\% & 74.50\% & 0.23\% & 76.27\% & 0.19\% & 74.75\% & 0.30\% & 77.04\% & 0.22\% \\ \hline
5 & 68.43\% & 0.21\% & 70.35\% & 0.16\% & 73.40\% & 0.28\% & 75.19\% & 0.27\% & 74.76\% & 0.22\% & 76.31\% & 0.07\% & 75.01\% & 0.27\% & 76.37\% & 0.26\% & 75.41\% & 0.18\% & 76.88\% & 0.16\% \\ \hline
10 & 69.84\% & 0.12\% & 70.50\% & 0.12\% & 73.98\% & 0.19\% & 74.93\% & 0.18\% & 75.22\% & 0.21\% & 76.17\% & 0.12\% & 75.59\% & 0.25\% & 76.29\% & 0.16\% & 75.57\% & 0.14\% & 76.62\% & 0.24\% \\ \hline
15 & 70.14\% & 0.25\% & 70.42\% & 0.22\% & 74.58\% & 0.25\% & 74.96\% & 0.21\% & 75.49\% & 0.18\% & 76.14\% & 0.20\% & 75.83\% & 0.21\% & 76.20\% & 0.17\% & 76.46\% & 0.19\% & 76.81\% & 0.19\% \\ \hline
20 & 70.61\% & 0.33\% & 70.65\% & 0.16\% & 74.82\% & 0.28\% & 74.91\% & 0.20\% & 75.82\% & 0.17\% & 75.59\% & 0.17\% & 75.89\% & 0.14\% & 76.20\% & 0.29\% & 76.69\% & 0.19\% & 76.84\% & 0.24\% \\ \hline
\end{tabular}}
\label{}
\vspace{0.5cm}
\vfill
\resizebox{\linewidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{|*{21}{l|}}
\hline
 & \multicolumn{20}{c|}{Ranked Distance $c$ - Discretisation} \\ \hline
k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
 & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} \\ \hline
1 & 0.668 & 0.0029 & 0.685 & 0.0039 & 0.773 & 0.0027 & 0.802 & 0.0029 & 0.796 & 0.0014 & 0.827 & 0.0014 & 0.801 & 0.0016 & 0.829 & 0.0014 & 0.806 & 0.0018 & 0.832 & 0.0009 \\ \hline
5 & 0.665 & 0.0030 & 0.681 & 0.0027 & 0.781 & 0.0026 & 0.804 & 0.0020 & 0.805 & 0.0018 & 0.830 & 0.0009 & 0.812 & 0.0014 & 0.832 & 0.0012 & 0.815 & 0.0008 & 0.834 & 0.0008 \\ \hline
10 & 0.677 & 0.0041 & 0.684 & 0.0031 & 0.785 & 0.0021 & 0.806 & 0.0018 & 0.808 & 0.0009 & 0.827 & 0.0015 & 0.818 & 0.0015 & 0.834 & 0.0013 & 0.818 & 0.0009 & 0.835 & 0.0011 \\ \hline
15 & 0.682 & 0.0044 & 0.680 & 0.0029 & 0.792 & 0.0021 & 0.806 & 0.0016 & 0.817 & 0.0016 & 0.828 & 0.0009 & 0.822 & 0.0017 & 0.832 & 0.0010 & 0.826 & 0.0012 & 0.836 & 0.0009 \\ \hline
20 & 0.686 & 0.0035 & 0.682 & 0.0028 & 0.796 & 0.0018 & 0.806 & 0.0016 & 0.821 & 0.0010 & 0.824 & 0.0015 & 0.827 & 0.0009 & 0.832 & 0.0020 & 0.830 & 0.0010 & 0.835 & 0.0013 \\ \hline
\end{tabular}}
\vspace{0.5cm}
\vfill
\resizebox{\linewidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{|*{21}{l|}}
\hline
 & \multicolumn{20}{c|}{Ranked Distance $c$ - No Discretisation} \\ \hline
k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
 & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} \\ \hline
1 & 67.03\% & 0.19\% & 69.75\% & 0.18\% & 71.88\% & 0.21\% & 74.17\% & 0.17\% & 73.58\% & 0.28\% & 75.15\% & 0.17\% & 73.42\% & 0.28\% & 74.25\% & 0.20\% & 74.41\% & 0.31\% & 74.72\% & 0.31\% \\ \hline
5 & 67.61\% & 0.24\% & 69.93\% & 0.21\% & 72.57\% & 0.39\% & 74.17\% & 0.31\% & 73.57\% & 0.21\% & 75.02\% & 0.20\% & 73.10\% & 0.19\% & 74.13\% & 0.12\% & 74.02\% & 0.23\% & 75.05\% & 0.26\% \\ \hline
10 & 68.26\% & 0.14\% & 70.06\% & 0.18\% & 72.77\% & 0.32\% & 74.01\% & 0.22\% & 74.27\% & 0.26\% & 74.90\% & 0.23\% & 73.87\% & 0.16\% & 74.46\% & 0.20\% & 73.88\% & 0.27\% & 74.87\% & 0.24\% \\ \hline
15 & 69.23\% & 0.29\% & 70.39\% & 0.32\% & 73.14\% & 0.18\% & 73.93\% & 0.18\% & 74.49\% & 0.21\% & 75.20\% & 0.24\% & 74.33\% & 0.24\% & 74.60\% & 0.20\% & 74.19\% & 0.20\% & 74.82\% & 0.19\% \\ \hline
20 & 69.73\% & 0.25\% & 71.18\% & 0.23\% & 73.90\% & 0.29\% & 73.98\% & 0.23\% & 74.82\% & 0.21\% & 75.07\% & 0.28\% & 74.19\% & 0.16\% & 74.71\% & 0.21\% & 74.52\% & 0.19\% & 74.74\% & 0.24\% \\ \hline
\end{tabular}}
\vspace{0.5cm}
\vfill
\resizebox{\linewidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{|*{21}{l|}}
\hline
 & \multicolumn{20}{c|}{Ranked Distance $c$ - No Discretisation} \\ \hline
k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
 & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} \\ \hline
1 & 0.653 & 0.0030 & 0.676 & 0.0032 & 0.771 & 0.0023 & 0.798 & 0.0022 & 0.794 & 0.0020 & 0.816 & 0.0014 & 0.802 & 0.0020 & 0.818 & 0.0009 & 0.807 & 0.0013 & 0.822 & 0.0017 \\ \hline
5 & 0.660 & 0.0040 & 0.679 & 0.0048 & 0.777 & 0.0026 & 0.797 & 0.0021 & 0.802 & 0.0013 & 0.817 & 0.0015 & 0.805 & 0.0014 & 0.820 & 0.0017 & 0.807 & 0.0017 & 0.824 & 0.0014 \\ \hline
10 & 0.661 & 0.0027 & 0.680 & 0.0041 & 0.780 & 0.0024 & 0.799 & 0.0013 & 0.806 & 0.0021 & 0.819 & 0.0010 & 0.810 & 0.0013 & 0.823 & 0.0011 & 0.810 & 0.0014 & 0.824 & 0.0012 \\ \hline
15 & 0.674 & 0.0038 & 0.682 & 0.0058 & 0.786 & 0.0012 & 0.799 & 0.0010 & 0.810 & 0.0010 & 0.821 & 0.0019 & 0.814 & 0.0014 & 0.824 & 0.0016 & 0.815 & 0.0013 & 0.825 & 0.0014 \\ \hline
20 & 0.678 & 0.0036 & 0.690 & 0.0023 & 0.794 & 0.0022 & 0.799 & 0.0019 & 0.812 & 0.0013 & 0.819 & 0.0013 & 0.816 & 0.0011 & 0.825 & 0.0007 & 0.819 & 0.0011 & 0.826 & 0.0013 \\ \hline
\end{tabular}}
\end{sidewaystable}

### 2014-10-29 vagrant/provision.sh
Add gnuplot to vagrant provision

@@ -6,7 +6,7 @@ echo "=== UPDATING PACKAGE MANAGER"
 apt-get update > /dev/null
 
 echo "=== INSTALLING TOOLS"
-apt-get install -y curl python-software-properties unzip git vim tmux > /dev/null
+apt-get install -y curl python-software-properties unzip git vim tmux gnuplot > /dev/null
 
 # OpenJDK 7 (for WEKA)
 echo "=== INSTALLING OPENJDK 7 JRE"

### 2014-10-30 code/weka/fast_summary.sh
Add a faster summarising script

@@ -1,0 +1,49 @@
+#!/usr/bin/env bash
+
+# A script that quicky calls summarise_results.sh and writes a CSV.
+
+usage_msg () {
+  echo "Usage: ./fast_summary.sh -n <num groups> -x <num runs> -f <input file>"
+}
+
+# Process command line arguments
+while [[ $# > 0 ]]; do
+  key=$1
+  shift
+  case $key in
+    -n)
+      NUMGROUPS="$1"
+      shift
+      ;;
+    -x)
+      NUMRUNS="$1"
+      shift
+      ;;
+    -f)
+      RESULTFILE="$1"
+      shift
+      ;;
+    *)
+      echo "Unknown option: $key"
+      ;;
+  esac
+done
+
+# Check that the inputs make sense
+if [ -z $NUMGROUPS ] || [ -z $NUMRUNS ] || [ -z $RESULTFILE ]; then
+  echo "Please specify all arguments."
+  usage_msg
+  exit 1
+fi
+if [ ! -e $RESULTFILE ]; then
+  echo "Input file does not exist."
+  exit 2
+fi
+
+NUMBER=^[1-9][0-9]*
+if [[ ! $NUMGROUPS =~ $NUMBER ]] || [[ ! $NUMRUNS =~ $NUMBER ]]; then
+  echo "Number of groups or runs is not a valid number."
+  exit 3
+fi
+
+eval "./summarise_results.sh -n $NUMGROUPS -x $NUMRUNS -f $RESULTFILE | python3.4 compute_ci.py | tr ' ' ',' > "$RESULTFILE.csv""

### 2014-10-30 thesis/appendix/tr-corr-threshold-features.tex
Complete all results tables for the trauma data set

@@
 \begin{table}[htbp]
-\caption{}
+\caption{Features selected from the discretised trauma data set using varying thresholds of Pearson's correlation coefficient.
+Features shown are in order from highest to lowest correlation with the class.}
 \resizebox{\linewidth}{!}{%
@@
 \begin{table}[htbp]
-\caption{}
+\caption{Features selected from the non-discretised trauma data set using varying thresholds of Pearson's correlation coefficient.
+Features shown are in order from highest to lowest correlation with the class.}
 \resizebox{\linewidth}{!}{%

### 2014-10-30 thesis/appendix/tr-ig-threshold-features.tex
Complete all results tables for the trauma data set

@@
 \begin{table}[htbp]
-\caption{}
+\caption{Features selected from the discretised trauma data set at varying thresholds of information gain. Features shown are in order from highest to lowest information gain.}
 \resizebox{\linewidth}{!}{%
@@
 0.15 & 1 & 1.28\% & iss \\ \hline
-0.2 & 0 & 0.00\% &  \\ \hline
-0.3 & 0 & 0.00\% &  \\ \hline
-0.5 & 0 & 0.00\% &  \\ \hline
+0.2 & 0 & 0.00\% & - \\ \hline
+0.3 & 0 & 0.00\% & - \\ \hline
+0.5 & 0 & 0.00\% & - \\ \hline
 \end{tabular}}
@@
 \begin{table}[htbp]
-\caption{}
+\caption{Features selected from the non-discretised trauma data set at varying thresholds of information gain. Features shown are in order from highest to lowest information gain.}
 \resizebox{\linewidth}{!}{%
@@
 0.15 & 1 & 1.28\% & iss \\ \hline
-0.2 & 0 & 0.00\% &  \\ \hline
-0.3 & 0 & 0.00\% &  \\ \hline
-0.5 & 0 & 0.00\% &  \\ \hline
+0.2 & 0 & 0.00\% & -  \\ \hline
+0.3 & 0 & 0.00\% & - \\ \hline
+0.5 & 0 & 0.00\% & - \\ \hline
 \end{tabular}}

### 2014-10-30 thesis/appendix/tr-nothreshold-features.tex
Complete all results tables for the trauma data set

@@
 \begin{table}[hbp]
-\caption{}
+\caption{Features selected using automatic methods which select the best subset and do not require a threshold (CFS and C4.5 wrapper), as well as using the manual methods: the features recommended by the domain expert and the baseline feature set from Dinh et.
+al.
+Features are not ranked with these feature selectors.}
 \begin{tabular}{|l|r|r|p{8.5cm}|}

### 2014-10-30 thesis/appendix/tr-oner-threshold-features.tex
Complete all results tables for the trauma data set

@@
 \begin{table}[htbp]
-\caption{}
+\caption{Features selected from the trauma data set using varying accuracy thresholds for the 1R feature selector. Features are shown in ranked order, from highest to lowest accuracy.}
 \resizebox{\linewidth}{!}{%
@@
 69.905 & 2 & 2.56\% & iss, upperlimbonly \\ \hline
-71.905 & 0 & 0.00\% &  \\ \hline
-74.405 & 0 & 0.00\% &  \\ \hline
-76.905 & 0 & 0.00\% &  \\ \hline
-81.905 & 0 & 0.00\% &  \\ \hline
-86.905 & 0 & 0.00\% &  \\ \hline
-96.905 & 0 & 0.00\% &  \\ \hline
+71.905 & 0 & 0.00\% & - \\ \hline
+74.405 & 0 & 0.00\% & - \\ \hline
+76.905 & 0 & 0.00\% & - \\ \hline
+81.905 & 0 & 0.00\% & - \\ \hline
+86.905 & 0 & 0.00\% & - \\ \hline
+96.905 & 0 & 0.00\% & - \\ \hline
 \end{tabular}}
@@
 71.905 & 1 & 1.28\% & triss \\ \hline
-74.405 & 0 & 0.00\% &  \\ \hline
-76.905 & 0 & 0.00\% &  \\ \hline
-81.905 & 0 & 0.00\% &  \\ \hline
-86.905 & 0 & 0.00\% &  \\ \hline
-96.905 & 0 & 0.00\% &  \\ \hline
+74.405 & 0 & 0.00\% & - \\ \hline
+76.905 & 0 & 0.00\% & - \\ \hline
+81.905 & 0 & 0.00\% & - \\ \hline
+86.905 & 0 & 0.00\% & - \\ \hline
+96.905 & 0 & 0.00\% & - \\ \hline
 \end{tabular}}

### 2014-10-30 thesis/appendix/tr-rankeddist-summary.tex
Complete all results tables for the trauma data set

@@
 \begin{sidewaystable}[htbp]
-\vspace{15cm}
-\caption{}
+\vspace{15.7cm}
+\caption{Results for the $k$-NN classifier with our proposed Ranked Distance approach on the trauma data set with and without discretisation, for varying numbers of numbers $k$ and decay constant $c$.}
 \medskip
@@
 \tabcolsep=2pt
-\begin{tabular}{|*{21}{l|}}
+\begin{tabular}{|l|*{10}{cc|}}
 \hline
- & \multicolumn{20}{c|}{Ranked Distance $c$ - Discretisation} \\ \hline
-k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
- & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} \\ \hline
-1 & 68.16\% & 0.27\% & 70.59\% & 0.30\% & 72.28\% & 0.30\% & 75.05\% & 0.29\% & 74.28\% & 0.25\% & 76.14\% & 0.24\% & 74.50\% & 0.23\% & 76.27\% & 0.19\% & 74.75\% & 0.30\% & 77.04\% & 0.22\% \\ \hline
-5 & 68.43\% & 0.21\% & 70.35\% & 0.16\% & 73.40\% & 0.28\% & 75.19\% & 0.27\% & 74.76\% & 0.22\% & 76.31\% & 0.07\% & 75.01\% & 0.27\% & 76.37\% & 0.26\% & 75.41\% & 0.18\% & 76.88\% & 0.16\% \\ \hline
-10 & 69.84\% & 0.12\% & 70.50\% & 0.12\% & 73.98\% & 0.19\% & 74.93\% & 0.18\% & 75.22\% & 0.21\% & 76.17\% & 0.12\% & 75.59\% & 0.25\% & 76.29\% & 0.16\% & 75.57\% & 0.14\% & 76.62\% & 0.24\% \\ \hline
-15 & 70.14\% & 0.25\% & 70.42\% & 0.22\% & 74.58\% & 0.25\% & 74.96\% & 0.21\% & 75.49\% & 0.18\% & 76.14\% & 0.20\% & 75.83\% & 0.21\% & 76.20\% & 0.17\% & 76.46\% & 0.19\% & 76.81\% & 0.19\% \\ \hline
+ & \multicolumn{20}{c|}{Decay constant $c$ - Discretisation} \\ \hline
+ & \multicolumn{ 2}{c|}{0.1} & \multicolumn{ 2}{c|}{0.2} & \multicolumn{ 2}{c|}{0.3} & \multicolumn{ 2}{c|}{0.4} & \multicolumn{ 2}{c|}{0.5} & \multicolumn{ 2}{c|}{0.6} & \multicolumn{ 2}{c|}{0.7} & \multicolumn{ 2}{c|}{0.8} & \multicolumn{ 2}{c|}{0.9} & \multicolumn{ 2}{c|}{1} \\
+k & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- \\ \hline
+  1 & 68.16\% & 0.27\% & 70.59\% & 0.30\% & 72.28\% & 0.30\% & 75.05\% & 0.29\% & 74.28\% & 0.25\% & 76.14\% & 0.24\% & 74.50\% & 0.23\% & 76.27\% & 0.19\% & 74.75\% & 0.30\% & 77.04\% & 0.22\% \\
+  5 & 68.43\% & 0.21\% & 70.35\% & 0.16\% & 73.40\% & 0.28\% & 75.19\% & 0.27\% & 74.76\% & 0.22\% & 76.31\% & 0.07\% & 75.01\% & 0.27\% & 76.37\% & 0.26\% & 75.41\% & 0.18\% & 76.88\% & 0.16\% \\
+  10 & 69.84\% & 0.12\% & 70.50\% & 0.12\% & 73.98\% & 0.19\% & 74.93\% & 0.18\% & 75.22\% & 0.21\% & 76.17\% & 0.12\% & 75.59\% & 0.25\% & 76.29\% & 0.16\% & 75.57\% & 0.14\% & 76.62\% & 0.24\% \\
+  15 & 70.14\% & 0.25\% & 70.42\% & 0.22\% & 74.58\% & 0.25\% & 74.96\% & 0.21\% & 75.49\% & 0.18\% & 76.14\% & 0.20\% & 75.83\% & 0.21\% & 76.20\% & 0.17\% & 76.46\% & 0.19\% & 76.81\% & 0.19\% \\
   20 & 70.61\% & 0.33\% & 70.65\% & 0.16\% & 74.82\% & 0.28\% & 74.91\% & 0.20\% & 75.82\% & 0.17\% & 75.59\% & 0.17\% & 75.89\% & 0.14\% & 76.20\% & 0.29\% & 76.69\% & 0.19\% & 76.84\% & 0.24\% \\ \hline
@@
 \tabcolsep=2pt
-\begin{tabular}{|*{21}{l|}}
+\begin{tabular}{|l|*{10}{cc|}}
 \hline
- & \multicolumn{20}{c|}{Ranked Distance $c$ - Discretisation} \\ \hline
-k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
- & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} \\ \hline
-1 & 0.668 & 0.0029 & 0.685 & 0.0039 & 0.773 & 0.0027 & 0.802 & 0.0029 & 0.796 & 0.0014 & 0.827 & 0.0014 & 0.801 & 0.0016 & 0.829 & 0.0014 & 0.806 & 0.0018 & 0.832 & 0.0009 \\ \hline
-5 & 0.665 & 0.0030 & 0.681 & 0.0027 & 0.781 & 0.0026 & 0.804 & 0.0020 & 0.805 & 0.0018 & 0.830 & 0.0009 & 0.812 & 0.0014 & 0.832 & 0.0012 & 0.815 & 0.0008 & 0.834 & 0.0008 \\ \hline
-10 & 0.677 & 0.0041 & 0.684 & 0.0031 & 0.785 & 0.0021 & 0.806 & 0.0018 & 0.808 & 0.0009 & 0.827 & 0.0015 & 0.818 & 0.0015 & 0.834 & 0.0013 & 0.818 & 0.0009 & 0.835 & 0.0011 \\ \hline
-15 & 0.682 & 0.0044 & 0.680 & 0.0029 & 0.792 & 0.0021 & 0.806 & 0.0016 & 0.817 & 0.0016 & 0.828 & 0.0009 & 0.822 & 0.0017 & 0.832 & 0.0010 & 0.826 & 0.0012 & 0.836 & 0.0009 \\ \hline
+ & \multicolumn{20}{c|}{Decay constant $c$ - Discretisation} \\ \hline
+ & \multicolumn{ 2}{c|}{0.1} & \multicolumn{ 2}{c|}{0.2} & \multicolumn{ 2}{c|}{0.3} & \multicolumn{ 2}{c|}{0.4} & \multicolumn{ 2}{c|}{0.5} & \multicolumn{ 2}{c|}{0.6} & \multicolumn{ 2}{c|}{0.7} & \multicolumn{ 2}{c|}{0.8} & \multicolumn{ 2}{c|}{0.9} & \multicolumn{ 2}{c|}{1} \\
+k & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- \\ \hline
+  1 & 0.668 & 0.0029 & 0.685 & 0.0039 & 0.773 & 0.0027 & 0.802 & 0.0029 & 0.796 & 0.0014 & 0.827 & 0.0014 & 0.801 & 0.0016 & 0.829 & 0.0014 & 0.806 & 0.0018 & 0.832 & 0.0009 \\
+  5 & 0.665 & 0.0030 & 0.681 & 0.0027 & 0.781 & 0.0026 & 0.804 & 0.0020 & 0.805 & 0.0018 & 0.830 & 0.0009 & 0.812 & 0.0014 & 0.832 & 0.0012 & 0.815 & 0.0008 & 0.834 & 0.0008 \\
+  10 & 0.677 & 0.0041 & 0.684 & 0.0031 & 0.785 & 0.0021 & 0.806 & 0.0018 & 0.808 & 0.0009 & 0.827 & 0.0015 & 0.818 & 0.0015 & 0.834 & 0.0013 & 0.818 & 0.0009 & 0.835 & 0.0011 \\
+  15 & 0.682 & 0.0044 & 0.680 & 0.0029 & 0.792 & 0.0021 & 0.806 & 0.0016 & 0.817 & 0.0016 & 0.828 & 0.0009 & 0.822 & 0.0017 & 0.832 & 0.0010 & 0.826 & 0.0012 & 0.836 & 0.0009 \\
   20 & 0.686 & 0.0035 & 0.682 & 0.0028 & 0.796 & 0.0018 & 0.806 & 0.0016 & 0.821 & 0.0010 & 0.824 & 0.0015 & 0.827 & 0.0009 & 0.832 & 0.0020 & 0.830 & 0.0010 & 0.835 & 0.0013 \\ \hline
@@
 \tabcolsep=2pt
-\begin{tabular}{|*{21}{l|}}
+\begin{tabular}{|l|*{10}{cc|}}
 \hline
- & \multicolumn{20}{c|}{Ranked Distance $c$ - No Discretisation} \\ \hline
-k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
- & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} \\ \hline
-1 & 67.03\% & 0.19\% & 69.75\% & 0.18\% & 71.88\% & 0.21\% & 74.17\% & 0.17\% & 73.58\% & 0.28\% & 75.15\% & 0.17\% & 73.42\% & 0.28\% & 74.25\% & 0.20\% & 74.41\% & 0.31\% & 74.72\% & 0.31\% \\ \hline
-5 & 67.61\% & 0.24\% & 69.93\% & 0.21\% & 72.57\% & 0.39\% & 74.17\% & 0.31\% & 73.57\% & 0.21\% & 75.02\% & 0.20\% & 73.10\% & 0.19\% & 74.13\% & 0.12\% & 74.02\% & 0.23\% & 75.05\% & 0.26\% \\ \hline
-10 & 68.26\% & 0.14\% & 70.06\% & 0.18\% & 72.77\% & 0.32\% & 74.01\% & 0.22\% & 74.27\% & 0.26\% & 74.90\% & 0.23\% & 73.87\% & 0.16\% & 74.46\% & 0.20\% & 73.88\% & 0.27\% & 74.87\% & 0.24\% \\ \hline
-15 & 69.23\% & 0.29\% & 70.39\% & 0.32\% & 73.14\% & 0.18\% & 73.93\% & 0.18\% & 74.49\% & 0.21\% & 75.20\% & 0.24\% & 74.33\% & 0.24\% & 74.60\% & 0.20\% & 74.19\% & 0.20\% & 74.82\% & 0.19\% \\ \hline
+ & \multicolumn{20}{c|}{Decay constant $c$ - No Discretisation} \\ \hline
+ & \multicolumn{ 2}{c|}{0.1} & \multicolumn{ 2}{c|}{0.2} & \multicolumn{ 2}{c|}{0.3} & \multicolumn{ 2}{c|}{0.4} & \multicolumn{ 2}{c|}{0.5} & \multicolumn{ 2}{c|}{0.6} & \multicolumn{ 2}{c|}{0.7} & \multicolumn{ 2}{c|}{0.8} & \multicolumn{ 2}{c|}{0.9} & \multicolumn{ 2}{c|}{1} \\
+k & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- \\ \hline
+  1 & 67.03\% & 0.19\% & 69.75\% & 0.18\% & 71.88\% & 0.21\% & 74.17\% & 0.17\% & 73.58\% & 0.28\% & 75.15\% & 0.17\% & 73.42\% & 0.28\% & 74.25\% & 0.20\% & 74.41\% & 0.31\% & 74.72\% & 0.31\% \\
+  5 & 67.61\% & 0.24\% & 69.93\% & 0.21\% & 72.57\% & 0.39\% & 74.17\% & 0.31\% & 73.57\% & 0.21\% & 75.02\% & 0.20\% & 73.10\% & 0.19\% & 74.13\% & 0.12\% & 74.02\% & 0.23\% & 75.05\% & 0.26\% \\
+  10 & 68.26\% & 0.14\% & 70.06\% & 0.18\% & 72.77\% & 0.32\% & 74.01\% & 0.22\% & 74.27\% & 0.26\% & 74.90\% & 0.23\% & 73.87\% & 0.16\% & 74.46\% & 0.20\% & 73.88\% & 0.27\% & 74.87\% & 0.24\% \\
+  15 & 69.23\% & 0.29\% & 70.39\% & 0.32\% & 73.14\% & 0.18\% & 73.93\% & 0.18\% & 74.49\% & 0.21\% & 75.20\% & 0.24\% & 74.33\% & 0.24\% & 74.60\% & 0.20\% & 74.19\% & 0.20\% & 74.82\% & 0.19\% \\
   20 & 69.73\% & 0.25\% & 71.18\% & 0.23\% & 73.90\% & 0.29\% & 73.98\% & 0.23\% & 74.82\% & 0.21\% & 75.07\% & 0.28\% & 74.19\% & 0.16\% & 74.71\% & 0.21\% & 74.52\% & 0.19\% & 74.74\% & 0.24\% \\ \hline
@@
 \tabcolsep=2pt
-\begin{tabular}{|*{21}{l|}}
+\begin{tabular}{|l|*{10}{cc|}}
 \hline
- & \multicolumn{20}{c|}{Ranked Distance $c$ - No Discretisation} \\ \hline
-k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
- & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} \\ \hline
-1 & 0.653 & 0.0030 & 0.676 & 0.0032 & 0.771 & 0.0023 & 0.798 & 0.0022 & 0.794 & 0.0020 & 0.816 & 0.0014 & 0.802 & 0.0020 & 0.818 & 0.0009 & 0.807 & 0.0013 & 0.822 & 0.0017 \\ \hline
-5 & 0.660 & 0.0040 & 0.679 & 0.0048 & 0.777 & 0.0026 & 0.797 & 0.0021 & 0.802 & 0.0013 & 0.817 & 0.0015 & 0.805 & 0.0014 & 0.820 & 0.0017 & 0.807 & 0.0017 & 0.824 & 0.0014 \\ \hline
-10 & 0.661 & 0.0027 & 0.680 & 0.0041 & 0.780 & 0.0024 & 0.799 & 0.0013 & 0.806 & 0.0021 & 0.819 & 0.0010 & 0.810 & 0.0013 & 0.823 & 0.0011 & 0.810 & 0.0014 & 0.824 & 0.0012 \\ \hline
-15 & 0.674 & 0.0038 & 0.682 & 0.0058 & 0.786 & 0.0012 & 0.799 & 0.0010 & 0.810 & 0.0010 & 0.821 & 0.0019 & 0.814 & 0.0014 & 0.824 & 0.0016 & 0.815 & 0.0013 & 0.825 & 0.0014 \\ \hline
+ & \multicolumn{20}{c|}{Decay constant $c$ - No Discretisation} \\ \hline
+ & \multicolumn{ 2}{c|}{0.1} & \multicolumn{ 2}{c|}{0.2} & \multicolumn{ 2}{c|}{0.3} & \multicolumn{ 2}{c|}{0.4} & \multicolumn{ 2}{c|}{0.5} & \multicolumn{ 2}{c|}{0.6} & \multicolumn{ 2}{c|}{0.7} & \multicolumn{ 2}{c|}{0.8} & \multicolumn{ 2}{c|}{0.9} & \multicolumn{ 2}{c|}{1} \\
+k & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- \\ \hline
+  1 & 0.653 & 0.0030 & 0.676 & 0.0032 & 0.771 & 0.0023 & 0.798 & 0.0022 & 0.794 & 0.0020 & 0.816 & 0.0014 & 0.802 & 0.0020 & 0.818 & 0.0009 & 0.807 & 0.0013 & 0.822 & 0.0017 \\
+  5 & 0.660 & 0.0040 & 0.679 & 0.0048 & 0.777 & 0.0026 & 0.797 & 0.0021 & 0.802 & 0.0013 & 0.817 & 0.0015 & 0.805 & 0.0014 & 0.820 & 0.0017 & 0.807 & 0.0017 & 0.824 & 0.0014 \\
+  10 & 0.661 & 0.0027 & 0.680 & 0.0041 & 0.780 & 0.0024 & 0.799 & 0.0013 & 0.806 & 0.0021 & 0.819 & 0.0010 & 0.810 & 0.0013 & 0.823 & 0.0011 & 0.810 & 0.0014 & 0.824 & 0.0012 \\
+  15 & 0.674 & 0.0038 & 0.682 & 0.0058 & 0.786 & 0.0012 & 0.799 & 0.0010 & 0.810 & 0.0010 & 0.821 & 0.0019 & 0.814 & 0.0014 & 0.824 & 0.0016 & 0.815 & 0.0013 & 0.825 & 0.0014 \\
   20 & 0.678 & 0.0036 & 0.690 & 0.0023 & 0.794 & 0.0022 & 0.799 & 0.0019 & 0.812 & 0.0013 & 0.819 & 0.0013 & 0.816 & 0.0011 & 0.825 & 0.0007 & 0.819 & 0.0011 & 0.826 & 0.0013 \\ \hline

### 2014-10-30 thesis/appendix/pt-corr-threshold-features.tex
All results are complete and tabulated nicely!

@@
 \begin{table}[htbp]
-\caption{}
+\caption{Features selected from the discretised general hospital data set using varying thresholds of Pearson's correlation coefficient.
+Features shown are in order from highest to lowest correlation with the class.}
 \resizebox{\linewidth}{!}{%
@@
 \begin{table}[htbp]
-\caption{}
+\caption{Features selected from the non-discretised general hospital data set using varying thresholds of Pearson's correlation coefficient.
+Features shown are in order from highest to lowest correlation with the class.}
 \resizebox{\linewidth}{!}{%

### 2014-10-30 thesis/appendix/pt-ig-threshold-features.tex
All results are complete and tabulated nicely!

@@
 \begin{table}[htbp]
-\caption{}
+\caption{Features selected from the discretised general hospital data set at varying thresholds of information gain. Features shown are in order from highest to lowest information gain.}
 \resizebox{\linewidth}{!}{%
@@
 0.3 & 2 & 14.29\% & EpisodeType, InpatientService \\ \hline
-0.5 & 0 & 0.00\% &  \\ \hline
+0.5 & 0 & 0.00\% & - \\ \hline
 \end{tabular}}
@@
 \begin{table}[htbp]
-\caption{}
+\caption{Features selected from the non-discretised general hospital data set at varying thresholds of information gain. Features shown are in order from highest to lowest information gain.}
 \resizebox{\linewidth}{!}{%

### 2014-10-30 thesis/appendix/pt-nothreshold-features.tex
All results are complete and tabulated nicely!

@@
-\begin{table}[h]
-\caption{}
+\begin{table}[htbp]
+\caption{Features selected using automatic methods which select the best subset and do not require a threshold (CFS and C4.5 wrapper).
+Features are not ranked with these feature selectors.
+The same features are selected by each method regardless of discretisation or not.}
 \begin{tabular}{|l|r|r|p{8.5cm}|}
 \hline
-\multicolumn{ 4}{|c|}{No Discretisation} \\ \hline
+\multicolumn{ 4}{|c|}{Discretisation/No Discretisation} \\ \hline
 Method & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
 CFS & 2 & 14.29\% & Episode Type, Main Diagnosis \\ \hline
-C4.5 Wrapper & 4 & 28.57\% & Education, Episode Type, Inpatient Service, Main Diagnosis \\ \hline
-\end{tabular}
-\vspace{1cm}
-\vfill
-\begin{tabular}{|l|r|r|p{8.5cm}|}
-\hline
-\multicolumn{ 4}{|c|}{Discretisation} \\ \hline
-Method & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
-CFS & 2 & 14.29\% & Episode Type, Main Diagnosis \\ \hline
-C4.5 Wrapper & 4 & 28.57\% & Education, Episode Type, Inpatient Service, Main Diagnosis \\ \hline
+C4.5 Wrapper & 4 & 28.57\% & Education, EpisodeType, InpatientService, MainDiagnosis \\ \hline
 \end{tabular}

### 2014-10-30 thesis/appendix/pt-oner-threshold-features.tex
All results are complete and tabulated nicely!

@@
 \begin{table}[htbp]
-\caption{}
+\caption{Features selected from the general hospital data set using varying accuracy thresholds for the 1R feature selector.
+Features are shown in ranked order, from highest to lowest accuracy.
+The same features are selected regardless of discretisation.}
 \resizebox{\linewidth}{!}{%
@@
 \hline
-\multicolumn{ 4}{|c|}{1R Accuracy - Discretisation} \\ \hline
+\multicolumn{ 4}{|c|}{1R Accuracy - Discretisation/No Discretisation} \\ \hline
 \multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
@@
 86.905 & 2 & 14.29\% & EpisodeType, InpatientService \\ \hline
-96.905 & 0 & 0.00\% &  \\ \hline
+96.905 & 0 & 0.00\% & - \\ \hline
 \end{tabular}}

### 2014-10-30 thesis/appendix/pt-rankeddist-summary.tex
All results are complete and tabulated nicely!

@@
 \begin{sidewaystable}[htbp]
-\vspace{15cm}
-\caption{}
+\vspace{15.7cm}
+\caption{Results for the $k$-NN classifier with our proposed Ranked Distance approach on the general hospital data set with and without discretisation, for varying numbers of numbers $k$ and decay constant $c$.}
 \medskip
@@
 \tabcolsep=2pt
-\begin{tabular}{|*{21}{l|}}
+\begin{tabular}{|l|*{10}{cc|}}
 \hline
- & \multicolumn{20}{c|}{Ranked Distance $c$ - Discretisation} \\ \hline
-k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
- & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} \\ \hline
+ & \multicolumn{20}{c|}{Decay constant $c$ - Discretisation} \\ \hline
+ & \multicolumn{ 2}{c|}{0.1} & \multicolumn{ 2}{c|}{0.2} & \multicolumn{ 2}{c|}{0.3} & \multicolumn{ 2}{c|}{0.4} & \multicolumn{ 2}{c|}{0.5} & \multicolumn{ 2}{c|}{0.6} & \multicolumn{ 2}{c|}{0.7} & \multicolumn{ 2}{c|}{0.8} & \multicolumn{ 2}{c|}{0.9} & \multicolumn{ 2}{c|}{1} \\
+k & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- \\ \hline
   1 & 92.77\% & 0.04\% & 96.45\% & 0.05\% & 94.72\% & 0.02\% & 96.86\% & 0.03\% & 94.45\% & 0.04\% & 96.71\% & 0.03\% & 94.62\% & 0.02\% & 96.71\% & 0.03\% & 94.38\% & 0.03\% & 96.51\% & 0.03\% \\ \hline
@@
 \tabcolsep=2pt
-\begin{tabular}{|*{21}{l|}}
+\begin{tabular}{|l|*{10}{cc|}}
 \hline
- & \multicolumn{20}{c|}{Ranked Distance $c$ - Discretisation} \\ \hline
-k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
- & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} \\ \hline
+ & \multicolumn{20}{c|}{Decay constant $c$ - Discretisation} \\ \hline
+ & \multicolumn{ 2}{c|}{0.1} & \multicolumn{ 2}{c|}{0.2} & \multicolumn{ 2}{c|}{0.3} & \multicolumn{ 2}{c|}{0.4} & \multicolumn{ 2}{c|}{0.5} & \multicolumn{ 2}{c|}{0.6} & \multicolumn{ 2}{c|}{0.7} & \multicolumn{ 2}{c|}{0.8} & \multicolumn{ 2}{c|}{0.9} & \multicolumn{ 2}{c|}{1} \\
+k & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- \\ \hline
   1 & 0.904 & 0.0009 & 0.957 & 0.0008 & 0.963 & 0.0005 & 0.984 & 0.0003 & 0.970 & 0.0005 & 0.988 & 0.0004 & 0.973 & 0.0004 & 0.990 & 0.0003 & 0.974 & 0.0003 & 0.990 & 0.0002 \\ \hline
@@
 \tabcolsep=2pt
-\begin{tabular}{|*{21}{l|}}
+\begin{tabular}{|l|*{10}{cc|}}
 \hline
- & \multicolumn{20}{c|}{Ranked Distance $c$ - No Discretisation} \\ \hline
-k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
- & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{Acc} & \multicolumn{ 1}{l|}{+/-} \\ \hline
+ & \multicolumn{20}{c|}{Decay constant $c$ - No Discretisation} \\ \hline
+ & \multicolumn{ 2}{c|}{0.1} & \multicolumn{ 2}{c|}{0.2} & \multicolumn{ 2}{c|}{0.3} & \multicolumn{ 2}{c|}{0.4} & \multicolumn{ 2}{c|}{0.5} & \multicolumn{ 2}{c|}{0.6} & \multicolumn{ 2}{c|}{0.7} & \multicolumn{ 2}{c|}{0.8} & \multicolumn{ 2}{c|}{0.9} & \multicolumn{ 2}{c|}{1} \\
+k & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- \\ \hline
   1 & 92.89\% & 0.04\% & 96.48\% & 0.04\% & 94.81\% & 0.04\% & 97.07\% & 0.02\% & 94.52\% & 0.01\% & 96.83\% & 0.03\% & 94.69\% & 0.03\% & 96.83\% & 0.02\% & 94.41\% & 0.02\% & 96.61\% & 0.01\% \\ \hline
@@
 \tabcolsep=2pt
-\begin{tabular}{|*{21}{l|}}
+\begin{tabular}{|l|*{10}{cc|}}
 \hline
- & \multicolumn{20}{c|}{Ranked Distance $c$ - No Discretisation} \\ \hline
-k & \multicolumn{ 2}{l|}{0.1} & \multicolumn{ 2}{l|}{0.2} & \multicolumn{ 2}{l|}{0.3} & \multicolumn{ 2}{l|}{0.4} & \multicolumn{ 2}{l|}{0.5} & \multicolumn{ 2}{l|}{0.6} & \multicolumn{ 2}{l|}{0.7} & \multicolumn{ 2}{l|}{0.8} & \multicolumn{ 2}{l|}{0.9} & \multicolumn{ 2}{l|}{1} \\ \hline
- & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} & \multicolumn{ 1}{l|}{AUC} & \multicolumn{ 1}{l|}{+/-} \\ \hline
+ & \multicolumn{20}{c|}{Decay constant $c$ - Discretisation} \\ \hline
+ & \multicolumn{ 2}{c|}{0.1} & \multicolumn{ 2}{c|}{0.2} & \multicolumn{ 2}{c|}{0.3} & \multicolumn{ 2}{c|}{0.4} & \multicolumn{ 2}{c|}{0.5} & \multicolumn{ 2}{c|}{0.6} & \multicolumn{ 2}{c|}{0.7} & \multicolumn{ 2}{c|}{0.8} & \multicolumn{ 2}{c|}{0.9} & \multicolumn{ 2}{c|}{1} \\
+k & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- \\ \hline
   1 & 0.905 & 0.0011 & 0.956 & 0.0007 & 0.965 & 0.0007 & 0.984 & 0.0003 & 0.973 & 0.0004 & 0.989 & 0.0003 & 0.975 & 0.0003 & 0.990 & 0.0003 & 0.977 & 0.0002 & 0.991 & 0.0003 \\ \hline

### 2014-10-31 thesis/results/overall-trauma-acc.tex
Add overall results for trauma data set

\begin{table}[htbp]
\caption{Overall accuracy of each classifier with a particular feature selection method on the trauma data set. `D' indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|l|cccccccc|}
\hline
Classifier & None & Baseline & Expert & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
ZeroR & 66.91 & 66.91 & 66.91 & 66.91 & 66.91 & 66.91 & 66.91 & 66.91 \\
1R & 71.95 & 66.91 & \begin{tabular}[c]{@{}c@{}}70.64\\ D\end{tabular} & \textbf{72.01} & 71.93 & \begin{tabular}[c]{@{}c@{}}71.91\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.95\\ t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.97\\ t=66.905\end{tabular} \\
NB & \begin{tabular}[c]{@{}c@{}}73.83\\ D\end{tabular} & 74.62 & \begin{tabular}[c]{@{}c@{}}73.19\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.4\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}75.8\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.85\\ D, t=0.125\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.45\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}72.37\\ D, t=69.905\end{tabular} \\
DT & \begin{tabular}[c]{@{}c@{}}75\\ D\end{tabular} & 74.38 & \begin{tabular}[c]{@{}c@{}}76.39\\ D\end{tabular} & 75.4 & \textbf{\begin{tabular}[c]{@{}c@{}}77.52\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}76.05\\ D, t=0.11\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.17\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.49\\ t=66.905\end{tabular} \\
LR & \begin{tabular}[c]{@{}c@{}}77.62\\ D\end{tabular} & 75.06 & \begin{tabular}[c]{@{}c@{}}76.45\\ D\end{tabular} & 74.14 & \begin{tabular}[c]{@{}c@{}}76.3\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.7\\ D, t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}77.65\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.4\\ D, t=66.905\end{tabular} \\
SVM & \begin{tabular}[c]{@{}c@{}}77.39\\ D\end{tabular} & 74.47 & \begin{tabular}[c]{@{}c@{}}75.67\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}72.57\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}75.26\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}77.42\\ D, t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.53\\ D, t=0.005\end{tabular}} & \begin{tabular}[c]{@{}c@{}}73.83\\ D, t=66.905\end{tabular} \\
1NN & \begin{tabular}[c]{@{}c@{}}71.22\\ D\end{tabular} & 70.28 & \begin{tabular}[c]{@{}c@{}}72.56\\ D\end{tabular} & 74.44 & \textbf{\begin{tabular}[c]{@{}c@{}}77.81\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.75\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.26\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.78\\ t=68.905\end{tabular} \\
20NN & \begin{tabular}[c]{@{}c@{}}75.9\\ D\end{tabular} & 73.59 & \begin{tabular}[c]{@{}c@{}}75.51\\ D\end{tabular} & 75.3 & \begin{tabular}[c]{@{}c@{}}76.48\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.26\\ D, t=0.125\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.56\\ D, t=0.005\end{tabular}} & \begin{tabular}[c]{@{}c@{}}73.56\\ t=68.905\end{tabular} \\
RD & \begin{tabular}[c]{@{}c@{}}76.92\\ D\end{tabular} & 69.68 & \begin{tabular}[c]{@{}c@{}}76.19\\ D\end{tabular} & 74.38 & \textbf{\begin{tabular}[c]{@{}c@{}}77.22\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}76.87\\ D, t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.82\\ D, t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.83\\ D, t=66.905\end{tabular} \\
K* & \begin{tabular}[c]{@{}c@{}}71.4\\ D\end{tabular} & 72.20 & \begin{tabular}[c]{@{}c@{}}75.08\\ D\end{tabular} & 75.58 & \textbf{\begin{tabular}[c]{@{}c@{}}77.27\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}75.07\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.08\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.45\\ t=68.905\end{tabular} \\
MLP & \begin{tabular}[c]{@{}c@{}}73.79\\ D\end{tabular} & 73.03 & \begin{tabular}[c]{@{}c@{}}73.45\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.38\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.66\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.65\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.03\\ D, t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.94\\ D, t=66.905\end{tabular} \\ \hline
\end{tabular}
}
\end{table}

### 2014-10-31 thesis/results/overall-trauma-auc.tex
Add overall results for trauma data set

\begin{table}[htbp]
\caption{Overall AUC performance of each classifier with a particular feature selection method on the trauma data set. `D' indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.}
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|cccccccc|}
\hline
Classifier & None & Baseline & Expert & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
ZeroR & 0.498 & 0.498 & 0.498 & 0.498 & 0.498 & 0.498 & 0.498 & 0.498 \\
1R & 0.67 & 0.5 & \begin{tabular}[c]{@{}c@{}}0.581\\ D\end{tabular} & 0.668 & 0.668 & \begin{tabular}[c]{@{}c@{}}0.669\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.669\\ t=0.005\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.67\\ t=69.905\end{tabular}} \\
NB & \begin{tabular}[c]{@{}c@{}}0.828\\ D\end{tabular} & 0.804 & \begin{tabular}[c]{@{}c@{}}0.806\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.815\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.827\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.829\\ D, t=0.125\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.832\\ D, t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.8\\ D, t=66.905\end{tabular} \\
DT & \begin{tabular}[c]{@{}c@{}}0.756\\ D\end{tabular} & 0.772 & \begin{tabular}[c]{@{}c@{}}0.786\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.793\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.806\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.799\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.766\\ D, t=66.905\end{tabular} \\
LR & \begin{tabular}[c]{@{}c@{}}0.844\\ D\end{tabular} & 0.812 & \begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.819\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.831\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.846\\ D, t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.844\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.811\\ D, t=66.905\end{tabular} \\
SVM & \begin{tabular}[c]{@{}c@{}}0.843\\ D\end{tabular} & 0.787 & 0.807 & \begin{tabular}[c]{@{}c@{}}0.782\\ D\end{tabular} & 0.806 & \textbf{\begin{tabular}[c]{@{}c@{}}0.844\\ D, t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.842\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.724\\ D, t=66.905\end{tabular} \\
1NN & \begin{tabular}[c]{@{}c@{}}0.726\\ D\end{tabular} & 0.719 & \begin{tabular}[c]{@{}c@{}}0.760\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.798\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.816\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.798\\ D, t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.768\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.778\\ t=68.905\end{tabular} \\
20NN & \begin{tabular}[c]{@{}c@{}}0.827\\ D\end{tabular} & 0.792 & \begin{tabular}[c]{@{}c@{}}0.808\\ D\end{tabular} & 0.814 & \begin{tabular}[c]{@{}c@{}}0.828\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.831\\ D, t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.832\\ D, t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.799\\ D, t=66.905\end{tabular} \\
RD & \begin{tabular}[c]{@{}c@{}}0.834\\ D\end{tabular} & 0.742 & \begin{tabular}[c]{@{}c@{}}0.807\\ D\end{tabular} & 0.81 & \begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.836\\ D, t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.834\\ D, t=0.005\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.791\\ D, t=66.905\end{tabular} \\
K* & \begin{tabular}[c]{@{}c@{}}0.768\\ D\end{tabular} & 0.774 & \begin{tabular}[c]{@{}c@{}}0.803\\ D\end{tabular} & 0.815 & \textbf{\begin{tabular}[c]{@{}c@{}}0.838\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.82\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ t=68.905\end{tabular} \\
MLP & \begin{tabular}[c]{@{}c@{}}0.803\\ D\end{tabular} & 0.781 & \begin{tabular}[c]{@{}c@{}}0.784\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.818\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.803\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.772\\ D, t=68.905\end{tabular} \\ \hline
\end{tabular}
}
\end{table}

### 2014-11-01 thesis/results/features-pt-acc.tex
Update results and discussion section

\begin{table}[htbp]
\caption{Comparison of accuracy (\%) between all feature selection methods for each classifier with and without discretisation, on the general hospital data set. The best accuracy for each feature selection method is highlighted in bold, and the best accuracy for each classifier is italicised.}
\label{tab:features-pt-acc}
\resizebox{!}{11cm}{%
\begin{tabular}{|l|cccccc|}
\hline
 & None & CFS & C4.5 Wrapper & Correlation & IG & 1R \\
\hline
 \textit{1R} &  &  &  &  &  &  \\
 Discretised & 92.88 & 92.88 & 92.88 & 92.88 & 92.88 & 92.88 \\
 Not discretised & 92.88 & 92.88 & 92.88 & 92.88 & 92.88 & 92.88 \\
 \textit{NB} &  &  &  &  &  &  \\
 Discretised & 93.66 & 93.06 & 93 & \textit{\begin{tabular}[c]{@{}c@{}}93.72\\ t=0.125\end{tabular}} & \begin{tabular}[c]{@{}c@{}}93.66\\ t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}93.66\\ t=74.405\end{tabular} \\
 Not discretised & 93.66 & 93.06 & 93 & \textit{\begin{tabular}[c]{@{}c@{}}93.72\\ t=0.125\end{tabular}} & \begin{tabular}[c]{@{}c@{}}93.66\\ t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}93.66\\ t=74.405\end{tabular} \\
 \textit{DT} &  &  &  &  &  &  \\
 Discretised & \textit{\textbf{98.23}} & 96.45 & \textbf{97.43} & \begin{tabular}[c]{@{}c@{}}97.75\\ t=0.125\end{tabular} & \begin{tabular}[c]{@{}c@{}}97.76\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}97.74\\ t=74.405\end{tabular} \\
 Not discretised & \textit{\textbf{98.23}} & 96.45 & \textbf{97.43} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}98.23\\ t=0\end{tabular}}} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}98.23\\ t=0.001\end{tabular}}} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}98.23\\ t=74.405\end{tabular}}} \\
 \textit{LR} &  &  &  &  &  &  \\
 Discretised & \textit{97.93} & \textbf{96.46} & 97.19 & \begin{tabular}[c]{@{}c@{}}97.86\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}97.86\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}97.83\\ t=71.905\end{tabular} \\
 Not discretised & \textit{97.93} & \textbf{96.46} & 97.19 & \textit{\begin{tabular}[c]{@{}c@{}}97.93\\ t=0.01\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}97.93\\ t=0.001\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}97.93\\ t=74.405\end{tabular}} \\
 \textit{SVM} &  &  &  &  &  &  \\
 Discretised & \textit{98.01} & \textbf{96.46} & 97.14 & \begin{tabular}[c]{@{}c@{}}97.97\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}97.97\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}97.96\\ t=74.405\end{tabular} \\
 Not discretised & \textit{98.01} & \textbf{96.46} & 97.14 & \textit{\begin{tabular}[c]{@{}c@{}}98.01\\ t=0.01\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}98.01\\ t=0.001\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}98.01\\ t=74.405\end{tabular}} \\
 \textit{1NN} &  &  &  &  &  &  \\
 Discretised & 96.93 & 96.45 & 97.36 & \textit{\begin{tabular}[c]{@{}c@{}}97.59\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}97.57\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}97.43\\ t=76.905\end{tabular} \\
 Not discretised & 96.77 & 96.45 & 97.36 & \textit{\begin{tabular}[c]{@{}c@{}}97.59\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}97.57\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}97.43\\ t=76.905\end{tabular} \\
 \textit{20NN} &  &  &  &  &  &  \\
 Discretised & 96.53 & 96.24 & 97.15 & \textit{\begin{tabular}[c]{@{}c@{}}97.28\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}97.14\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}96.54\\ t=74.405\end{tabular} \\
 Not discretised & 96.85 & 96.24 & 97.15 & \textit{\begin{tabular}[c]{@{}c@{}}97.3\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}97.15\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}96.86\\ t=68.905\end{tabular} \\
 \textit{RD} &  &  &  &  &  &  \\
 Discretised & 97.27 & 79.1 & 95.45 & \begin{tabular}[c]{@{}c@{}}97.28\\ t=0.01\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}97.3\\ t=0\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}97.3\\ t=66.905\end{tabular}} \\
 Not discretised & 97.34 & 79.1 & 95.45 & \begin{tabular}[c]{@{}c@{}}97.3\\ t=0\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}97.36\\ t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}97.34\\ t=68.905\end{tabular} \\
 \textit{K*} &  &  &  &  &  &  \\
 Discretised & 97.22 & 92.94 & 96.53 & \textit{\begin{tabular}[c]{@{}c@{}}97.46\\ t=0.05\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}97.46\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}97.22\\ t=74.405\end{tabular} \\
 Not discretised & 97.36 & 92.94 & 96.53 & \textit{\begin{tabular}[c]{@{}c@{}}97.46\\ t=0.05\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}97.46\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}97.36\\ t=74.405\end{tabular} \\
 \textit{MLP} &  &  &  &  &  &  \\
 Discretised & 98.14 & 96.45 & 97.35 & \textit{\begin{tabular}[c]{@{}c@{}}98.22\\ t=0.05\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}98.22\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}98.14\\ t=74.405\end{tabular} \\
 Not discretised & 98.16 & 96.45 & 97.35 & \textit{\begin{tabular}[c]{@{}c@{}}98.22\\ t=0.05\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}98.22\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}98.16\\ t=74.405\end{tabular} \\ \hline
 \end{tabular}
}
\end{table}

### 2014-11-01 thesis/results/features-pt-auc.tex
Update results and discussion section

\begin{table}[htbp]
\caption{Comparison of AUC between all feature selection methods for each classifier with and without discretisation, on the general hospital data set. The best AUC for each feature selection method is highlighted in bold, and the best AUC for each classifier is italicised.}
\label{tab:features-pt-auc}
\resizebox{!}{11cm}{%
\begin{tabular}{|l|cccccc|}
 \hline
 & None & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
 \textit{1R} &  &  &  &  &  &  \\
 Discretised & 0.858 & 0.858 & 0.858 & 0.858 & 0.858 & 0.858 \\
 Not discretised & 0.858 & 0.858 & 0.858 & 0.858 & 0.858 & 0.858 \\
 \textit{NB} &  &  &  &  &  &  \\
 Discretised & 0.979 & \textit{0.982} & 0.983 & \begin{tabular}[c]{@{}c@{}}0.981\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.979\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.979\\ t=74.405\end{tabular} \\
 Not discretised & 0.976 & 0.982 & \textit{0.983} & \begin{tabular}[c]{@{}c@{}}0.981\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.979\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.979\\ t=74.405\end{tabular} \\
 \textit{DT} &  &  &  &  &  &  \\
 Discretised & 0.982 & 0.981 & 0.984 & \begin{tabular}[c]{@{}c@{}}0.983\\ t=0.2\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.989\\ t=0.001\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.989\\ t=74.405\end{tabular}} \\
 Not discretised & \textit{0.989} & 0.981 & 0.984 & \textit{\begin{tabular}[c]{@{}c@{}}0.989\\ t=0.01\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.989\\ t=0.001\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.989\\ t=74.405\end{tabular}} \\
 \textit{LR} &  &  &  &  &  &  \\
 Discretised & \textit{\textbf{0.994}} & 0.981 & 0.99 & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=0.1\end{tabular}}} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=0.05\end{tabular}}} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=74.405\end{tabular}}} \\
 Not discretised & \textit{\textbf{0.994}} & 0.981 & 0.99 & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=0.1\end{tabular}}} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=0.05\end{tabular}}} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=74.405\end{tabular}}} \\
 \textit{SVM} &  &  &  &  &  &  \\
 Discretised & \textit{\textbf{0.994}} & 0.971 & 0.981 & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=0.1\end{tabular}}} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=0.01\end{tabular}}} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=74.405\end{tabular}}} \\
 Not discretised & \textit{\textbf{0.994}} & 0.971 & 0.981 & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=0.1\end{tabular}}} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=0.01\end{tabular}}} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=74.405\end{tabular}}} \\
 \textit{1NN} &  &  &  &  &  &  \\
 Discretised & 0.979 & 0.982 & \textit{0.991} & \textit{\begin{tabular}[c]{@{}c@{}}0.991\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.990\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.989\\ t=76.905\end{tabular} \\
 Not discretised & 0.973 & 0.982 & \textit{0.991} & \textit{\begin{tabular}[c]{@{}c@{}}0.991\\ t=0.15\end{tabular}} & \scell{0.990\\t=0.1} & \begin{tabular}[c]{@{}c@{}}0.989\\ t=76.905\end{tabular} \\
 \textit{20NN} &  &  &  &  &  &  \\
 Discretised & \textit{0.992} & 0.981 & 0.991 & \textit{\begin{tabular}[c]{@{}c@{}}0.992\\ t=0.15\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.992\\ t=0.05\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.992\\ t=74.405\end{tabular}} \\
 Not discretised & \textit{0.992} & 0.981 & 0.991 & \textit{\begin{tabular}[c]{@{}c@{}}0.992\\ t=0.15\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.992\\ t=0.05\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.992\\ t=74.405\end{tabular}} \\
 \textit{RD} &  &  &  &  &  &  \\
 Discretised & \textit{0.993} & 0.794 & 0.986 & \textit{\begin{tabular}[c]{@{}c@{}}0.993\\ t=0.05\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.993\\ t=0.005\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.993\\ t=74.405\end{tabular}} \\
 Not discretised & \textit{0.993} & 0.794 & 0.986 & \textit{\begin{tabular}[c]{@{}c@{}}0.993\\ t=0.05\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.993\\ t=0.005\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.993\\ t=74.405\end{tabular}} \\
 \textit{K*} &  &  &  &  &  &  \\
 Discretised & 0.993 & 0.98 & 0.989 & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=0.1\end{tabular}}} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=0.05\end{tabular}}} & \begin{tabular}[c]{@{}c@{}}0.993\\ t=74.405\end{tabular} \\
 Not discretised & \textit{\textbf{0.994}} & 0.98 & 0.989 & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=0.1\end{tabular}}} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=0.05\end{tabular}}} & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.994\\ t=74.405\end{tabular}}} \\
 \textit{MLP} &  &  &  &  &  &  \\
 Discretised & 0.99 & \textbf{0.983} & \textbf{0.992} & \textit{\begin{tabular}[c]{@{}c@{}}0.993\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.991\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.990\\ t=76.905\end{tabular} \\
 Not discretised & 0.99 & \textbf{0.983} & \textbf{0.992} & \textit{\begin{tabular}[c]{@{}c@{}}0.993\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.991\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.990\\ t=76.905\end{tabular} \\ \hline
 \end{tabular}
}
\end{table}

### 2014-11-01 thesis/results/features-tr-acc.tex
Update results and discussion section

\begin{table}[htbp]
\caption{Comparison of accuracy (\%) between all feature selection methods for each classifier with and without discretisation, on the trauma data set. The best accuracy for each feature selection method is highlighted in bold, and the best accuracy for each classifier is italicised.}
\label{tab:features-tr-acc}
\resizebox{!}{11cm}{%
\begin{tabular}{|l|cccccccc|}
\hline
 & \multicolumn{1}{l}{None} & Baseline & Expert & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
\textit{1R} &  &  &  &  &  &  &  &  \\
Discretised & 69.74 & 66.91 & \textit{70.64} & 69.63 & 70.64 & \textit{\begin{tabular}[c]{@{}c@{}}70.64\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}69.9\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}69.77\\ t=67.905\end{tabular} \\
Not discretised & 71.95 & 66.91 & 70.61 & \textit{72.01} & 71.93 & \begin{tabular}[c]{@{}c@{}}71.91\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.95\\ t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.97\\ t=66.905\end{tabular} \\
\textit{NB} &  &  &  &  &  &  &  &  \\
Discretised & 73.83 & 74.62 & 73.19 & 73.4 & \textit{75.8} & \begin{tabular}[c]{@{}c@{}}74.85\\ t=0.125\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.45\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}72.37\\ t=69.905\end{tabular} \\
Not discretised & 67.25 & \textit{74.62} & 68.31 & 64.77 & 62.66 & \begin{tabular}[c]{@{}c@{}}71.91\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}67.65\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}66.91\\ t=71.905\end{tabular} \\
\textit{DT} &  &  &  &  &  &  &  &  \\
Discretised & 75 & 74.38 & 76.39 & 74.12 & \textit{77.52} & \begin{tabular}[c]{@{}c@{}}76.05\\ t=0.125\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.17\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}72.71\\ t=69.905\end{tabular} \\
Not discretised & 74.58 & 74.38 & 76.04 & 75.4 & 75.79 & \textit{\begin{tabular}[c]{@{}c@{}}75.93\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}75.74\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.49\\ t=66.905\end{tabular} \\
\textit{LR} &  &  &  &  &  &  &  &  \\
Discretised & \textbf{77.62} & \textbf{75.06} & \textbf{76.45} & 73.99 & 76.3 & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}77.7\\ t=0\end{tabular}}} & \textbf{\begin{tabular}[c]{@{}c@{}}77.65\\ t=0\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}74.4\\ t=66.905\end{tabular}} \\
Not discretised & 77.1 & \textbf{75.06} & 74.53 & 74.14 & 73.66 & \begin{tabular}[c]{@{}c@{}}77.09\\ t=0.01\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}77.13\\ t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}73.57\\ t=66.905\end{tabular} \\
\textit{SVM} &  &  &  &  &  &  &  &  \\
Discretised & 77.39 & 74.47 & 75.67 & 72.57 & 75.26 & \begin{tabular}[c]{@{}c@{}}77.42\\ t=0.05\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}77.53\\ t=0.005\end{tabular}} & \begin{tabular}[c]{@{}c@{}}73.83\\ t=66.905\end{tabular} \\
Not discretised & 76.72 & 74.47 & 74.48 & 70.59 & 73.58 & \begin{tabular}[c]{@{}c@{}}76.8\\ t=0.01\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}76.83\\ t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}72.27\\ t=66.905\end{tabular} \\
\textit{1NN} &  &  &  &  &  &  &  &  \\
Discretised & 71.22 & 70.28 & 72.56 & 73.84 & \textit{\textbf{77.81}} & \begin{tabular}[c]{@{}c@{}}74.75\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.26\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}72.72\\ t=69.905\end{tabular} \\
Not discretised & 67.4 & 70.28 & 68.77 & 74.44 & \textit{74.94} & \begin{tabular}[c]{@{}c@{}}72.94\\ t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.44\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.78\\ t=68.905\end{tabular} \\
\textit{20NN} &  &  &  &  &  &  &  &  \\
Discretised & 75.9 & 73.59 & 75.51 & 73.76 & 76.48 & \begin{tabular}[c]{@{}c@{}}76.26\\ t=0.125\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}76.56\\ t=0.005\end{tabular}} & \begin{tabular}[c]{@{}c@{}}73.52\\ t=66.905\end{tabular} \\
Not discretised & 73.48 & 73.59 & 72.11 & \textit{75.3} & 74.59 & \begin{tabular}[c]{@{}c@{}}74.99\\ t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.42\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.56\\ t=68.905\end{tabular} \\
\textit{RD} &  &  &  &  &  &  &  &  \\
Discretised & 76.92 & 69.68 & 76.19 & 73.39 & \textit{77.22} & \begin{tabular}[c]{@{}c@{}}76.87\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.82\\ t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.82\\ t=66.905\end{tabular} \\
Not discretised & 74.97 & 69.68 & 74.58 & 74.38 & 75.35 & \begin{tabular}[c]{@{}c@{}}76.05\\ t=0.1\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}76.16\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}73.71\\ t=68.905\end{tabular} \\
\textit{K*} &  &  &  &  &  &  &  &  \\
Discretised & 71.4 & 72.2 & 75.08 & 74.43 & \textit{77.27} & \begin{tabular}[c]{@{}c@{}}75.07\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.08\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}72.61\\ t=67.905\end{tabular} \\
Not discretised & 67.34 & 72.2 & 72.6 & \textit{\textbf{75.58}} & 75.48 & \begin{tabular}[c]{@{}c@{}}74.77\\ t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.15\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.45\\ t=68.905\end{tabular} \\
\textit{MLP} &  &  &  &  &  &  &  &  \\
Discretised & 73.79 & 73.03 & 73.45 & 73.38 & \textit{76.66} & \begin{tabular}[c]{@{}c@{}}74.41\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.03\\ t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.94\\ t=66.905\end{tabular} \\
Not discretised & 71.87 & 73.03 & 72.31 & 73.04 & 73.35 & \textit{\begin{tabular}[c]{@{}c@{}}74.65\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}72.63\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}70.99\\ t=66.905\end{tabular} \\ \hline
 \end{tabular}
}
\end{table}

### 2014-11-01 thesis/results/features-tr-auc.tex
Update results and discussion section

\begin{table}[htbp]
\caption{Comparison of AUC between all feature selection methods for each classifier with and without discretisation, on the trauma data set. The best AUC for each feature selection method is highlighted in bold, and the best AUC for each classifier is italicised.}
\label{tab:features-tr-auc}
\resizebox{!}{11cm}{%
\begin{tabular}{|l|cccccccc|}
 \hline
 & None & Baseline & Expert & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
\textit{1R} &  &  &  &  &  &  &  &  \\
Discretised & 0.569 & 0.5 & 0.581 & 0.569 & 0.581 & \begin{tabular}[c]{@{}c@{}}0.572\\ t=0.1\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.581\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.571\\ t=68.905\end{tabular} \\
Not discretised & 0.67 & 0.5 & 0.581 & 0.668 & 0.668 & \begin{tabular}[c]{@{}c@{}}0.669\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.669\\ t=0.005\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.67\\ t=69.905\end{tabular}} \\
\textit{NB} &  &  &  &  &  &  &  &  \\
Discretised & 0.828 & 0.804 & 0.806 & 0.815 & 0.827 & \begin{tabular}[c]{@{}c@{}}0.829\\ t=0.125\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.832\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.8\\ t=66.905\end{tabular} \\
Not discretised & 0.812 & 0.804 & 0.793 & 0.801 & 0.786 & \begin{tabular}[c]{@{}c@{}}0.815\\ t=0.15\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.817\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.784\\ t=66.905\end{tabular} \\
\textit{DT} &  &  &  &  &  &  &  &  \\
Discretised & 0.756 & 0.772 & 0.786 & 0.793 & \textit{0.806} & \begin{tabular}[c]{@{}c@{}}0.799\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.766\\ t=66.905\end{tabular} \\
Not discretised & 0.727 & 0.772 & 0.762 & 0.782 & 0.784 & \textit{\begin{tabular}[c]{@{}c@{}}0.783\\ t=0.2\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.757\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.756\\ t=69.905\end{tabular} \\
\textit{LR} &  &  &  &  &  &  &  &  \\
Discretised & \textbf{0.844} & \textbf{0.812} & \textbf{0.825} & \textbf{0.819} & 0.831 & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.846\\ t=0.1\end{tabular}}} & \textbf{\begin{tabular}[c]{@{}c@{}}0.844\\ t=0\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}0.811\\ t=66.905\end{tabular}} \\
Not discretised & \textit{0.837} & \textbf{0.812} & 0.812 & 0.812 & 0.809 & \textit{\begin{tabular}[c]{@{}c@{}}0.837\\ t=0\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.837\\ t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.798\\ t=66.905\end{tabular} \\
\textit{SVM} &  &  &  &  &  &  &  &  \\
Discretised & 0.843 & 0.787 & 0.767 & 0.782 & 0.796 & \textit{\begin{tabular}[c]{@{}c@{}}0.844\\ t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.842\\ t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.724\\ t=66.905\end{tabular} \\
Not discretised & \textit{0.836} & 0.787 & 0.807 & 0.778 & 0.806 & \textit{\begin{tabular}[c]{@{}c@{}}0.836\\ t=0\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.836\\ t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.704\\ t=66.905\end{tabular} \\
\textit{1NN} &  &  &  &  &  &  &  &  \\
Discretised & 0.726 & 0.719 & 0.76 & 0.798 & \textit{0.816} & \begin{tabular}[c]{@{}c@{}}0.798\\ t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.763\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.771\\ t=68.905\end{tabular} \\
Not discretised & 0.651 & 0.719 & 0.653 & 0.772 & 0.756 & \begin{tabular}[c]{@{}c@{}}0.752\\ t=0.3\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.768\\ t=0.1\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.778\\ t=68.905\end{tabular}} \\
\textit{20NN} &  &  &  &  &  &  &  &  \\
Discretised & 0.827 & 0.792 & 0.808 & 0.811 & 0.828 & \begin{tabular}[c]{@{}c@{}}0.831\\ t=0.1\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.832\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.799\\ t=66.905\end{tabular} \\
Not discretised & 0.799 & 0.792 & 0.77 & \textit{0.814} & 0.803 & \begin{tabular}[c]{@{}c@{}}0.812\\ t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.81\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ t=68.905\end{tabular} \\
\textit{RD} &  &  &  &  &  &  &  &  \\
Discretised & 0.834 & 0.742 & 0.807 & 0.806 & 0.825 & \textit{\begin{tabular}[c]{@{}c@{}}0.836\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.834\\ t=0.005\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.791\\ t=66.905\end{tabular} \\
Not discretised & 0.826 & 0.742 & 0.805 & 0.81 & 0.811 & \begin{tabular}[c]{@{}c@{}}0.83\\ t=0.1\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.831\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.788\\ t=68.905\end{tabular} \\
\textit{K*} &  &  &  &  &  &  &  &  \\
Discretised & 0.768 & 0.774 & 0.803 & 0.813 & \textit{\textbf{0.838}} & \begin{tabular}[c]{@{}c@{}}0.82\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.782\\ t=67.905\end{tabular} \\
Not discretised & 0.732 & 0.774 & 0.784 & \textit{0.815} & 0.809 & \begin{tabular}[c]{@{}c@{}}0.811\\ t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.779\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ t=68.905\end{tabular} \\
\textit{MLP} &  &  &  &  &  &  &  &  \\
Discretised & 0.803 & 0.781 & 0.784 & 0.802 & \textit{0.825} & \begin{tabular}[c]{@{}c@{}}0.807\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.803\\ t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.772\\ t=68.905\end{tabular} \\
Not discretised & 0.78 & 0.781 & 0.774 & 0.799 & 0.804 & \textit{\begin{tabular}[c]{@{}c@{}}0.818\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.788\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.752\\ t=68.905\end{tabular} \\ \hline
\end{tabular}
}
\end{table}

### 2014-11-01 thesis/results/overall-portugal-acc.tex
Update results and discussion section

\begin{table}[htbp]
\label{tab:overall-pt-acc}
\caption{Overall best accuracy of each classifier with a particular feature selection method on the general hospital data set. * indicates that results were identical regardless of discretisation. D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied. The absence of * or D means that not discretising produced a better result. The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
\begin{tabular}{|l|cccccc|}
\hline
Classifier & None & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
ZeroR & 75.64* & 75.64* & 75.64* & 75.64* & 75.64* & 75.64* \\
1R & 92.88* & 92.88* & 92.88* & 92.88* & 92.88* & 92.88* \\
NB & 93.66* & 93.06* & 93* & \textbf{\begin{tabular}[c]{@{}c@{}}93.72*\\ t=0.125\end{tabular}} & \begin{tabular}[c]{@{}c@{}}93.66*\\ t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}93.66*\\ t=74.405\end{tabular} \\
DT & \textbf{98.23*} & 96.45* & 97.43* & \begin{tabular}[c]{@{}c@{}}98.23\\ t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}98.23\\ t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}98.23\\ t=74.405\end{tabular} \\
LR & \textbf{97.93*} & 96.46* & 97.19* & \begin{tabular}[c]{@{}c@{}}97.93\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}97.93\\ t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}97.93\\ t=74.405\end{tabular} \\
SVM & \textbf{98.01*} & 96.46* & 97.14* & \begin{tabular}[c]{@{}c@{}}98.01\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}98.01\\ t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}98.01\\ t=74.405\end{tabular} \\
1NN & \begin{tabular}[c]{@{}c@{}}96.93\\ D\end{tabular} & 96.45* & 97.36* & \textbf{\begin{tabular}[c]{@{}c@{}}97.59*\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}97.57*\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}97.43*\\ t=76.905\end{tabular} \\
20NN & 96.85 & 96.24* & 97.15* & \textbf{\begin{tabular}[c]{@{}c@{}}97.3\\ T=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}97.15\\ T=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}96.86\\ T=68.905\end{tabular} \\
RD & 97.34 & 79.1* & 95.45* & \begin{tabular}[c]{@{}c@{}}97.3\\ t=0\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}97.36\\ t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}97.34\\ t=68.905\end{tabular} \\
K* & 97.36 & 92.94* & 96.53* & \textbf{\begin{tabular}[c]{@{}c@{}}97.46*\\ t=0.05\end{tabular}} & \begin{tabular}[c]{@{}c@{}}97.46*\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}97.36\\ t=74.405\end{tabular} \\
MLP & 98.16 & 96.45* & 97.35* & \textbf{\begin{tabular}[c]{@{}c@{}}98.22*\\ t=0.05\end{tabular}} & \begin{tabular}[c]{@{}c@{}}98.22*\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}98.16\\ t=74.405\end{tabular} \\ \hline
\end{tabular}
\end{table}

### 2014-11-01 thesis/results/overall-portugal-auc.tex
Update results and discussion section

\begin{table}[htbp]
\label{tab:overall-pt-auc}
\caption{Overall best AUC of each classifier with a particular feature selection method on the general hospital data set. * indicates that results were identical regardless of discretisation. D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied. The absence of * or D means that not discretising produced a better result. The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
\begin{tabular}{|l|cccccc|}
\hline
Classifier & None & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
ZeroR & 0.5* & 0.5* & 0.5* & 0.5* & 0.5* & 0.5* \\
1R & 0.858* & 0.858* & 0.858* & 0.858* & 0.858* & 0.858* \\
NB & \begin{tabular}[c]{@{}c@{}}0.979\\ D\end{tabular} & 0.982* & \textbf{0.983*} & \begin{tabular}[c]{@{}c@{}}0.981*\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.979*\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.979*\\ t=74.405\end{tabular} \\
DT & \textbf{0.989} & 0.981* & 0.984* & \begin{tabular}[c]{@{}c@{}}0.989\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.989*\\ t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.989*\\ t=74.405\end{tabular} \\
LR & 0.994* & 0.981* & 0.990* & \begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.05\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.994*\\ t=74.405\end{tabular} \\
SVM & 0.994* & 0.971* & 0.981* & \textbf{\begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.994*\\ t=74.405\end{tabular} \\
1NN & \begin{tabular}[c]{@{}c@{}}0.979\\ D\end{tabular} & 0.982* & \textbf{0.991*} & \begin{tabular}[c]{@{}c@{}}0.991*\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.990*\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.989*\\ t=76.905\end{tabular} \\
20NN & 0.992* & 0.981* & 0.991* & \textbf{\begin{tabular}[c]{@{}c@{}}0.992*\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.992*\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.992*\\ t=74.405\end{tabular} \\
RD & 0.993* & 0.794* & 0.986* & \textbf{\begin{tabular}[c]{@{}c@{}}0.993*\\ t=0.05\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.993*\\ t=0.005\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.993*\\ t=74.405\end{tabular} \\
K* & 0.994 & 0.980* & 0.989* & \begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.05\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.994\\ t=74.405\end{tabular} \\
MLP & 0.990* & 0.983* & 0.992* & \textbf{\begin{tabular}[c]{@{}c@{}}0.993*\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.991*\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.990*\\ t=76.905\end{tabular} \\ \hline
\end{tabular}
\end{table}

### 2014-11-01 thesis/results/overall-trauma-acc.tex
Update results and discussion section

@@
 \begin{table}[htbp]
-\caption{Overall accuracy of each classifier with a particular feature selection method on the trauma data set.
-`D' indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.}
+\label{tab:overall-tr-acc}
+\caption{Overall best accuracy of each classifier with a particular feature selection method on the trauma data set.
+* indicates that results were identical regardless of discretisation.
+D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.
+The absence of * or D means that not discretising produced a better result.
+The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
 \resizebox{\linewidth}{!}{%
@@
 Classifier & None & Baseline & Expert & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
-ZeroR & 66.91 & 66.91 & 66.91 & 66.91 & 66.91 & 66.91 & 66.91 & 66.91 \\
-1R & 71.95 & 66.91 & \begin{tabular}[c]{@{}c@{}}70.64\\ D\end{tabular} & \textbf{72.01} & 71.93 & \begin{tabular}[c]{@{}c@{}}71.91\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.95\\ t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.97\\ t=66.905\end{tabular} \\
-NB & \begin{tabular}[c]{@{}c@{}}73.83\\ D\end{tabular} & 74.62 & \begin{tabular}[c]{@{}c@{}}73.19\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.4\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}75.8\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.85\\ D, t=0.125\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.45\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}72.37\\ D, t=69.905\end{tabular} \\
-DT & \begin{tabular}[c]{@{}c@{}}75\\ D\end{tabular} & 74.38 & \begin{tabular}[c]{@{}c@{}}76.39\\ D\end{tabular} & 75.4 & \textbf{\begin{tabular}[c]{@{}c@{}}77.52\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}76.05\\ D, t=0.11\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.17\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.49\\ t=66.905\end{tabular} \\
-LR & \begin{tabular}[c]{@{}c@{}}77.62\\ D\end{tabular} & 75.06 & \begin{tabular}[c]{@{}c@{}}76.45\\ D\end{tabular} & 74.14 & \begin{tabular}[c]{@{}c@{}}76.3\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.7\\ D, t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}77.65\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.4\\ D, t=66.905\end{tabular} \\
-SVM & \begin{tabular}[c]{@{}c@{}}77.39\\ D\end{tabular} & 74.47 & \begin{tabular}[c]{@{}c@{}}75.67\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}72.57\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}75.26\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}77.42\\ D, t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.53\\ D, t=0.005\end{tabular}} & \begin{tabular}[c]{@{}c@{}}73.83\\ D, t=66.905\end{tabular} \\
-1NN & \begin{tabular}[c]{@{}c@{}}71.22\\ D\end{tabular} & 70.28 & \begin{tabular}[c]{@{}c@{}}72.56\\ D\end{tabular} & 74.44 & \textbf{\begin{tabular}[c]{@{}c@{}}77.81\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.75\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.26\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.78\\ t=68.905\end{tabular} \\
-20NN & \begin{tabular}[c]{@{}c@{}}75.9\\ D\end{tabular} & 73.59 & \begin{tabular}[c]{@{}c@{}}75.51\\ D\end{tabular} & 75.3 & \begin{tabular}[c]{@{}c@{}}76.48\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.26\\ D, t=0.125\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.56\\ D, t=0.005\end{tabular}} & \begin{tabular}[c]{@{}c@{}}73.56\\ t=68.905\end{tabular} \\
-RD & \begin{tabular}[c]{@{}c@{}}76.92\\ D\end{tabular} & 69.68 & \begin{tabular}[c]{@{}c@{}}76.19\\ D\end{tabular} & 74.38 & \textbf{\begin{tabular}[c]{@{}c@{}}77.22\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}76.87\\ D, t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.82\\ D, t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.83\\ D, t=66.905\end{tabular} \\
-K* & \begin{tabular}[c]{@{}c@{}}71.4\\ D\end{tabular} & 72.20 & \begin{tabular}[c]{@{}c@{}}75.08\\ D\end{tabular} & 75.58 & \textbf{\begin{tabular}[c]{@{}c@{}}77.27\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}75.07\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.08\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.45\\ t=68.905\end{tabular} \\
-MLP & \begin{tabular}[c]{@{}c@{}}73.79\\ D\end{tabular} & 73.03 & \begin{tabular}[c]{@{}c@{}}73.45\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.38\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.66\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.65\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.03\\ D, t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.94\\ D, t=66.905\end{tabular} \\ \hline
+ZeroR & 66.91* & 66.91* & 66.91* & 66.91* & 66.91* & 66.91* & 66.91* & 66.91* \\
+1R & 71.95 & 66.91* & \begin{tabular}[c]{@{}c@{}}70.64\\ D\end{tabular} & \textbf{72.01} & 71.93 & \begin{tabular}[c]{@{}c@{}}71.91\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.95\\ t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.97\\ t=66.905\end{tabular} \\
+NB & \begin{tabular}[c]{@{}c@{}}73.83\\ D\end{tabular} & 74.62* & \begin{tabular}[c]{@{}c@{}}73.19\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.4\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}75.8\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.85\\ D, t=0.125\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.45\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}72.37\\ D, t=69.905\end{tabular} \\
+DT & \begin{tabular}[c]{@{}c@{}}75\\ D\end{tabular} & 74.38* & \begin{tabular}[c]{@{}c@{}}76.39\\ D\end{tabular} & 75.4 & \textbf{\begin{tabular}[c]{@{}c@{}}77.52\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}76.05\\ D, t=0.11\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.17\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.49\\ t=66.905\end{tabular} \\
+LR & \begin{tabular}[c]{@{}c@{}}77.62\\ D\end{tabular} & 75.06* & \begin{tabular}[c]{@{}c@{}}76.45\\ D\end{tabular} & 74.14 & \begin{tabular}[c]{@{}c@{}}76.3\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.7\\ D, t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}77.65\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.4\\ D, t=66.905\end{tabular} \\
+SVM & \begin{tabular}[c]{@{}c@{}}77.39\\ D\end{tabular} & 74.47* & \begin{tabular}[c]{@{}c@{}}75.67\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}72.57\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}75.26\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}77.42\\ D, t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.53\\ D, t=0.005\end{tabular}} & \begin{tabular}[c]{@{}c@{}}73.83\\ D, t=66.905\end{tabular} \\
+1NN & \begin{tabular}[c]{@{}c@{}}71.22\\ D\end{tabular} & 70.28* & \begin{tabular}[c]{@{}c@{}}72.56\\ D\end{tabular} & 74.44 & \textbf{\begin{tabular}[c]{@{}c@{}}77.81\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.75\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.26\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.78\\ t=68.905\end{tabular} \\
+20NN & \begin{tabular}[c]{@{}c@{}}75.9\\ D\end{tabular} & 73.59* & \begin{tabular}[c]{@{}c@{}}75.51\\ D\end{tabular} & 75.3 & \begin{tabular}[c]{@{}c@{}}76.48\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.26\\ D, t=0.125\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.56\\ D, t=0.005\end{tabular}} & \begin{tabular}[c]{@{}c@{}}73.56\\ t=68.905\end{tabular} \\
+RD & \begin{tabular}[c]{@{}c@{}}76.92\\ D\end{tabular} & 69.68* & \begin{tabular}[c]{@{}c@{}}76.19\\ D\end{tabular} & 74.38 & \textbf{\begin{tabular}[c]{@{}c@{}}77.22\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}76.87\\ D, t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.82\\ D, t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.83\\ D, t=66.905\end{tabular} \\
+K* & \begin{tabular}[c]{@{}c@{}}71.4\\ D\end{tabular} & 72.20* & \begin{tabular}[c]{@{}c@{}}75.08\\ D\end{tabular} & 75.58 & \textbf{\begin{tabular}[c]{@{}c@{}}77.27\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}75.07\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.08\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.45\\ t=68.905\end{tabular} \\
+MLP & \begin{tabular}[c]{@{}c@{}}73.79\\ D\end{tabular} & 73.03* & \begin{tabular}[c]{@{}c@{}}73.45\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.38\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.66\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.65\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.03\\ D, t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.94\\ D, t=66.905\end{tabular} \\ \hline
 \end{tabular}

### 2014-11-01 thesis/results/overall-trauma-auc.tex
Update results and discussion section

@@
 \begin{table}[htbp]
-\caption{Overall AUC performance of each classifier with a particular feature selection method on the trauma data set.
-`D' indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.}
+\label{tab:overall-tr-auc}
+\caption{Overall best AUC of each classifier with a particular feature selection method on the trauma data set.
+* indicates that results were identical regardless of discretisation.
+D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.
+The absence of * or D means that not discretising produced a better result.
+The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
 \resizebox{\textwidth}{!}{%
@@
 Classifier & None & Baseline & Expert & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
-ZeroR & 0.498 & 0.498 & 0.498 & 0.498 & 0.498 & 0.498 & 0.498 & 0.498 \\
-1R & 0.67 & 0.5 & \begin{tabular}[c]{@{}c@{}}0.581\\ D\end{tabular} & 0.668 & 0.668 & \begin{tabular}[c]{@{}c@{}}0.669\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.669\\ t=0.005\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.67\\ t=69.905\end{tabular}} \\
-NB & \begin{tabular}[c]{@{}c@{}}0.828\\ D\end{tabular} & 0.804 & \begin{tabular}[c]{@{}c@{}}0.806\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.815\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.827\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.829\\ D, t=0.125\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.832\\ D, t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.8\\ D, t=66.905\end{tabular} \\
-DT & \begin{tabular}[c]{@{}c@{}}0.756\\ D\end{tabular} & 0.772 & \begin{tabular}[c]{@{}c@{}}0.786\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.793\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.806\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.799\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.766\\ D, t=66.905\end{tabular} \\
-LR & \begin{tabular}[c]{@{}c@{}}0.844\\ D\end{tabular} & 0.812 & \begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.819\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.831\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.846\\ D, t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.844\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.811\\ D, t=66.905\end{tabular} \\
-SVM & \begin{tabular}[c]{@{}c@{}}0.843\\ D\end{tabular} & 0.787 & 0.807 & \begin{tabular}[c]{@{}c@{}}0.782\\ D\end{tabular} & 0.806 & \textbf{\begin{tabular}[c]{@{}c@{}}0.844\\ D, t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.842\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.724\\ D, t=66.905\end{tabular} \\
-1NN & \begin{tabular}[c]{@{}c@{}}0.726\\ D\end{tabular} & 0.719 & \begin{tabular}[c]{@{}c@{}}0.760\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.798\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.816\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.798\\ D, t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.768\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.778\\ t=68.905\end{tabular} \\
-20NN & \begin{tabular}[c]{@{}c@{}}0.827\\ D\end{tabular} & 0.792 & \begin{tabular}[c]{@{}c@{}}0.808\\ D\end{tabular} & 0.814 & \begin{tabular}[c]{@{}c@{}}0.828\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.831\\ D, t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.832\\ D, t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.799\\ D, t=66.905\end{tabular} \\
-RD & \begin{tabular}[c]{@{}c@{}}0.834\\ D\end{tabular} & 0.742 & \begin{tabular}[c]{@{}c@{}}0.807\\ D\end{tabular} & 0.81 & \begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.836\\ D, t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.834\\ D, t=0.005\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.791\\ D, t=66.905\end{tabular} \\
-K* & \begin{tabular}[c]{@{}c@{}}0.768\\ D\end{tabular} & 0.774 & \begin{tabular}[c]{@{}c@{}}0.803\\ D\end{tabular} & 0.815 & \textbf{\begin{tabular}[c]{@{}c@{}}0.838\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.82\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ t=68.905\end{tabular} \\
-MLP & \begin{tabular}[c]{@{}c@{}}0.803\\ D\end{tabular} & 0.781 & \begin{tabular}[c]{@{}c@{}}0.784\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.818\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.803\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.772\\ D, t=68.905\end{tabular} \\ \hline
+ZeroR & 0.498* & 0.498* & 0.498* & 0.498* & 0.498* & 0.498* & 0.498* & 0.498* \\
+1R & 0.67 & 0.5* & 0.581* & 0.668 & 0.668 & \begin{tabular}[c]{@{}c@{}}0.669\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.669\\ t=0.005\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.67\\ t=69.905\end{tabular}} \\
+NB & \begin{tabular}[c]{@{}c@{}}0.828\\ D\end{tabular} & 0.804* & \begin{tabular}[c]{@{}c@{}}0.806\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.815\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.827\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.829\\ D, t=0.125\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.832\\ D, t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.8\\ D, t=66.905\end{tabular} \\
+DT & \begin{tabular}[c]{@{}c@{}}0.756\\ D\end{tabular} & 0.772* & \begin{tabular}[c]{@{}c@{}}0.786\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.793\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.806\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.799\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.766\\ D, t=66.905\end{tabular} \\
+LR & \begin{tabular}[c]{@{}c@{}}0.844\\ D\end{tabular} & 0.812* & \begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.819\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.831\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.846\\ D, t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.844\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.811\\ D, t=66.905\end{tabular} \\
+SVM & \begin{tabular}[c]{@{}c@{}}0.843\\ D\end{tabular} & 0.787* & 0.807 & \begin{tabular}[c]{@{}c@{}}0.782\\ D\end{tabular} & 0.806 & \textbf{\begin{tabular}[c]{@{}c@{}}0.844\\ D, t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.842\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.724\\ D, t=66.905\end{tabular} \\
+1NN & \begin{tabular}[c]{@{}c@{}}0.726\\ D\end{tabular} & 0.719* & \begin{tabular}[c]{@{}c@{}}0.760\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.798\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.816\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.798\\ D, t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.768\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.778\\ t=68.905\end{tabular} \\
+20NN & \begin{tabular}[c]{@{}c@{}}0.827\\ D\end{tabular} & 0.792* & \begin{tabular}[c]{@{}c@{}}0.808\\ D\end{tabular} & 0.814 & \begin{tabular}[c]{@{}c@{}}0.828\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.831\\ D, t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.832\\ D, t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.799\\ D, t=66.905\end{tabular} \\
+RD & \begin{tabular}[c]{@{}c@{}}0.834\\ D\end{tabular} & 0.742* & \begin{tabular}[c]{@{}c@{}}0.807\\ D\end{tabular} & 0.81 & \begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.836\\ D, t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.834\\ D, t=0.005\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.791\\ D, t=66.905\end{tabular} \\
+K* & \begin{tabular}[c]{@{}c@{}}0.768\\ D\end{tabular} & 0.774* & \begin{tabular}[c]{@{}c@{}}0.803\\ D\end{tabular} & 0.815 & \textbf{\begin{tabular}[c]{@{}c@{}}0.838\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.82\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ t=68.905\end{tabular} \\
+MLP & \begin{tabular}[c]{@{}c@{}}0.803\\ D\end{tabular} & 0.781* & \begin{tabular}[c]{@{}c@{}}0.784\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.818\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.803\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.772\\ D, t=68.905\end{tabular} \\ \hline
 \end{tabular}

### 2014-11-01 thesis/results/results.tex
Add overall results for trauma data set
Update results and discussion section

@@
 
-The results go here.
+\section{Overall Performance}
+\subsection{Trauma Data Set}
+Tables \ref{tab:overall-tr-acc} and \ref{tab:overall-tr-auc} show the best
+accuracy and AUC results for each of the feature selection methods for each
+classifier. The best accuracy was 77.81\%, achieved using the $k$-NN algorithm
+with 1 nearest neighbour, a discretised data set, and feature selection using a
+C4.5 wrapper. The best AUC of 0.846 was achieved using logistic regression with
+features selected by a correlation coefficient threshold of 0.1, and again
+using the discretised data set.
+
+\input{results/overall-trauma-acc.tex}
+\input{results/overall-trauma-auc.tex}
+
+Recall that in order for our work to be comparable to that of previous work on
+trauma LOS prediction by Dinh et al. \cite{Dinh2013a}, we use the same data set
+that they used, and also tested the features they used as one of the manual
+feature selection methods. From Table \ref{tab:overall-tr-acc} we can see that
+the best accuracy achieved using their features is 75.06\% with logistic
+regression. It is worth noting that our best accuracy, using $k$-NN with 1
+nearest neighbour, improves upon the accuracy that they achieved by 2.75\%
+while using only 11 features (compared to the 19 they used). These 11 features
+make up only 14.1\% of all features in the data set, and performed better than
+the 11 features selected manually by the domain expert.
+
+Despite our inclusion of accuracy as an evaluation metric, the AUC is what Dinh
+et al. used to evaluate the ability of their logistic regression classifier to
+discriminate between the two LOS classes. Using their features, the best AUC
+was 0.812 with logistic regression, and our best result was 0.846, also with
+logistic regression. However, this was using the discretised data set with 29
+features selected by correlation coefficient at a threshold of 0.1.
+Although our logistic regression classifier exhibited an AUC improvement of
+0.034, it required roughly 50\% more features than the baseline of 19 features.
+
+Although $k$-NN with 1 nearest neighbour had the highest accuracy (77.81\%) out
+of all combinations of classifiers and feature selection methods for this data
+set, we should mention that logistic regression came a very close second with
+77.7\%, outperforming more sophisticated approaches such as SVM and MLP. It
+also achieved a higher AUC, and all while being faster to train.
+
+We would also like to highlight the importance of discretisation on improving
+performance: in Tables \ref{tab:overall-tr-acc} and
+\ref{tab:overall-tr-auc} the majority of the quoted results are from applying
+the learning algorithm and feature selection method on the discretised data.
+This indicates that discretisation produced a better result than applying the
+same classifier and feature selector on the non-discretised data.
+
+\subsection{General Data Set}
+For the general hospital data set, we report the best accuracy and AUC results
+for each combination of classifier and feature selection method in Tables
+\ref{tab:overall-pt-acc} and \ref{tab:overall-pt-auc}. The most accurate
+classifier was the C4.5 decision tree using all 14 features.
+This result was the same
+regardless of whether or not discretisation was first applied to the data set.
+The best AUC result was achieved by three classifiers: logistic regression, SVM
+and K*, with an AUC of 0.994. Logistic regression and K* achieved this AUC with
+an information gain feature selector with threshold 0.05, and SVM with a
+correlation coefficient selector at a 0.1 threshold. We consider logistic
+regression and K* to be superior to the SVM in this case because they achieved
+the same AUC with only 8 features, whereas the SVM required 10.
+
+\input{results/overall-portugal-acc.tex}
+\input{results/overall-portugal-auc.tex}
+
+Due to the general nature of this data set, and because we have defined a
+classification problem rather than a regression one, it is difficult to find a
+valid baseline for comparison. However, we can still make some remarks about
+how this data set performed differently with the same methods that were used
+for the trauma LOS data, which will give us valuable insight into the
+suitability of classifiers and learning algorithms for predicting the LOS for
+various medical domains. In addition, in the following sections it will be
+valuable to discuss and compare the important predictive features in this
+general data set with features that other researchers have found to be
+useful in predicting length of hospital stay.
+
+We note that, unlike for the trauma data set, discretisation does not improve
+the accuracy or the AUC of almost all of the classifier and feature selection
+combinations. This is likely due to the presence of only one numeric feature
+(\texttt{prev\_adm}) which, when discretised, was transformed into a nominal
+feature with only one value. Classifiers using the discretised
+\texttt{prev\_adm} were likely to perform worse because the single value does
+not help predict the class (since it will be the same value for both classes)
+and is simply noise. Classifiers which used a reduced set of features that
+excluded \texttt{prev\_adm} would have the same performance regardless of
+discretisation because all other features were nominal and would not have been
+affected by discretisation.
+
+\section{Comparison of Feature Selection Methods}
+\subsection{Trauma Data Set}
+Tables \ref{tab:features-tr-acc} and \ref{tab:features-tr-auc} show the
+best accuracy and AUC achieved with each classifier and feature method, with
+both discretised and non-discretised results. We did not include the results of
+Zero-R as they do not vary with discretisation or feature selection.
+The classifier that achieved the
+highest accuracy and AUC over most of the feature selection methods was
+logistic regression, with accuracies ranging from 74.4\% with 1R feature
+selection to 77.7\% with correlation coefficient feature selection. The AUC
+performance of logistic regression ranged from 0.811 to 0.846, again with
+1R and correlation feature selection respectively.
+\input{results/features-tr-acc.tex}
+\input{results/features-tr-auc.tex}
+
+As indicated by italics in the tables, the best accuracy and AUC results for
+each classifier resulted from a reduced feature set obtained through a
+particular feature selection method. There was at least one feature selector
+for each classifier that managed to outperform the same classifier with
+no feature selection. This is likely due to the large number of features in
+the original data -- after cleaning and preliminary removal of redundant
+features, we were still left with over 50. Many of these would still likely
+be irrelevant to predicting the class.
+Out of 20 combinations of classifiers
+and discretisation (the 10 classifiers with and 10 without
+discretisation), the C4.5 wrapper method resulted in the best
+accuracy for 7 of these combinations, followed by 5 using information gain
+and 4 using the correlation coefficient and very few for the other methods.
+Under no combination was 1R the best feature selection method to use, as other
+methods produced better accuracy.
+
+Focussing our attention on the AUC tells a slightly different story. Here we
+find that the correlation coefficient and the information gain are the feature
+selectors that give the best AUC for most of the classifiers, and perform
+equally well when used in conjunction with logistic regression and SVMs.
+Interestingly, although the C4.5
+wrapper method selects features which allow classifiers to achieve good
+accuracy, only in 2 out of the 20 combinations of classifiers did it give the
+best AUC result out of all the feature selectors.
+
+Note that regardless of whether we consider accuracy or AUC to be more
+important in evaluating how our classifiers performed, neither the baseline
+feature set from Dinh et al. \cite{Dinh2013a} nor the features suggested by
+the domain expert (both of which were manually selected) yielded better
+results than automatic feature selection in all but two cases. We did find
+that expert selection produced the best accuracy result for 1R (equal with
+correlation coefficient selection, but used less features) and that the
+baseline feature set produced the best accuracy for Na\"{i}ve Bayes. Neither
+feature selection method produced the best AUC result for any classifier.
+This is a surprising result because intuitively and in the literature it
+has been stated that in a highly specialised field such as medicine, the domain
+experts would be best-equipped to make judgements about which features to
+use \cite{Witten2005}, and these should give reasonable results. However,
+our results show that in attempting to classify patients according to their
+LOS, we should consider both manual and automatic approaches.
+
+Recall that $k$-NN with 1 nearest neighbour and C4.5 wrapper feature
+selection produced the best accuracy (77.81\%) for the trauma data set, and
+logistic regression with correlation coefficient feature selection at a
+threshold of 0.1 gave the best AUC (0.846). These two feature selection
+methods selected 11 and 29 features respectively, and the features that
+were chosen in common between them were: \texttt{operation}, \texttt{bp},
+\texttt{iss}, \texttt{lowerlimbnopelvis}, \texttt{age}, \texttt{lowerlimbany}
+and \texttt{icu}. This adds evidence to the work of Dinh et al. who found
+that \texttt{iss}, \texttt{age} and \texttt{operation} were important
+distinguishers of LOS $\leq$ or $>$ 2 days \cite{Dinh2013a}. We also
+confirm the findings of Gabbe et al. who found that \texttt{age},
+\texttt{iss} and \texttt{bp} were useful in predicting the length of
+hospital stay of blunt trauma patients \cite{Gabbe2005}.
+
+Although we were able to achieve higher accuracy and AUC than the baseline,
+our best approaches also have a number of limitations which need to be pointed
+out. The first point to note is that using a wrapper method of feature
+selection is slower than using a filter approach (or manually picking the
+features before training), as the learning algorithm used as the wrapper
+needs to be invoked at each of the folds of cross-validation. Additionally,
+the $k$-NN algorithm becomes slower as the data set increases in the number
+of examples. We did not face such a problem during our work, but this will
+have to be addressed for larger data sets. It is also common to use more
+than one nearest neighbour, so care needs to be taken to find a suitable number
+of neighbours for a given problem. Finally, we achieved an improvement
+in AUC of 0.034 but this was with 29 features, which is still a fairly large
+number. The importance of using a fewer number of features is twofold: not
+only does this speed up the training time, it also means that we are able to
+make predictions without requiring a lot of information about a patient,
+which is a crucial consideration if this is to be implemented for use by
+physicians.
+
+\subsection{General Data Set}
+Tables \ref{tab:features-pt-acc} and \ref{tab:features-pt-auc} compare the
+best accuracy and AUC of each classifier between all the feature selectors
+used on the general hospital data set. Again, Zero-R is not included as it
+does not vary with discretisation or feature selection. For accuracy, the
+best classifier across all feature selectors except CFS was C4.5, and even
+then it
+was only 0.01\% less accurate than logistic regression and the SVM. Both
+logistic regression and SVM achieved 0.994 AUC using no feature selection
+as well as correlation, information gain and 1R at the same thresholds. Whereas
+in the trauma data set, the MLP did not perform so well compared to many of the
+methods, here it produced the best AUC with CFS and C4.5 wrapper feature
+selection, resulting in 0.983 and 0.992 AUC respectively.
+\input{results/features-pt-acc.tex}
+\input{results/features-pt-auc.tex}
+
+While in the trauma data set we were able to consistently improve the accuracy
+and AUC of each classifier using at least one feature selection method,
+this data set seems to exhibit somewhat different
+behaviour. There are many more situations where the use of feature selection
+makes little improvement to the performance (accuracy or AUC) of a classifier,
+such as for the C4.5 decision tree, logistic regression and SVM if we consider
+accuracy, and logistic regression, $k$-NN, Ranked Distance and K* if we
+consider AUC. With all of these classifiers, we found that using CFS and the
+C4.5 wrapper method \textit{decreased} their accuracy and AUC. The other three
+methods (correlation coefficient, information gain and 1R) resulted in
+performance that was only just as good as without feature selection, and in
+some cases worse. Often, the feature selection that resulted in an accuracy
+or AUC that was as good as the one without feature selection was achieved at
+a threshold which did not discard any features, which would plainly give us
+identical results.
+
+There are a number of results which are notable, however. Logistic regression
+and K*
+were still able to achieve a mean AUC of 0.994 with features selected using
+information gain at a 0.05 threshold. This meant that it only used 8 out of
+the original 14 features, cutting down the number of features by nearly half.
+We mentioned that the MLP had the highest discriminating ability when used with
+CFS and C4.5 wrapper selection, with AUC 0.983 and 0.992 respectively. While
+this is not as high as the best result of 0.994, it is worth pointing out that
+CFS only selected 2 features and C4.5 wrapper, 4 features. This means that
+using only 2 features, we were still able to discriminate between the two LOS
+classes almost as well as if we used all of the features. Moreover, using
+only 6 out of 14 features and $k$-NN with 1 or 20 nearest neighbours, we can
+still achieve an accuracy that is less than 1\% worse than the best result
+of 98.23\%.
+
+These results are perhaps not surprising when we consider the characteristics
+of the features. The 1R classifier was able to achieve 92.88\% accuracy and
+0.858 AUC after ten-fold cross-validation with only a single feature. The best
+performing feature for the trauma data set using 1R only resulted in 72.01\%
+accuracy and 0.67 AUC at best. Using 1R at the 74.405\% accuracy threshold,
+we did not discard any features from the general hospital data set, but all
+features were discarded from the trauma data. This explains why feature
+selection was not effective at improving performance for the general hospital
+setting, but also why removing features did not drastically reduce the
+performance of classifiers -- each feature was individually predictive of the
+class, so removing certain features decreased accuracy or AUC but only by a
+small amount, since the other features were still able to predict the class
+to a high degree of accuracy. This is in contrast to the trauma data, where
+the removal of features improved performance and where most of the features
+had a predictive accuracy (indicated by the 1R feature selector) of
+\textit{less than} the Zero-R classifier.
+
+Out of all four automatic feature selection methods, the single feature that
+was deemed to be the most important in predicting LOS was \texttt{EpisodeType}.
+Although this was a general hospital-wide data set, the importance of this
+attribute indicates that condition-specific factors should be taken into
+account when predicting the LOS for any medical domain. Additionally, the
+\texttt{InpatientService} feature was also considered important in predicting
+the class, a finding that was also noted by \cite{Lucas2009}, although they
+considered the classification of LOS $\leq$ and > 3 days.
+
+\section{Comparison of Classifiers}
+\subsection{Trauma Data Set}
+
+\subsection{General Data Set}
+
+\section{Effect of Discretisation}
+
+\section{Summary}

### 2014-11-02 thesis/results/results.tex
Add code for all trauma plots

@@
 \texttt{InpatientService} feature was also considered important in predicting
-the class, a finding that was also noted by \cite{Lucas2009}, although they
-considered the classification of LOS $\leq$ and > 3 days.
+the class, a finding that was also noted by Lucas et al. \cite{Lucas2009},
+although they considered the classification of LOS $\leq$ and $>$ 3 days.
 
 \section{Comparison of Classifiers}
+\begin{figure}[htbp]
+\begin{subfigure}{.48\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-nofs-acc.eps}
+\caption{1a}
+\label{}
+\end{subfigure}%
+\begin{subfigure}{.55\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-nofs-auc.eps}
+\caption{1b}
+\label{}
+\end{subfigure}
+
+\begin{subfigure}{.48\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-dinh-acc.eps}
+\caption{1a}
+\label{}
+\end{subfigure}%
+\begin{subfigure}{.55\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-dinh-auc.eps}
+\caption{1b}
+\label{}
+\end{subfigure}
+
+\begin{subfigure}{.48\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-expert-acc.eps}
+\caption{1a}
+\label{}
+\end{subfigure}%
+\begin{subfigure}{.55\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-expert-auc.eps}
+\caption{1b}
+\label{}
+\end{subfigure}
+
+\begin{subfigure}{.48\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-cfs-acc.eps}
+\caption{1a}
+\label{}
+\end{subfigure}%
+\begin{subfigure}{.55\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-cfs-auc.eps}
+\caption{1b}
+\label{}
+\end{subfigure}
+\caption{plots of}
+\label{}
+\end{figure}
+
+\begin{figure}
+\ContinuedFloat
+
+\begin{subfigure}{.48\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-wrapper-acc.eps}
+\caption{1a}
+\label{}
+\end{subfigure}%
+\begin{subfigure}{.55\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-wrapper-auc.eps}
+\caption{1b}
+\label{}
+\end{subfigure}
+\caption{plots of}
+\label{}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{subfigure}{\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-corr.eps}
+\caption{}
+\label{}
+\end{subfigure}
+\caption{}
+\label{}
+\end{figure}
+
+\begin{figure}[htbp]
+\ContinuedFloat
+\begin{subfigure}{\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-ig.eps}
+\caption{}
+\label{}
+\end{subfigure}
+\end{figure}
+
+\begin{figure}[htbp]
+\ContinuedFloat
+\begin{subfigure}{\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-oner.eps}
+\caption{}
+\label{}
+\end{subfigure}
+\end{figure}
+
 \subsection{Trauma Data Set}

### 2014-11-02 thesis/litreview/litreview.tex
Fix broken table and figure labels in latex

@@
-\chapter{Predicting the LOS: A Review} \label{chap:litreview}
+\chapter{Predicting the Length of Stay: A Review} \label{chap:litreview}
 
@@
 \begin{figure}[h]
-\caption{}
+\caption{A graphical representation of a neural network with one hidden layer,
+used by Tu \& Guerriere in predicting LOS.}
 \centering
@@
 can be found where the margin between two different records is as large as
-possible. In figure \ref{svm} from Yoo et al. \citep{Yoo2012}, hyperplane 2
-is the better plane because its margin is as large as possible. 
+possible. In Figure \ref{fig:svm-example} from Yoo et al. \citep{Yoo2012},
+hyperplane 2 is the better plane because its margin is as large as possible. 
 \begin{figure}[h]
-\label{svm}
-\caption{}
+\caption{An example of finding the largest margin hyperplane for a two-class
+support vector machine classifier.}
+\label{fig:svm-example}
 \centering

### 2014-11-02 thesis/method/classification.tex
Fix broken table and figure labels in latex

@@
-\chapter{LOS Classification Using Supervised Learning}
+\chapter{Length of Stay Classification Using Supervised Learning}
   \label{chap:classification}
@@
 
-\section{0R and 1R}
+\section{Zero-R and 1R}
 \label{sec:zeror}
-Short for ``0-rule'' and ``1-rule'' respectively, 0R and 1R are
-very simple classifiers. 0R uses no features to predict the class, and
+Short for ``0-rule'' and ``1-rule'' respectively, Zero-R and 1R are
+very simple classifiers. Zero-R uses no features to predict the class, and
 1R only uses one.
 
-Given a data set, the 0R classification algorithm will find the class that
+Given a data set, the Zero-R classification algorithm will find the class that
 occurs the most often, and predict this class for all new examples. For LOS
@@
 ``LOS $>$ 2 days'', and feature vectors (which represent patients in this
-scenario) were more commonly labelled ``LOS $\leq$ 2 days'', then 0R will
+scenario) were more commonly labelled ``LOS $\leq$ 2 days'', then Zero-R will
 predict all unseen patients with ``LOS $\leq$ 2 days''.
-The prediction decision of 0R is very easy to understand,
+The prediction decision of Zero-R is very easy to understand,
 and its results also provide an apt baseline for evaluating the more
@@
 for this classifier,
-and if there is a tie, it is broken randomly. As with 0R, the prediction
+and if there is a tie, it is broken randomly. As with Zero-R, the prediction
 decision of 1R is transparent and it provides another baseline upon which
@@
 noting down the error rate. 1R then chooses the feature that produces rules
-with the lowest total error rate.
+with the highest total accuracy.
 
-\section{Na\"{i}ve Bayes (NB)}
+\section{Na\"{i}ve Bayes}
 Unlike 1R, which considers only one feature in the feature vector, Na\"{i}ve
@@
 in our work both as a simple classifier in its own right, and also as a more
-sophisticated baseline to 0R and 1R.
+sophisticated baseline to Zero-R and 1R.
 
-\section{C4.5 Decision Tree (DT)}
+\section{C4.5 Decision Tree}
 \label{sec:c45}
@@
 
-\section{Logistic Regression (LR)}
-Logistic regression is a regression technique used for when the dependent
+\section{Logistic Regression}
+Logistic regression (LR) is a regression technique used for when the dependent
 variable is dichotomous or two-valued, and thus it is a suitable candidate to
@@
 Training the logistic regression classifier, then, is a matter of finding the
-$\beta_i$s that will maximise the likelihood (conditional probability of the
+$\beta_i$ that will maximise the likelihood (conditional probability of the
 data given the values for $\beta_i$) of the given data set. This is done using
-numerical methods, where estimates of the $\beta_i$s are chosen, the likelihood
+numerical methods, where estimates of the $\beta_i$ are chosen, the likelihood
 computed, and the $\beta_i$ are updated, and so on: this process is repeated
@@
 
-To classify a feature vector $x = (x_1,x_2,\ldots,x_n)$, we substitute the
-$x_i$'s into the above equation in place of the $F_i$'s, which gives us a
+To classify a feature vector $\mathbf{x} = (x_1,x_2,\ldots,x_n)$, we substitute the
+$x_i$ into the above equation in place of the $F_i$, which gives us a
 probability. If this probability is above a particular threshold, then we
@@
 
-\section{Support Vector Machine (SVM)}
+\section{Support Vector Machine}
 Support vector machines (SVMs) are an example of a supervised learning algorithm
@@
 \begin{figure}[h]
+\caption{An example of a 2-dimensional SVM.}
 \label{fig:svm}
-\caption{}
 \centering
@@
 
-\section{$k$-Nearest Neighbour ($k$-NN)}
+\section{$k$-Nearest Neighbour}
 In this section we will describe the $k$-nearest neighbour ($k$-NN) algorithm
@@
 
-\subsection{Extensions}
+\subsection{Existing and Proposed Extensions}
 There have been many extensions proposed to the $k$-NN algorithm described
@@
 $k$-NN algorithm. We will focus on two extensions to the Euclidean distance:
-K* and our contribution, Ranked Distance.
+K* and our proposed distance function, Ranked Distance.
 
@@
 
-\subsubsection{Ranked Distance}
+\subsubsection{Proposed Approach: Ranked Distance}
 \label{sec:rankeddist}
@@
 \paragraph{Mathematical formulation}
-Given two instances $x = (x_1,x_2,\ldots,x_n)$ and $y = (y_1,y_2,\ldots,y_n)$,
-the distance between them is given by:
+Given two instances $\mathbf{x} = (x_1,x_2,\ldots,x_n)$ and
+$\mathbf{y} = (y_1,y_2,\ldots,y_n)$, the distance between them is given by:
 \begin{equation*}
@@
 
-\noindent We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}}$ and
-$f(\mathrm{Rank}) = n-\mathrm{Rank}$.
+\noindent We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}^c}$, where
+$c$ varies from 0 to 1.
 \end{enumerate}
 
-\section{Multi-layer Perceptron (MLP)}
+\section{Multi-layer Perceptron}
 Multi-layer perceptrons (MLPs) are a generalised form of the
@@
 \begin{figure}[h]
+\caption{An MLP with one hidden layer.}
 \label{fig:mlp}
-\caption{}
 \centering
@@
 be evaluating with respect to our classification problem, from the simple
-0R and 1R to Na\"{i}ve Bayes, decision trees, logistic regression,
+Zero-R and 1R to Na\"{i}ve Bayes, decision trees, logistic regression,
 $k$-nearest neighbour (and extensions K* and Ranked Distance), to the more

### 2014-11-02 thesis/method/evaluation.tex
Fix broken table and figure labels in latex

@@
 \subsubsection{Repository}
-All code relating to this work, including the \LaTeX code for this document,
+All code relating to this work, including the \LaTeX\ code for this document,
 as well as results from our evaluations are available publicly online at
@@
 two measures we will use are the \textit{accuracy} and the \textit{area under
-the receiver operating characteristic (ROC) curve}, commonly abbreviated AUC
-(for \textit{area under the curve}).
+the receiver operating characteristic (ROC) curve}, commonly abbreviated AUC.
 
@@
 The accuracy can also be expressed more simply and intuitively as
-$\dfrac{\text{number of correct classifications}}{\text{total number of examples}}$
+the number of correct classifications out of the total number of examples,
 but the terms introduced above will be necessary to understand the
@@
 
-\subsection{Area Under the Curve (AUC)}
+\subsection{Area Under the Curve}
 The area under the receiver operating characteristic (ROC) curve, often
@@
 \begin{figure}[h]
+\caption{Two ROC curves on the same graph. Note that the straight line from
+the bottom left corner to the top right corner indicates an AUC of 0.5. Curves
+which head towards the top left corner indicate increasing AUC.}
 \label{fig:sampleroc}
-\caption{}
 \centering
@@
 \subsection{Confidence Intervals}
-In order to compare our results with that of Dinh et. al. \citep{Dinh2013a},
-we are especially interested in the mean cross-validated AUC of our
-classifiers as well as
-the upper and lower bounds of the 95\% confidence interval associated with this
-mean. The intervals are constructed using this argument:
+In order to compare our results with our baseline, the work of Dinh et. al.
+\citep{Dinh2013a}, as well as quantify the level of confidence we have in
+our performance measures, we constructed 95\% confidence intervals for the
+mean accuracy and AUC.
 
-For each classifier, let the population of all of its ten-fold cross-validated
-AUC scores be some random variable $X$, which follows a distribution whose
+For each classifier, let the population of all of its accuracy (or AUC)
+scores be some random variable $X$, which follows a distribution whose
 parameters (such as mean and standard deviation) we do not know.
-The Central Limit Theorem states that given a random
+The Central Limit Theorem from statistics states that given a random
 sample of $n$ items from $X$, namely $X_1,X_2,\dots,X_n$, the sample mean
@@
 is approximately normally distributed when $n$ is at least roughly around 30
-and the $X_i$'s are independent and identically distributed.
-Consequently, we can construct a $(1-\alpha)$\% confidence interval for
+and the $X_i$'s are independent and identically distributed \cite{Phipps2001}.
+
+Consequently, we can construct a $(1-\alpha)$\% normal confidence interval for
 $\bar{x}$ -- where $\alpha$ is commonly called the
-\textit{significance level} -- as follows:
+\textit{significance level} -- as follows \cite{Wasserman2003}:
 \begin{equation*}
@@
 of the standard normal variable $Z \sim \mathcal{N}(0,1)$ with cumulative
-probability $1-\alpha/2$. Since we have taken 40 independent samples of the
-ten-fold cross-validated AUC, the Central Limit Theorem holds and we can
-compute the mean and standard deviation of our 40 samples in order to
-construct a 95\% confidence interval for the mean AUC of each classifier.
+probability $1-\alpha/2$.
+
+These confidence intervals can be constructed
+
+In our work, we repeated the ten-fold stratified cross-validation procedure
+ten times, leading to $n=100$ samples with which to obtain our confidence
+estimates.
 

### 2014-11-02 thesis/method/selection.tex
Fix broken table and figure labels in latex

@@
 
-\subsection{Correlation-based Feature Selection (CFS)}
+\subsection{Correlation-based Feature Selection}
 Correlation-based feature selection (CFS) is a method developed by Hall
@@
 
-\subsection{Pearson Correlation Coefficient (PCC)}
+\subsection{Pearson Correlation Coefficient}
 The Pearson correlation coefficient measures the degree of linear correlation
@@
 
-\subsection{Information Gain (IG)}
+\subsection{Information Gain}
 Recall from our discussion of decision trees in in Section \ref{sec:c45} that
@@
 \subsection{Wrapper-based Feature Selection}
-The feature selection methods described above (CFS, Pearson correlataion and
+The feature selection methods described above (CFS, Pearson correlation and
 1R) have been described as \textit{filter} approaches to attribute selection.
@@
 \begin{figure}[h]
+\caption{Filter and wrapper approaches to feature selection. Note the role of
+the learning algorithm in the wrapper method.}
 \label{fig:wrapper}
-\caption{}
 \centering

### 2014-11-02 thesis/results/pt-plots.tex
Update results chapter

\begin{figure}[htbp]
\begin{subfigure}{.48\textwidth}
\includegraphics[width=\textwidth]{images/results/pt-nofs-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.55\textwidth}
\includegraphics[width=\textwidth]{images/results/pt-nofs-auc.eps}
\caption{}
\label{}
\end{subfigure}

\begin{subfigure}{.48\textwidth}
\includegraphics[width=\textwidth]{images/results/pt-cfs-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.55\textwidth}
\includegraphics[width=\textwidth]{images/results/pt-cfs-auc.eps}
\caption{}
\label{}
\end{subfigure}

\begin{subfigure}{.48\textwidth}
\includegraphics[width=\textwidth]{images/results/pt-wrapper-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.55\textwidth}
\includegraphics[width=\textwidth]{images/results/pt-wrapper-auc.eps}
\caption{}
\label{}
\end{subfigure}
\caption{Comparison of accuracy and AUC between classifiers for each non-threshold feature selection method, grouped by whether or not discretisation was applied.}
\label{fig:pt-nothreshold}
\end{figure}

\begin{figure}[htbp]
\begin{subfigure}{\textwidth}
\includegraphics[width=\textwidth]{images/results/pt-corr.eps}
\caption{}
\label{fig:pt-threshold-corr}
\end{subfigure}
\ContinuedFloat
\begin{subfigure}{\textwidth}
\includegraphics[width=\textwidth]{images/results/pt-ig.eps}
\caption{}
\label{}
\end{subfigure}
\caption{Comparison of accuracy and AUC between classifiers for feature selectors that discard features based on a threshold. Solid lines correspond to a result without discretisation, and dashed lines correspond to a result with discretisation.}
\label{fig:pt-threshold}
\end{figure}

\begin{figure}[htbp]
\ContinuedFloat
\begin{subfigure}{\textwidth}
\includegraphics[width=\textwidth]{images/results/pt-oner.eps}
\caption{}
\label{}
\end{subfigure}
\caption{Comparison of accuracy and AUC between classifiers for feature selectors that discard features based on a threshold. Solid lines correspond to a result without discretisation, and dashed lines correspond to a result with discretisation.}
\end{figure}

### 2014-11-02 thesis/results/results.tex
Update results chapter

--- 2014-11-01 thesis/results/results.tex
@@
 0.034, it required roughly 50\% more features than the baseline of 19 features.
+If we are willing to sacrifice a little discriminating power, we are able to
+achieve 0.838 AUC with only 11 features using the K* classifier.
 
@@
 77.7\%, outperforming more sophisticated approaches such as SVM and MLP. It
-also achieved a higher AUC, and all while being faster to train.
+also achieved a higher AUC, and has the added advantages of being faster to
+train, mathematically well-understood and widely applied in medicine.
 
@@
 regression and K* to be superior to the SVM in this case because they achieved
-the same AUC with only 8 features, whereas the SVM required 10.
+the same AUC with only 8 features, whereas the SVM required 10. Being able to
+maintain predictive accuracy and the ability to discriminate with less features
+means that the classifier is easier to understand. More importantly, we are
+able to make a decision with less complete information, a characteristic
+which is particularly useful in medicine.
 
@@
 discretisation because all other features were nominal and would not have been
-affected by discretisation.
+affected by discretising.
 
@@
 \texttt{iss} and \texttt{bp} were useful in predicting the length of
-hospital stay of blunt trauma patients \cite{Gabbe2005}.
+hospital stay of blunt trauma patients \cite{Gabbe2005}. Our feature
+selection results also
+support the findings of McGonigal et al. \cite{McGonigal1993} who used
+\texttt{age} and
+\texttt{iss} to predict the probability of survival for trauma patients, a
+related problem in the same medical domain.
+Our observations suggest the use of feature selection not only as a
+pre-processing step to improve the accuracy or AUC of a classifier, but also
+as a tool used separately that could help extract insights from medical data.
+
+It is interesting to note that feature selection improved the performance of
+some classifiers and hindered the performance of others, when compared to the
+same classifier without feature selection. Both 1NN and K* showed improved
+accuracy and AUC as the threshold for correlation, information gain and 1R
+accuracy was increased -- observe the dotted and solid lines corresponding
+to these classifiers in Figure \ref{fig:tr-threshold} sloping upward initially
+-- but this again decreased when a certain threshold was passed. C4.5 also
+benefitted from feature reduction, which reaffirms what we expect. On the other
+hand, correlation, information gain and 1R did not improve the classification
+accuracy or AUC of logistic regression or SVM, and produced mixed results for
+the other classifiers.
 
@@
 These results are perhaps not surprising when we consider the characteristics
-of the features. The 1R classifier was able to achieve 92.88\% accuracy and
-0.858 AUC after ten-fold cross-validation with only a single feature. The best
+of the features. For this data set, the 1R classifier was able to achieve
+92.88\% accuracy and
+0.858 AUC after ten-fold cross-validation with only a single feature. On the
+other hand, the best
 performing feature for the trauma data set using 1R only resulted in 72.01\%
@@
 we did not discard any features from the general hospital data set, but all
-features were discarded from the trauma data. This explains why feature
+features were discarded from the trauma data. This shows that the predictive
+ability of the features in the general hospital data are much higher than
+that of the trauma features. This explains why feature
 selection was not effective at improving performance for the general hospital
-setting, but also why removing features did not drastically reduce the
-performance of classifiers -- each feature was individually predictive of the
-class, so removing certain features decreased accuracy or AUC but only by a
+setting, but also did not drastically reduce performance: each feature
+was individually predictive of the
+class so removing certain features decreased accuracy or AUC but only by a
 small amount, since the other features were still able to predict the class
@@
 \texttt{InpatientService} feature was also considered important in predicting
-the class, a finding that was also noted by \cite{Lucas2009}, although they
-considered the classification of LOS $\leq$ and > 3 days.
+the class, a finding that was also noted by Lucas et al. \cite{Lucas2009},
+although they considered the classification of LOS $\leq$ and $>$ 3 days.
 
@@
 \subsection{Trauma Data Set}
+Figures \ref{fig:tr-nothreshold} and \ref{fig:tr-threshold} compare the
+accuracy and AUC of all classifiers under each feature selection method for
+the trauma data set.
+The methods that do not select features based upon a specified threshold
+(manual methods as well as CFS, C4.5 wrapper and no feature selection)
+are graphed in Figure \ref{fig:tr-nothreshold}, and the methods that select
+based upon thresholds (correlation, information gain and 1R) are graphed
+in Figure \ref{fig:tr-threshold}, showing varying accuracy and AUC as the
+threshold is changed.
+
+\input{results/tr-plots.tex}
+
+In terms of classification accuracy, we did not find a lot of variation
+between all classifiers, with Zero-R producing 66.91\% accuracy and 77.81\%
+being the best we achieved -- an improvement of just over 10\%. The AUC of
+the 11 classifiers we tested ranged from 0.498 with Zero-R to 0.846 with
+logistic regression. Logistic regression, the \textit{de facto} method
+used in LOS classification, performed consistently better than the others
+across all feature
+selection methods, but only when the feature set was not heavily reduced by
+the feature selector: in these cases, such as the results from Figure
+\ref{fig:tr-nothreshold-cfs-acc} showing classifier performance with features
+selected by CFS, C4.5 decision trees and nearest
+neighbour methods have higher accuracy than logistic regression. This is also
+apparent from the graphs in Figure \ref{fig:tr-threshold}: for all three
+feature selection methods, as we increase the threshold, logistic regression
+drops in performance after the initial few thresholds and is outperformed by
+C4.5 and the various nearest neighbour classifiers ($k$-NN with 1 and 20
+neighbours, Ranked Distance and K*). This is true for both accuracy and AUC.
+
+We found that in most cases, the SVM did not achieve a higher accuracy or
+AUC than logistic regression with the same features even though SVMs are
+considered the state-of-the-art in classification algorithms. Additionally,
+as the feature set was reduced, the discriminating ability of SVMs (as
+indicated by the AUC) decreased more than their accuracy: this is particularly
+noticeable in Figure \ref{fig:tr-threshold-oner}, where the AUC of the SVM
+drops sharply to below the AUC of even the 1R classifier at the 67.905\%
+threshold, where there are
+only 4 features. SVMs have not been used in a LOS classification problem like
+ours, so there will need to be further work carried out in order to assess the
+suitability of using SVMs to classify patient LOS.
+
+Although the MLP has been investigated in studies attempting to identify
+factors associated with longer LOS or mortality for trauma patients
+\cite{Hunter2000,McGonigal1993}, leading to improved results with the neural
+network, we did not see the MLP outperform all other classifiers with respect
+to accuracy or AUC for any feature selection method for this data set. In fact,
+the MLP often performed no better than any of the nearest neighbour
+classifiers, which use a straightfoward mechanism of classifying unseen
+examples. Even if the MLP had managed to achieve superior accuracy or AUC, the
+``black box'' nature of its prediction decisions are a barrier for their
+widespread use in medical decision-making.
+
+Nearest neighbour approaches appear to have performed well on this data set in
+both accuracy and AUC, especially in comparison to more sophisticated
+classifiers and when the feature set is significantly reduced. From the graphs
+in Figure \ref{fig:tr-threshold}, we can see that all the nearest neighbours
+approaches perform similarly when the number of features is decreased, but our
+Ranked Distance approach performs better than the other NN algorithms. This
+improvement is not apparent at smaller feature sets, where using the standard
+$k$-NN algorithm with 1 or 20 neighbours performs slightly better.
+The performance of NN we observed in our results is
+an interesting finding because nearest neighbours have not been commonly
+applied in predicting the LOS, but there is intuitive appeal in the
+case-by-case nature of how these algorithms classify new examples that should
+be explored further. Unlike MLPs and SVMs, prediction decision of nearest
+neighbours are more transparent and understandable. However, the major drawback
+of NN algorithms is their storage requirement and the computational time
+required to search for the nearest neighbours to make the classification.
+
+There is a limitation to our findings that needs to be noted: we did not
+attempt to tune the parameters of the classifiers in order to find the ones
+that resulted in the best performance, so our statements are based on the
+results of the default parameters of each classifier.
+MLPs in particular are sensitive to
+the network architecture, so it is likely that its accuracy and AUC could be
+improved if the parameters were selected to optimise its performance on this
+data set.
 
 \subsection{General Data Set}
+In Figures \ref{fig:pt-nothreshold} and \ref{fig:pt-threshold}, we compare the
+performance of the classifiers we used on the general hospital data set. In
+contrast to the trauma data, we were able to achieve very high accuracy and
+AUC on this data set using the exact same classifiers and feature selection
+methods. Instead of only a 10\% improvement from Zero-R by the best classifier
+in the trauma data, we were able to achieve an accuracy of 98.23\%, which is
+an increase of over 20\% from the Zero-R accuracy on this data set of 75.64\%.
+Additionally, we also managed to observe very high ($>0.97$) AUC figures for
+all classifiers except Zero-R and 1R.
+
+\input{results/pt-plots.tex}
+
+Logistic regression and the SVM performed equal best in discriminating ability,
+along with K* with an AUC of 0.994. We consider the performance of K* to be
+superior because it was able to achieve this with 8 features out of 14, and
+additionally it does not require any optimising of its parameters.
+The AUC result of both logistic regression and the SVM were achieved with all
+14 features. Although this is not a large number of features, simplifying the
+inputs to a classifier is beneficial because it reduces the time needed for
+training and classifying new examples. However, as indicated by the graphs in
+Figure \ref{fig:pt-threshold} of AUC against threshold, and the bar charts in
+Figure \ref{fig:pt-nothreshold}, the AUC of all classifiers except Zero-R and
+1R was very close, around 0.98-0.99.
+
+Notably, the MLP performed better than logistic regression on this data set
+than on the trauma data, achieving better AUC with CFS and C4.5 wrapper feature
+selection and better accuracy over all feature selection methods. This is a
+surprising result as CFS and C4.5 wrapper select only 2 and 4 features
+respectively. However, the major drawback of training the MLP with this data
+set was the amount of time needed for one run of ten-fold cross-validation.
+Even with only 14 features, the classifier required several hours to train.
+Although we did not attempt to quantify the exact training time, it was by far
+the slowest learning algorithm.
+This is because the data set contained over 17000 training examples, which have
+to be processed several times by the MLP until all of the weights between its
+connections do not change more than a certain threshold.
+
+Notice that in Figure \ref{fig:pt-threshold-corr}, both the accuracy and AUC
+of Ranked Distance drop much more than for the other classifiers. This is at
+a correlation coefficient threshold of 0.2, which reduces the feature set to
+4 features. We noted in the discussion on classifiers for the trauma data set
+that Ranked Distance performed worse for smaller feature sets than for larger
+ones, and this is another example of this phenomenon. Recall that in Ranked
+Distance, features are first ranked according to some similarity metric (we
+have used the correlation coefficient), and then assigned a weight according
+to their rank. The function we have used assigns rapidly decaying weights to
+the features, which results in greater importance being placed on a few of
+the features with the highest rank. Further work needs to be carried out to
+investigate the cause of the poor performance on smaller feature sets, but
+it is not altogether surprising: Ranked Distance performs best by using the
+weighted distance contributions of all features, and by significantly reducing
+their number we can expect to sharply reduce its ability to differentiate
+between neighbours. There is also scope for investigating other weight
+assignment functions which would work better in reduced-feature situations.
 
 \section{Effect of Discretisation}
+We touched briefly upon the effect of discretisation on the performance of the
+classifiers used, but we would like to point out a few other things not
+mentioned earlier. Firstly,
+although discretisation improved the best accuracy and AUC for many
+classifiers trained on the trauma data, this was only when using a larger
+proportion of features: note the dotted lines in
+Figures \ref{fig:tr-threshold-corr}, \ref{fig:tr-threshold-ig} and
+\ref{fig:tr-threshold-oner} are above their solid line counterparts of the
+same colour on the very left of the graph (where no features have been
+discarded), but as we increase the threshold and more features are discarded,
+the classifier using the discretised data set does not always perform better.
+Notably, discretisation improves the accuracy and AUC of nearest neighbour
+classifiers when few features have been removed, but as the number of features
+decreases, these approaches perform worse with discretised features.
+
+Although the results from discretisation with varying thresholds of the
+correlation, information gain and 1R feature selectors are mixed, the graphs
+in Figure \ref{fig:tr-nothreshold} show improvements in AUC and accuracy for
+most classifiers in most of the situations, and show that discretisation is a
+viable pre-processing technique for data mining problem. This improvement is
+not present in the general hospital data set due to the nature of the data set:
+as mentioned earlier, only one feature is discretised, which transforms it into
+a single value. This single value does not help the classifier distinguish
+between classes, and hence degrades performance.
+
+Discretisation consistently improved the accuracy and AUC of the Na\"{i}ve
+Bayes classifier, which suggests that it is more suitable for the data than the
+classifier's assumption that continuous-valued features follow a normal
+distribution.
+
+Supervised discretisation such as the one we have used is very specific to the
+data set, so it is difficult to draw general conclusions that will apply to
+other learning problems. However, we have not yet seen this technique being
+used in LOS prediction problems, and we believe it is worth investigating
+when approaching a new learning problem.
 
 \section{Summary}
+In this section we have presented and discussed the main findings of our
+work, examining the overall picture, and then detailing comparisons between
+classifiers and feature selection methods in detail before concluding with a
+few remarks on discretisation.
+
+We found that irrespective of discretisation or feature selection, logistic
+regression still performed better than other learning algorithms, which
+reaffirms its use as the \textit{de facto} classifier in predictive medicine.
+However, we were also able to achieve results that were almost as good using
+nearest neighbour approaches combined with feature selection techniques,
+neither of which have been investigated thoroughly in previous work. Our
+proposed Ranked Distance approach was able to improve the accuracy and AUC of
+the standard $k$-NN algorithm with the same number of neighbours, but we found
+that its performance suffered when used in conjunction with feature selection.

### 2014-11-02 thesis/results/tr-plots.tex
Update results chapter

\begin{figure}[htbp]
\begin{subfigure}{.48\textwidth}
\includegraphics[width=\textwidth]{images/results/tr-nofs-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.55\textwidth}
\includegraphics[width=\textwidth]{images/results/tr-nofs-auc.eps}
\caption{}
\label{}
\end{subfigure}

\begin{subfigure}{.48\textwidth}
\includegraphics[width=\textwidth]{images/results/tr-dinh-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.55\textwidth}
\includegraphics[width=\textwidth]{images/results/tr-dinh-auc.eps}
\caption{}
\label{}
\end{subfigure}

\begin{subfigure}{.48\textwidth}
\includegraphics[width=\textwidth]{images/results/tr-expert-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.55\textwidth}
\includegraphics[width=\textwidth]{images/results/tr-expert-auc.eps}
\caption{}
\label{}
\end{subfigure}

\begin{subfigure}{.48\textwidth}
\includegraphics[width=\textwidth]{images/results/tr-cfs-acc.eps}
\caption{}
\label{fig:tr-nothreshold-cfs-acc}
\end{subfigure}%
\begin{subfigure}{.55\textwidth}
\includegraphics[width=\textwidth]{images/results/tr-cfs-auc.eps}
\caption{}
\label{}
\end{subfigure}
\caption{Comparison of accuracy and AUC between classifiers for each non-threshold feature selection method, grouped by whether or not discretisation was applied.}
\label{fig:tr-nothreshold}
\end{figure}

\begin{figure}
\ContinuedFloat

\begin{subfigure}{.48\textwidth}
\includegraphics[width=\textwidth]{images/results/tr-wrapper-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.55\textwidth}
\includegraphics[width=\textwidth]{images/results/tr-wrapper-auc.eps}
\caption{}
\label{}
\end{subfigure}
\caption{Comparison of accuracy and AUC between classifiers for each non-threshold feature selection method, grouped by whether or not discretisation was applied.}
\label{}
\end{figure}

\begin{figure}[htbp]
\begin{subfigure}{\textwidth}
\includegraphics[width=\textwidth]{images/results/tr-corr.eps}
\caption{}
\label{fig:tr-threshold-corr}
\end{subfigure}
\caption{Comparison of accuracy and AUC between classifiers for feature selectors that discard features based on a threshold. Solid lines correspond to a result without discretisation, and dashed lines correspond to a result with discretisation.}
\label{fig:tr-threshold}
\end{figure}

\begin{figure}[htbp]
\ContinuedFloat
\begin{subfigure}{\textwidth}
\includegraphics[width=\textwidth]{images/results/tr-ig.eps}
\caption{}
\label{fig:tr-threshold-ig}
\end{subfigure}

\begin{subfigure}{\textwidth}
\includegraphics[width=\textwidth]{images/results/tr-oner.eps}
\caption{}
\label{fig:tr-threshold-oner}
\end{subfigure}
\caption{Comparison of accuracy and AUC between classifiers for feature selectors that discard features based on a threshold. Solid lines correspond to a result without discretisation, and dashed lines correspond to a result with discretisation.}
\end{figure}

### 2014-11-02 thesis/results/results.tex
Finish results chapter

@@
 that its performance suffered when used in conjunction with feature selection.
+
+In both the trauma and general hospital data sets, we were not able to
+consistently improve the accuracy and AUC of all
+classifiers with feature selection but we were able to sacrifice some
+predictive power in order to use a drastically reduced feature set: for trauma
+we found that K* with 11 features selected by a C4.5 wrapper performed only
+slightly worse in discriminating between the two LOS classes than logistic
+regression using 29 features, and for the hospital-wide data we achieved
+0.983 AUC using only 2 features out of 14 using the MLP, 0.011 less than the
+best result from logistic regression which required all 14 features. It is
+therefore important to consider the trade-off between predictive ability,
+speed and ease of understanding when making judgements about classifiers.
+
+Finally, we pointed out that although discretisation was able to improve
+classifier performance, this effect was less noticeable when more features
+were discarded from feature selection. Discretisation has not been used
+extensively in the LOS prediction literature, so our work will serve as a
+starting point for further investigation.

### 2014-11-02 thesis/thanks/thanks.tex
Update acknowledgements

@@
 Hospital, without whom the research in this thesis would not have been
-possible. Thank you,
-Michael, for providing access to the data we needed for this study, and for
-patiently answering all my questions about the data set and your previous
-experiments.
+possible. Thank you for providing access to the data we needed for this study,
+and for patiently answering all my questions about the data set and your
+previous experiments.
 
-Thank you to my close friends Daniel Buckmaster, Stephen Caraher and Christine
-Au Yeung for the selfless encouragement, help and sanity checks that have
-made this year-long journey possible.
+Thank you also to Professor Paulo Cortez from the University of Minho in
+Portugal for providing an extensive, independent data set from all the way
+across the world.
 
-And finally, but no less significant -- to my parents Caixiu and Ning, and
-my sister Lily -- thank you for tirelessly supporting me throughout this year
-and in everything that I do.
+Thank you to all of my friends, in particular Daniel Buckmaster, Stephen
+Caraher and Christine Au Yeung, for the selfless encouragement, help and good
+times which have made made this year-long journey possible.
+
+And, last but not least -- to my parents Caixiu and Ning, and
+my sister Lily -- thank you for tirelessly supporting me throughout this year,
+especially towards the end. This would not have been possible without all of
+you.

### 2014-11-02 thesis/conclusion/conclusion.tex
Update structure of conclusions
Add conclusions and directions for future work

@@
-\chapter{Conclusions} \label{chap:conclusion}
+\chapter{Conclusions and Future Work} \label{chap:conclusion}
 
-The conclusion goes here.
+In this thesis we conducted a systematic and empirical evaluation of a number
+of classification algorithms and feature selection methods, as well as the
+effect of discretisation, for predicting the length of hospital stay of trauma
+patients. In particular, we introduced a new distance function, Ranked
+Distance, for the standard
+$k$-nearest neighbours algorithm and compared its accuracy and AUC with ten
+other classifiers -- Zero-R, 1R, Na\"{i}ve Bayes, C4.5 decision tree, logistic
+regression, support vector machine, $k$-nearest neighbour with $k=1$ and 20,
+K* and multi-layer perceptron -- several of which (for example K*) have not
+been used in existing work on LOS prediction. We also investigated a variety of
+feature selection methods -- CFS, C4.5 wrapper, correlation coefficient,
+information gain, 1R and expert selection -- encompassing both manual and
+automatic methods, many of which have not been applied to the problem of LOS
+prediction. Our experiments were conducted on two data sets, one from the
+trauma ward of a Sydney hospital and the other a general hospital-wide data
+set from Portugal.
+
+\section{Conclusions}
+The best accuracy we achieved on the trauma data was 77.81\% with the $k$-NN
+classifier using 1 nearest neighbour and 11 out of 78 features. This represents
+an improvement over previous work done by Dinh et al. on the same data set,
+where they managed to achieve 75.06\% with logistic regression. More
+importantly, we were able to improve upon their AUC of 0.812 with a result of
+0.846, also using logistic regression but first discretising numeric features
+and reducing the feature set to 29 features.
+If we consider an even smaller number of features, then K*
+gives an AUC of 0.838 with 11 features selected by the C4.5 wrapper, which
+improves upon the result of Dinh et al. while using a little more than half
+the number of features. The drawback of $k$-NN and K* is that they require all
+examples to be stored and then searched during classification, which becomes
+storage- and computation-expensive as the data set increases in size.
+
+We were able to achieve substantially better results on the hospital-wide data,
+with a best accuracy of 98.23\% using a C4.5 decision tree and best AUC of
+0.994 achieved equally by logistic regression, SVM and K*. Like with the
+trauma data, we were also able to sacrifice a small amount of accuracy or AUC
+in order to use substantially less features: the MLP was able to achieve
+96.45\% accuracy and 0.983 AUC while using only 2 out of 14 features. In both
+these data sets, investigating feature selection was not only a way to
+improve the performance of our classifiers, but also allowed us to see the
+trade-off between the two (sometimes conflicting) objectives of increasing
+predictive power and reducing feature set size.
+
+An important contribution of our work is the use of several nearest neighbour
+algorithms for predicting the LOS class of a patient. We found that these
+techniques can achieve good accuracy and discriminating ability if a suitable
+set of features is found. In addition, we proposed a variation on the standard
+nearest neighbours distance function, in order to take the relative importance
+of each feature into account when determining the nearest neighbours of an
+example. We found that this achieved better results than the other nearest
+neighbour algorithms we evaluated, but only when most of the original feature
+set was used.
+
+Although we were able to improve on the work of Dinh et al. and confirm the
+importance of some features they and others found in predicting the LOS for
+trauma patients, we did not find a single classifier or feature selector
+that was superior to the others but instead observed different combinations
+giving better results than others in certain situations. This highlights the
+importance of understanding the data and investigating a range of classifiers
+and feature selection techniques when approaching any medical data mining
+problem, for which our work will become a catalogue.
+However, we did find that logistic regression tended to perform well
+on both data sets and across a number of feature selection methods, supporting
+its common use in deriving classifiers for various medical phenomena.
+
+\section{Future Work}
+In this thesis we have systematically evaluated a large number of combinations
+of classifiers and feature selection methods while also testing the effect of
+supervised discretisation, and have thus created a catalogue of examples of
+the application of data mining techniques to LOS classification. Our work on
+this topic has raised a number of avenues for futher research, one of which is
+to re-formulate the binary classification problem as a \textit{multi-class}
+classification problem. Our current categorisation of the LOS outcome is very
+crude: we can only classify patients into two categories. In order to better
+assist physicians in making decisions, we could split the classification into
+three or more classes, such as 1-3 days, 4-10 days and $>$11 days, and
+investigate how different classifiers and features perform for this slightly
+different problem.
+
+We mentioned in the discussion in Chapter \ref{chap:results} that we did not
+fine-tune the parameters of the classifiers that we evaluated. There is room
+for us to optimise these parameters and test whether these result in
+significant improvements in accuracy and AUC, and explore reasons why this
+did or did not help. We could also broaden our investigation by additionally
+using another data set, preferably from a different medical domain.
+
+For the $k$-NN algorithm using our Ranked Distance approach, we only considered
+one kind of weight assignment function, but it is not the only way to assign
+these weights. There is room to investigate other types of functions, as well
+as using weights obtained from another algorithm (such as logistic regression).
+In addition to assigning the weights, we have also only investigated one method
+of determining the relative importance of features: the correlation coefficient.
+This leaves much room to use other ways of ranking features, such as the
+information gain and 1R accuracy (which we have used in our work as feature
+selectors).
+
+Another area of potential future work is to investigate \textit{ensembles} of
+classifiers, which take the predictions of more than one classifier (of the
+same or different kind) and combine them to make a final prediction. There are
+numerous ways in which this can be done: an example is to take the majority
+class that is predicted as the final prediction for our ensemble. Ensembles
+have often been demonstrated to perform better than any single classifier,
+and there is a wide range of methods and combinations we could evaluate in
+order to assess their performance relative to the single classifiers we have
+used in our work.

### 2014-11-02 thesis/thanks/thanks.tex
Minor fix to acknowledgements

@@
 my sister Lily -- thank you for tirelessly supporting me throughout this year,
-especially towards the end. This would not have been possible without all of
-you.
+especially towards the deadline. This would not have been possible without all
+of you.

### 2014-11-02 thesis/abstract/abstract.tex
Add abstract

\chapter*{Abstract}

Predicting the length of hospital stay for injured patients involves making a
judgement about how long we believe they will need to be hospitalised, so that
doctors and hospital staff can plan the appropriate course of treatment for the
patient to ensure they recover in the quickest possible time. This ensures the
most efficient use of limited hospital resources. However, previous studies
have used a limited number of statistical techniques such as logistic
regression in order to predict the length of stay.

Our work systematically
evaluates the effect of discretisation on all combinations of 11 classifiers
using 5 different feature selection methods. One of these classifiers is a
novel contribution of our work. Furthermore, our evaluation includes 4 feature
selection methods and 3 classifiers which have to date not been used in
predicting the length of stay, compared alongside commonly used
techniques. We also compare two independent data sets.

Our results show that while logistic regression tended to perform best out of
many combinations, it is possible to achieve to achieve an accuracy and an area
under the receiver operating characteristic curve that is less than 2\% worse
than this best result using only around 15\% of the total features. We were
also able to improve upon the baseline accuracy and area under the curve by
2.75\% and 0.034 respectively, and recommend further investigation into the
use of nearest neighbour approaches and feature selection to future studies
of length of stay prediction.

### 2014-11-03 thesis/introduction/introduction.tex
Added a todo note to the introduction
Rewrite introduction

--- 2014-09-20 thesis/introduction/introduction.tex
@@
 are dealing with classification; otherwise, if we wish to predict a numeric
-quantity, such as the value of the stock market at some future data, then
+quantity, such as the value of the stock market at some future date, then
 it is a regression problem;
@@
 
-\subsection{Problem Statement}
+The technical methods of data mining are drawn substantially from a field of
+computer science known as \textit{machine learning}, which is dedicated to the
+study of systems that can learn structural descriptions from data examples.
+In this thesis we will focus on the task of predictive modelling, also
+known as \textit{supervised learning} in the field of machine learning.
+Specifically,
+our problem will be one of classification. We will introduce the necessary
+terminology in the next section as we describe the problem, as well as
+throughout this thesis as necessary.
 
-\subsection{Previous Work}
+\section{Problem Statement}
+Consider a data set of patients admitted to the trauma ward of a hospital.
+This data set looks like a very large table: there are many columns, each one
+describing some aspect of a patient (for example, age); there are also many
+rows, each corresponding to a particular patient that was admitted. We will
+refer to the columns as \textit{features} or \textit{attributes}, so-named
+because they describe some particular aspect of the rows, which are often
+called \textit{instances}, \textit{samples}, or \textit{training examples}.
+Concretely, as an example, our data set could contain
+two features (age and gender) for five patients: as a table, this means that
+we have two columns, one for age and another for gender, with five rows
+containing the age and gender data for each patient. Attributes can be numeric
+(such as age) and potentially take on an infinite number of different values,
+or categorical (such as gender) and can only have one of a limited, pre-defined
+list of values.
+
+In a general supervised learning problem, we wish to predict a value for some
+desired quantity given some values for the attributes. The relationship between
+the features and the desired quantity can be \textit{learned} from using data
+that contains known values for the quantity that we wish to predict: this is
+what makes the problem \textit{supervised}.
+Our problem is a binary \textit{classification} task in which we
+predict one of two possible categories for a patient: a length of stay (LOS) of
+less than or equal to 2 days, or a LOS of greater than 2 days. These outcomes
+are also called \textit{classes} or \textit{class labels}.
+
+To do this, we use our data set of trauma patients with \textit{known}
+values for the LOS and apply various \textit{learning} or
+\textit{classification algorithms} to deduce
+a relationship between the features and the LOS. Once the learning algorithm
+or \textit{classifier}
+has ``learned'' this relationship, it is then able to predict the class of
+any patient given a set of features describing him or her.
+Note that the LOS is simply another attribute of our data set; however, because
+it is the quantity that we are interested in predicting, it is also called the
+\textit{target}, \textit{outcome}, or \textit{class}.
+
+\section{Motivation}
+The LOS is an important measure of resource utilisation in hospitals around
+the world \citep{Ng2006},
+and many studies have been carried out to investigate whether the
+LOS of a patient in various medical domains -- after surgery, burns, or
+intensive care treatment to name a few -- can be predicted accurately.
+Regardless of the
+specifics of the problem, accurate LOS prediction results in better planning
+and management of limited resources such as staff, equipment and medical
+supplies. This is particularly critical as hospitals often face the problem
+of limited funding \citep{Walczak2003}.
+As we will see in the next chapter, there has been limited
+application of a variety of data mining techniques to predict the LOS in
+various medical domains, including trauma. Our work will aim to bridge this
+gap by providing insights on the application of a wide range of data mining
+techniques to LOS prediction.
 
 \section{Contributions}
+The work presented in this thesis aims to address several gaps in the current
+research into LOS prediction, by building upon the work of Dinh et al.
+\citep{Dinh2013a}. We systematically and empirically evaluated a comprehensive
+catalogue of 11 classification
+algorithms and 5 feature selection methods, as well as the effect of
+discretisation, for predicting the LOS of trauma patients. 3 of the
+classifiers and 4 of the feature selection methods have not been applied in
+LOS prediction to date. Additionally, we are also not aware of any previous
+work using the discretisation method that we evaluate in this thesis.
 
-\section{Results and Evaluation}
+We also proposed a new distance function, Ranked Distance, for use with one
+of the classification algorithms that we evaluate, and showed that this
+performed better than the standard distance function in certain situations.
+
+Finally, we conducted our evaluations on two independent data sets from
+different medical domains: one from trauma and the other a general hospital
+data set. Not only is the evaluation of a second data set in one work
+uncommon in the literature, but the inclusion of more than one data set
+allowed us to draw additional insights about the classifiers and feature
+selections we used which would not have been possible with only one set of
+data.
+
+\section{Thesis Outline}
+This thesis is structured as follows. Chapter \ref{chap:litreview} introduces
+the major methods in supervised learning as used in past work on LOS
+prediction. We explore the major algorithms, summarise the current state of
+research, and identify where our work fits within this area.
+
+Chapters \ref{chap:classification}, \ref{chap:preprocess} and
+\ref{chap:selection} describe in detail the classification algorithms, and
+feature preprocessing and selection methods used in our work. This leads into
+the process of our evaluation, which is detailed in Chapter
+\ref{chap:evaluation}.
+
+Our findings are presented and discussed in Chapter \ref{chap:results}.
+We conclude in Chapter \ref{chap:conclusion}, with indications for future
+work in this area.

### 2014-11-03 thesis/litreview/litreview.tex
Small updates to litreview

@@
 particular outcome by an equation of the form
-\begin{equation}
+\begin{equation*}
 \mathrm{log}\left(\frac{p}{1-p}\right) = \beta_0 + \beta_1X_1 + \beta_2X_2 + \dots + \beta_nX_n
-\end{equation}
+\end{equation*}
 where $p$ is the probability of the desired outcome, $\beta_i$ are coefficients
@@
 can be tested using measures such as the area under the receiver operating
-characteristic (ROC) curve (AUC) and the Hosmer-Lemeshow statistic.
+characteristic (ROC) curve (AUC or AUROC) and the Hosmer-Lemeshow statistic.
 
@@
 6.314 days at post-treatment, where MAE is defined as
-\begin{equation}
-MAE = \frac{\sum_{i=1}^n |\mathrm{PredictedLOS}_i-\mathrm{ActualLOS}_i|}{n}
-\end{equation}
+\begin{equation*}
+MAE = \dfrac{1}{n}\sum_{i=1}^n |\mathrm{PredictedLOS}_i-\mathrm{ActualLOS}_i|
+\end{equation*}
 and PredictedLOS$_i$ is the predicted length of stay for the $i$-th instance,
@@
 application \citep{Kapadia2000}:
-\begin{center}
+
+\begin{table}[h]
 \begin{tabular}{l|llll}
@@
 \end{tabular}
-\end{center}
+\caption{}
+\end{table}
 From the matrix, we can say that the probability of transitioning from state
@@
 \begin{figure}[h]
+\centering
+\includegraphics[scale=0.95]{images/a2-nn}
 \caption{A graphical representation of a neural network with one hidden layer,
 used by Tu \& Guerriere in predicting LOS.}
-\centering
-\includegraphics[scale=0.95]{images/a2-nn}
 \end{figure}
@@
 \begin{figure}[h]
-\caption{An example of finding the largest margin hyperplane for a two-class
-support vector machine classifier.}
 \label{fig:svm-example}
@@
 \includegraphics[scale=1.05]{images/a2-svm}
+\caption{An example of finding the largest margin hyperplane for a two-class
+support vector machine classifier.}
 \end{figure}

### 2014-11-03 thesis/abstract/abstract.tex
A whole bunch of updates including source code listing in the appendix

@@
 Predicting the length of hospital stay for injured patients involves making a
-judgement about how long we believe they will need to be hospitalised, so that
-doctors and hospital staff can plan the appropriate course of treatment for the
-patient to ensure they recover in the quickest possible time. This ensures the
+judgement about how long they need to be hospitalised, so that doctors and
+hospital staff can plan the appropriate course of treatment for the patient
+to ensure they recover in the quickest possible time. This also ensures the
 most efficient use of limited hospital resources. However, previous studies
 have used a limited number of statistical techniques such as logistic
-regression in order to predict the length of stay.
+regression in order to predict the length of stay. Also, they have used only
+manual feature selection and have not investigated the effect of automatic
+feature selection on classifier performance.
 
-Our work systematically
-evaluates the effect of discretisation on all combinations of 11 classifiers
-using 5 different feature selection methods. One of these classifiers is a
-novel contribution of our work. Furthermore, our evaluation includes 4 feature
-selection methods and 3 classifiers which have to date not been used in
-predicting the length of stay, compared alongside commonly used
-techniques. We also compare two independent data sets.
+Our work systematically evaluates and compares the performance of a number of
+state-of-the-art classification algorithms  and feature selection methods,
+and also assesses the effect of feature discretisation for all combinations
+of classifiers and feature selectors.  Our evaluation were conducted on two
+datasets, one from a trauma ward in a major Sydney hospital and another one
+from a general hospital in Portugal. We propose a new classification algorithm
+called Ranked Distance Nearest Neighbor that considers the
+relative importance of each feature when determining the nearest neighbours.
 
-Our results show that while logistic regression tended to perform best out of
-many combinations, it is possible to achieve to achieve an accuracy and an area
-under the receiver operating characteristic curve that is less than 2\% worse
-than this best result using only around 15\% of the total features. We were
-also able to improve upon the baseline accuracy and area under the curve by
-2.75\% and 0.034 respectively, and recommend further investigation into the
-use of nearest neighbour approaches and feature selection to future studies
-of length of stay prediction.
+For the trauma dataset, we were able to achieve an accuracy of 77.81\% with
+1-nearest neighbour and an area under the curve of 0.846 with logistic
+regression. This represents an improvement of 2.75\% and 0.034 respectively
+over the previous best result on the same dataset. For the
+general hospital dataset, the best results were 98.23\% accuracy with a
+decision tree and 0.994 area under the curve with logistic regression,
+support vector machine and K*. We also show that our proposed Ranked
+Distance Nearest Neighbour improves accuracy and area under the curve over
+traditional nearest neighbour algorithms. We comprehensively discuss
+the results and suggest new directions for future studies.

### 2014-11-03 thesis/conclusion/conclusion.tex
A whole bunch of updates including source code listing in the appendix

@@
 effect of discretisation, for predicting the length of hospital stay of trauma
-patients. In particular, we introduced a new distance function, Ranked
-Distance, for the standard
-$k$-nearest neighbours algorithm and compared its accuracy and AUC with ten
+patients. In particular, we introduced a new classification algorithm, Ranked
+Distance Nearest Neighbours, and compared its accuracy and AUC with ten
 other classifiers -- Zero-R, 1R, Na\"{i}ve Bayes, C4.5 decision tree, logistic
@@
 classifier using 1 nearest neighbour and 11 out of 78 features. This represents
-an improvement over previous work done by Dinh et al. on the same data set,
+an improvement over previous work done by Dinh et al. \cite{Dinh2013a}
+on the same data set,
 where they managed to achieve 75.06\% with logistic regression. More
@@
 gives an AUC of 0.838 with 11 features selected by the C4.5 wrapper, which
-improves upon the result of Dinh et al. while using a little more than half
+improves upon the result of Dinh et al. \cite{Dinh2013a} while using a little
+more than half
 the number of features. The drawback of $k$-NN and K* is that they require all
@@
 techniques can achieve good accuracy and discriminating ability if a suitable
-set of features is found. In addition, we proposed a variation on the standard
-nearest neighbours distance function, in order to take the relative importance
+set of features is found. In addition, we proposed a new algorithm, Ranked
+Distance Nearest Neighbours, which is an extension of the standard nearest
+neighbours classifier that takes the relative importance
 of each feature into account when determining the nearest neighbours of an
@@
 
-Although we were able to improve on the work of Dinh et al. and confirm the
+Although we were able to improve on the work of Dinh et al. \cite{Dinh2013a}
+and confirm the
 importance of some features they and others found in predicting the LOS for
@@
 
-For the $k$-NN algorithm using our Ranked Distance approach, we only considered
+In our proposed Ranked Distance Nearest Neighbours algorithm, we only considered
 one kind of weight assignment function, but it is not the only way to assign

### 2014-11-03 thesis/introduction/introduction.tex
A whole bunch of updates including source code listing in the appendix

@@
 research into LOS prediction, by building upon the work of Dinh et al.
-\citep{Dinh2013a}. We systematically and empirically evaluated a comprehensive
-catalogue of 11 classification
-algorithms and 5 feature selection methods, as well as the effect of
-discretisation, for predicting the LOS of trauma patients. 3 of the
-classifiers and 4 of the feature selection methods have not been applied in
-LOS prediction to date. Additionally, we are also not aware of any previous
-work using the discretisation method that we evaluate in this thesis.
+\citep{Dinh2013a}. Our contributions can be summarised as follows:
 
-We also proposed a new distance function, Ranked Distance, for use with one
-of the classification algorithms that we evaluate, and showed that this
-performed better than the standard distance function in certain situations.
+\begin{enumerate}
+\item We systematically evaluated the performance of a catalogue of eleven
+classification algorithms and five feature selection methods for predicting the
+LOS of trauma patients. Three of the classifiers and four of the feature
+selection methods have not been applied in LOS prediction to date, and only
+one has been applied in our medical domain.
 
-Finally, we conducted our evaluations on two independent data sets from
-different medical domains: one from trauma and the other a general hospital
-data set. Not only is the evaluation of a second data set in one work
-uncommon in the literature, but the inclusion of more than one data set
-allowed us to draw additional insights about the classifiers and feature
-selections we used which would not have been possible with only one set of
-data.
+\item We assessed the effect of feature discretisation on classifier
+performance. This has not been studied in previous work.
+
+\item We proposed a new algorithm, Ranked Distance Nearest Neighbour, and
+showed that it performed better than the standard nearest neighbour algorithm
+in certain situations.
+
+\item We conducted our evaluations on two independent data sets from different
+medical domains: one from a trauma ward in a major Sydney hospital and the other
+from a general hospital in Portugal. The collection, processing and
+understanding of the data involved collaboration with medical experts (Dr
+Michael Dinh from the Royal Prince Alfred Hospital in Sydney) and researchers
+from Portugal (Associate Professor Paulo Cortez from the University of Minho).
+Not only is the evaluation of a second data set in one work uncommon in the
+literature, the inclusion of more than one data set allowed us to draw
+additional insights about the classifiers and feature selections we used
+which would not have been possible with only one set of data.
+
+\item We comprehensively analysed and discussed various aspects of the results
+(such as comparison across classification algorithms, feature selection
+methods, performance measures, datasets and discretisation). Our results showed
+an improvement over the best result on the same datasets.
+\end{enumerate}
 

### 2014-11-03 thesis/method/evaluation.tex
Add example ROC curve to evaluation
A whole bunch of updates including source code listing in the appendix

@@
 comparable to the literature and suitable for our particular learning problem.
-This chapter discusses the data sets, procedure and performance measures used
+This chapter discusses the procedure and performance measures used
 in evaluating our work.
-
-\section{Data Sets}
-In our work we obtained two independent sets of LOS data, one specifically for
-trauma and the other a general, hospital-wide set.
-
-\subsection{Trauma LOS Data}
-\subsubsection{Collection}
-The data set used consisted of trauma registry data from the trauma centre
-at the Royal Prince Alfred Hospital, a major trauma centre in New South Wales,
-Australia. It covered all adult (age 15 and over) inpatient admissions to the
-trauma centre from 2007--2011.
-
-All patients were first admitted to the trauma ward until discharged
-or transferred to an appropriate unit within the hospital. A single trained
-data manager recorded a variety of attributes about the admitted patient,
-such as age, gender, blood pressure, mechanism of injury and body regions
-that were injured. We received this data in a Microsoft Excel spreadsheet.
-
-\subsubsection{Characteristics}
-There were 2546 patient records in the data set we received, comprising of 79
-features, one of which was the target variable \texttt{los48}.
-\texttt{los48} was a binary variable -- that is, it could only take two
-values, 0 or 1 -- with 1 indicating that the patient stayed two days or less,
-and 0 indicating a stay of greater than 2 days.
-
-\subsection{General Hospital LOS Data}
-By collaborating with our colleagues from the University of Minho in Portugal,
-we were able to obtain LOS data for all inpatient hospitalisations from
-2000--2013 at the Hospital das Foras Armadas in Lisboa, Portugal. The data
-contained 26462 inpatient records spanning all medical departments within the
-hospital, comprising 15 features that were not specific to any medical domain.
 
@@
 \label{sec:code}
-\todo{attach code in appendix?}
 Leading and trailing whitespace was removed and feature names transformed using
-code written in Python to manipulate the CSV. All runs of WEKA were executed by
+code written in Python to manipulate the CSV exported from Microsoft Excel.
+All runs of WEKA were executed by
 a program we wrote in Bash (a Unix shell) script. This program was able to run
@@
 
-In order to implement our novel Ranked Distance function (Section
+In order to implement our novel Ranked Distance Nearest Neighbour algorithm,
 \ref{sec:rankeddist}), we created our own class within the WEKA Java package
@@
 development environment we used was Eclipse \todo{insert version}, and the
-source was built using the 1.7 Java runtime. \todo{double check}
+source was built using the 1.7.0 Java runtime.
+
+The source code listing of the most important programs we wrote are displayed
+in Appendix \ref{app:code}.
 
@@
 \begin{figure}[h]
+\centering
+\includegraphics{images/method/example-roc.eps}
 \caption{Two ROC curves on the same graph. Note that the straight line from
@@
 \label{fig:sampleroc}
-\centering
-%\includegraphics[scale=0.95]{images/a2-nn}
 \end{figure}

### 2014-11-03 thesis/method/preprocess.tex
A whole bunch of updates including source code listing in the appendix

@@
 
-In this chapter give an overview of preprocessing and describe the steps we
-took on our data.
+In this chapter we introduce the data sets and give an overview of
+preprocessing, describing the specific steps we took on our data.
+
+\section{Data Sets}
+In our work we obtained two independent sets of LOS data, one specifically for
+trauma and the other a general, hospital-wide set.
+
+\subsection{Trauma LOS Data}
+\subsubsection{Collection}
+The data set used consisted of trauma registry data from the trauma centre
+at the Royal Prince Alfred Hospital, a major trauma centre in New South Wales,
+Australia. It covered all adult (age 15 and over) inpatient admissions to the
+trauma centre from 2007--2011.
+
+All patients were first admitted to the trauma ward until discharged
+or transferred to an appropriate unit within the hospital. A single trained
+data manager recorded a variety of attributes about the admitted patient,
+such as age, gender, blood pressure, mechanism of injury and body regions
+that were injured. We received this data in a Microsoft Excel spreadsheet.
+
+\subsubsection{Characteristics}
+There were 2546 patient records in the data set we received, comprising of 79
+features, one of which was the target variable \texttt{los48}.
+\texttt{los48} was a binary variable -- that is, it could only take two
+values, 0 or 1 -- with 1 indicating that the patient stayed two days or less,
+and 0 indicating a stay of greater than 2 days.
+
+\subsubsection{Features}
+
+\subsection{General Hospital LOS Data}
+By collaborating with our colleagues from the University of Minho in Portugal,
+we were able to obtain LOS data for all inpatient hospitalisations from
+2000--2013 at the Hospital das Foras Armadas in Lisboa, Portugal. The data
+contained 26462 inpatient records spanning all medical departments within the
+hospital, comprising 15 features that were not specific to any medical domain.
+
+\subsubsection{Features}
 

### 2014-11-03 thesis/thanks/thanks.tex
A whole bunch of updates including source code listing in the appendix

@@
 Caraher and Christine Au Yeung, for the selfless encouragement, help and good
-times which have made made this year-long journey possible.
+times which have made this year-long journey possible.
 

### 2014-11-03 thesis/results/features-selected-nothreshold.tex
Add tables from feature selection

\begin{table}[htbp]
\resizebox{\textwidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{llcccccccc}
\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{4}{c}{No discretisation} & \multicolumn{4}{c}{Discretisation} \\
\hline
 &  & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} \\
 Type & Method & \multicolumn{1}{l}{Selected} & \multicolumn{1}{l}{\% of total} & \multicolumn{1}{l}{Selected} & \multicolumn{1}{l}{\% of total} & Selected & \% of total & Selected & \% of total \\
 \hline
 \multirow{2}{*}{Manual} & Baseline & 19 & 24.36 & - & - & 19 & 24.36 & - & - \\
  & Expert Selection & 11 & 14.10 & - & - & 11 & 14.10 & - & - \\
  \multirow{2}{*}{Automatic} & CFS & 9 & 11.54 & 2 & 14.29 & 9 & 11.54 & 2 & 14.29 \\
   & C4.5 Wrapper & 10 & 12.82 & 4 & 28.57 & 11 & 14.10 & 4 & 28.57 \\
   \end{tabular}
}
\caption{A listing of the number of features selected by non-threshold feature selectors for both data sets, with and without discretisation.}
\label{tab:features-selected-nothreshold}
\end{table}

### 2014-11-03 thesis/results/features-selected-threshold.tex
Add tables from feature selection

\begin{table}[htbp]
\resizebox{\textwidth}{!}{%
\begin{tabular}{lcccccccc}
\multicolumn{9}{c}{Correlation} \\
 & \multicolumn{4}{c}{No discretisation} & \multicolumn{4}{c}{Discretisation} \\
 \hline
  & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} \\
  Threshold & Selected & \% of total & Selected & \% of total & Selected & \% of total & Selected & \% of total \\
  \hline
  0 & 58 & 74.36 & 14 & 100.00 & 57 & 73.08 & 13 & 92.86 \\
  0.01 & 55 & 70.51 & 14 & 100.00 & 54 & 69.23 & 13 & 92.86 \\
  0.05 & 41 & 52.56 & 12 & 85.71 & 41 & 52.56 & 12 & 85.71 \\
  0.1 & 27 & 34.62 & 10 & 71.43 & 29 & 37.18 & 10 & 71.43 \\
  0.125 & 19 & 24.36 & 7 & 50.00 & 20 & 25.64 & 7 & 50.00 \\
  0.15 & 15 & 19.23 & 6 & 42.86 & 12 & 15.38 & 6 & 42.86 \\
  0.2 & 9 & 11.54 & 4 & 28.57 & 7 & 8.97 & 4 & 28.57 \\
  0.3 & 2 & 2.56 & 1 & 7.14 & 1 & 1.28 & 1 & 7.14 \\
  0.5 & 0 & 0.00 & 1 & 7.14 & 0 & 0.00 & 1 & 7.14 \\
  0.75 & 0 & 0.00 & 1 & 7.14 & 0 & 0.00 & 1 & 7.14 \\
   &  &  &  &  &  &  &  &  \\
   \multicolumn{9}{c}{IG} \\
    & \multicolumn{4}{c}{No discretisation} & \multicolumn{4}{c}{Discretisation} \\
    \hline
     & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} \\
     Threshold & Selected & \% of total & Selected & \% of total & Selected & \% of total & Selected & \% of total \\
     \hline
     0 & 57 & 73.08 & 13 & 92.86 & 57 & 73.08 & 14 & 100.00 \\
     0.001 & 47 & 60.26 & 13 & 92.86 & 47 & 60.26 & 14 & 100.00 \\
     0.005 & 33 & 42.31 & 12 & 85.71 & 33 & 42.31 & 12 & 85.71 \\
     0.01 & 27 & 34.62 & 12 & 85.71 & 27 & 34.62 & 12 & 85.71 \\
     0.05 & 5 & 6.41 & 8 & 57.14 & 4 & 5.13 & 8 & 57.14 \\
     0.1 & 2 & 2.56 & 7 & 50.00 & 2 & 2.56 & 7 & 50.00 \\
     0.15 & 1 & 1.28 & 5 & 35.71 & 1 & 1.28 & 5 & 35.71 \\
     0.2 & 0 & 0.00 & 3 & 21.43 & 0 & 0.00 & 3 & 21.43 \\
     0.3 & 0 & 0.00 & 2 & 14.29 & 0 & 0.00 & 2 & 14.29 \\
     0.5 & 0 & 0.00 & 0 & 0.00 & 0 & 0.00 & 0 & 0.00 \\
      &  &  &  &  &  &  &  &  \\
      \multicolumn{9}{c}{1R} \\
       & \multicolumn{4}{c}{No discretisation} & \multicolumn{4}{c}{Discretisation} \\
       \hline
        & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} \\
        Threshold & Selected & \% of total & Selected & \% of total & Selected & \% of total & Selected & \% of total \\
        \hline
        66.905 & 55 & 70.51 & 14 & 100.00 & 59 & 75.64 & 14 & 100.00 \\
        67.905 & 4 & 5.13 & 14 & 100.00 & 4 & 5.13 & 14 & 100.00 \\
        68.905 & 3 & 3.85 & 14 & 100.00 & 2 & 2.56 & 14 & 100.00 \\
        69.905 & 3 & 3.85 & 14 & 100.00 & 2 & 2.56 & 14 & 100.00 \\
        71.905 & 1 & 1.28 & 14 & 100.00 & 0 & 0.00 & 14 & 100.00 \\
        74.405 & 0 & 0.00 & 14 & 100.00 & 0 & 0.00 & 14 & 100.00 \\
        76.905 & 0 & 0.00 & 6 & 42.86 & 0 & 0.00 & 6 & 42.86 \\
        81.905 & 0 & 0.00 & 3 & 21.43 & 0 & 0.00 & 3 & 21.43 \\
        86.905 & 0 & 0.00 & 2 & 14.29 & 0 & 0.00 & 2 & 14.29 \\
        96.905 & 0 & 0.00 & 0 & 0.00 & 0 & 0.00 & 0 & 0.00
        \end{tabular}
}
\caption{A listing of the number of features selected by threshold for correlation, information gain and 1R for both data sets, with and without discretisation.}
\label{tab:features-selected-threshold}
\end{table}

### 2014-11-03 thesis/appendix/code.tex
Lots of changes

\chapter{Source Code Listings}
\label{app:code}

In this appendix we display the source code of the most important scripts we
wrote in our to conduct our experiments. These are described in more detail
in Section \ref{sec:code}. To download the latest version of the code, please
visit \url{https://github.com/tianyupu/hons-thesis}.

\lstset{
  language = Python,
  texcl = false, % Parse comments as TeX
  backgroundcolor = \color{white},
  basicstyle = \footnotesize\ttfamily\singlespacing,
  commentstyle = \color{mygreen},
  keywordstyle = \color{blue},
  stringstyle = \color{mymauve},
  numberstyle = \tiny\color{mygray},
  numbers = left, % possible values are (none, left, right)
  numbersep = 5pt, % how far the line-numbers are from the code
  rulecolor = \color{black},
  stepnumber = 1, % the step between two line-numbers
  showstringspaces = false, % underline spaces within strings?
  literate = {~} {$\sim$}{1}, % make ~ nice
  literate = {_} {\textunderscore}{1}, % make _ nice
  breaklines = true,
  prebreak = \raisebox{0ex}[0ex][0ex]
  {\color{red}\ensuremath{\space\hookleftarrow}},
  postbreak = \raisebox{0ex}[0ex][0ex]
  {\color{red}\ensuremath{\hookrightarrow\space}},
  tabsize = 2
}

\section{Data Cleaning and Whitespace Removal}
\label{sec:python-preprocess}
\lstinputlisting{../code/sklearn/util.py}

\section{Automating WEKA and Saving Results}
\label{sec:run-weka}
\lstset{language=bash}
\lstinputlisting{../code/weka/run_weka.sh}

\subsection{WEKA Configuration File}
The configuration files that were used in the above script consisted of WEKA
classes in a text file, one per line with as many as we wanted to test in one
experiment. A typical file looked like this:
\lstset{language={}}
\lstinputlisting{../code/weka/weka-config/all_clfs_default}

\section{Summarising a Large Number of Results}
\label{sec:summarise-script}
\lstset{language=bash}
\lstinputlisting{../code/weka/summarise_results.sh}

\subsection{Computing Confidence Intervals from Summarised Results}
\lstset{language=Python}
\lstinputlisting{../code/weka/compute_ci.py}

\section{Ranked Distance Implementation}
\label{sec:rankeddist-code}
\lstset{language=Java}
\lstinputlisting{../code/weka/RankedDistance.java}

### 2014-11-03 thesis/appendix/paper.tex
Lots of changes

@@
-\chapter{Artificial Intelligence in Medicine -- Submitted Paper} \label{app:paper}
+\chapter{Artificial Intelligence in Medicine -- Draft Paper} \label{app:paper}
+
+We are currently in the process of drafting a paper of our results to submit to
+Artificial Intelligence in Medicine, a top journal for research that uses
+artificial intelligence in fields such as medicine, human biology and health
+care. Here we have attached our draft, which will be soon ready for submission.
+
+\clearpage
+
+\setlength{\multicolsep}{0pt}
+\setlength{\topskip}{\fontcharht\font`B }
+\setlength{\parskip}{0pt}
+%[
+\begin{center}
+\section*{Improving the Prediction of Length of Hospital Stay for Trauma Patients}
+\end{center}
+\begin{multicols}{2}
+\begin{center}
+Tianyu Pu, Irena Koprinska \\
+School of Information Technologies \\
+University of Sydney, NSW, 2006 \\
+\texttt{\{tianyu.pu,irena.koprinska\}@sydney.edu.au}
+\end{center}
+\vfill
+\columnbreak
+\begin{center}
+Michael Dinh \\
+Trauma Services, Royal Prince Alfred Hospital \\
+Sydney, NSW, 2042 \\
+\texttt{dinh.mm@gmail.com}
+\end{center}
+\end{multicols}
+%]
+
+\begin{multicols}{2}
+\subsection*{Abstract}
+Predicting the length of hospital stay for injured patients involves making a
+judgement about how long they need to be hospitalised, so that doctors and
+hospital staff can plan the appropriate course of treatment for the patient
+to ensure they recover in the quickest possible time. However, previous studies
+have used a limited number of statistical techniques such as logistic
+regression in order to predict the length of stay, and have not considered
+automatic methods of feature selection.
+
+Our work systematically evaluates and compares the performance of a number of
+state-of-the-art classification algorithms and feature selection methods,
+and also assesses the effect of feature discretisation for all combinations
+of classifiers and feature selectors. We propose a new classification algorithm
+called Ranked Distance Nearest Neighbor that considers the
+relative importance of each feature when determining the nearest neighbours.
+
+We were able to achieve improved accuracy and area
+under the curve for both data sets, and we also show that our proposed Ranked
+Distance Nearest Neighbour improves accuracy and area under the curve over
+traditional nearest neighbour algorithms.
+
+\subsection*{Introduction}
+Hospitals often have limited funding, and therefore managing and utilising
+their resources efficiently has always been a priority \cite{Walczak2003}.
+One key measurement of hospital activity, health care utilisation, and hospital
+cost is the patient length of stay \citep{Omachonu2004,Ng2006}. The length of
+stay has been used in a wide variety of medical domains, such as burns
+ \citep{Yang2010}, intensive
+care \citep{Tu1993,Harper2005,Perez2006,Dybowski1996},
+psychiatry \citep{Lowell1997}, and acute pancreatitis \citep{Pofahl1998}, in
+order to assist hospitals in allocating beds and other patient management
+resources as efficiently as possible. However, given the many different
+conditions that hospitals manage, the length of stay of a patient is
+not easy to predict \citep{Walczak2003}, especially upon a patient's admission.
+Many methods have been explored in the literature, from the realm of statistics
+to techniques in data mining, and they have been explored for many different
+medical domains.
+
+One of the more well-understood and widely used statistical methods for
+predicting the length of stay is \textit{logistic regression} \citep{Tu1996}.
+Logistic regression models are straightforward to construct and use via the
+help of statistical software packages such as SAS, and
+also make the contribution of each independent variable to the final
+prediction explicit. Thus they is easy to interpret when compared to more
+``black box'' approaches such as artificial neural networks \citep{Adams2012}.
+A wide variety of other statistical approaches have been proposed, notably:
+various scoring systems from univariate analysis \citep{Adams2012,Lavoie2005},
+linear regression \citep{Yang2010}, exponential models \citep{Clark2007},
+survival analysis \citep{Vasilakis2005}, and Markov
+models \citep{Perez2006,Jain1989,Kapadia2000}.
+
+Out of the development of the various statistically-motivated models, data
+mining techniques, particularly artificial neural networks,
+began to become more widely adopted following a number of
+early applications to length of stay prediction in intensive
+care \citep{Tu1993,Dybowski1996}, in addition to disease diagnosis and mortality
+prediction \citep{Silva2006}. Artificial neural networks are able to detect
+complex and non-linear relationships between input and output variables and can
+be developed using many different techniques. Much work has been done in using
+neural networks to predict patient length of stay in various medical conditions
+such as acute pancreatitis \citep{Walczak2003}, gastroenteritis \citep{Ng2006},
+surgical intensive care \citep{Buchman1994,Tu1993} and
+psychiatry \citep{Lowell1997}. Other data mining techniques have also been
+applied, such as support vector machines and tree-based
+methods \citep{Harper2005}.
+
+Given the benefits to the hospital of accurate length of stay prediction,
+this research aims to build upon improving length of stay prediction for trauma
+patients by utilising state-of-the-art data mining techniques. Being able to
+predict the length of stay for trauma patients is especially crucial because of
+the scarcity of patient care resource in this area \citep{Yang2010}.
+Additionally,
+despite the wealth of research into length of stay prediction in various
+medical domains, there has been little work done in generalising models from
+one domain to another, or to predict a combination of outcomes (not only
+length of stay, but mortality, or rehabilitation requirement). Being able to
+do this would mean that hospital staff are better equipped to handle patient
+cases and prepare for possible critical outcomes well in advance, resulting
+in time and resources saved for both hospital and patient.
+
+\subsection*{Problem Statement}
+Our problem is a binary \textit{classification} task in which we
+predict one of two possible categories for a patient: a length of stay (LOS) of
+less than or equal to 2 days, or a LOS of greater than 2 days.
+
+\subsection*{Ranked Distance Nearest Neighbour}
+\paragraph{Intuition}
+As we mentioned earlier, all features are considered equally in a $k$-NN
+classifier, which implies that all features are equally important in predicting
+the class. In practice, however, this is not always the case: to predict the
+LOS of a trauma patient, the severity of their injury should affect the final
+LOS more than whether or not they can speak English (which could be just two
+of many features that are recorded about a patient). Our contribution is a
+new NN algorithm that uses a modified distance function which takes into
+account the relative importance of all features called Ranked Distance
+Nearest Neighbour (RD).
+
+\paragraph{Mathematical formulation}
+Given two instances $\mathbf{x} = (x_1,x_2,\ldots,x_n)$ and
+$\mathbf{y} = (y_1,y_2,\ldots,y_n)$, the distance between them is given by:
+\begin{equation*}
+\mathrm{Distance} = \sum_{i=1}^n w_i |x_i-y_i|
+\end{equation*}
+where the $w_i$s weight the contribution of the $i$-th feature to the overall
+distance that is computed by the $k$-NN algorithm. We assume that all feature
+values have been normalised, as described above.
+
+\paragraph{Assignment of weights}
+The weights $w_i$ can be tuned to match the particular problem that is being
+investigated. For our LOS classification problem, we will consider one way
+of weighting features.
+First, we compute the correlation coefficients as described in the point above.
+This gives us a ranking of the importance of each feature to the prediction of
+the class, with 1 being the highest rank (indicating the most importance) and
+$n$ being the lowest rank (indicating the least importance).
+Instead of using the values of the correlation directly to weight
+the distance contribution of each feature, we specify a function:
+\begin{equation*}
+f : \mathrm{Rank} \rightarrow \mathbb{R}, \mathrm{Rank} \in \{1,2,\ldots,n\}
+\end{equation*}
+that describes
+how the weights of the features vary with the rank. This function should
+inituitively be non-increasing and should produce lower values when the rank
+number is greater, meaning that less important features have a lower weight
+in the distance calculation.
+Note that choosing $f(\mathrm{Rank}) = k$ for any non-zero constant $k$
+results in all features contributing equally to the distance calculation.
+We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}^c}$, where
+$c$ varies from 0 to 1.
+
+\subsection*{Feature Selection}
+Here we introduce the feature selection algorithms we evaluate in our work.
+Several of these have not been investigated in previous work on length of
+stay prediction.
+\subsubsection*{Correlation-based Feature Selection}
+Correlation-based feature selection (CFS) is a method developed by Hall
+\cite{Hall2000} which looks for a subset of features that have high correlation
+with the class but low intercorrelation among the features. In this case,
+correlation between two features is measured using what is called the
+\textit{symmetric uncertainty}:
+\begin{equation*}
+U(A,B) = 2\dfrac{H(A)+H(B)-H(A,B)}{H(A)+H(B)}
+\end{equation*}
+where $A$ and $B$ are two nominal attributes and $H$ is the entropy function.
+The entropies are calculated from the probabilities of each feature value, and
+$H(A,B)$, the joint entropy of $A$ and $B$, is calculated from the
+probabilities of each combination of values of $A$ and $B$. Note that $U(A,B)$
+is always between 0 and 1.
+CFS conducts a search through subsets of all features in order to find the
+one feature subset that has the best merit.
+
+\subsubsection*{Pearson Correlation Coefficient}
+The Pearson correlation coefficient measures the degree of linear correlation
+between two variables. It is always in the range $[-1,1]$, where -1 is total
+negative correlation, 0 is no (linear) correlation, and 1 is total positive
+correlation. We compute the Pearson correlation between each feature and the
+class using:
+\begin{equation*}
+r_i = \dfrac{\sum_{i=1}^m (x_i-\bar{x})(c-\bar{c})}{\sqrt{(\sum_{i=1}^m x_i-\bar{x})(\sum_{i=1}^m c-\bar{c})}}
+\end{equation*}
+where $m$ is the number of training examples, $x_i$ is the value of feature
+$i$, and $\bar{x}$ and $\bar{c}$ are the arithmetic averages of the feature
+values and the class values respectively.
+
+\subsubsection*{Information Gain}
+We can also measure the IG of each feature as follows.
+If we have a class $C$ and a feature $F$, the entropy of the
+class before and after observing the feature is:
+\begin{equation*}
+\begin{aligned}
+& \mathrm{entropy}(C) = - \sum_{c \in C} p(c) \mathrm{log}_2 p(c) \\
+& \mathrm{entropy}(C|F) = - \sum_{f \in F} p(f) \sum_{c \in C} p(c|f) \mathrm{log}_2 p(c|f)
+\end{aligned}
+\end{equation*}
+The decrease in entropy reflects the gain in information provided by the
+feature and is given by:
+\begin{equation*}
+IG(C|F) = \mathrm{entropy}(C) - \mathrm{entropy}(C|F)
+\end{equation*}
+
+We can compute the IG for each feature and then select subsets of features that
+are greater than some threshold we specify, similarly to the correlation
+coefficient method of feature selection. We choose a similar set of thresholds
+because of the high numbers of features with very low information gain.
+
+\subsubsection*{1R}
+Another way to evaluate the relevance of features is to use the idea behind the
+1R classifier described in Section. Recall that the 1R
+classifier constructs simple 1-rules based on each feature and uses the feature
+that produces the lowest error rate (that is, the highest accuracy) on the data
+set to classify unseen examples. These accuracy rates can be calculated for
+each feature, which are then used to rank the features from highest to lowest
+accuracy.
+
+We can also select a fixed number of features or all features above a certain
+1R accuracy threshold. This threshold is a slightly different quantity to the
+threshold of the correlation coefficient and information gain methods, as it
+is an accuracy, expressed in percentages. Our first threshold is the Zero-R
+accuracy for the trauma data set, which is the lower of the two Zero-R
+accuracies on our two data sets. We increase this percentage by in increasing
+increments to obtain a range of thresholds up to nearly 97\% accuracy (which
+discards all features for both data sets). Using the same thresholds across
+both data sets allows us to make valid comparisons of the effect of feature
+selection on performance.
+
+\subsubsection*{Wrapper-based Feature Selection}
+The feature selection methods described above (CFS, Pearson correlation and
+1R) have been described as \textit{filter} approaches to attribute selection.
+Kohavi and John proposed another approach, the \textit{wrapper} approach:
+instead of performing feature selection independently of the learning
+algorithm using some measure of ``relevance'',
+the learning algorithm itself is used to evaluate the goodness of
+subsets of features \cite{Kohavi1997}.
+
+A key advantage of the wrapper approach is that the selected feature sets are
+specifically tailored to a learning algorithm, which ensures the best
+performance of that algorithm given the data set. However, wrapper approaches
+can also be prohibitively slow if the data or feature sets are large and the
+chosen learning algorithm expensive to train (such as support vector machines
+and multi-layer perceptrons). We will evaluate the use of C4.5 decision trees
+in selecting features using the wrapper approach.
+Using decision trees to select features is relatively fast as due to the
+divide and conquer nature of the learning algorithm, and the tendency of the
+trees to select fewer features for learning than other algorithms. Like CFS,
+wrapper methods search for the best subset of features and do not require a
+threshold to discard attributes.
+
+\subsection*{Evaluation Method}
+Our evaluation was carried out using a trauma-specific data set
+containing 2546 records collected from 2007--2011 collected at a major hospital
+in Sydney, Australia.
+\subsubsection*{Classifiers}
+We tested the performance of Zero-R, 1R, Na\"{i}ve Bayes, C4.5 decision trees,
+support vector machines, multi-layer perceptrons, $k$-nearest neighbour with 1
+and 20 nearest neighbours, K* and logistic regression, as well as our proposed
+Ranked Distance Nearest Neighbour.
+
+\subsubsection*{Procedure}
+The steps that we took to evaluate the performance of the various classifiers
+and feature selection methods were:
+\begin{enumerate}
+\item For each data set, we evaluated the effect of the 4 automatic feature
+selection methods on each of the 8 classifiers listed above, giving us 32
+configurations for each data set, over
+4 sets. Evaluation was performed using ten runs of \textit{ten-fold stratified
+cross-validation}.
+\item For the two trauma LOS data sets, we additionally evaluated the two
+manual feature selection methods with all 8 classifiers. Recall from Section
+\ref{sec:manual} that one of the feature sets is from the key previous work
+on trauma LOS prediction that we use as a baseline, and the other feature set
+is a set of features included through consultation with a domain expert.
+\end{enumerate}
+
+\subsection*{Results and Discussion}
+The best accuracy was 77.81\%, achieved using the $k$-NN algorithm
+with 1 nearest neighbour, a discretised data set, and feature selection using a
+C4.5 wrapper. The best AUC of 0.846 was achieved using logistic regression with
+features selected by a correlation coefficient threshold of 0.1, and again
+using the discretised data set.
+
+\begin{figure*}[htbp]
+\begin{subfigure}{.48\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-nofs-acc.eps}
+\caption{}
+\label{}
+\end{subfigure}%
+\begin{subfigure}{.55\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-nofs-auc.eps}
+\caption{}
+\label{}
+\end{subfigure}
+
+\begin{subfigure}{.48\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-dinh-acc.eps}
+\caption{}
+\label{}
+\end{subfigure}%
+\begin{subfigure}{.55\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-dinh-auc.eps}
+\caption{}
+\label{}
+\end{subfigure}
+
+\begin{subfigure}{.48\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-expert-acc.eps}
+\caption{}
+\label{}
+\end{subfigure}%
+\begin{subfigure}{.55\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-expert-auc.eps}
+\caption{}
+\label{}
+\end{subfigure}
+
+\begin{subfigure}{.48\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-cfs-acc.eps}
+\caption{}
+\label{fig:tr-nothreshold-cfs-acc}
+\end{subfigure}%
+\begin{subfigure}{.55\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-cfs-auc.eps}
+\caption{}
+\label{}
+\end{subfigure}
+\caption[]{Comparison of accuracy and AUC between classifiers for each non-threshold feature selection method, grouped by whether or not discretisation was applied.}
+\label{fig:tr-nothreshold}
+\end{figure*}
+
+\begin{figure*}
+\ContinuedFloat
+
+\begin{subfigure}{.48\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-wrapper-acc.eps}
+\caption{}
+\label{}
+\end{subfigure}%
+\begin{subfigure}{.55\textwidth}
+\includegraphics[width=\textwidth]{images/results/tr-wrapper-auc.eps}
+\caption{}
+\label{}
+\end{subfigure}
+\caption[]{Comparison of accuracy and AUC between classifiers for each non-threshold feature selection method, grouped by whether or not discretisation was applied.}
+\label{}
+\end{figure*}
+
+It is worth noting that our best accuracy, using $k$-NN with 1
+nearest neighbour, shows a statistically significant improvement upon the accuracy
+that they achieved by 2.75\%
+while using only 11 features (compared to the 19 they used). These 11 features
+make up only 14.1\% of all features in the data set, and performed better than
+the 11 features selected manually by the domain expert.
+
+Despite our inclusion of accuracy as an evaluation metric, the AUC is what Dinh
+et al. \cite{Dinh2013a}
+used to evaluate the ability of their logistic regression classifier to
+discriminate between the two LOS classes. Using their features, the best AUC
+was 0.812 with logistic regression, and our best result was 0.846, also with
+logistic regression. However, this was using the discretised data set with 29
+features selected by correlation coefficient at a threshold of 0.1.
+Although our logistic regression classifier exhibited an AUC improvement of
+0.034, it required roughly 50\% more features than the baseline of 19 features.
+If we are willing to sacrifice a little discriminating power, we are able to
+achieve 0.838 AUC with only 11 features using the K* classifier.
+
+Although $k$-NN with 1 nearest neighbour had the highest accuracy (77.81\%) out
+of all combinations of classifiers and feature selection methods for this data
+set, we should mention that logistic regression came a very close second with
+77.7\%, outperforming more sophisticated approaches such as SVM and MLP. It
+also achieved a higher AUC, and has the added advantages of being faster to
+train, mathematically well-understood and widely applied in medicine.
+
+The best accuracy and AUC results for
+each classifier resulted from a reduced feature set obtained through a
+particular feature selection method. There was at least one feature selector
+for each classifier that managed to outperform the same classifier with
+no feature selection. This is likely due to the large number of features in
+the original data -- after cleaning and preliminary removal of redundant
+features, we were still left with over 50. Many of these would still likely
+be irrelevant to predicting the class.
+Out of 20 combinations of classifiers
+and discretisation (the 10 classifiers with and 10 without
+discretisation), the C4.5 wrapper method resulted in the best
+accuracy for 7 of these combinations, followed by 5 using information gain
+and 4 using the correlation coefficient and very few for the other methods.
+Under no combination was 1R the best feature selection method to use, as other
+methods produced better accuracy.
+
+Focussing our attention on the AUC reveals slightly different patterns. Here we
+find that the correlation coefficient and the information gain are the feature
+selectors that give the best AUC for most of the classifiers, and perform
+equally well when used in conjunction with logistic regression and SVMs.
+Interestingly, although the C4.5
+wrapper method selects features which allow classifiers to achieve good
+accuracy, only in 2 out of the 20 combinations of classifiers did it give the
+best AUC result out of all the feature selectors.
+
+Recall that $k$-NN with 1 nearest neighbour and C4.5 wrapper feature
+selection produced the best accuracy (77.81\%) for the trauma data set, and
+logistic regression with correlation coefficient feature selection at a
+threshold of 0.1 gave the best AUC (0.846). These two feature selection
+methods selected 11 and 29 features respectively, and the features that
+were chosen in common between them were: \texttt{operation}, \texttt{bp},
+\texttt{iss}, \texttt{lowerlimbnopelvis}, \texttt{age}, \texttt{lowerlimbany}
+and \texttt{icu}. This adds evidence to the work of Dinh et al. who found
+that \texttt{iss}, \texttt{age} and \texttt{operation} were important
+distinguishers of LOS $\leq$ or $>$ 2 days \cite{Dinh2013a}. We also
+confirm the findings of Gabbe et al. who found that \texttt{age},
+\texttt{iss} and \texttt{bp} were useful in predicting the length of
+hospital stay of blunt trauma patients \cite{Gabbe2005}. Our feature
+selection results also
+support the findings of McGonigal et al. \cite{McGonigal1993} who used
+\texttt{age} and
+\texttt{iss} to predict the probability of survival for trauma patients, a
+related problem in the same medical domain.
+Our observations suggest the use of feature selection not only as a
+pre-processing step to improve the accuracy or AUC of a classifier, but also
+as a tool used separately that could help extract insights from medical data.
+
+Although we were able to achieve higher accuracy and AUC than the baseline,
+our best approaches also have a number of limitations which need to be pointed
+out. The first point to note is that using a wrapper method of feature
+selection is slower than using a filter approach (or manually picking the
+features before training), as the learning algorithm used as the wrapper
+needs to be invoked at each of the folds of cross-validation. Additionally,
+the $k$-NN algorithm becomes slower as the data set increases in the number
+of examples. We did not face such a problem during our work, but this will
+have to be addressed for larger data sets. It is also common to use more
+than one nearest neighbour, so care needs to be taken to find a suitable number
+of neighbours for a given problem. Finally, we achieved an improvement
+in AUC of 0.034 but this was with 29 features, which is still a fairly large
+number. The importance of using a fewer number of features is twofold: not
+only does this speed up the training time, it also means that we are able to
+make predictions without requiring a lot of information about a patient,
+which is a crucial consideration if this is to be implemented for use by
+physicians.
+
+\subsection*{Conclusions}
+The best accuracy we achieved on the trauma data was 77.81\% with the $k$-NN
+classifier using 1 nearest neighbour and 11 out of 78 features. This represents
+an improvement over previous work done by Dinh et al. \cite{Dinh2013a}
+on the same data set,
+where they managed to achieve 75.06\% with logistic regression. More
+importantly, we were able to improve upon their AUC of 0.812 with a result of
+0.846, also using logistic regression but first discretising numeric features
+and reducing the feature set to 29 features.
+If we consider an even smaller number of features, then K*
+gives an AUC of 0.838 with 11 features selected by the C4.5 wrapper, which
+improves upon the result of Dinh et al. \cite{Dinh2013a} while using a little
+more than half
+the number of features. The drawback of $k$-NN and K* is that they require all
+examples to be stored and then searched during classification, which becomes
+storage- and computation-expensive as the data set increases in size.
+
+An important contribution of our work is the use of several nearest neighbour
+algorithms for predicting the LOS class of a patient. We found that these
+techniques can achieve good accuracy and discriminating ability if a suitable
+set of features is found. In addition, we proposed a new algorithm, Ranked
+Distance Nearest Neighbours, which is an extension of the standard nearest
+neighbours classifier that takes the relative importance
+of each feature into account when determining the nearest neighbours of an
+example. We found that this achieved better results than the other nearest
+neighbour algorithms we evaluated, but only when most of the original feature
+set was used.
+
+\end{multicols}

### 2014-11-03 thesis/litreview/litreview.tex
A whole bunch of updates including source code listing in the appendix
Lots of changes

@@
 found to have 95\% sensitivity and 24\% specificity, with an area under the ROC
-curve (AUC) of 0.62. This indicates that the model is not able to discriminate
+curve (AUC) of 0.62. This indicates that the model is not able to accurately
+discriminate
 length of stay cases of less than 5 days and equal to 5 days (i.e., the two
-desired outcomes) very well, and only does
+desired outcomes) very well, performing only
 slightly better than simply random guessing (which has AUC 0.50).
@@
 In a similar study, Dinh et al. sought to predict whether or not trauma
-patients could be predicted to stay in hospital \textit{less than 2 days} or
-\textit{greater than 2 days} \citep{Dinh2013a}.
+patients could be predicted to stay in hospital for \textit{less than 2 days}
+or \textit{greater than 2 days} \citep{Dinh2013a}.
 They also used a logistic regression model and
 selected their features using a stepwise selection algorithm with the same
-$P$-value threshold (0.05). Even though we cannot compare their result to that
-of Gabbe et al. above because the models were tested on completely different
+$P$-value threshold (0.05). Even though we cannot directly compare their result
+to that
+of Gabbe et al. above, because the models were tested on completely different
 data sets, it is interesting to note that the model obtained by Dinh et al.
@@
 the two values for the outcome. Again, while we cannot directly compare, the
-better performance could be due to the availability of more relevant features
-to the outcome: Gabbe et al. included very few features specific to trauma
-such as place of injury and mechanism of injury \citep{Gabbe2005} whereas Dinh
-et al. also had variables for certain kinds of injuries (such as if it was
-penetrating), specific body regions that were affected, and whether operations
-were required. Given the nature of the length of stay, it would seem that Dinh
+better performance could be due to the presence of features in the data more
+relevant to the outcome: the data of Gabbe et al. included very few features
+specific to trauma \citep{Gabbe2005} whereas Dinh
+et al. had variables indicating the kind of injury (such as penetrating trauma,
+the bodily regions that were affected, and whether operations
+were required). Given the nature of the length of stay it would seem that Dinh
 et al. derived a model that had included features with more explanatory power
@@
 Buchman et al., who used logistic regression to predict the probability of a
-patient in a surgical intensive care unit staying greater than a
+patient in a surgical intensive care unit staying greater than one
 week \citep{Buchman1994}.
@@
 While logistic regression is simple to use and its results straightforward
-to interpret, the above approaches have described a coarse two-class prediction
-problem, namely that of predicting the length of stay less than or greater than
-a certain threshold. A number of \textit{linear} regression models have been
-explored in the literature that examine the task of predicting length of stay
+to interpret, the above approaches describe a coarse two-class prediction
+problem, namely that of predicting length of stay dichotomised by an
+arbitrarily chosen threshold.
+A number of \textit{linear} regression models have been
+explored in the literature that examine the problem of predicting length of stay
 as a continuous quantity.
@@
 to predict length of stay as a continous variable could give us more accurate
-and useful insight and facilitate better resource allocation.
+and useful insights, and facilitate better resource allocation.
 
@@
 they did not perform any feature selection when constructing their models and
-found that this gave better performance against models constructed using
+found that this gave better performance compared with models constructed using
 correlation-based feature selection\footnote{Abbreviated CFS, correlation-based
@@
 at the admission, acute, and post-treatment stages, they were able to decrease
-their predictions' mean absolute error (MAE) from 9.532 days at admission, to
-6.314 days at post-treatment, where MAE is defined as
+the mean absolute error (MAE) of their predictions from 9.532 days at admission,
+to 6.314 days at post-treatment, where MAE is defined as
 \begin{equation*}
@@
 \end{tabular}
-\caption{}
+\caption{A transition probability matrix from Kapadia et al. for the
+movement of patients through a pediatric intensive care unit.}
 \end{table}
@@
 data, which allowed them to predict neonatal length of stay across heterogenous
-sub-groups (short-stay and long-stay). The idea is that these two sub-groups
+sub-groups (short-stay and long-stay). Their idea is that these two sub-groups
 have different length of stay patterns, and the length of stay of any patient
@@
 identify which input factors contributed to which component (short or long
-stay) and derive useful insight for hospitals.
+stay) and thereby derive a useful insight for hospitals.
 
@@
 understandable and useful to the data owner \citep{Hand2001}. Compared to
-traditional statistical, it is a relatively recent field that emerged in the
-1990s, and has
+traditional statistical techniques, it is a relatively recent field, having
+emerged in the 1990s, and has
 given researchers a variety of new tools to analyse and understand biomedical
@@
 treatment of the techniques and their applications. Here we will describe the
-major methods in the context of their application to predicting hospital length
-of stay.
+major methods in the context of their applicability to the problem of
+predicting length of hospital stay.
 
@@
 \begin{figure}[h]
-\label{fig:svm-example}
 \centering
@@
 support vector machine classifier.}
+\label{fig:svm-example}
 \end{figure}

### 2014-11-03 thesis/method/classification.tex
Add SVM diagram
Add MLP diagram
Fix strange scaling of SVM, MLP and feature selection diagrams
Lots of changes

@@
 \begin{equation*}
-\mathrm{gain}(feature) = \mathrm{info}(\text{child node}) - \mathrm{info}(\text{parent node})
+\mathrm{gain}(feature) = \mathrm{info}(\text{split}) - \mathrm{info}(\text{parent})
 \end{equation*}
@@
 This gives the SVM the best generalising ability to classify new examples.
+\begin{figure}[h]
+\includegraphics{images/method/svm-example.pdf}
+\caption{An example of a 2-dimensional SVM. The solid line indicates the
+maximum margin hyperplane.}
+\label{fig:svm}
+\end{figure}
 
@@
 cannot find a hyperplane until it finds a suitable one.
-
-\begin{figure}[h]
-\caption{An example of a 2-dimensional SVM.}
-\label{fig:svm}
-\centering
-%\includegraphics[scale=0.95]{images/a2-nn}
-\end{figure}
 
@@
 two classes the new example belongs to.
-
-\section{$k$-Nearest Neighbour}
-In this section we will describe the $k$-nearest neighbour ($k$-NN) algorithm
-and discuss some extensions which we use in our work. Additionally, we will
-describe our contribution, which is an extension of the distance metric for
-finding the nearest neighbours.
-
-\subsection{Basic Algorithm}
-The $k$-nearest neighbour classifier is an example of an
-\textit{instance-based} learning algorithm. This is because it does not, unlike
-the other classifiers described in this section, attempt to deduce or
-generalise a relationship between the features and the class: it simply stores
-all training examples and classifies an unseen example by finding the $k$
-``nearest'' examples (or \textit{instances}) and assigning the majority class
-to the new example. In the simplest form of the algorithm, $k$ is chosen to be
-1 and the Euclidean distance is used to measure the closeness of neighbours:
-\begin{equation*}
-\mathrm{Distance} = \sqrt{(x_1-y_1)^2 + (x_2-y_2)^2 + \ldots + (x_n-y_n)^2}
-\end{equation*}
-where $x_i,y_i$ are the values of the $i$-th feature for two different feature
-vectors (or instances).
-
-Like the Na\"{i}ve Bayes classifier, $k$-NN allows all features to be taken
-into account equally when classifying a new example. However, since feature
-values often span different ranges and units of measurement, a particularly
-large range of values for a feature would contribute much more to the distance
-than another feature that spans a lower set of values.
-To avoid this, it is important to \textit{normalise} the
-values of each numeric feature before using $k$-NN: for each feature $i$, we
-calculate the normalised value of the feature for each training example. This
-is done using the following relationship:
-\begin{equation*}
-a_i = s\left(\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right) + t
-\end{equation*}
-where $a_i$ is the normalised value, $v_i$ is the actual value in the data
-set, max $v_i$ and min $v_i$ are taken over all training examples, $s$ is a
-scaling factor $t$ indicates how much to translate the range by. If $s=1$
-and $t=0$, then the range of normalised values will be in $[0,1]$.
-
-\subsection{Existing and Proposed Extensions}
-There have been many extensions proposed to the $k$-NN algorithm described
-above. These usually seek to find a more suitable distance metric for the
-problem domain at hand, or to reduce the storage requirements of the original
-$k$-NN algorithm. We will focus on two extensions to the Euclidean distance:
-K* and our proposed distance function, Ranked Distance.
-
-\subsubsection{K*: Instance-based Classifier with Entropic Distance Function}
-Cleary and Trigg developed K*, an instance-based classifier that uses entropy
-to measure the distances between instances \cite{Cleary1995}. They consider
-the distance between two instances to be the complexity of transforming one
-into another. Define a sequence of transformations on an instance:
-\begin{equation*}
-\bar{t}(a) = t_n(t_{n-1}(\ldots t_1(a)\ldots)), \bar{t} \in T
-\end{equation*}
-where $\bar{t} = t_1,\ldots,t_n$. We can define a probability function $p$
-that gives the probabilities of these transformations occurring, and then we
-can define a probability function that gives the probabilities of all paths
-(transformations) from example $a$ to $b$:
-\begin{equation*}
-P^*(b|a) = \sum_{\bar{t}\in T:\bar{t}(a)=b} p(\bar{t})
-\end{equation*}
-Within this framework, we can define transformations for both real-valued and
-categorical features, which allows K* to work with all types of features.
-The K* function is then defined as:
-\begin{equation*}
-K^*(b|a) = -\mathrm{log}_2P^*(b|a)
-\end{equation*}
-
-To calculate the probability of an instance $a$ being in a class $C$, we sum
-the probabilities from $a$ to each instance of $C$, and repeat for each class.
-The predicted class will be the one with the highest probability:
-\begin{equation*}
-P^*(C|a) = \sum_{b\in C} P^*(b|a)
-\end{equation*}
-
-We use K* as a comparison to the basic NN algorithm and also as a comparison
-to $k$-NN with our distance function.
-
-\subsubsection{Proposed Approach: Ranked Distance}
-\label{sec:rankeddist}
-\paragraph{Intuition}
-As we mentioned earlier, all features are considered equally in a $k$-NN
-classifier, which implies that all features are equally important in predicting
-the class. In practice, however, this is not always the case: to predict the
-LOS of a trauma patient, the severity of their injury should affect the final
-LOS more than whether or not they can speak English (which could be just two
-of many features that are recorded about a patient). Our contribution is thus
-a modified distance function to take the relative importance of attributes into
-account.
-
-\paragraph{Mathematical formulation}
-Given two instances $\mathbf{x} = (x_1,x_2,\ldots,x_n)$ and
-$\mathbf{y} = (y_1,y_2,\ldots,y_n)$, the distance between them is given by:
-\begin{equation*}
-\mathrm{Distance} = \sum_{i=1}^n w_i |x_i-y_i|
-\end{equation*}
-where the $w_i$s weight the contribution of the $i$-th feature to the overall
-distance that is computed by the $k$-NN algorithm. We assume that all feature
-values have been normalised, as described above.
-
-\paragraph{Assignment of weights}
-The weights $w_i$ can be tuned to match the particular problem that is being
-investigated. For our LOS classification problem, we will consider two ways
-of weighting features:
-\begin{enumerate}
-\item The weight of each feature is the magnitude of the correlation
-coefficient
-between it and the class, meaning that features that are more highly correlated
-with the class will contribute more to the distance. The correlation
-coefficient $r_i$ for feature $i$ with the class $c$ can be computed by:
-\begin{equation*}
-r_i = \dfrac{\sum_{i=1}^m (x_i-\bar{x})(c-\bar{c})}{\sqrt{(\sum_{i=1}^m x_i-\bar{x})(\sum_{i=1}^m c-\bar{c})}}
-\end{equation*}
-where $m$ is the number of training examples, $x_i$ is the value of feature
-$i$, and $\bar{x}$ and $\bar{c}$ are the arithmetic averages of the feature
-values and the class values respectively. 
-\item We compute the correlation coefficients as described in the point above.
-This gives us a ranking of the importance of each feature to the prediction of
-the class, with 1 being the highest rank (indicating the most importance) and
-$n$ being the lowest rank (indicating the least importance).
-Instead of using the values of the correlation directly to weight
-the distance contribution of each feature, we specify a function:
-\begin{equation*}
-f : \mathrm{Rank} \rightarrow \mathbb{R}, \mathrm{Rank} \in \{1,2,\ldots,n\}
-\end{equation*}
-that describes
-how the weights of the features vary with the rank. This function should
-inituitively be non-increasing and should produce lower values when the rank
-number is greater, meaning that less important features have a lower weight
-in the distance calculation.
-Note that choosing $f(\mathrm{Rank}) = k$ for any non-zero constant $k$
-results in all features contributing equally to the distance calculation.
-
-\noindent We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}^c}$, where
-$c$ varies from 0 to 1.
-\end{enumerate}
 
@@
 \begin{figure}[h]
-\caption{An MLP with one hidden layer.}
+\centering
+\includegraphics{images/method/mlp.pdf}
+\caption{An MLP with one hidden layer. There are $n$ input nodes, $k$ nodes
+in the hidden layer, and one output node.}
 \label{fig:mlp}
-\centering
-%\includegraphics[scale=0.95]{images/a2-nn}
 \end{figure}
@@
 
+\section{$k$-Nearest Neighbour}
+In this section we will describe the $k$-nearest neighbour ($k$-NN) algorithm
+and discuss some extensions which we use in our work. Additionally, we will
+describe our contribution, which is an extension of the distance metric for
+finding the nearest neighbours.
+
+\subsection{Basic Algorithm}
+The $k$-nearest neighbour classifier is an example of an
+\textit{instance-based} learning algorithm. This is because it does not, unlike
+the other classifiers described in this section, attempt to deduce or
+generalise a relationship between the features and the class: it simply stores
+all training examples and classifies an unseen example by finding the $k$
+``nearest'' examples (or \textit{instances}) and assigning the majority class
+to the new example. In the simplest form of the algorithm, $k$ is chosen to be
+1 and the Euclidean distance is used to measure the closeness of neighbours:
+\begin{equation*}
+\mathrm{Distance} = \sqrt{(x_1-y_1)^2 + (x_2-y_2)^2 + \ldots + (x_n-y_n)^2}
+\end{equation*}
+where $x_i,y_i$ are the values of the $i$-th feature for two different feature
+vectors (or instances).
+
+Like the Na\"{i}ve Bayes classifier, $k$-NN allows all features to be taken
+into account equally when classifying a new example. However, since feature
+values often span different ranges and units of measurement, a particularly
+large range of values for a feature would contribute much more to the distance
+than another feature that spans a lower set of values.
+To avoid this, it is important to \textit{normalise} the
+values of each numeric feature before using $k$-NN: for each feature $i$, we
+calculate the normalised value of the feature for each training example. This
+is done using the following relationship:
+\begin{equation*}
+a_i = s\left(\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right) + t
+\end{equation*}
+where $a_i$ is the normalised value, $v_i$ is the actual value in the data
+set, max $v_i$ and min $v_i$ are taken over all training examples, $s$ is a
+scaling factor and $t$ indicates how much to translate the range by. If $s=1$
+and $t=0$, then the range of normalised values will be in $[0,1]$.
+
+\subsection{Existing and Proposed Extensions}
+There have been many extensions proposed to the $k$-NN algorithm described
+above. These usually seek to find a more suitable distance metric for the
+problem domain at hand, or to reduce the storage requirements of the original
+$k$-NN algorithm. We will focus on two extensions to the Euclidean distance:
+K* and our proposed algorithm, Ranked Distance.
+
+\subsubsection{K*: Instance-based Classifier with Entropic Distance Function}
+Cleary and Trigg developed K*, an instance-based classifier that uses entropy
+to measure the distances between instances \cite{Cleary1995}. They consider
+the distance between two instances to be the complexity of transforming one
+into another. Define a sequence of transformations on an instance:
+\begin{equation*}
+\bar{t}(a) = t_n(t_{n-1}(\ldots t_1(a)\ldots)), \bar{t} \in T
+\end{equation*}
+where $\bar{t} = t_1,\ldots,t_n$. We can define a probability function $p$
+that gives the probabilities of these transformations occurring, and then we
+can define a probability function that gives the probabilities of all paths
+(transformations) from example $a$ to $b$:
+\begin{equation*}
+P^*(b|a) = \sum_{\bar{t}\in T:\bar{t}(a)=b} p(\bar{t})
+\end{equation*}
+Within this framework, we can define transformations for both real-valued and
+categorical features, which allows K* to work with all types of features.
+The K* function is then defined as:
+\begin{equation*}
+K^*(b|a) = -\mathrm{log}_2P^*(b|a)
+\end{equation*}
+
+To calculate the probability of an instance $a$ being in a class $C$, we sum
+the probabilities from $a$ to each instance of $C$, and repeat for each class.
+The predicted class will be the one with the highest probability:
+\begin{equation*}
+P^*(C|a) = \sum_{b\in C} P^*(b|a)
+\end{equation*}
+
+We use K* as a comparison to the basic NN algorithm and also as a comparison
+to our proposed Ranked Distance Nearest Neighbours algorithm.
+
+\subsubsection{Proposed Approach: Ranked Distance Nearest Neighour}
+\label{sec:rankeddist}
+\paragraph{Intuition}
+As we mentioned earlier, all features are considered equally in a $k$-NN
+classifier, which implies that all features are equally important in predicting
+the class. In practice, however, this is not always the case: to predict the
+LOS of a trauma patient, the severity of their injury should affect the final
+LOS more than whether or not they can speak English (which could be just two
+of many features that are recorded about a patient). Our contribution is a
+new NN algorithm that uses a modified distance function which takes into
+account the relative importance of all features called Ranked Distance
+Nearest Neighbour (RD).
+
+\paragraph{Mathematical formulation}
+Given two instances $\mathbf{x} = (x_1,x_2,\ldots,x_n)$ and
+$\mathbf{y} = (y_1,y_2,\ldots,y_n)$, the distance between them is given by:
+\begin{equation*}
+\mathrm{Distance} = \sum_{i=1}^n w_i |x_i-y_i|
+\end{equation*}
+where the $w_i$s weight the contribution of the $i$-th feature to the overall
+distance that is computed by the $k$-NN algorithm. We assume that all feature
+values have been normalised, as described above.
+
+\paragraph{Assignment of weights}
+The weights $w_i$ can be tuned to match the particular problem that is being
+investigated. For our LOS classification problem, we will consider one way
+of weighting features.
+First, we compute the correlation coefficients using:
+\begin{equation*}
+r_i = \dfrac{\sum_{i=1}^m (x_i-\bar{x})(c-\bar{c})}{\sqrt{(\sum_{i=1}^m x_i-\bar{x})(\sum_{i=1}^m c-\bar{c})}}
+\end{equation*}
+where $m$ is the number of training examples, $x_i$ is the value of feature
+$i$, and $\bar{x}$ and $\bar{c}$ are the arithmetic averages of the feature
+values and the class values respectively. 
+This gives us a ranking of the importance of each feature to the prediction of
+the class, with 1 being the highest rank (indicating the most importance) and
+$n$ being the lowest rank (indicating the least importance).
+Instead of using the values of the correlation directly to weight
+the distance contribution of each feature, we specify a function:
+\begin{equation*}
+f : \mathrm{Rank} \rightarrow \mathbb{R}, \mathrm{Rank} \in \{1,2,\ldots,n\}
+\end{equation*}
+that describes
+how the weights of the features vary with the rank. This function should
+inituitively be non-increasing and should produce lower values when the rank
+number is greater, meaning that less important features have a lower weight
+in the distance calculation.
+Note that choosing $f(\mathrm{Rank}) = k$ for any non-zero constant $k$
+results in all features contributing equally to the distance calculation.
+We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}^c}$, where
+$c$ varies from 0 to 1.
+
 \section{Summary}
@@
 Zero-R and 1R to Na\"{i}ve Bayes, decision trees, logistic regression,
-$k$-nearest neighbour (and extensions K* and Ranked Distance), to the more
-sophisticated support vector machines and multi-layer perceptrons. The reason
+$k$-nearest neighbour (and K*), to the more
+sophisticated support vector machines and multi-layer perceptrons. We also
+described our proposed new algorithm, Ranked Distance Nearest Neighbours.
+The reason
 for our inclusion of so many algorithms is two-fold. Firstly, many of

### 2014-11-03 thesis/method/evaluation.tex
Lots of changes

@@
 \item For each data set, we evaluated the effect of the 4 automatic feature
-selection methods on each of the 8 classifiers described in Chapter
-\ref{chap:classification}, giving us 32 configurations for each data set, over
+selection methods on each of the 11 classifiers described in Chapter
+\ref{chap:classification}, giving us 44 configurations for each data set, over
 4 sets. Evaluation was performed using \textit{ten-fold stratified
@@
 Leading and trailing whitespace was removed and feature names transformed using
-code written in Python to manipulate the CSV exported from Microsoft Excel.
-All runs of WEKA were executed by
+code written in Python (language version 3.4.1) to manipulate the CSV exported
+from Microsoft Excel. All runs of WEKA were executed by
 a program we wrote in Bash (a Unix shell) script. This program was able to run
@@
 order to use the new class, as well as building the entire WEKA source. The
-development environment we used was Eclipse \todo{insert version}, and the
-source was built using the 1.7.0 Java runtime.
+development environment we used was Eclipse 4.2.1, and the
+source was built using version 1.7.0 Update 17 of the Java Development Kit.
 
@@
 \includegraphics{images/method/example-roc.eps}
-\caption{Two ROC curves on the same graph. Note that the straight line from
-the bottom left corner to the top right corner indicates an AUC of 0.5. Curves
-which head towards the top left corner indicate increasing AUC.}
+\caption{Three ROC curves on the same graph. Curves
+which head towards the top left corner indicate increasing AUC. In this
+example, the curve labelled `No feature selection' has the highest AUC.}
 \label{fig:sampleroc}
@@
 $\bar{x}$
-is approximately normally distributed when $n$ is at least roughly around 30
+is approximately normally distributed when $n$ is at least approximately 30
 and the $X_i$'s are independent and identically distributed \cite{Phipps2001}.
@@
 \begin{equation*}
-  \left[\bar{x} - z_{1-\alpha/2}\dfrac{s}{\sqrt{n}},
+  \left[\bar{x} - z_{1-\alpha/2}\dfrac{s}{\sqrt{n}},\text{ }
     \bar{x} + z_{1-\alpha/2}\dfrac{s}{\sqrt{n}}\right]
@@
 
-These confidence intervals can be constructed
-
+These confidence intervals can be constructed by computing the average and
+standard deviation of performance estimates of at least three ten-fold
+cross-validation runs per classifier.
 In our work, we repeated the ten-fold stratified cross-validation procedure
@@
 \subsection{Tests for Statistical Significance}
-To decide whether or not the difference in values of each measure between
-classifiers and feature selection methods was statistically significant,
-paired $t$-tests at a significance level of 0.05 were conducted.
-\todo{elaborate}
+We conducted paired $t$-tests in order to determine whether or not our
+improvements in performance were statistically significant at the 95\% level.
+This is done by comparing the results from ten-fold cross-validation, which
+are means. We are also able to take advantage of the paired test because we
+can use the same way of creating the cross-validation folds for different
+classifiers, giving us a more accurate result.
 
@@
 we obtained these figures using stratified ten-fold cross-validation. Paired
-$t$-tests were conducted to test for statistical significance.
+$t$-tests were conducted to test for statistical significance of improvements
+in accuracy and AUC.

### 2014-11-03 thesis/method/feature-descriptions.tex
Lots of changes

\renewcommand{\arraystretch}{0.7}
\begin{longtable}{lp{13cm}}
\textbf{Name} & \textbf{Description} \\ 
\hline
id & The id of the current record (numeric) \\ 
age & Age of patient in years (numeric) \\ 
sex & Patient gender (nominal) \\ 
bp & Systolic blood pressure in mmHg (numeric) \\ 
bp1 & Whether or not the patient had a normal systolic blood pressure between 90-180 mmHg (nominal) \\ 
pr1 & Whether or not the patient had a normal pulse rate between 50-110 BPM (nominal) \\ 
rr1 & Whether or not the patient had a normal respiratory rate of 10-24 breaths per minute (nominal) \\ 
normalvitals & Whether or not the patient had bp, rr and pr normal (nominal) \\ 
pr & Pulse rate in BPM (numeric) \\ 
rr & Respiratory rate in breaths per minute (numeric) \\ 
gcs1 & Whether or not the patient had a Glasgow Coma Score of 14 or 15 (nominal) \\ 
gcs & Glasgow Coma Score on arrival (numeric) \\ 
iculos & Length of ICU stay in days (numeric) \\ 
icu & Whether or not the patient stayed in the ICU (nominal) \\ 
outcome & Whether the patient survived or died (nominal) \\ 
disp\_from\_ed & Disposition from hospital (nominal) \\ 
disp\_from\_hosp & Discharge destination from hospital (nominal) \\ 
iss8 & Whether or not the patient had an Injury Severity Score 8 or less (nominal) \\ 
iss & Injury Severity Score (numeric) \\ 
mechanism & Mechanism of injury (nominal) \\ 
male & Whether or not the patient was male (nominal) \\ 
age65 & Whether or not the patient was older than 65 years (nominal) \\ 
arrivalmode & How the patient arrived at the hospital (nominal) \\ 
transfer & Whether or not the patient was transferred from another hospital (nominal) \\ 
traumatier & The level of trauma team activation (nominal) \\ 
traumateam & Whether or not the trauma team was required (nominal) \\ 
penetrating & Whether or not the injury was penetrating (nominal) \\ 
died & Whether or not the patient died in hospital (nominal) \\ 
eddc & Patient was discharged within expected time (nominal) \\ 
rehab & Whether or not the patient was discharged from hospital to rehabiliation (nominal) \\ 
mechcode & Category of injury mechanism (nominal) \\ 
fall & Mechanism of injury was a fall (nominal) \\ 
headany & Whether or not there was any head injury (nominal) \\ 
faceany & Whether or not there was any face injury (nominal) \\ 
headface & Presence of any injury to head or face (nominal) \\ 
neckany & Any injury to neck (nominal) \\ 
chestany & Any injury to chest (nominal) \\ 
abdoany & Any injury to abdomen (nominal) \\ 
spineany & Any injury to spine (nominal) \\ 
upperlimbany & Any injury to upper limb (nominal) \\ 
lowerlimbany & Any injury to lower limb (nominal) \\ 
bodyregions & Number of bodyregions injured (numeric) \\ 
externalany & Any external burns injury (nominal) \\ 
head3 & Any major head injury (nominal) \\ 
face3 & Any major face injury (nominal) \\ 
neck3 & Any major neck injury (nominal) \\ 
chest3 & Any major chest injury (nominal) \\ 
abdo3 & Any major abdominal injury (nominal) \\ 
spine3 & Any major spine injury (nominal) \\ 
upper3 & Any major upper limb injury (nominal) \\ 
lower3 & Any major lower limb injury (nominal) \\ 
ext3 & Any major external injury (nominal) \\ 
laparotomy & Whether a laparotomy was performed (nominal) \\ 
thoracotomy & Whether a thoracotomy was performed (nominal) \\ 
neurosurgery & Whether any neurosurgery was performed (nominal) \\ 
anypelvis & Any injury to pelvic region (nominal) \\ 
pelvic\_fracture & Whether or not there was any pelvic fracture (nominal) \\ 
upper\_limbop & Whether an operation was performed on the upper limb (nominal) \\ 
lower\_limbop & Whether an operation was performed on the lower limb (nominal) \\ 
married & Whether or not the patient is married (nominal) \\ 
english & Whether or not the patient speaks English (nominal) \\ 
charlson & Charlson Comorbidity Index (numeric) \\ 
triss & Calculated probability of survival (numeric) \\ 
mentalhealth & Whether the patient suffered any mental health problems (nominal) \\ 
alcohol & Whether alcohol was involved in the injury (nominal) \\ 
comorbidity & Whether or not the patient had comorbidities (nominal) \\ 
operation & Whether or not an operation was performed in the hospital (nominal) \\ 
upperlimboponly & An operation was performed only on the upper limb (nominal) \\ 
upperlimbonly & Any injury to the upper limb only (nominal) \\ 
headminor & Minor or no head injury (nominal) \\ 
lowerlimbnopelvis & Any injury to the lower limb, excluding pelvic injuries (nominal) \\ 
any\_heart & Any heart disease (nominal) \\ 
any\_liver & Any liver disease (nominal) \\ 
any\_kidney & Any kidney disease (nominal) \\ 
any\_cancer & Whether the patient has or had cancer (nominal) \\ 
any\_dm & Any diabetes mellitus (nominal) \\ 
any\_stroketia & Whether or not the patient had suffered a stroke or transient ischemic attack (nominal) \\ 
los & Length of hospital stay in days (numeric) \\ 
los48 & Whether or not the length of hospital stay was less than or equal to 2 days (nominal) \\ 
\caption{A listing of all features from the trauma data set and what they represent.}
\label{tab:feature-desc-tr}
\end{longtable}
\renewcommand{\arraystretch}{1}

\begin{table}[htbp]
\begin{tabular}{ll}
\textbf{Name} & \textbf{Description} \\ 
\hline
Sex & Patient gender (nominal) \\ 
Age & Age at the time of admission, separated into categories (nominal) \\ 
Education & Highest level of education completed (nominal) \\ 
Marital Status & Marital status (nominal) \\ 
Episode Type & Patient type of episode (nominal) \\ 
Inpatient Service & Physical inpatient service (nominal) \\ 
Medical Specialty & Medical specialty associated with patient condition (nominal) \\ 
Origin Episode Type & Origin episode type that caused hospitalisation (nominal) \\ 
Admission Month & Month admitted to hospital (nominal) \\ 
Admission Day & Day of week admitted to hospital (nominal) \\ 
Admission Hour & Hour of the day patient was admitted to hospital (nominal) \\ 
Main Procedure & Type of main procedure performed (nominal) \\ 
Main Diagnosis & Main diagnosis of patient (nominal) \\ 
Previous Admissions & Number of times patient was previously admitted (numeric) \\ 
LOS48 & Whether or not the length of hospital stay was less than or equal to 2 days (nominal) \\ 
\end{tabular}
\caption{A listing of all features from the general hospital data set and what they represent.}
\label{tab:feature-desc-pt}
\end{table}

### 2014-11-03 thesis/method/preprocess.tex
Lots of changes

@@
 In our work we obtained two independent sets of LOS data, one specifically for
-trauma and the other a general, hospital-wide set.
+trauma and the other a general, hospital-wide data set.
 
-\subsection{Trauma LOS Data}
+\subsection{Trauma}
 \subsubsection{Collection}
@@
 \subsubsection{Features}
+The features present in the original data set are listed and briefly described
+in Table \ref{tab:feature-desc-tr}.
 
-\subsection{General Hospital LOS Data}
+\subsection{General Hospital}
 By collaborating with our colleagues from the University of Minho in Portugal,
@@
 2000--2013 at the Hospital das Foras Armadas in Lisboa, Portugal. The data
-contained 26462 inpatient records spanning all medical departments within the
+contained 17547 inpatient records spanning all medical departments within the
 hospital, comprising 15 features that were not specific to any medical domain.
+One of these features was the class, \texttt{LOS48}, indicating a stay of
+less than or equal to 2 days.
 
 \subsubsection{Features}
+The features present in the original data set are listed and briefly described
+in Table \ref{tab:feature-desc-pt}.
+
+\input{method/feature-descriptions.tex}
 
@@
 
+\subsection{Trauma Data}
 \paragraph{Incorrect spelling and punctuation.}
@@
 fields in the data set contained leading or trailing whitespace, which was
-removed with a few lines of Python code (see appendix \todo{[add reference
-to appendix]}). All attribute names were converted to lowercase. In the case
+removed with a few lines of Python code (see Appendix
+\ref{sec:python-preprocess}).
+All attribute names were converted to lowercase. In the case
 where attribute names consisted of several words, the spaces were replaced
@@
 distinguished from the numeric.
+
+\subsection{General Hospital Data}
+The general hospital data we received had already been preprocessed. The only
+preprocessing here was to
+undo the logarithm transform that had been applied to the \texttt{Previous
+Admissions} feature.
 
@@
 sub-interval. The decision rule is then to split if the information gain for
-that split exceeds the MDL value: $\mathrm{gain} > \mathrm{MDL}$.
+that split exceeds the MDL value.
 
@@
 The data was discretised according to the above method using the supervised
-discretisation filter in WEKA. We keep both the original data set (without
+discretisation filter in WEKA. We kept both the original data set (without
 discretisation) and the discretised data set for comparison.

### 2014-11-03 thesis/method/selection.tex
Add filter vs wrapper diagram
Fix strange scaling of SVM, MLP and feature selection diagrams
A whole bunch of updates including source code listing in the appendix
Lots of changes

@@
 
-To evaluate the ``goodness'' of a subset of features with the class $C$, we
-compute:
+To evaluate the ``goodness'' or \textit{merit} of a subset of features with
+the class $C$, we compute:
 \begin{equation}
@@
 becomes $n$ and the denominator $\sqrt{n^2}$ if the subset has $n$ features.
-\todo{elaborate a bit more}
+
+CFS conducts a search through subsets of all features in order to find the
+one feature subset that has the best merit.
 
@@
 some subset of these either by selecting a fixed number, or selecting all
-features that have a correlation above a specified threshold.
+features that have a correlation above a specified threshold. We start with a
+threshold of 0 and gradually increase this by larger amounts. This is because
+a large number of features are usually discarded in the lower thresholds, and
+we would like to investigate a range of resulting subset sizes.
 
@@
 We can compute the IG for each feature and then select subsets of features that
-are greater than some threshold we specify.
+are greater than some threshold we specify, similarly to the correlation
+coefficient method of feature selection. We choose a similar set of thresholds
+because of the high numbers of features with very low information gain.
 
@@
 We can also select a fixed number of features or all features above a certain
-1R accuracy threshold.
+1R accuracy threshold. This threshold is a slightly different quantity to the
+threshold of the correlation coefficient and information gain methods, as it
+is an accuracy, expressed in percentages. Our first threshold is the Zero-R
+accuracy for the trauma data set, which is the lower of the two Zero-R
+accuracies on our two data sets. We increase this percentage by in increasing
+increments to obtain a range of thresholds up to nearly 97\% accuracy (which
+discards all features for both data sets). Using the same thresholds across
+both data sets allows us to make valid comparisons of the effect of feature
+selection on performance.
 
@@
 \begin{figure}[h]
+\centering
+\includegraphics[scale=0.9]{images/method/feature-selection.pdf}
 \caption{Filter and wrapper approaches to feature selection. Note the role of
@@
 \label{fig:wrapper}
-\centering
-%\includegraphics[scale=0.95]{images/a2-nn}
 \end{figure}
@@
 divide and conquer nature of the learning algorithm, and the tendency of the
-trees to select fewer features for learning than other algorithms.
+trees to select fewer features for learning than other algorithms. Like CFS,
+wrapper methods search for the best subset of features and do not require a
+threshold to discard attributes.
 

### 2014-11-03 thesis/results/features-pt-acc.tex
Lots of changes

@@
 \begin{table}[htbp]
-\caption{Comparison of accuracy (\%) between all feature selection methods for each classifier with and without discretisation, on the general hospital data set.
-The best accuracy for each feature selection method is highlighted in bold, and the best accuracy for each classifier is italicised.}
-\label{tab:features-pt-acc}
 \resizebox{!}{11cm}{%
-\begin{tabular}{|l|cccccc|}
-\hline
+\begin{tabular}{l|cccccc}
  & None & CFS & C4.5 Wrapper & Correlation & IG & 1R \\
@@
  Discretised & 98.14 & 96.45 & 97.35 & \textit{\begin{tabular}[c]{@{}c@{}}98.22\\ t=0.05\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}98.22\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}98.14\\ t=74.405\end{tabular} \\
- Not discretised & 98.16 & 96.45 & 97.35 & \textit{\begin{tabular}[c]{@{}c@{}}98.22\\ t=0.05\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}98.22\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}98.16\\ t=74.405\end{tabular} \\ \hline
+ Not discretised & 98.16 & 96.45 & 97.35 & \textit{\begin{tabular}[c]{@{}c@{}}98.22\\ t=0.05\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}98.22\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}98.16\\ t=74.405\end{tabular} \\
  \end{tabular}
 }
+\caption{Comparison of accuracy (\%) between all feature selection methods for each classifier with and without discretisation, on the general hospital data set.
+The best accuracy for each feature selection method is highlighted in bold, and the best accuracy for each classifier is italicised.}
+\label{tab:features-pt-acc}
 \end{table}

### 2014-11-03 thesis/results/features-pt-auc.tex
Lots of changes

@@
 \begin{table}[htbp]
-\caption{Comparison of AUC between all feature selection methods for each classifier with and without discretisation, on the general hospital data set.
-The best AUC for each feature selection method is highlighted in bold, and the best AUC for each classifier is italicised.}
-\label{tab:features-pt-auc}
 \resizebox{!}{11cm}{%
-\begin{tabular}{|l|cccccc|}
- \hline
+\begin{tabular}{l|cccccc}
  & None & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
@@
  Discretised & 0.99 & \textbf{0.983} & \textbf{0.992} & \textit{\begin{tabular}[c]{@{}c@{}}0.993\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.991\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.990\\ t=76.905\end{tabular} \\
- Not discretised & 0.99 & \textbf{0.983} & \textbf{0.992} & \textit{\begin{tabular}[c]{@{}c@{}}0.993\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.991\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.990\\ t=76.905\end{tabular} \\ \hline
+ Not discretised & 0.99 & \textbf{0.983} & \textbf{0.992} & \textit{\begin{tabular}[c]{@{}c@{}}0.993\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.991\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.990\\ t=76.905\end{tabular} \\
  \end{tabular}
 }
+\caption{Comparison of AUC between all feature selection methods for each classifier with and without discretisation, on the general hospital data set.
+The best AUC for each feature selection method is highlighted in bold, and the best AUC for each classifier is italicised.}
+\label{tab:features-pt-auc}
 \end{table}

### 2014-11-03 thesis/results/features-tr-acc.tex
Lots of changes

@@
 \begin{table}[htbp]
-\caption{Comparison of accuracy (\%) between all feature selection methods for each classifier with and without discretisation, on the trauma data set.
-The best accuracy for each feature selection method is highlighted in bold, and the best accuracy for each classifier is italicised.}
-\label{tab:features-tr-acc}
 \resizebox{!}{11cm}{%
-\begin{tabular}{|l|cccccccc|}
-\hline
+\begin{tabular}{l|cccccccc}
  & \multicolumn{1}{l}{None} & Baseline & Expert & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
@@
 Discretised & 73.79 & 73.03 & 73.45 & 73.38 & \textit{76.66} & \begin{tabular}[c]{@{}c@{}}74.41\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.03\\ t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.94\\ t=66.905\end{tabular} \\
-Not discretised & 71.87 & 73.03 & 72.31 & 73.04 & 73.35 & \textit{\begin{tabular}[c]{@{}c@{}}74.65\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}72.63\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}70.99\\ t=66.905\end{tabular} \\ \hline
+Not discretised & 71.87 & 73.03 & 72.31 & 73.04 & 73.35 & \textit{\begin{tabular}[c]{@{}c@{}}74.65\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}72.63\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}70.99\\ t=66.905\end{tabular} \\
  \end{tabular}
 }
+\caption{Comparison of accuracy (\%) between all feature selection methods for each classifier with and without discretisation, on the trauma data set.
+The best accuracy for each feature selection method is highlighted in bold, and the best accuracy for each classifier is italicised.}
+\label{tab:features-tr-acc}
 \end{table}

### 2014-11-03 thesis/results/features-tr-auc.tex
Lots of changes

@@
 \begin{table}[htbp]
-\caption{Comparison of AUC between all feature selection methods for each classifier with and without discretisation, on the trauma data set.
-The best AUC for each feature selection method is highlighted in bold, and the best AUC for each classifier is italicised.}
-\label{tab:features-tr-auc}
 \resizebox{!}{11cm}{%
-\begin{tabular}{|l|cccccccc|}
- \hline
+\begin{tabular}{l|cccccccc}
  & None & Baseline & Expert & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
@@
 Discretised & 0.803 & 0.781 & 0.784 & 0.802 & \textit{0.825} & \begin{tabular}[c]{@{}c@{}}0.807\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.803\\ t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.772\\ t=68.905\end{tabular} \\
-Not discretised & 0.78 & 0.781 & 0.774 & 0.799 & 0.804 & \textit{\begin{tabular}[c]{@{}c@{}}0.818\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.788\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.752\\ t=68.905\end{tabular} \\ \hline
+Not discretised & 0.78 & 0.781 & 0.774 & 0.799 & 0.804 & \textit{\begin{tabular}[c]{@{}c@{}}0.818\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.788\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.752\\ t=68.905\end{tabular} \\
 \end{tabular}
 }
+\caption{Comparison of AUC between all feature selection methods for each classifier with and without discretisation, on the trauma data set.
+The best AUC for each feature selection method is highlighted in bold, and the best AUC for each classifier is italicised.}
+\label{tab:features-tr-auc}
 \end{table}

### 2014-11-03 thesis/results/overall-portugal-acc.tex
Fix broken table and figure labels in latex
Lots of changes

@@
 \begin{table}[htbp]
-\label{tab:overall-pt-acc}
-\caption{Overall best accuracy of each classifier with a particular feature selection method on the general hospital data set.
-* indicates that results were identical regardless of discretisation.
-D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.
-The absence of * or D means that not discretising produced a better result.
-The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
-\begin{tabular}{|l|cccccc|}
-\hline
+\begin{tabular}{l|cccccc}
 Classifier & None & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
@@
 K* & 97.36 & 92.94* & 96.53* & \textbf{\begin{tabular}[c]{@{}c@{}}97.46*\\ t=0.05\end{tabular}} & \begin{tabular}[c]{@{}c@{}}97.46*\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}97.36\\ t=74.405\end{tabular} \\
-MLP & 98.16 & 96.45* & 97.35* & \textbf{\begin{tabular}[c]{@{}c@{}}98.22*\\ t=0.05\end{tabular}} & \begin{tabular}[c]{@{}c@{}}98.22*\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}98.16\\ t=74.405\end{tabular} \\ \hline
+MLP & 98.16 & 96.45* & 97.35* & \textbf{\begin{tabular}[c]{@{}c@{}}98.22*\\ t=0.05\end{tabular}} & \begin{tabular}[c]{@{}c@{}}98.22*\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}98.16\\ t=74.405\end{tabular} \\
 \end{tabular}
+\caption{Overall best accuracy of each classifier with a particular feature selection method on the general hospital data set.
+* indicates that results were identical regardless of discretisation.
+D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.
+The absence of * or D means that not discretising produced a better result.
+The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
+\label{tab:overall-pt-acc}
 \end{table}

### 2014-11-03 thesis/results/overall-portugal-auc.tex
Fix broken table and figure labels in latex
Lots of changes

@@
 \begin{table}[htbp]
-\label{tab:overall-pt-auc}
-\caption{Overall best AUC of each classifier with a particular feature selection method on the general hospital data set.
-* indicates that results were identical regardless of discretisation.
-D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.
-The absence of * or D means that not discretising produced a better result.
-The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
-\begin{tabular}{|l|cccccc|}
-\hline
+\begin{tabular}{l|cccccc}
 Classifier & None & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
@@
 K* & 0.994 & 0.980* & 0.989* & \begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.05\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.994\\ t=74.405\end{tabular} \\
-MLP & 0.990* & 0.983* & 0.992* & \textbf{\begin{tabular}[c]{@{}c@{}}0.993*\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.991*\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.990*\\ t=76.905\end{tabular} \\ \hline
+MLP & 0.990* & 0.983* & 0.992* & \textbf{\begin{tabular}[c]{@{}c@{}}0.993*\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.991*\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.990*\\ t=76.905\end{tabular} \\
 \end{tabular}
+\caption{Overall best AUC of each classifier with a particular feature selection method on the general hospital data set.
+* indicates that results were identical regardless of discretisation.
+D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.
+The absence of * or D means that not discretising produced a better result.
+The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
+\label{tab:overall-pt-auc}
 \end{table}

### 2014-11-03 thesis/results/overall-trauma-acc.tex
Fix broken table and figure labels in latex
Lots of changes

@@
 \begin{table}[htbp]
-\label{tab:overall-tr-acc}
-\caption{Overall best accuracy of each classifier with a particular feature selection method on the trauma data set.
-* indicates that results were identical regardless of discretisation.
-D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.
-The absence of * or D means that not discretising produced a better result.
-The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
 \resizebox{\linewidth}{!}{%
-\begin{tabular}{|l|cccccccc|}
-\hline
+\begin{tabular}{l|cccccccc}
 Classifier & None & Baseline & Expert & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
@@
 K* & \begin{tabular}[c]{@{}c@{}}71.4\\ D\end{tabular} & 72.20* & \begin{tabular}[c]{@{}c@{}}75.08\\ D\end{tabular} & 75.58 & \textbf{\begin{tabular}[c]{@{}c@{}}77.27\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}75.07\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.08\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.45\\ t=68.905\end{tabular} \\
-MLP & \begin{tabular}[c]{@{}c@{}}73.79\\ D\end{tabular} & 73.03* & \begin{tabular}[c]{@{}c@{}}73.45\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.38\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.66\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.65\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.03\\ D, t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.94\\ D, t=66.905\end{tabular} \\ \hline
+MLP & \begin{tabular}[c]{@{}c@{}}73.79\\ D\end{tabular} & 73.03* & \begin{tabular}[c]{@{}c@{}}73.45\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.38\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.66\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.65\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.03\\ D, t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.94\\ D, t=66.905\end{tabular} \\
 \end{tabular}
 }
+\caption{Overall best accuracy of each classifier with a particular feature selection method on the trauma data set.
+* indicates that results were identical regardless of discretisation.
+D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.
+The absence of * or D means that not discretising produced a better result.
+The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
+\label{tab:overall-tr-acc}
 \end{table}

### 2014-11-03 thesis/results/overall-trauma-auc.tex
Fix broken table and figure labels in latex
Lots of changes

@@
 \begin{table}[htbp]
-\label{tab:overall-tr-auc}
-\caption{Overall best AUC of each classifier with a particular feature selection method on the trauma data set.
-* indicates that results were identical regardless of discretisation.
-D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.
-The absence of * or D means that not discretising produced a better result.
-The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
 \resizebox{\textwidth}{!}{%
-\begin{tabular}{|l|cccccccc|}
-\hline
+\begin{tabular}{l|cccccccc}
 Classifier & None & Baseline & Expert & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
@@
 K* & \begin{tabular}[c]{@{}c@{}}0.768\\ D\end{tabular} & 0.774* & \begin{tabular}[c]{@{}c@{}}0.803\\ D\end{tabular} & 0.815 & \textbf{\begin{tabular}[c]{@{}c@{}}0.838\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.82\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ t=68.905\end{tabular} \\
-MLP & \begin{tabular}[c]{@{}c@{}}0.803\\ D\end{tabular} & 0.781* & \begin{tabular}[c]{@{}c@{}}0.784\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.818\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.803\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.772\\ D, t=68.905\end{tabular} \\ \hline
+MLP & \begin{tabular}[c]{@{}c@{}}0.803\\ D\end{tabular} & 0.781* & \begin{tabular}[c]{@{}c@{}}0.784\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.818\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.803\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.772\\ D, t=68.905\end{tabular} \\
 \end{tabular}
 }
+\caption{Overall best AUC of each classifier with a particular feature selection method on the trauma data set.
+* indicates that results were identical regardless of discretisation.
+D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.
+The absence of * or D means that not discretising produced a better result.
+The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
+\label{tab:overall-tr-auc}
 \end{table}

### 2014-11-03 thesis/results/results.tex
Add tables from feature selection
Lots of changes

@@
 \section{Overall Performance}
-\subsection{Trauma Data Set}
+\subsection{Trauma Data}
 Tables \ref{tab:overall-tr-acc} and \ref{tab:overall-tr-auc} show the best
@@
 regression. It is worth noting that our best accuracy, using $k$-NN with 1
-nearest neighbour, improves upon the accuracy that they achieved by 2.75\%
+nearest neighbour, shows a statistically significant improvement upon the accuracy
+that they achieved by 2.75\%
 while using only 11 features (compared to the 19 they used). These 11 features
 make up only 14.1\% of all features in the data set, and performed better than
-the 11 features selected manually by the domain expert.
+the 11 features selected manually by the domain expert. The number of features
+selected by each method for each data set is listed in Tables
+\ref{tab:features-selected-nothreshold} and
+\ref{tab:features-selected-threshold}, and the full details of the selected
+features can be found in Appendices \ref{app:features-tr} and
+\ref{app:features-pt}.
 
 Despite our inclusion of accuracy as an evaluation metric, the AUC is what Dinh
-et al. used to evaluate the ability of their logistic regression classifier to
+et al. \cite{Dinh2013a}
+used to evaluate the ability of their logistic regression classifier to
 discriminate between the two LOS classes. Using their features, the best AUC
@@
 
-\subsection{General Data Set}
+\subsection{General Hospital Data}
 For the general hospital data set, we report the best accuracy and AUC results
@@
 \section{Comparison of Feature Selection Methods}
-\subsection{Trauma Data Set}
+\subsection{Trauma Data}
 Tables \ref{tab:features-tr-acc} and \ref{tab:features-tr-auc} show the
@@
 
-Focussing our attention on the AUC tells a slightly different story. Here we
+\input{results/features-selected-nothreshold.tex}
+\input{results/features-selected-threshold.tex}
+
+Focussing our attention on the AUC reveals slightly different patterns. Here we
 find that the correlation coefficient and the information gain are the feature
@@
 
-\subsection{General Data Set}
+\subsection{General Hospital Data}
 Tables \ref{tab:features-pt-acc} and \ref{tab:features-pt-auc} compare the
@@
 \section{Comparison of Classifiers}
-\subsection{Trauma Data Set}
+\subsection{Trauma Data}
 Figures \ref{fig:tr-nothreshold} and \ref{fig:tr-threshold} compare the
@@
 C4.5 and the various nearest neighbour classifiers ($k$-NN with 1 and 20
-neighbours, Ranked Distance and K*). This is true for both accuracy and AUC.
+neighbours and K*), as well as our Ranked Distance algorithm.
+This is true for both accuracy and AUC.
 
@@
 approaches perform similarly when the number of features is decreased, but our
-Ranked Distance approach performs better than the other NN algorithms. This
+Ranked Distance algorithm showed a statistically significant improvement over
+other NN algorithms. This
 improvement is not apparent at smaller feature sets, where using the standard
@@
 
-\subsection{General Data Set}
+\subsection{General Hospital Data}
 In Figures \ref{fig:pt-nothreshold} and \ref{fig:pt-threshold}, we compare the
@@
 Notice that in Figure \ref{fig:pt-threshold-corr}, both the accuracy and AUC
-of Ranked Distance drop much more than for the other classifiers. This is at
+of Ranked Distance NN drop much more than for the other classifiers. This is at
 a correlation coefficient threshold of 0.2, which reduces the feature set to
 4 features. We noted in the discussion on classifiers for the trauma data set
-that Ranked Distance performed worse for smaller feature sets than for larger
+that Ranked Distance NN performed worse for smaller feature sets than for larger
 ones, and this is another example of this phenomenon. Recall that in Ranked
-Distance, features are first ranked according to some similarity metric (we
+Distance NN, features are first ranked according to some similarity metric (we
 have used the correlation coefficient), and then assigned a weight according
@@
 investigate the cause of the poor performance on smaller feature sets, but
-it is not altogether surprising: Ranked Distance performs best by using the
+it is not altogether surprising: Ranked Distance NN performs best by using the
 weighted distance contributions of all features, and by significantly reducing
@@
 neither of which have been investigated thoroughly in previous work. Our
-proposed Ranked Distance approach was able to improve the accuracy and AUC of
+proposed Ranked Distance algorithm was able to improve upon the accuracy and AUC of
 the standard $k$-NN algorithm with the same number of neighbours, but we found

### 2014-11-03 code/weka/RankedDistance.java

@@ -1,24 +1,17 @@
-/**
- * 
- */
 package weka.core;
 
+import java.util.Collections;
 import java.util.Enumeration;
 import java.util.Vector;
 
 import weka.attributeSelection.ASEvaluation;
 import weka.attributeSelection.ASSearch;
 import weka.attributeSelection.CorrelationAttributeEval;
-import weka.attributeSelection.InfoGainAttributeEval;
 import weka.attributeSelection.Ranker;
 import weka.core.Instance;
 import weka.core.Instances;
 import weka.core.NormalizableDistance;
 import weka.core.RevisionUtils;
-import weka.core.TechnicalInformation;
-import weka.core.TechnicalInformationHandler;
-import weka.core.TechnicalInformation.Field;
-import weka.core.TechnicalInformation.Type;
 import weka.core.neighboursearch.PerformanceStats;
 
 /**
@@ -27,7 +20,7 @@ import weka.core.neighboursearch.PerformanceStats;
  */
 public class RankedDistance
 extends NormalizableDistance
-implements TechnicalInformationHandler {
+implements Cloneable {
 
 	private static final long serialVersionUID = 1068606253458807903L;
 
@@ -36,7 +29,7 @@ implements TechnicalInformationHandler {
 
 	/** Holds the "merit" of the ranked attributes. */
 	protected double[] m_RankedMerits;
-	
+
 	/** Holds the weighting of the ranked attributes used by this distance function */
 	protected double[] m_RankWeights;
 
@@ -45,12 +38,15 @@ implements TechnicalInformationHandler {
 
 	/** The class that evaluates the goodness of the feature set. */
 	protected ASEvaluation m_AttributeEvaluator;
-	
+
 	/** How much to update a weight if we find it important. */
 	protected static double m_WeightUpdateFactor = 0.0001;
-	
+
 	/** The degree of the decay function. */
-	protected static double m_DecayDegree = 1;
+	protected double m_DecayDegree = 1;
+
+	/** Whether or not to use a linear decay function. */
+	protected boolean m_UseLinear = false;
 
 	/**
 	 * 
@@ -73,7 +69,7 @@ implements TechnicalInformationHandler {
 	protected void setDecayDegree(double decayDegree) {
 		m_DecayDegree = decayDegree;
 	}
-	
+
 	/**
 	 * Gets the decay degree.
 	 * @return the decay degree as a double
[…]

### 2014-11-04 thesis/abstract/abstract.tex
Lots of changes
Final version of thesis

@@
 most efficient use of limited hospital resources. However, previous studies
-have used a limited number of statistical techniques such as logistic
+have only used a limited number of statistical techniques such as logistic
 regression in order to predict the length of stay. Also, they have used only
@@
 from a general hospital in Portugal. We propose a new classification algorithm
-called Ranked Distance Nearest Neighbor that considers the
+called Ranked Distance Nearest Neighbour that considers the
 relative importance of each feature when determining the nearest neighbours.
@@
 For the trauma dataset, we were able to achieve an accuracy of 77.81\% with
-1-nearest neighbour and an area under the curve of 0.846 with logistic
+1-nearest neighbour and an area under the receiver operating characteristic
+curve of 0.846 with logistic
 regression. This represents an improvement of 2.75\% and 0.034 respectively

### 2014-11-04 thesis/appendix/code.tex
Final version of thesis

@@
 
-In this appendix we display the source code of the most important scripts we
+In this appendix we display the source code of the most important programs we
 wrote in our to conduct our experiments. These are described in more detail
@@
 \lstset{
-  language = Python,
+  language = Java,
   texcl = false, % Parse comments as TeX
@@
 
+\section{Ranked Distance Implementation}
+\label{sec:rankeddist-code}
+\lstset{language=Java}
+\lstinputlisting{../code/weka/RankedDistance.java}
+
 \section{Data Cleaning and Whitespace Removal}
+\lstset{language=Python}
 \label{sec:python-preprocess}
@@
 \subsection{Computing Confidence Intervals from Summarised Results}
+\label{sec:compute-ci}
 \lstset{language=Python}
 \lstinputlisting{../code/weka/compute_ci.py}
-
-\section{Ranked Distance Implementation}
-\label{sec:rankeddist-code}
-\lstset{language=Java}
-\lstinputlisting{../code/weka/RankedDistance.java}

### 2014-11-04 thesis/conclusion/conclusion.tex
Final version of thesis

@@
 patients. In particular, we introduced a new classification algorithm, Ranked
-Distance Nearest Neighbours, and compared its accuracy and AUC with ten
+Distance Nearest Neighbour, and compared its accuracy and AUC with ten
 other classifiers -- Zero-R, 1R, Na\"{i}ve Bayes, C4.5 decision tree, logistic
@@
 
-In our proposed Ranked Distance Nearest Neighbours algorithm, we only considered
+In our proposed Ranked Distance Nearest Neighbour algorithm, we only considered
 one kind of weight assignment function, but it is not the only way to assign

### 2014-11-04 thesis/introduction/introduction.tex
Final version of thesis

@@
 selection methods have not been applied in LOS prediction to date, and only
-one has been applied in our medical domain.
+one has been applied in our medical domain;
 
 \item We assessed the effect of feature discretisation on classifier
-performance. This has not been studied in previous work.
+performance. This has not been studied in previous work;
 
@@
 showed that it performed better than the standard nearest neighbour algorithm
-in certain situations.
+in certain situations;
 
@@
 Not only is the evaluation of a second data set in one work uncommon in the
-literature, the inclusion of more than one data set allowed us to draw
+literature, its inclusion enabled us to draw
 additional insights about the classifiers and feature selections we used
-which would not have been possible with only one set of data.
+which would not have been possible with only one set of data;
 
@@
 Our findings are presented and discussed in Chapter \ref{chap:results}.
-We conclude in Chapter \ref{chap:conclusion}, with indications for future
-work in this area.
+We conclude in Chapter \ref{chap:conclusion}, with indications for possible
+avenues of future research.

### 2014-11-04 thesis/litreview/litreview.tex
Final version of thesis

@@
 One key measurement of hospital activity, health care utilisation, and hospital
-cost is the patient length of stay \citep{Omachonu2004,Ng2006}. The length of
+cost is the patient length of stay \citep{Ng2006,Omachonu2004}. The length of
 stay has been used in a wide variety of medical domains, such as burns
  \citep{Yang2010}, intensive
-care \citep{Tu1993,Harper2005,Perez2006,Dybowski1996},
+care \citep{Dybowski1996,Harper2005,Perez2006,Tu1993},
 psychiatry \citep{Lowell1997}, and acute pancreatitis \citep{Pofahl1998}, in
@@
 also make the contribution of each independent variable to the final
-prediction explicit. Thus they is easy to interpret when compared to more
+prediction explicit. Thus they are easier to interpret when compared to more
 ``black box'' approaches such as artificial neural networks \citep{Adams2012}.
@@
 survival analysis \citep{Vasilakis2005}, and Markov
-models \citep{Perez2006,Jain1989,Kapadia2000}.
+models \citep{Jain1989,Kapadia2000,Perez2006}.
 
@@
 early applications to length of stay prediction in intensive
-care \citep{Tu1993,Dybowski1996}, in addition to disease diagnosis and mortality
+care \citep{Dybowski1996,Tu1993}, in addition to disease diagnosis and mortality
 prediction \citep{Silva2006}. Artificial neural networks are able to detect
@@
 one domain to another, or to predict a combination of outcomes (not only
-length of stay, but mortality, or rehabilitation requirement). Being able to
+length of stay, but mortality or rehabilitation requirement). Being able to
 do this would mean that hospital staff are better equipped to handle patient
@@
 were initially included in the model and then progressively eliminated if their
-likelihood ratio (LR) test $P$-value was $>0.15$. Features could only be
-re-included if $P<0.05$. For length of stay
+likelihood ratio (LR) test $P$-value was greater than 0.15. Features could
+only be re-included if $P<0.05$. For length of stay
 prediction, this left them with 9 features out of the 15 that they originally
@@
 given researchers a variety of new tools to analyse and understand biomedical
-datasets \citep{Bellazzi2008,Yoo2012}. Bellazzi \& Zupan \citep{Bellazzi2008}
+datasets \citep{Bellazzi2008,Yoo2012}. Bellazzi and Zupan \citep{Bellazzi2008}
 have written a
 comprehensive review paper of data mining techniques in clinical prediction
-settings, and we refer the reader to their article for a more detailed
+settings, and we refer readers to their article for a more detailed
 treatment of the techniques and their applications. Here we will describe the
@@
 One of the earliest applications of neural networks to
-length of stay prediction was by Tu \& Guerriere, who used a neural network
+length of stay prediction was by Tu and Guerriere, who used a neural network
 with the following structure \citep{Tu1993} to predict length of stay in
@@
 \item Neural networks are able to model arbitrarily complex relations between
-independent and dependent variables \citep{Tu1996,Buchman1994}, especially if
+independent and dependent variables \citep{Buchman1994,Tu1996}, especially if
 the relation is of a non-linear nature
 \item Neural networks are able to handle models where the features (independent
-variables) are not completely independent of each other
+variables) are not completely independent of each other \cite{Tu1996}
 \item Neural networks can be ``trained'' in different ways, which changes their
@@
 by `trial and error' ''.}
-and learning rate\footnote{A constant that is set by the
+and learning rate\footnote{A constant set by the
 network developer that governs how much the connection weights should adjust at
-each iteration of the training algorithm.}), as well as the difficulty in
+each iteration of the training algorithm.}), as well as difficulty in
 inducing sensible rules from the model (even by domain
@@
 an improvement from the ``trial and error'' method Tu \& Guerriere used. It is
-interesting to briefly discuss their method even though they were interested
+interesting to briefly discuss their method even though their focus was
 in predicting the outcome of critically ill patients in intensive care, rather
@@
 
-Additionally, Walczak et al. also examined not only backpropagation, which
+Additionally, Walczak et al. examined not only backpropagation, which
 is the training algorithm used by most of the applications, but also radial
@@
 healthcare domain is that the rules it ``learns'' are very simple to extract
-from the resulting tree, unlike the ``black box'' that neural networks commonly
-are. Harper found that CART performed almost as well as neural networks in
+from the resulting tree, unlike the ``black box'' nature of neural network
+predictions.
+Harper found that CART performed almost as well as neural networks in
 predicting intensive care length of stay, and much better in predicting
@@
 possible. In Figure \ref{fig:svm-example} from Yoo et al. \citep{Yoo2012},
-hyperplane 2 is the better plane because its margin is as large as possible. 
+hyperplane 2 is the better plane because its margin is as large as possible,
+meaning that the SVM has the best generalising ability to classify unseen
+examples.
 \begin{figure}[h]
@@
 investigating a number of the best data mining techniques, i.e. using more
-than just neural networks. There is also value in investigating the performance
-of such a model on data sets from different locations, since most studies are
+than just neural networks. An investigation of the performance
+of such a model on data sets from different locations is also warranted, since
+most studies are
 only validated on one or two data sets. Additionally, length of stay prediction

### 2014-11-04 thesis/method/classification.tex
Final version of thesis

@@
 noting down the error rate. 1R then chooses the feature that produces rules
-with the highest total accuracy.
+with the highest total accuracy across all discretised feature values.
 
@@
 probability can be computed by
-counting the frequency with which the value is associated with a class in the
+counting the frequency with which a feature value is associated with a class in the
 training data. After these probabilities are computed, classifying a feature
@@
 belongs in the $i$-th class. Classifying the feature vector involves
-calculating this probability for each for each class and choosing the class
+calculating this probability for each class and choosing the class
 with the highest probability.
@@
 above equation zero. A feature vector that could otherwise have been highly
-likely to belong in class $i$ is now counted as having zero probability of
+likely to belong in class $i$ is then counted as having zero probability of
 belonging in the class. In order to ensure that each probability is non-zero,
 a small constant is added to all probabilities in a process called
-\textit{Laplace smoothing}, which is commonly used for NB classifiers.
+\textit{Laplace smoothing}, which is now commonly used for NB classifiers.
 
@@
 
-The only thing left to specify
-now is how to decide which attribute to split on. Note that when splitting,
+We now only need to specify how to decide which attribute to split on at
+each step. Note that when splitting,
 we would like to create subsets that are as ``pure''
@@
 created subsets were as pure as possible. The measure of purity used in
-learning decision trees is the \textit{information}, and is measured in
+learning decision trees is called \textit{information}, and is measured in
 \textit{bits}. At any node (or subset) of the tree, this measure represents
@@
 for humans to interpret, the latter of which is particularly valuable in a
-clinical environment. We will use the C4.5 tree induction algorithm, which is
-considered the state-of-the-art and handles different types of feature values
-as well as missing values.
+clinical environment. We will use the C4.5 tree induction algorithm
+\cite{Quinlan1993}, which is considered the state-of-the-art in decision trees
+and handles different types of feature values as well as missing values.
 
@@
 Previous work on trauma LOS prediction by
-Dinh et. al. \cite{Dinh2013a}, as well as many attempts at modelling the LOS
-in various medical domains, have all used logistic regression, often as a
+Dinh et. al. \cite{Dinh2013a} as well as many attempts at modelling the LOS
+in various medical domains have all used logistic regression, often as a
 starting point.
@@
 \end{equation*}
-For LOS classification, this can be rearranged to:
+For our LOS classification problem, this can be rearranged to:
 \begin{equation*}
@@
 numerical methods, where estimates of the $\beta_i$ are chosen, the likelihood
-computed, and the $\beta_i$ are updated, and so on: this process is repeated
+computed, the $\beta_i$ updated accordingly, and so on: this process is repeated
 until parameter estimates converge or do not change more than a given threshold
@@
 \caption{An example of a 2-dimensional SVM. The solid line indicates the
-maximum margin hyperplane.}
+maximum margin hyperplane, and the dotted line indicates the margin.}
 \label{fig:svm}
@@
 of the hyperplane, and all black points lie on the other. In practice, this is
-not common, and often data cannot be separated cleanly by a hyperplane.
+not common, and often the data cannot be separated cleanly by a hyperplane.
 Note that the points closest to the margin are those that are most likely to be
-ambiguous when the data cannot be cleanly separated. The SVM will progressively
-remove more and more of these ambiguous points from consideration when it
-cannot find a hyperplane until it finds a suitable one.
+ambiguous when the data cannot be cleanly separated. If it cannot find a
+hyperplane, the SVM will progressively
+remove more and more of these ambiguous points from consideration 
+until it finds a suitable one.
 
@@
 practically possible with data sets containing even a modest number of features
-(around 10), due to the computational complexity of estimating a large number
+(around 10) due to the computational complexity of estimating a large number
 of coefficients that are introduced by the transformation.
@@
 $\mathbf{w}\cdot\mathbf{x_i} + b = 0$ is the hyperplane that the SVM finds.
-The $s_i$'s allow the SVM to ignore some examples if it cannot find a suitable
+The $s_i$ allow the SVM to ignore some examples if it cannot find a suitable
 hyperplane, and $C$ (sometimes called the SVM's complexity constant) is a
@@
 SVM involves solving this optimisation problem and obtaining values for
-$\mathbf{w}$ and $b$.
+$\mathbf{w}$ and $b$, which define the hyperplane.
 
@@
 \caption{An MLP with one hidden layer. There are $n$ input nodes, $k$ nodes
-in the hidden layer, and one output node.}
+in the hidden layer, and one output node which contains the prediction of
+the network.}
 \label{fig:mlp}
@@
 function $E=\frac{1}{2}(\text{actual} - \text{predicted})^2$ -- the one most
-commonly used -- which is
+commonly used in practice -- which is
 minimised in the gradient descent process. Updating the weights for each
@@
 \end{equation*}
-where $a_i$ is the normalised value, $v_i$ is the actual value in the data
-set, max $v_i$ and min $v_i$ are taken over all training examples, $s$ is a
-scaling factor and $t$ indicates how much to translate the range by. If $s=1$
+where $a_i$ is the normalised feature value, $v_i$ is the actual value in the
+data set, max $v_i$ and min $v_i$ are taken over all training examples, $s$ is
+a scaling factor and $t$ indicates how much to translate the range by. If $s=1$
 and $t=0$, then the range of normalised values will be in $[0,1]$.
@@
 problem domain at hand, or to reduce the storage requirements of the original
-$k$-NN algorithm. We will focus on two extensions to the Euclidean distance:
-K* and our proposed algorithm, Ranked Distance.
+$k$-NN algorithm (since the classifier makes no attempt to generalise from the
+examples, all of them must be stored).
+We will focus on two algorithms that extend the standard Euclidean distance:
+K* and our proposed approach, Ranked Distance Nearest Neighbour.
 
@@
 We use K* as a comparison to the basic NN algorithm and also as a comparison
-to our proposed Ranked Distance Nearest Neighbours algorithm.
+to our proposed Ranked Distance Nearest Neighbour algorithm. Additionally,
+K* has not been yet been applied in predicting LOS, and its performance in this
+domain is worth investigating.
 
@@
 classifier, which implies that all features are equally important in predicting
-the class. In practice, however, this is not always the case: to predict the
+the class. In practice, however, this is not always the case: for example,
+to predict the
 LOS of a trauma patient, the severity of their injury should affect the final
@@
 of many features that are recorded about a patient). Our contribution is a
-new NN algorithm that uses a modified distance function which takes into
-account the relative importance of all features called Ranked Distance
-Nearest Neighbour (RD).
+new NN algorithm, Ranked Distance Nearest Neighbour (RD), which uses a modified
+distance function that takes into account the relative importance of each
+feature in predicting the class.
 
@@
 \begin{equation*}
-f : \mathrm{Rank} \rightarrow \mathbb{R}, \mathrm{Rank} \in \{1,2,\ldots,n\}
+f : \mathrm{Rank} \rightarrow \mathbb{R}, \text{ }\mathrm{Rank} \in \{1,2,\ldots,n\}
 \end{equation*}
@@
 how the weights of the features vary with the rank. This function should
-inituitively be non-increasing and should produce lower values when the rank
-number is greater, meaning that less important features have a lower weight
-in the distance calculation.
+inituitively be non-increasing: that is, it should produce lower values when
+the rank number is greater (which indicates a lower rank), meaning that less
+important features have a lower weight in the final distance calculation.
 Note that choosing $f(\mathrm{Rank}) = k$ for any non-zero constant $k$
 results in all features contributing equally to the distance calculation.
-We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}^c}$, where
+We will use $f(\mathrm{Rank}) = \dfrac{1}{\mathrm{Rank}^c}$, where
 $c$ varies from 0 to 1.
@@
 sophisticated support vector machines and multi-layer perceptrons. We also
-described our proposed new algorithm, Ranked Distance Nearest Neighbours.
+described our proposed new algorithm, Ranked Distance Nearest Neighbour.
 The reason
@@
 classifier and give us more insight into the nature of the data set and of the
-problem.
+problem itself.

### 2014-11-04 thesis/method/evaluation.tex
Final version of thesis

@@
 This chapter discusses the procedure and performance measures used
-in evaluating our work.
+in our evaluation.
 
@@
 \ref{chap:preprocess}, giving us two data sets: one with and one without
-discretisation. The general LOS data from Portugal was already cleaned when we
-received it; however, we discretised the continuous \texttt{los} variable to
-the same categories as for the trauma data (2 days or less, greater than 2
-days), normalised the numeric attribute \texttt{prev\_admissions}, and derived
-both a discretised and non-discretised data set. This resulted in 4 sets of
-data.
-\item For each data set, we evaluated the effect of the 4 automatic feature
+discretisation. This was also done for the general hospital data set, resulting
+in 4 sets of data.
+\item For each data set, we evaluated the effect of the 5 automatic feature
 selection methods on each of the 11 classifiers described in Chapter
-\ref{chap:classification}, giving us 44 configurations for each data set, over
-4 sets. Evaluation was performed using \textit{ten-fold stratified
+\ref{chap:classification}, giving us 55 configurations for each data set, over
+the 4 sets obtained in the previous step.
+Evaluation was performed using \textit{ten-fold stratified
 cross-validation} (described below).
 \item For the two trauma LOS data sets, we additionally evaluated the two
-manual feature selection methods with all 8 classifiers. Recall from Section
+manual feature selection methods with all 11 classifiers. Recall from Section
 \ref{sec:manual} that one of the feature sets is from the key previous work
 on trauma LOS prediction that we use as a baseline, and the other feature set
-is a set of features included through consultation with a domain expert.
+is a set of features derived through consultation with a domain expert.
 \end{enumerate}
@@
 from Microsoft Excel. All runs of WEKA were executed by
-a program we wrote in Bash (a Unix shell) script. This program was able to run
+a program we wrote in Bash (a Unix shell) script, the source of which can be
+found in Appendix \ref{sec:run-weka}. This program was able to run
 WEKA with the following features:
@@
 interval (also discussed in the next section) in a single output that we could
-then save as CSV.
+then save as CSV. These can be found in Appendix \ref{sec:summarise-script} and
+\ref{sec:compute-ci}.
 
-In order to implement our novel Ranked Distance Nearest Neighbour algorithm,
+In order to implement our novel Ranked Distance Nearest Neighbour algorithm
+(introduced in Section
 \ref{sec:rankeddist}), we created our own class within the WEKA Java package
@@
 source was built using version 1.7.0 Update 17 of the Java Development Kit.
-
-The source code listing of the most important programs we wrote are displayed
-in Appendix \ref{app:code}.
+The source for this can be found in Appendix \ref{sec:rankeddist-code}.
 
@@
 dividing the graph diagonally in half. Figure \ref{fig:sampleroc} shows
-two ROC curves on the same graph:
+three ROC curves on the same graph:
 
@@
 discriminate between instances of two classes. Recall from Chapter
-\ref{chap:classification} that 0R will always predict the majority class --
+\ref{chap:classification} that Zero-R will always predict the majority class --
 this means that it will be evaluated positively on the basis of accuracy
@@
 In this section we described our method of evaluating the performance of each
-classifier and feature selection method that we use. We used accuracy and the
+classifier and feature selection method that we used, including the code
+written and software used. We used accuracy and the
 area under the curve to measure two different aspects of our classifiers, and

### 2014-11-04 thesis/method/preprocess.tex
Final version of thesis

@@
 \section{Data Sets}
-In our work we obtained two independent sets of LOS data, one specifically for
-trauma and the other a general, hospital-wide data set.
+In our work we obtained two independent sets of LOS data through
+collaboration with medical experts and researchers in Portugal: one
+specifically for trauma and the other a general, hospital-wide data set.
 
@@
 values, 0 or 1 -- with 1 indicating that the patient stayed two days or less,
-and 0 indicating a stay of greater than 2 days.
+and 0 indicating a stay of longer than 2 days.
 
@@
 data entry
-\item Leading or trailing whitespace, and other garbage characters,
-perhaps as a result of data entry or
-the system that stores the data
+\item Leading or trailing whitespace and other garbage characters,
+perhaps as a result of data entry or the system that stores the data
 \item Missing values for various features or for the class label
@@
 the data was then exported into CSV (comma-separated value) format, so that
-we could continue to process it programmatically rather than manually without
-having to use Excel. Many
-fields in the data set contained leading or trailing whitespace, which was
-removed with a few lines of Python code (see Appendix
+we could continue to process it programmatically, without the need for
+Excel. Many fields in the data set contained leading or trailing whitespace,
+which was removed with a few lines of Python code (see Appendix
 \ref{sec:python-preprocess}).
@@
 WEKA, a suite of machine learning algorithms implemented in Java
-\cite{Hall2009}. Each feature was carefully checked to ensure that their
+\cite{Hall2009}. Each feature was carefully checked to ensure that its
 type was specified correctly -- that is, the nominal attributes were
@@
 undo the logarithm transform that had been applied to the \texttt{Previous
-Admissions} feature.
+Admissions} feature, because it had been previously applied during the course
+of other work conducted on the data set. This was done using $e^x-1$, where
+$x$ is the value of the feature.
 
@@
 range. This transformation was applied to each value of all numeric features
+of both data sets
 using WEKA, and was performed after all cleaning steps described above.
@@
 unsupervised discretisation when class information about the training examples
-is known. And the method proposed by Fayyad and Irani is one of the best known
-general supervised discretisation techniques \cite{Witten2005}.
+is known -- and the method proposed by Fayyad and Irani is one of the best
+known general supervised discretisation techniques \cite{Witten2005}.
 
 \subsection{Entropy-based Discretisation}
-In Chapter \ref{chap:classification} we outlined the induction of a decision
+In Section \ref{sec:c45} we outlined the induction of a decision
 tree on a data set, using the information gain as the criterion for deciding
@@
 the split point in $\mathrm{log}_2(N-1)$ bits (where $N$ is the number of
-instances), then the classes of those above and below it. If we do not split,
+instances) and subsequently the classes of those above and below it. If we do not split,
 then each instance's label must be encoded. Mathematically, given $N$ examples
@@
 that any potential split that minimises the information value will never occur
-between two examples of the same class. This means that we only need to
+between two examples of the same class \cite{Witten2005}.
+This means that we only need to
 consider splits between regions of different classes, reducing the number of
@@
 continuous-valued data that ensures all features take on the same range of
-values. This is important for ensuring that classifiers sensitive to the
-relative range of values between features give valid outputs.
+values. This is important for ensuring that classifiers which are sensitive to
+the relative range of values between features give valid outputs.
 
 We concluded with outlining discretisation, in particular Fayyad and Irani's
-entropy-based method. Such a technique has not been used in LOS prediction
-and by including it in our work, we can evaluate its effectiveness in this
-domain and let our results serve as a baseline and reference for future work.
+supervised entropy-based method. Such a technique has not been used in LOS
+prediction and by including it in our work, we can evaluate its effectiveness
+in this domain and let our results serve as a baseline and reference for
+future work.

### 2014-11-04 thesis/method/selection.tex
Final version of thesis

@@
 
-Our data set contains 78 features aside from the class, which is a fairly high
-number. Many of these attributes
+One of our data sets (the trauma-specific one) contains 78 features aside from
+the class, which is a fairly large number. Many of these attributes
 are derived from others in the set: for example, whether or not a patient is
 over 65 years of age (the \texttt{age65} feature) is directly dependent on the
-value of their \texttt{age} attribute.
+value of their \texttt{age} attribute. The other general hospital data set
+contains 14 features, which is significantly less. For both data sets, we would
+like to see if we can maintain or improve classifier performance by reducing
+the number of features used.
 
@@
 After eliminating redundant or inapplicable features as detailed above, we
-considered four \textit{automatic} (correlation-based
-feature selection, Pearson correlation, 1R and wrapper selection
+considered five \textit{automatic} (correlation-based
+feature selection, Pearson correlation, information gain, 1R and wrapper selection
 using C4.5) and two \textit{manual} feature selection methods (domain expert
@@
 \section{Automatic Methods}
-Automatic feature selection methods are useful for when we do not have a deep
-understanding of the underlying domain and what each attribute really means to
-the class. Even if we are well-versed in the domain, evaluating a few different
-methods of automatic feature selection can give us new insight into the
-learning problem at hand. Here we will describe a few of the methods we used.
+Automatic feature selection methods are useful when we do not have a deep
+understanding of the underlying domain and the exact relationship of each
+feature to the class. Even if we are well-versed in the domain, evaluating a
+few different methods of automatic feature selection can give us new insight
+into the learning problem at hand, and could also prove useful in extracting
+new knowledge. Here we will describe a few of the methods we used.
 
@@
 each feature, which are then used to rank the features from highest to lowest
-accuracy. The worst accuracy for any feature is the 0R accuracy, which is the
-proportion of the majority class in the data set. This is because if 1R finds
-that the accuracy of the rule constructed using a feature is worse than 0R,
-then it will not use that rule and predict the majority class instead.
+accuracy. The worst accuracy for any feature is the Zero-R accuracy, which is
+the proportion of the majority class in the data set. This is because if 1R
+finds that the accuracy of the rule constructed using a feature is worse than
+Zero-R, then it will not use that rule and predict the majority class instead.
 
@@
 The feature selection methods described above (CFS, Pearson correlation and
-1R) have been described as \textit{filter} approaches to attribute selection.
+1R) have been described as \textit{filter} approaches to attribute selection,
+whereby the entire feature set is first ``filtered'' by some evaluation metric
+and the reduced features used in a learning algorithm.
 Kohavi and John proposed another approach, the \textit{wrapper} approach:
@@
 wrapper methods search for the best subset of features and do not require a
-threshold to discard attributes.
+threshold to decide which features to select.
 
@@
 worthwhile to compare the performance of classifiers using features selected
-automatically and manually.
+automatically and manually. We were able to obtain two sets of features
+selected manually for the trauma data set.
 
@@
 \subsection{Domain Expert Selection}
-Consultation with a domain expert yielded another feature set that we could
+Consultation with a domain expert yielded another feature set that we
 use to compare with the baseline and with the other automatic methods.

### 2014-11-04 thesis/results/results.tex
Final version of thesis

@@
 that \texttt{iss}, \texttt{age} and \texttt{operation} were important
-distinguishers of LOS $\leq$ or $>$ 2 days \cite{Dinh2013a}. We also
+distinguishers of LOS ``less than or equal to 2 days'' or ``greater than 2 days'' \cite{Dinh2013a}. We also
 confirm the findings of Gabbe et al. who found that \texttt{age},
@@
 this is not as high as the best result of 0.994, it is worth pointing out that
-CFS only selected 2 features and C4.5 wrapper, 4 features. This means that
+CFS and C4.5 selected only selected 2 and 4 features, respectively. This means that
 using only 2 features, we were still able to discriminate between the two LOS
@@
 the class, a finding that was also noted by Lucas et al. \cite{Lucas2009},
-although they considered the classification of LOS $\leq$ and $>$ 3 days.
+although they considered the classification of LOS less than and greater than
+3 days.
 

### 2014-11-04 code/weka/compute_ci.py

@@ -63,5 +63,5 @@ for i in range(len(headings)):
         results = []
         for col in zip(*sample):
             m, u = compute_ci(col)
-            results.append('{0:.3f}({1:.5f})'.format(m, u))
+            results.append('{0:.3f}+/-{1:.5f}'.format(m, u))
         print(' '.join(results))

### 2014-11-07 presentation/slides.tex
Add skeleton slides

\documentclass[compress]{beamer}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{verbatim}

% Load the usyd template
\input{usyd-beamer/usyd-beamer.tex}

\begin{document}
\usydtitleframe{Data Mining for Predicting Length of Hospital Stay}{}{Tianyu Pu \\
  Supervisor: Irena Koprinska}

\section{Introduction}
\subsection{}
\begin{frame}{Thought exercise}
Imagine that you have just been admitted to hospital...
\end{frame}

\subsection{}
\begin{frame}{Outline of talk}
The problem

Previous work

Our approach

Results

Conclusions and future work
\end{frame}

\section{The Problem}
\subsection{}
\begin{frame}{The problem}
\end{frame}

\section{Background}
\subsection{}
\begin{frame}{Common approach: logistic regression}
\end{frame}

\subsection{}
\begin{frame}{Artificial neural networks}
\end{frame}

\subsection{}
\begin{frame}{Feature selection}
\end{frame}

\section{Approach}
\subsection{}
\begin{frame}{Ranked Distance Nearest Neighbour}
\end{frame}

\subsection{}
\begin{frame}{Overview of our approach}
\end{frame}

\section{Results}
\subsection{}
\begin{frame}{Trauma data set}
\end{frame}

\subsection{}
\begin{frame}{General hospital data set}
\end{frame}

\section{Conclusions}
\subsection{}
\begin{frame}{Conclusions}
\end{frame}

\end{document}

### 2014-11-08 presentation/slides.tex
Add draft presentation slides

@@
 \documentclass[compress]{beamer}
+\usepackage{epstopdf}
 \usepackage{graphicx}
 \usepackage{lmodern}
+\usepackage{tikz}
 \usepackage{verbatim}
+
+\usetikzlibrary{decorations.pathreplacing,calc}
+
+\newcommand{\tikzmark}[2][-3pt]{\tikz[remember picture, overlay, baseline=-2.0ex]\node[#1](#2){};}
+
+\tikzset{brace/.style={decorate, decoration={brace}},
+ brace mirrored/.style={decorate, decoration={brace,mirror}},
+}
+
+\newcounter{brace}
+\setcounter{brace}{0}
+\newcommand{\drawbrace}[3][brace]{%
+ \refstepcounter{brace}
+  \tikz[remember picture, overlay]\draw[#1] (#2.center)--(#3.center)node[pos=0.5, name=brace-\thebrace]{};
+}
+
+\newcounter{arrow}
+\setcounter{arrow}{0}
+\newcommand{\drawcurvedarrow}[3][]{%
+ \refstepcounter{arrow}
+  \tikz[remember picture, overlay]\draw (#2.center)edge[#1]node[coordinate,pos=0.5, name=arrow-\thearrow]{}(#3.center);
+}
+
+% #1 options, #2 position, #3 text 
+\newcommand{\annote}[3][]{%
+ \tikz[remember picture, overlay]\node[#1] at (#2) {#3};
+}
+
+\renewcommand*\footnoterule{} % remove footnote rule
+
+\definecolor{emerald}{RGB}{0,166,79}
 
@@
 
+\setbeamerfont{footnote}{size=\tiny}
+
 \begin{document}
-\usydtitleframe{Data Mining for Predicting Length of Hospital Stay}{}{Tianyu Pu \\
+\usydtitleframe{Data Mining Methods for Predicting Length of Hospital Stay}{}{\textbf{Tianyu Pu} Honours Student \\
   Supervisor: Irena Koprinska}
@@
 \subsection{}
-\begin{frame}{Thought exercise}
+\begin{frame}{Introduction}
 Imagine that you have just been admitted to hospital...
+  \pause
+
+\vspace{0.5cm}
+\textbf{Why do we care?}
 \end{frame}
@@
 
-Previous work
+\vspace{0.5cm}
+Background and previous work
 
+\vspace{0.5cm}
 Our approach
 
-Results
+\vspace{0.5cm}
+Results and discussion
 
+\vspace{0.5cm}
 Conclusions and future work
@@
 \begin{frame}{The problem}
+Given
+\begin{center}
+\begin{tabular}{cccc|c}
+\tikzmark[]{x}age & gender & heart rate & ... & length of stay (LOS)\tikzmark[]{y} \\
+\hline
+\tikzmark[xshift=-0.80ex, yshift=-1ex]{a}45 & F & 122 & ... & $>$ 2 days \\
+23 & M & 98 & ... & $\leq$ 2 days \\
+\multicolumn{4}{c|}{\tikzmark[xshift=-13.3ex, yshift=-2ex]{b}...} & ... \\
+\end{tabular}
+\end{center}
+\drawbrace[brace mirrored, thick, gray]{y}{x}
+\drawbrace[brace mirrored, thick, gray]{a}{b}
+\annote[yshift=1.5ex, gray]{brace-1}{\scriptsize features}
+\annote[left, gray]{brace-2}{\scriptsize examples}
+
+\vspace{0.5cm}
+Predict LOS category for new patient
+\begin{center}
+\begin{tabular}{cccc|c}
+age & gender & heart rate & ... & length of stay (LOS)  \\
+\hline
+34 & F & 101 & ... & ? \\
+\end{tabular}
+\end{center}
 \end{frame}
 
-\section{Background}
+\section{Background and Previous Work}
 \subsection{}
-\begin{frame}{Common approach: logistic regression}
+\begin{frame}{Prediction}
+\begin{center}
+\includegraphics<1>{learning.pdf}
+\includegraphics<2>{learning-focus2.pdf}
+\end{center}
 \end{frame}
@@
 \subsection{}
-\begin{frame}{Artificial neural networks}
+\begin{frame}{Feature discretisation and selection}
+\textbf{Discretisation}
+\begin{itemize}
+\item Convert numeric features into categorical features
+\item \textit{Supervised} discretisation
+\item Why?
+\end{itemize}
+
+\vspace{0.5cm}
+
+  \pause
+\textbf{Selection}
+\begin{itemize}
+\item Find a subset of features that are the most `relevant'
+\item Two ways: \textit{manual} and \textit{automatic}
+\item Why?
+\end{itemize}
 \end{frame}
@@
 \subsection{}
-\begin{frame}{Feature selection}
+\begin{frame}{Learning algorithms}
+\textbf{Logistic regression}
+
+One of the most widely used approaches in LOS prediction
+\begin{equation*}
+\begin{aligned}
+\text{log}\left(\dfrac{\text{probability of LOS}\leq\text{2 days}}{\text{probability of LOS}>\text{2 days}}\right) &= b_0 + b_1\text{age} + b_2\text{gender} + \ldots
+\end{aligned}
+\end{equation*}
+  \pause
+
+\textbf{Others}
+
+Artificial neural networks, ...
 \end{frame}
@@
 \subsection{}
-\begin{frame}{Ranked Distance Nearest Neighbour}
+\begin{frame}{Proposed approach (1)}
+Store all examples and do not try to generalise
+
+Classify new example by looking at $k$ most `similar' (`nearest') examples
+\begin{itemize}
+\item \textit{Nearest neighbour} approach
+\end{itemize}
+
+\vspace{0.5cm}
+\textbf{Similarity}
+
+Typical approach: \textit{Euclidean distance}
+
+\begin{equation*}
+\sqrt{(x_1-y_1)^2 + (x_2-y_2)^2 + \ldots}
+\end{equation*}
 \end{frame}
@@
 \subsection{}
-\begin{frame}{Overview of our approach}
+\begin{frame}{Proposed approach (2)}
+Our approach: \textit{Ranked distance}
+\begin{equation*}
+w_1|x_1-y_1| + w_2|x_2-y_2| + \ldots
+\end{equation*}
+
+\vspace{0.5cm}
+\textbf{Intuition}
+
+The relative importance of a feature should determine its contribution to the similarity
+\end{frame}
+
+\subsection{}
+\begin{frame}{Evaluation}
+\textbf{Two datasets}: trauma and general hospital LOS data
+
+\vspace{0.5cm}
+
+\textbf{Procedure}:
+\begin{enumerate}
+\item Preprocess and discretise datasets $\implies$ 4 sets of data
+\item Systematically evaluate 11 learning algorithms with 5 \textit{automatic} and 2 \textit{manual} feature selection methods
+\end{enumerate}
+
+\vspace{0.5cm}
+
+  \pause
+\textbf{Performance measures}: \textit{accuracy} and \textit{area under the curve (AUC)}
 \end{frame}
@@
 \subsection{}
+\begin{frame}{Summary of best results}
+\textbf{Trauma dataset}
+\begin{center}
+\begin{tabular}{l|cc}
+ & Baseline & Our best result \\
+\hline
+Accuracy (\%) & 75.06 (19) & \textcolor{emerald}{77.81} (\textcolor{emerald}{11})\footnote{1-NN, C4.5 wrapper, discretised} \\
+AUC & 0.812 (19) & \textcolor{emerald}{0.846} (\textcolor{sydneyred}{29})\footnote{Logistic regression, no feature selection, discretised} \\
+\end{tabular}
+\end{center}
+
+\pause
+\textbf{General hospital dataset}
+\begin{center}
+\begin{tabular}{l|c}
+ & Our best result \\
+\hline
+Accuracy (\%) & 98.23 (14)\footnote{C4.5 decision tree, no feature selection or discretisation} \\
+AUC & 0.994 (14)\footnote{Logistic regression, SVM and K*, no feature selection or discretisation} \\
+\end{tabular}
+\end{center}
+\end{frame}
+
+\subsection{}
 \begin{frame}{Trauma data set}
+\includegraphics[width=0.93\textwidth]{tr-corr.eps}
 \end{frame}
@@
 \begin{frame}{General hospital data set}
+\begin{center}
+\includegraphics[width=0.92\textwidth]{pt-corr.eps}
+\end{center}
+\end{frame}
+
+\subsection{}
+\begin{frame}{Discussion}
+Limitations of nearest neighbour methods
+
+\vspace{0.5cm}
+Trade-off between predictive power and number of features
+
+\vspace{0.5cm}
+Know your data!
 \end{frame}
@@
 \begin{frame}{Conclusions}
+Comprehensive and systematic evaluation
+
+\vspace{0.5cm}
+We improved accuracy by 2.75\% and AUC by 0.034 on the same dataset using nearest neighbour methods
+
+\pause
+\vspace{0.5cm}
+\textbf{Future work}
+\begin{itemize}
+\item Ensembles
+\item Different similarity measures
+\item Implementation for doctors
+\end{itemize}
 \end{frame}

### 2014-12-23 presentation/summer-slides.tex
Add slides for summer scholarship presentation

--- 2014-11-08 presentation/slides.tex
@@
 \begin{document}
-\usydtitleframe{Data Mining Methods for Predicting Length of Hospital Stay}{}{\textbf{Tianyu Pu} Honours Student \\
-  Supervisor: Irena Koprinska}
+\usydtitleframe{Predicting the Length of Stay in Hospital Using Data Mining Methods}{}{\textbf{Tianyu Pu} \\
+  Supervisors: Irena Koprinska \& Kalina Yacef}
 
@@
 \textbf{Why do we care?}
-\end{frame}
-
-\subsection{}
-\begin{frame}{Outline of talk}
-The problem
-
-\vspace{0.5cm}
-Background and previous work
-
-\vspace{0.5cm}
-Our approach
-
-\vspace{0.5cm}
-Results and discussion
-
-\vspace{0.5cm}
-Conclusions and future work
 \end{frame}
@@
 \item \textit{Supervised} discretisation
-\item Why?
 \end{itemize}
@@
 \item Two ways: \textit{manual} and \textit{automatic}
-\item Why?
 \end{itemize}
@@
 
-\section{Approach}
 \subsection{}
-\begin{frame}{Proposed approach (1)}
-Store all examples and do not try to generalise
-
-Classify new example by looking at $k$ most `similar' (`nearest') examples
-\begin{itemize}
-\item \textit{Nearest neighbour} approach
-\end{itemize}
+\begin{frame}{Where We're At}
+Comprehensive evaluation of feature selection, discretisation and learning algorithms
+\pause
 
 \vspace{0.5cm}
-\textbf{Similarity}
-
-Typical approach: \textit{Euclidean distance}
-
-\begin{equation*}
-\sqrt{(x_1-y_1)^2 + (x_2-y_2)^2 + \ldots}
-\end{equation*}
+Proposed new nearest-neighbour approach
 \end{frame}
 
+\section{Next Steps}
 \subsection{}
-\begin{frame}{Proposed approach (2)}
-Our approach: \textit{Ranked distance}
-\begin{equation*}
-w_1|x_1-y_1| + w_2|x_2-y_2| + \ldots
-\end{equation*}
-
-\vspace{0.5cm}
-\textbf{Intuition}
-
-The relative importance of a feature should determine its contribution to the similarity
-\end{frame}
-
-\subsection{}
-\begin{frame}{Evaluation}
-\textbf{Two datasets}: trauma and general hospital LOS data
-
-\vspace{0.5cm}
-
-\textbf{Procedure}:
-\begin{enumerate}
-\item Preprocess and discretise datasets $\implies$ 4 sets of data
-\item Systematically evaluate 11 learning algorithms with 5 \textit{automatic} and 2 \textit{manual} feature selection methods
-\end{enumerate}
-
-\vspace{0.5cm}
-
-  \pause
-\textbf{Performance measures}: \textit{accuracy} and \textit{area under the curve (AUC)}
-\end{frame}
-
-\section{Results}
-\subsection{}
-\begin{frame}{Summary of best results}
-\textbf{Trauma dataset}
-\begin{center}
-\begin{tabular}{l|cc}
- & Baseline & Our best result \\
-\hline
-Accuracy (\%) & 75.06 (19) & \textcolor{emerald}{77.81} (\textcolor{emerald}{11})\footnote{1-NN, C4.5 wrapper, discretised} \\
-AUC & 0.812 (19) & \textcolor{emerald}{0.846} (\textcolor{sydneyred}{29})\footnote{Logistic regression, no feature selection, discretised} \\
-\end{tabular}
-\end{center}
-
-\pause
-\textbf{General hospital dataset}
-\begin{center}
-\begin{tabular}{l|c}
- & Our best result \\
-\hline
-Accuracy (\%) & 98.23 (14)\footnote{C4.5 decision tree, no feature selection or discretisation} \\
-AUC & 0.994 (14)\footnote{Logistic regression, SVM and K*, no feature selection or discretisation} \\
-\end{tabular}
-\end{center}
-\end{frame}
-
-\subsection{}
-\begin{frame}{Trauma data set}
-\includegraphics[width=0.93\textwidth]{tr-corr.eps}
-\end{frame}
-
-\subsection{}
-\begin{frame}{General hospital data set}
-\begin{center}
-\includegraphics[width=0.92\textwidth]{pt-corr.eps}
-\end{center}
-\end{frame}
-
-\subsection{}
-\begin{frame}{Discussion}
-Limitations of nearest neighbour methods
-
-\vspace{0.5cm}
-Trade-off between predictive power and number of features
-
-\vspace{0.5cm}
-Know your data!
-\end{frame}
-
-\section{Conclusions}
-\subsection{}
-\begin{frame}{Conclusions}
-Comprehensive and systematic evaluation
-
-\vspace{0.5cm}
-We improved accuracy by 2.75\% and AUC by 0.034 on the same dataset using nearest neighbour methods
+\begin{frame}{Next Steps}
+Feedback from domain experts
 
@@
 \vspace{0.5cm}
-\textbf{Future work}
-\begin{itemize}
-\item Ensembles
-\item Different similarity measures
-\item Implementation for doctors
-\end{itemize}
+Draft papers for publication
+\pause
+
+\vspace{0.5cm}
+Investigate ensembles and refinement of nearest-neighbour method
 \end{frame}

### 2014-12-23 papers/paper.tex
Add skeleton of NN paper

--- 2014-11-03 thesis/appendix/paper.tex
@@
-\chapter{Artificial Intelligence in Medicine -- Draft Paper} \label{app:paper}
+\documentclass[10pt]{article}
 
-We are currently in the process of drafting a paper of our results to submit to
-Artificial Intelligence in Medicine, a top journal for research that uses
-artificial intelligence in fields such as medicine, human biology and health
-care. Here we have attached our draft, which will be soon ready for submission.
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{graphicx}
+\usepackage{multicol}
+\usepackage{natbib}
+\usepackage{subcaption}
 
-\clearpage
+\title{Nearest Neighbour Methods for Length of Stay Prediction}
+\author{
+  Tianyu Pu, Irena Koprinska, Kalina Yacef\\
+  School of Information Technologies\\
+  University of Sydney, NSW, 2006\\
+  \texttt{\{tianyu.pu,irena.koprinska,kalina.yacef\}@sydney.edu.au}
+  \and
+  Michael Dinh\\
+  Trauma Services, Royal Prince Alfred Hospital\\
+  Sydney, NSW, 2042\\
+  \texttt{dinh.mm@gmail.com}
+}
 
-\setlength{\multicolsep}{0pt}
-\setlength{\topskip}{\fontcharht\font`B }
-\setlength{\parskip}{0pt}
-%[
-\begin{center}
-\section*{Improving the Prediction of Length of Hospital Stay for Trauma Patients}
-\end{center}
-\begin{multicols}{2}
-\begin{center}
-Tianyu Pu, Irena Koprinska \\
-School of Information Technologies \\
-University of Sydney, NSW, 2006 \\
-\texttt{\{tianyu.pu,irena.koprinska\}@sydney.edu.au}
-\end{center}
-\vfill
-\columnbreak
-\begin{center}
-Michael Dinh \\
-Trauma Services, Royal Prince Alfred Hospital \\
-Sydney, NSW, 2042 \\
-\texttt{dinh.mm@gmail.com}
-\end{center}
-\end{multicols}
-%]
+\date{\today}
 
-\begin{multicols}{2}
-\subsection*{Abstract}
+\begin{document}
+\maketitle
+
+\begin{abstract}
 Predicting the length of hospital stay for injured patients involves making a
@@
 traditional nearest neighbour algorithms.
+\end{abstract}
 
-\subsection*{Introduction}
+\section{Introduction}
 Hospitals often have limited funding, and therefore managing and utilising
@@
 
-\subsection*{Problem Statement}
+\section{Problem Statement}
 Our problem is a binary \textit{classification} task in which we
@@
 
-\subsection*{Ranked Distance Nearest Neighbour}
-\paragraph{Intuition}
+\section{Ranked Distance Nearest Neighbour}
+\subsection{Intuition}
 As we mentioned earlier, all features are considered equally in a $k$-NN
@@
 
-\paragraph{Mathematical formulation}
+\subsection{Mathematical formulation}
 Given two instances $\mathbf{x} = (x_1,x_2,\ldots,x_n)$ and
@@
 
-\paragraph{Assignment of weights}
+\subsection{Assignment of weights}
 The weights $w_i$ can be tuned to match the particular problem that is being
@@
 
-\subsection*{Feature Selection}
+\section{Feature Selection}
 Here we introduce the feature selection algorithms we evaluate in our work.
@@
 stay prediction.
-\subsubsection*{Correlation-based Feature Selection}
+\subsection{Correlation-based Feature Selection}
 Correlation-based feature selection (CFS) is a method developed by Hall
@@
 
-\subsubsection*{Pearson Correlation Coefficient}
+\subsection{Pearson Correlation Coefficient}
 The Pearson correlation coefficient measures the degree of linear correlation
@@
 
-\subsubsection*{Information Gain}
+\subsection{Information Gain}
 We can also measure the IG of each feature as follows.
@@
 
-\subsubsection*{1R}
+\subsection{1R}
 Another way to evaluate the relevance of features is to use the idea behind the
@@
 
-\subsubsection*{Wrapper-based Feature Selection}
+\subsection{Wrapper-based Feature Selection}
 The feature selection methods described above (CFS, Pearson correlation and
@@
 
-\subsection*{Evaluation Method}
+\section{Evaluation Method}
 Our evaluation was carried out using a trauma-specific data set
@@
 in Sydney, Australia.
-\subsubsection*{Classifiers}
+\subsection{Classifiers}
 We tested the performance of Zero-R, 1R, Na\"{i}ve Bayes, C4.5 decision trees,
@@
 
-\subsubsection*{Procedure}
+\subsection{Procedure}
 The steps that we took to evaluate the performance of the various classifiers
@@
 
-\subsection*{Results and Discussion}
+\section{Results and Discussion}
 The best accuracy was 77.81\%, achieved using the $k$-NN algorithm
@@
 \begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-nofs-acc.eps}
+\includegraphics[width=\textwidth]{../thesis/images/results/tr-nofs-acc.eps}
 \caption{}
@@
 \begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-nofs-auc.eps}
+\includegraphics[width=\textwidth]{../thesis/images/results/tr-nofs-auc.eps}
 \caption{}
@@
 \begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-dinh-acc.eps}
+\includegraphics[width=\textwidth]{../thesis/images/results/tr-dinh-acc.eps}
 \caption{}
@@
 \begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-dinh-auc.eps}
+\includegraphics[width=\textwidth]{../thesis/images/results/tr-dinh-auc.eps}
 \caption{}
@@
 \begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-expert-acc.eps}
+\includegraphics[width=\textwidth]{../thesis/images/results/tr-expert-acc.eps}
 \caption{}
@@
 \begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-expert-auc.eps}
+\includegraphics[width=\textwidth]{../thesis/images/results/tr-expert-auc.eps}
 \caption{}
@@
 \begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-cfs-acc.eps}
+\includegraphics[width=\textwidth]{../thesis/images/results/tr-cfs-acc.eps}
 \caption{}
@@
 \begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-cfs-auc.eps}
+\includegraphics[width=\textwidth]{../thesis/images/results/tr-cfs-auc.eps}
 \caption{}
@@
 \begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-wrapper-acc.eps}
+\includegraphics[width=\textwidth]{../thesis/images/results/tr-wrapper-acc.eps}
 \caption{}
@@
 \begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-wrapper-auc.eps}
+\includegraphics[width=\textwidth]{../thesis/images/results/tr-wrapper-auc.eps}
 \caption{}
@@
 
-\subsection*{Conclusions}
+\section{Conclusions}
 The best accuracy we achieved on the trauma data was 77.81\% with the $k$-NN
@@
 
-\end{multicols}
+\bibliographystyle{abbrv}{
+  \bibliography{../thesis/references}
+}
+
+\end{document}

### 2014-12-23 papers/paper.tex
Rearrange paper structure, change Abstract to Summary

@@
-\documentclass[10pt]{article}
+\documentclass{article}
 
@@
 \usepackage{graphicx}
+\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue,filecolor=blue,urlcolor=blue,pdfborder={0 0 0}]{hyperref}
+\usepackage[utf8]{inputenc}
 \usepackage{multicol}
@@
 
+\renewcommand{\abstractname}{Summary}
+
 \begin{abstract}
-Predicting the length of hospital stay for injured patients involves making a
-judgement about how long they need to be hospitalised, so that doctors and
-hospital staff can plan the appropriate course of treatment for the patient
-to ensure they recover in the quickest possible time. However, previous studies
-have used a limited number of statistical techniques such as logistic
-regression in order to predict the length of stay, and have not considered
-automatic methods of feature selection.
+Objective:
 
-Our work systematically evaluates and compares the performance of a number of
-state-of-the-art classification algorithms and feature selection methods,
-and also assesses the effect of feature discretisation for all combinations
-of classifiers and feature selectors. We propose a new classification algorithm
-called Ranked Distance Nearest Neighbor that considers the
-relative importance of each feature when determining the nearest neighbours.
+Materials and methods:
 
-We were able to achieve improved accuracy and area
-under the curve for both data sets, and we also show that our proposed Ranked
-Distance Nearest Neighbour improves accuracy and area under the curve over
-traditional nearest neighbour algorithms.
+Results:
+
+Conclusions:
+
+Keywords:
+
 \end{abstract}
@@
 
-\section{Problem Statement}
-Our problem is a binary \textit{classification} task in which we
-predict one of two possible categories for a patient: a length of stay (LOS) of
-less than or equal to 2 days, or a LOS of greater than 2 days.
+\section{Nearest Neighbour Classification}
+In this section we will describe the $k$-nearest neighbour ($k$-NN) algorithm
+and discuss some extensions which we use in our work. Additionally, we will
+describe our contribution, which is an extension of the distance metric for
+finding the nearest neighbours.
 
-\section{Ranked Distance Nearest Neighbour}
-\subsection{Intuition}
+\subsection{Basic Algorithm}
+The $k$-nearest neighbour classifier is an example of an
+\textit{instance-based} learning algorithm. This is because it does not, unlike
+the other classifiers described in this section, attempt to deduce or
+generalise a relationship between the features and the class: it simply stores
+all training examples and classifies an unseen example by finding the $k$
+``nearest'' examples (or \textit{instances}) and assigning the majority class
+to the new example. In the simplest form of the algorithm, $k$ is chosen to be
+1 and the Euclidean distance is used to measure the closeness of neighbours:
+\begin{equation*}
+\mathrm{Distance} = \sqrt{(x_1-y_1)^2 + (x_2-y_2)^2 + \ldots + (x_n-y_n)^2}
+\end{equation*}
+where $x_i,y_i$ are the values of the $i$-th feature for two different feature
+vectors (or instances).
+
+Like the Na\"{i}ve Bayes classifier, $k$-NN allows all features to be taken
+into account equally when classifying a new example. However, since feature
+values often span different ranges and units of measurement, a particularly
+large range of values for a feature would contribute much more to the distance
+than another feature that spans a lower set of values.
+To avoid this, it is important to \textit{normalise} the
+values of each numeric feature before using $k$-NN: for each feature $i$, we
+calculate the normalised value of the feature for each training example. This
+is done using the following relationship:
+\begin{equation*}
+a_i = s\left(\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right) + t
+\end{equation*}
+where $a_i$ is the normalised feature value, $v_i$ is the actual value in the
+data set, max $v_i$ and min $v_i$ are taken over all training examples, $s$ is
+a scaling factor and $t$ indicates how much to translate the range by. If $s=1$
+and $t=0$, then the range of normalised values will be in $[0,1]$.
+
+\subsection{Existing and Proposed Extensions}
+There have been many extensions proposed to the $k$-NN algorithm described
+above. These usually seek to find a more suitable distance metric for the
+problem domain at hand, or to reduce the storage requirements of the original
+$k$-NN algorithm (since the classifier makes no attempt to generalise from the
+examples, all of them must be stored).
+We will focus on two algorithms that extend the standard Euclidean distance:
+K* and our proposed approach, Ranked Distance Nearest Neighbour.
+
+\subsubsection{K*: Instance-based Classifier with Entropic Distance Function}
+Cleary and Trigg developed K*, an instance-based classifier that uses entropy
+to measure the distances between instances \cite{Cleary1995}. They consider
+the distance between two instances to be the complexity of transforming one
+into another. Define a sequence of transformations on an instance:
+\begin{equation*}
+\bar{t}(a) = t_n(t_{n-1}(\ldots t_1(a)\ldots)), \bar{t} \in T
+\end{equation*}
+where $\bar{t} = t_1,\ldots,t_n$. We can define a probability function $p$
+that gives the probabilities of these transformations occurring, and then we
+can define a probability function that gives the probabilities of all paths
+(transformations) from example $a$ to $b$:
+\begin{equation*}
+P^*(b|a) = \sum_{\bar{t}\in T:\bar{t}(a)=b} p(\bar{t})
+\end{equation*}
+Within this framework, we can define transformations for both real-valued and
+categorical features, which allows K* to work with all types of features.
+The K* function is then defined as:
+\begin{equation*}
+K^*(b|a) = -\mathrm{log}_2P^*(b|a)
+\end{equation*}
+
+To calculate the probability of an instance $a$ being in a class $C$, we sum
+the probabilities from $a$ to each instance of $C$, and repeat for each class.
+The predicted class will be the one with the highest probability:
+\begin{equation*}
+P^*(C|a) = \sum_{b\in C} P^*(b|a)
+\end{equation*}
+
+We use K* as a comparison to the basic NN algorithm and also as a comparison
+to our proposed Ranked Distance Nearest Neighbour algorithm. Additionally,
+K* has not been yet been applied in predicting LOS, and its performance in this
+domain is worth investigating.
+
+\subsubsection{Ranked Distance Nearest Neighbour}
+\paragraph{Intuition}
 As we mentioned earlier, all features are considered equally in a $k$-NN
@@
 
-\subsection{Mathematical formulation}
+\paragraph{Mathematical formulation}
 Given two instances $\mathbf{x} = (x_1,x_2,\ldots,x_n)$ and
@@
 
-\subsection{Assignment of weights}
+\paragraph{Assignment of weights}
 The weights $w_i$ can be tuned to match the particular problem that is being
@@
 
-\section{Feature Selection}
+\section{Evaluation}
+\subsection{Data Set}
+Our evaluation was carried out using a trauma-specific data set
+containing 2546 records collected from 2007--2011 collected at a major hospital
+in Sydney, Australia.
+
+\subsection{Feature Selection}
 Here we introduce the feature selection algorithms we evaluate in our work.
@@
 stay prediction.
-\subsection{Correlation-based Feature Selection}
-Correlation-based feature selection (CFS) is a method developed by Hall
-\cite{Hall2000} which looks for a subset of features that have high correlation
-with the class but low intercorrelation among the features. In this case,
-correlation between two features is measured using what is called the
-\textit{symmetric uncertainty}:
-\begin{equation*}
-U(A,B) = 2\dfrac{H(A)+H(B)-H(A,B)}{H(A)+H(B)}
-\end{equation*}
-where $A$ and $B$ are two nominal attributes and $H$ is the entropy function.
-The entropies are calculated from the probabilities of each feature value, and
-$H(A,B)$, the joint entropy of $A$ and $B$, is calculated from the
-probabilities of each combination of values of $A$ and $B$. Note that $U(A,B)$
-is always between 0 and 1.
-CFS conducts a search through subsets of all features in order to find the
-one feature subset that has the best merit.
 
-\subsection{Pearson Correlation Coefficient}
+\subsubsection{Pearson Correlation Coefficient}
 The Pearson correlation coefficient measures the degree of linear correlation
@@
 
-\subsection{Information Gain}
+\subsubsection{Information Gain}
 We can also measure the IG of each feature as follows.
@@
 
-\subsection{1R}
-Another way to evaluate the relevance of features is to use the idea behind the
-1R classifier described in Section. Recall that the 1R
-classifier constructs simple 1-rules based on each feature and uses the feature
-that produces the lowest error rate (that is, the highest accuracy) on the data
-set to classify unseen examples. These accuracy rates can be calculated for
-each feature, which are then used to rank the features from highest to lowest
-accuracy.
-
-We can also select a fixed number of features or all features above a certain
-1R accuracy threshold. This threshold is a slightly different quantity to the
-threshold of the correlation coefficient and information gain methods, as it
-is an accuracy, expressed in percentages. Our first threshold is the Zero-R
-accuracy for the trauma data set, which is the lower of the two Zero-R
-accuracies on our two data sets. We increase this percentage by in increasing
-increments to obtain a range of thresholds up to nearly 97\% accuracy (which
-discards all features for both data sets). Using the same thresholds across
-both data sets allows us to make valid comparisons of the effect of feature
-selection on performance.
-
-\subsection{Wrapper-based Feature Selection}
+\subsubsection{Wrapper-based Feature Selection}
 The feature selection methods described above (CFS, Pearson correlation and
@@
 
-\section{Evaluation Method}
-Our evaluation was carried out using a trauma-specific data set
-containing 2546 records collected from 2007--2011 collected at a major hospital
-in Sydney, Australia.
-\subsection{Classifiers}
+\subsection{Classification Algorithms}
 We tested the performance of Zero-R, 1R, Na\"{i}ve Bayes, C4.5 decision trees,
@@
 
-\subsection{Procedure}
+\subsection{Experimental Setup and Procedure}
 The steps that we took to evaluate the performance of the various classifiers
@@
 
+\section*{Acknowledgements}
+
 \bibliographystyle{abbrv}{

### 2015-02-02 papers/paper.tex
Paper: slowly coming along, paragraph by paragraph

@@
 their resources efficiently has always been a priority \cite{Walczak2003}.
-One key measurement of hospital activity, health care utilisation, and hospital
-cost is the patient length of stay \citep{Omachonu2004,Ng2006}. The length of
+One key measurement of hospital activity, health care utilisation and hospital
+cost is the patient length of stay \citep{Omachonu2004,Ng2006}.
+The length of
 stay has been used in a wide variety of medical domains, such as burns
@@
 Logistic regression models are straightforward to construct and use via the
-help of statistical software packages such as SAS, and
+help of statistical software packages such as
+SAS\footnote{SAS Institute, Cary, NC, USA}, and
 also make the contribution of each independent variable to the final
-prediction explicit. Thus they is easy to interpret when compared to more
+prediction explicit. Thus they are easier to interpret when compared to more
 ``black box'' approaches such as artificial neural networks \citep{Adams2012}.
@@
 mining techniques, particularly artificial neural networks,
-began to become more widely adopted following a number of
+began to be more widely adopted following a number of
 early applications to length of stay prediction in intensive
@@
 
-Given the benefits to the hospital of accurate length of stay prediction,
-this research aims to build upon improving length of stay prediction for trauma
-patients by utilising state-of-the-art data mining techniques. Being able to
-predict the length of stay for trauma patients is especially crucial because of
-the scarcity of patient care resource in this area \citep{Yang2010}.
-Additionally,
-despite the wealth of research into length of stay prediction in various
-medical domains, there has been little work done in generalising models from
-one domain to another, or to predict a combination of outcomes (not only
-length of stay, but mortality, or rehabilitation requirement). Being able to
-do this would mean that hospital staff are better equipped to handle patient
-cases and prepare for possible critical outcomes well in advance, resulting
-in time and resources saved for both hospital and patient.
+One of the major drawbacks of sophisticated data mining techniques is the high
+level of difficulty required in interpreting their predictions and the factors
+affecting those predictions. This is
+particularly true of artificial neural networks and the state-of-the-art
+support vector machine, which are able to achieve very high accuracy but can be
+non-intuitive to understand. Medical domain experts who use the predictions of
+trained learning algorithms need to understand how the algorithm reached its
+decision before they are able to confidently use it to aid their clinical
+decision-making.
+
+In this paper we present the application of nearest neighbour methods
+for predicting the length of stay category for patients admitted to a trauma
+ward. The categories are \textit{less than or equal to 2 days} or
+\textit{greater than 2 days}, which have been used in previous work on length
+of stay prediction in the same medical domain \citep{Dinh2013a}.
+Nearest neighour methods have not been studied in predicting the length of stay,
+and we believe they are much more intuitive to
+understand than techniques such as support vector machines.
+Additionally, they can be adjusted to suit particular medical domains by
+changing their similarity function. We compare the performance of the basic
+nearest neighbour algorithm, as well as an extension (K*), against the
+performance
+of previous work which used logistic regression. We also propose a novel
+extension to the standard nearest neighbour algorithm, Ranked Distance, and
+test its performance against the other classifiers. In addition, we apply a
+range of feature selection methods, as they have not been
+thoroughly studied in existing work. Finally, we compare the performance of
+all of these with the state-of-the-art support vector machines and the
+commonly-used artificial neural networks.
+
+We found that the standard 1-nearest neighbour classifier achieved the highest
+accuracy of 77.81\%, an improvement of 2.75\% over previous work using the same
+data set carried out by Dinh et al.\citep{Dinh2013a}.
 

### 2015-02-03 papers/paper.tex
Finish introduction

@@
 In this paper we present the application of nearest neighbour methods
-for predicting the length of stay category for patients admitted to a trauma
+in predicting the length of stay category for patients admitted to a trauma
 ward. The categories are \textit{less than or equal to 2 days} or
@@
 test its performance against the other classifiers. In addition, we apply a
-range of feature selection methods, as they have not been
+range of feature selection methods and systematically investigate the effect
+of discretisation, as they have not been
 thoroughly studied in existing work. Finally, we compare the performance of
-all of these with the state-of-the-art support vector machines and the
+these with the state-of-the-art support vector machines and the
 commonly-used artificial neural networks.
 
-We found that the standard 1-nearest neighbour classifier achieved the highest
-accuracy of 77.81\%, an improvement of 2.75\% over previous work using the same
-data set carried out by Dinh et al.\citep{Dinh2013a}.
+We found that the standard nearest neighbour classifier using only 1 nearest
+neighbour achieved a ten-fold cross-validated accuracy of 77.81\%, the highest
+out of all the methods. This is an improvement of 2.75\% over previous work
+carried out by Dinh et al. \citep{Dinh2013a} using the same data set. Also, the
+nearest neighbour classifiers almost always outperformed the logistic
+regression, support vector machines and neural networks when the number of
+features was gradually reduced. This means that predictions are able to be made
+more quickly, and will also be more intuitive to interpret and understand for
+physicians.
+
+In Section \ref{sec:nn}, we outline the basics of nearest neighbour
+classification and describe both the existing K* extension and our proposed
+extension, Ranked Distance. Section \ref{sec:eval} describes the method used
+to evaluate the performance of these classifiers, and Section \ref{sec:results}
+proceeds to discuss the results. We present concluding remarks in Section
+\ref{sec:conclusions}.
 
 \section{Nearest Neighbour Classification}
+\label{sec:nn}
 In this section we will describe the $k$-nearest neighbour ($k$-NN) algorithm
@@
 \section{Evaluation}
+\label{sec:eval}
 \subsection{Data Set}
@@
 \section{Results and Discussion}
+\label{sec:results}
 The best accuracy was 77.81\%, achieved using the $k$-NN algorithm
@@
 \section{Conclusions}
+\label{sec:conclusions}
 The best accuracy we achieved on the trauma data was 77.81\% with the $k$-NN

### 2015-02-04 papers/paper.tex
Move feature selection into its own section, add subheadings

@@
 classification and describe both the existing K* extension and our proposed
-extension, Ranked Distance. Section \ref{sec:eval} describes the method used
+extension, Ranked Distance. Section \ref{sec:features} outlines the feature
+selection methods we used. Section \ref{sec:eval} describes the method used
 to evaluate the performance of these classifiers, and Section \ref{sec:results}
@@
 \label{sec:nn}
-In this section we will describe the $k$-nearest neighbour ($k$-NN) algorithm
-and discuss some extensions which we use in our work. Additionally, we will
-describe our contribution, which is an extension of the distance metric for
-finding the nearest neighbours.
 
@@
 The $k$-nearest neighbour classifier is an example of an
-\textit{instance-based} learning algorithm. This is because it does not, unlike
-the other classifiers described in this section, attempt to deduce or
+\textit{instance-based} learning algorithm. This is because it does not attempt
+to deduce or
 generalise a relationship between the features and the class: it simply stores
@@
 
-\section{Evaluation}
-\label{sec:eval}
-\subsection{Data Set}
-Our evaluation was carried out using a trauma-specific data set
-containing 2546 records collected from 2007--2011 collected at a major hospital
-in Sydney, Australia.
-
-\subsection{Feature Selection}
+\section{Feature Selection}
+\label{sec:features}
 Here we introduce the feature selection algorithms we evaluate in our work.
@@
 
+\subsection{Automatic Selection}
 \subsubsection{Pearson Correlation Coefficient}
@@
 threshold to discard attributes.
+
+\subsection{Manual Selection}
+
+\section{Evaluation}
+\label{sec:eval}
+\subsection{Data Set}
+Our evaluation was carried out using a trauma-specific data set
+containing 2546 records collected from 2007--2011 collected at a major hospital
+in Sydney, Australia.
+
+\subsection{Discretisation}
 

### 2015-02-05 papers/paper.tex
Finish feature selection section of paper

@@
 Here we introduce the feature selection algorithms we evaluate in our work.
-Several of these have not been investigated in previous work on length of
-stay prediction.
+Several of these have not been investigated in previous work on LOS prediction.
+The role of feature selection is to reduce the number of attributes taken into
+consideration by a learning algorithm, which improves running speed and can
+make the resulting learned relationship easier to understand. We used both
+automatic and manual methods of selecting features.
 
 \subsection{Automatic Selection}
+Automatic feature selection involves calculating the ``relevance'' of each
+feature to predicting the class and only keeping features for use in training
+if their relevance is above some threshold. We consider the Pearson correlation
+coefficient, information gain and the wrapper approach.
+
 \subsubsection{Pearson Correlation Coefficient}
 The Pearson correlation coefficient measures the degree of linear correlation
-between two variables. It is always in the range $[-1,1]$, where -1 is total
+between two variables. It is always in the range $[-1,1]$, where $-1$ is total
 negative correlation, 0 is no (linear) correlation, and 1 is total positive
@@
 \subsubsection{Information Gain}
-We can also measure the IG of each feature as follows.
+We can also measure the information gain (IG) of each feature as follows.
 If we have a class $C$ and a feature $F$, the entropy of the
@@
 We can compute the IG for each feature and then select subsets of features that
-are greater than some threshold we specify, similarly to the correlation
+are greater than some threshold we specify, similar to the correlation
 coefficient method of feature selection. We choose a similar set of thresholds
@@
 \subsubsection{Wrapper-based Feature Selection}
-The feature selection methods described above (CFS, Pearson correlation and
-1R) have been described as \textit{filter} approaches to attribute selection.
+The feature selection methods described above (Pearson correlation and IG) have
+been described as \textit{filter} approaches to
+attribute selection, because the feature set is reduced \textit{before} it
+is given to the learning algorithm.
 Kohavi and John proposed another approach, the \textit{wrapper} approach:
 instead of performing feature selection independently of the learning
-algorithm using some measure of ``relevance'',
+algorithm using some measure of ``relevance,''
 the learning algorithm itself is used to evaluate the goodness of
@@
 divide and conquer nature of the learning algorithm, and the tendency of the
-trees to select fewer features for learning than other algorithms. Like CFS,
-wrapper methods search for the best subset of features and do not require a
+trees to select fewer features for learning than other algorithms.
+Wrapper methods search for the best subset of features and do not require a
 threshold to discard attributes.
@@
 \subsection{Manual Selection}
+All of the above automatic feature selection methods use some notion of
+``relevance,'' such as information gain or linear correlation,
+in order to determine the best subset of features to select.
+Depending on the problem at hand, these definitions of relevance may or may
+not work well with the data. In the LOS prediction literature, studies
+have enlisted the help of medical domain experts to assist in feature
+selection, due to their experience and understanding of the underlying
+problem and the effect of the features on the outcome. It will therefore be
+worthwhile to compare the performance of classifiers using features selected
+automatically and manually.
+
+\subsubsection{Domain Expert Selection}
+Consultation with a domain expert yielded a feature set that we
+use to compare with the baseline and with the other automatic methods.
+
+\subsection{Baseline Feature Set}
+The key previous work on trauma LOS prediction was by Dinh et. al.
+\cite{Dinh2013a}, who used a specific feature set from the data set that we
+use. We would like to use the features they selected as a baseline for
+evaluating the other feature selection methods we describe.
 
@@
 \subsection{Discretisation}
+Fayyad
+and Irani's entropy-based method of discretisation \cite{Fayyad1993}:
+to discretise an attribute,
+each possible split point is considered and the information gain from that
+split is computed. The split point with the highest information gain is
+therefore chosen. Once the first split is determined, the algorithm continues
+recursively in the newly created upper and lower ranges.
+
+The \textit{minimum description length (MDL)} principle is used to determine
+the stopping criterion. The MDL value is calculated by adding the information
+values from two cases: split or do not split. If we split, then we can encode
+the split point in $\mathrm{log}_2(N-1)$ bits (where $N$ is the number of
+instances) and subsequently the classes of those above and below it. If we do not split,
+then each instance's label must be encoded. Mathematically, given $N$ examples
+with entropy $E$, sub-intervals $k_1$ and $k_2$ with the entropy of examples
+in each sub-interval denoted $E_1$ and $E_2$:
+\begin{equation*}
+\mathrm{MDL} = \dfrac{\mathrm{log}_2(N-1)}{N} + \dfrac{\mathrm{log}_2(3^k-2) - kE + k_1E_1 + k_2E_2}{N}
+\end{equation*}
+The first term is the information needed for the splitting point, and the
+second is a correction necessary for encoding which classes belong to which
+sub-interval. The decision rule is then to split if the information gain for
+that split exceeds the MDL value.
+
+Since this requires exhaustive searching of the full range of feature values
+to find potential split points, this method of supervised discretisation can
+be time-consuming given large enough data sets. However, it has been proven
+that any potential split that minimises the information value will never occur
+between two examples of the same class \cite{Witten2005}.
+This means that we only need to
+consider splits between regions of different classes, reducing the number of
+possible points to consider.
+
+The data was discretised according to the above method using the supervised
+discretisation filter in WEKA. We kept both the original data set (without
+discretisation) and the discretised data set for comparison.
 

### 2015-02-07 papers/paper.tex
Small rearrangements, fleshing out the evaluation section

@@
 is done using the following relationship:
-\begin{equation*}
+\begin{equation}
+\label{eq:normalise}
 a_i = s\left(\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right) + t
-\end{equation*}
+\end{equation}
 where $a_i$ is the normalised feature value, $v_i$ is the actual value in the
@@
 \subsection{Data Set}
-Our evaluation was carried out using a trauma-specific data set
-containing 2546 records collected from 2007--2011 collected at a major hospital
-in Sydney, Australia.
+The data set used consisted of trauma registry data from the trauma centre
+at the Royal Prince Alfred Hospital, a major trauma centre in New South Wales,
+Australia. It covered all adult (age 15 and over) inpatient admissions to the
+trauma centre from 2007--2011. 
+All patients were first admitted to the trauma ward until discharged
+or transferred to an appropriate unit within the hospital. A single trained
+data manager recorded a variety of attributes about the admitted patient,
+such as age, gender, blood pressure, mechanism of injury and body regions
+that were injured.
 
-\subsection{Discretisation}
-Fayyad
-and Irani's entropy-based method of discretisation \cite{Fayyad1993}:
-to discretise an attribute,
-each possible split point is considered and the information gain from that
-split is computed. The split point with the highest information gain is
-therefore chosen. Once the first split is determined, the algorithm continues
-recursively in the newly created upper and lower ranges.
+There were 2546 patient records in the data set we received, comprising of 79
+features, one of which was the target variable \texttt{los48}.
+\texttt{los48} was a binary variable -- that is, it could only take two
+values, 0 or 1 -- with 1 indicating that the patient stayed two days or less,
+and 0 indicating a stay of longer than 2 days.
 
-The \textit{minimum description length (MDL)} principle is used to determine
-the stopping criterion. The MDL value is calculated by adding the information
-values from two cases: split or do not split. If we split, then we can encode
-the split point in $\mathrm{log}_2(N-1)$ bits (where $N$ is the number of
-instances) and subsequently the classes of those above and below it. If we do not split,
-then each instance's label must be encoded. Mathematically, given $N$ examples
-with entropy $E$, sub-intervals $k_1$ and $k_2$ with the entropy of examples
-in each sub-interval denoted $E_1$ and $E_2$:
-\begin{equation*}
-\mathrm{MDL} = \dfrac{\mathrm{log}_2(N-1)}{N} + \dfrac{\mathrm{log}_2(3^k-2) - kE + k_1E_1 + k_2E_2}{N}
-\end{equation*}
-The first term is the information needed for the splitting point, and the
-second is a correction necessary for encoding which classes belong to which
-sub-interval. The decision rule is then to split if the information gain for
-that split exceeds the MDL value.
+\subsection{Preprocessing}
+The data set was cleaned by correcting spelling and punctuation
+inconsistencies and removing extraneous whitespace. Examples which had missing
+values for at least one feature were removed, in line with previous work on the
+same data set.
 
-Since this requires exhaustive searching of the full range of feature values
-to find potential split points, this method of supervised discretisation can
-be time-consuming given large enough data sets. However, it has been proven
-that any potential split that minimises the information value will never occur
-between two examples of the same class \cite{Witten2005}.
-This means that we only need to
-consider splits between regions of different classes, reducing the number of
-possible points to consider.
-
-The data was discretised according to the above method using the supervised
-discretisation filter in WEKA. We kept both the original data set (without
-discretisation) and the discretised data set for comparison.
+Continuous-valued features were then normalised into the $[0,1]$ range as
+described in Section \ref{sec:nn} above, substituting $s=2$ and $t=-1$ into
+Equation \ref{eq:normalise}. These continuous features were then discretised
+using Fayyad and Irani's method of supervised discretisation
+\citep{Fayyad1993}. We kept both the original data set (normalised, without
+discretisation) and the discretised (and normalised) data set for comparison.
 
 \subsection{Classification Algorithms}
-We tested the performance of Zero-R, 1R, Na\"{i}ve Bayes, C4.5 decision trees,
+We tested the performance of
 support vector machines, multi-layer perceptrons, $k$-nearest neighbour with 1
 and 20 nearest neighbours, K* and logistic regression, as well as our proposed
-Ranked Distance Nearest Neighbour.
+Ranked Distance Nearest Neighbour -- a total of 7 different classifiers.
+
+\subsection{Statistical Significance}
 
@@
 \begin{enumerate}
-\item For each data set, we evaluated the effect of the 4 automatic feature
-selection methods on each of the 8 classifiers listed above, giving us 32
-configurations for each data set, over
-4 sets. Evaluation was performed using ten runs of \textit{ten-fold stratified
-cross-validation}.
-\item For the two trauma LOS data sets, we additionally evaluated the two
-manual feature selection methods with all 8 classifiers. Recall from Section
-\ref{sec:manual} that one of the feature sets is from the key previous work
-on trauma LOS prediction that we use as a baseline, and the other feature set
-is a set of features included through consultation with a domain expert.
+\item For both the discretised and non-discretised data set, we trained all 7
+classifiers without feature selection. Evaluation was performed using ten runs
+of \textit{ten-fold stratified cross-validation}.
+\item Then, for each data set, we evaluated the effect of the 3 automatic
+feature
+selection methods on each of the 7 classifiers listed above, giving us 21
+configurations for the discretised and non-discretised data set.
+\item In order to ensure that the baseline in the literature was comparable
+to our work, we used the baseline features and applied their method, logistic
+regression, to obtain a measure of performance on the non-discretised data set.
+We also evaluated this feature set against the rest of the 6 classifiers, and
+repeated this using the discretised data.
+\item Additionally, the expert-selected features were tested against all 7
+classifiers, with both the discretised and non-discretised data.
 \end{enumerate}
@@
 \end{subfigure}
-
-\begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{../thesis/images/results/tr-cfs-acc.eps}
-\caption{}
-\label{fig:tr-nothreshold-cfs-acc}
-\end{subfigure}%
-\begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{../thesis/images/results/tr-cfs-auc.eps}
-\caption{}
-\label{}
-\end{subfigure}
-\caption[]{Comparison of accuracy and AUC between classifiers for each non-threshold feature selection method, grouped by whether or not discretisation was applied.}
-\label{fig:tr-nothreshold}
-\end{figure*}
-
-\begin{figure*}
-\ContinuedFloat
 

### 2015-02-08 papers/paper.tex
Add more content to results and discussion section

@@
 \subsection{Basic Algorithm}
-The $k$-nearest neighbour classifier is an example of an
+The $k$-nearest neighbour ($k$-NN) classifier is an example of an
 \textit{instance-based} learning algorithm. This is because it does not attempt
@@
 We tested the performance of
-support vector machines, multi-layer perceptrons, $k$-nearest neighbour with 1
-and 20 nearest neighbours, K* and logistic regression, as well as our proposed
-Ranked Distance Nearest Neighbour -- a total of 7 different classifiers.
+support vector machines (SVMs), multi-layer perceptrons (MLPs), $k$-nearest
+neighbour with 1
+and 20 nearest neighbours, K* and logistic regression (LR), as well as our
+proposed
+Ranked Distance Nearest Neighbour (RD) -- a total of 7 different classifiers.
 
@@
 \label{sec:results}
+
+\subsection{Overall Performance}
 The best accuracy was 77.81\%, achieved using the $k$-NN algorithm
@@
 C4.5 wrapper. The best AUC of 0.846 was achieved using logistic regression with
-features selected by a correlation coefficient threshold of 0.1, and again
+features selected by a correlation coefficient threshold of 0.1, again
 using the discretised data set.
@@
 
+Recall that in order for our work to be comparable to that of previous work on
+trauma LOS prediction by Dinh et al. \cite{Dinh2013a}, we use the same data set
+that they used, and also tested the features they used as one of the manual
+feature selection methods.
+From Table \ref{tab:overall-tr-acc} we can see that
+the best accuracy achieved using their features is 75.06\% with logistic
+regression.
 It is worth noting that our best accuracy, using $k$-NN with 1
-nearest neighbour, shows a statistically significant improvement upon the accuracy
-that they achieved by 2.75\%
+nearest neighbour, shows a statistically significant improvement of 2.75\% upon
+the accuracy achieved by Dinh et al.
 while using only 11 features (compared to the 19 they used). These 11 features
@@
 
+\subsection{Comparison of Feature Selection Methods}
 The best accuracy and AUC results for
@@
 best AUC result out of all the feature selectors.
+
+Note that regardless of whether we consider accuracy or AUC to be more
+important in evaluating how our classifiers performed, neither the baseline
+feature set from Dinh et al. \cite{Dinh2013a} nor the features suggested by
+the domain expert (both of which were manually selected) yielded better
+results than automatic feature selection in all but two cases. We did find
+that expert selection produced the best accuracy result for 1R (equal with
+correlation coefficient selection, but used less features) and that the
+baseline feature set produced the best accuracy for Na\"{i}ve Bayes. Neither
+feature selection method produced the best AUC result for any classifier.
+This is a surprising result because intuitively and in the literature it
+has been stated that in a highly specialised field such as medicine, the domain
+experts would be best-equipped to make judgements about which features to
+use \cite{Witten2005}, and these should give reasonable results. However,
+our results show that in attempting to classify patients according to their
+LOS, we should consider both manual and automatic approaches.
 
@@
 
+\subsection{Comparison of Classifiers}
+Figures \ref{fig:tr-nothreshold} and \ref{fig:tr-threshold} compare the
+accuracy and AUC of all classifiers under each feature selection method for
+the trauma data set.
+The methods that do not select features based upon a specified threshold
+(manual methods as well as CFS, C4.5 wrapper and no feature selection)
+are graphed in Figure \ref{fig:tr-nothreshold}, and the methods that select
+based upon thresholds (correlation, information gain and 1R) are graphed
+in Figure \ref{fig:tr-threshold}, showing varying accuracy and AUC as the
+threshold is changed.
+
+In terms of classification accuracy, we did not find a lot of variation
+between all classifiers, with Zero-R producing 66.91\% accuracy and 77.81\%
+being the best we achieved -- an improvement of just over 10\%. The AUC of
+the 11 classifiers we tested ranged from 0.498 with Zero-R to 0.846 with
+logistic regression. Logistic regression, the \textit{de facto} method
+used in LOS classification, performed consistently better than the others
+across all feature
+selection methods, but only when the feature set was not heavily reduced by
+the feature selector: in these cases, such as the results from Figure
+\ref{fig:tr-nothreshold-cfs-acc} showing classifier performance with features
+selected by CFS, C4.5 decision trees and nearest
+neighbour methods have higher accuracy than logistic regression. This is also
+apparent from the graphs in Figure \ref{fig:tr-threshold}: for all three
+feature selection methods, as we increase the threshold, logistic regression
+drops in performance after the initial few thresholds and is outperformed by
+C4.5 and the various nearest neighbour classifiers ($k$-NN with 1 and 20
+neighbours and K*), as well as our Ranked Distance algorithm.
+This is true for both accuracy and AUC.
+
+We found that in most cases, the SVM did not achieve a higher accuracy or
+AUC than logistic regression with the same features even though SVMs are
+considered the state-of-the-art in classification algorithms. Additionally,
+as the feature set was reduced, the discriminating ability of SVMs (as
+indicated by the AUC) decreased more than their accuracy: this is particularly
+noticeable in Figure \ref{fig:tr-threshold-oner}, where the AUC of the SVM
+drops sharply to below the AUC of even the 1R classifier at the 67.905\%
+threshold, where there are
+only 4 features. SVMs have not been used in a LOS classification problem like
+ours, so there will need to be further work carried out in order to assess the
+suitability of using SVMs to classify patient LOS.
+
+Although the MLP has been investigated in studies attempting to identify
+factors associated with longer LOS or mortality for trauma patients
+\cite{Hunter2000,McGonigal1993}, leading to improved results with the neural
+network, we did not see the MLP outperform all other classifiers with respect
+to accuracy or AUC for any feature selection method for this data set. In fact,
+the MLP often performed no better than any of the nearest neighbour
+classifiers, which use a straightfoward mechanism of classifying unseen
+examples. Even if the MLP had managed to achieve superior accuracy or AUC, the
+``black box'' nature of its prediction decisions are a barrier for their
+widespread use in medical decision-making.
+
+Nearest neighbour approaches appear to have performed well on this data set in
+both accuracy and AUC, especially in comparison to more sophisticated
+classifiers and when the feature set is significantly reduced. From the graphs
+in Figure \ref{fig:tr-threshold}, we can see that all the nearest neighbours
+approaches perform similarly when the number of features is decreased, but our
+Ranked Distance algorithm showed a statistically significant improvement over
+other NN algorithms. This
+improvement is not apparent at smaller feature sets, where using the standard
+$k$-NN algorithm with 1 or 20 neighbours performs slightly better.
+The performance of NN we observed in our results is
+an interesting finding because nearest neighbours have not been commonly
+applied in predicting the LOS, but there is intuitive appeal in the
+case-by-case nature of how these algorithms classify new examples that should
+be explored further. Unlike MLPs and SVMs, prediction decision of nearest
+neighbours are more transparent and understandable. However, the major drawback
+of NN algorithms is their storage requirement and the computational time
+required to search for the nearest neighbours to make the classification.
+
+There is a limitation to our findings that needs to be noted: we did not
+attempt to tune the parameters of the classifiers in order to find the ones
+that resulted in the best performance, so our statements are based on the
+results of the default parameters of each classifier.
+MLPs in particular are sensitive to
+the network architecture, so it is likely that its accuracy and AUC could be
+improved if the parameters were selected to optimise its performance on this
+data set.
+
 \section{Conclusions}

### 2015-02-09 papers/paper.tex
Cut out some of the results and discussion section

@@
 no feature selection. This is likely due to the large number of features in
-the original data -- after cleaning and preliminary removal of redundant
+the original data: after cleaning and preliminary removal of redundant
 features, we were still left with over 50. Many of these would still likely
@@
 and 4 using the correlation coefficient and very few for the other methods.
-Under no combination was 1R the best feature selection method to use, as other
-methods produced better accuracy.
+\texttt{TODO: recount the numbers in this paragraph.}
 
@@
 Note that regardless of whether we consider accuracy or AUC to be more
-important in evaluating how our classifiers performed, neither the baseline
+important in evaluating classifiers, neither the baseline
 feature set from Dinh et al. \cite{Dinh2013a} nor the features suggested by
 the domain expert (both of which were manually selected) yielded better
-results than automatic feature selection in all but two cases. We did find
-that expert selection produced the best accuracy result for 1R (equal with
-correlation coefficient selection, but used less features) and that the
-baseline feature set produced the best accuracy for Na\"{i}ve Bayes. Neither
+results than automatic feature selection in all but two cases. Neither
 feature selection method produced the best AUC result for any classifier.
@@
 Recall that $k$-NN with 1 nearest neighbour and C4.5 wrapper feature
-selection produced the best accuracy (77.81\%) for the trauma data set, and
+selection produced the best accuracy (77.81\%) for the data set, and
 logistic regression with correlation coefficient feature selection at a
@@
 of examples. We did not face such a problem during our work, but this will
-have to be addressed for larger data sets. It is also common to use more
+need to be addressed for larger data sets. It is also common to use more
 than one nearest neighbour, so care needs to be taken to find a suitable number
 of neighbours for a given problem. Finally, we achieved an improvement
-in AUC of 0.034 but this was with 29 features, which is still a fairly large
+in AUC of 0.034 but with 29 features, which is still a fairly large
 number. The importance of using a fewer number of features is twofold: not
@@
 Figures \ref{fig:tr-nothreshold} and \ref{fig:tr-threshold} compare the
-accuracy and AUC of all classifiers under each feature selection method for
-the trauma data set.
+accuracy and AUC of all classifiers under each feature selection method.
 The methods that do not select features based upon a specified threshold
-(manual methods as well as CFS, C4.5 wrapper and no feature selection)
+(manual and baseline as well as C4.5 wrapper and no feature selection)
 are graphed in Figure \ref{fig:tr-nothreshold}, and the methods that select
-based upon thresholds (correlation, information gain and 1R) are graphed
+based upon thresholds (correlation and information gain) are graphed
 in Figure \ref{fig:tr-threshold}, showing varying accuracy and AUC as the
@@
 \ref{fig:tr-nothreshold-cfs-acc} showing classifier performance with features
-selected by CFS, C4.5 decision trees and nearest
+selected by C4.5 decision trees and nearest
 neighbour methods have higher accuracy than logistic regression. This is also
-apparent from the graphs in Figure \ref{fig:tr-threshold}: for all three
+apparent from the graphs in Figure \ref{fig:tr-threshold}: for both
 feature selection methods, as we increase the threshold, logistic regression
@@
 considered the state-of-the-art in classification algorithms. Additionally,
-as the feature set was reduced, the discriminating ability of SVMs (as
-indicated by the AUC) decreased more than their accuracy: this is particularly
-noticeable in Figure \ref{fig:tr-threshold-oner}, where the AUC of the SVM
-drops sharply to below the AUC of even the 1R classifier at the 67.905\%
-threshold, where there are
-only 4 features. SVMs have not been used in a LOS classification problem like
+as the feature set was reduced, both the discriminating ability of SVMs
+(indicated by the AUC) and accuracy decreased noticeably. SVMs have not been
+used in a LOS classification problem like
 ours, so there will need to be further work carried out in order to assess the
@@
 both accuracy and AUC, especially in comparison to more sophisticated
-classifiers and when the feature set is significantly reduced. From the graphs
+classifiers and when the feature set was significantly reduced. From the graphs
 in Figure \ref{fig:tr-threshold}, we can see that all the nearest neighbours

### 2015-02-10 papers/paper.tex
Small tidying changes to paper

@@
 
-Like the Na\"{i}ve Bayes classifier, $k$-NN allows all features to be taken
+$k$-NN allows all features to be taken
 into account equally when classifying a new example. However, since feature
@@
 feature set from Dinh et al. \cite{Dinh2013a} nor the features suggested by
-the domain expert (both of which were manually selected) yielded better
-results than automatic feature selection in all but two cases. Neither
+the domain expert yielded better
+results than automatic feature selection in all but \texttt{two?} cases. Neither
 feature selection method produced the best AUC result for any classifier.
-This is a surprising result because intuitively and in the literature it
+This is a surprising result because intuitively -- and in the literature -- it
 has been stated that in a highly specialised field such as medicine, the domain

### 2015-02-10 papers/paper.tex
Add paragraph to conclusion

@@
 
+Although we were able to improve on the work of Dinh et al. \cite{Dinh2013a}
+and confirm the
+importance of some features they and others found in predicting the LOS for
+trauma patients, we did not find a single classifier or feature selector
+that was superior to the others but instead observed different combinations
+giving better results than others in certain situations. This highlights the
+importance of understanding the data and investigating a range of classifiers
+and feature selection techniques when approaching any medical data mining
+problem, for which our work will become a catalogue.
+However, we did find that logistic regression tended to perform well
+on both data sets and across a number of feature selection methods, supporting
+its common use in deriving classifiers for various medical phenomena.
+
 \section*{Acknowledgements}

### 2015-02-14 papers/paper.tex
Rearrange paragraphs in results/discussion section

@@
 Recall that in order for our work to be comparable to that of previous work on
-trauma LOS prediction by Dinh et al. \cite{Dinh2013a}, we use the same data set
-that they used, and also tested the features they used as one of the manual
-feature selection methods.
+trauma LOS prediction by Dinh et al. \cite{Dinh2013a}, we use the same data
+set and also test the features they used as the baseline feature set.
 From Table \ref{tab:overall-tr-acc} we can see that
@@
 also achieved a higher AUC, and has the added advantages of being faster to
-train, mathematically well-understood and widely applied in medicine.
+train and widely applied in medicine.
 
@@
 each classifier resulted from a reduced feature set obtained through a
-particular feature selection method. There was at least one feature selector
-for each classifier that managed to outperform the same classifier with
-no feature selection. This is likely due to the large number of features in
-the original data: after cleaning and preliminary removal of redundant
-features, we were still left with over 50. Many of these would still likely
-be irrelevant to predicting the class.
-Out of 20 combinations of classifiers
+particular feature selection method. Out of 20 combinations of classifiers
 and discretisation (the 10 classifiers with and 10 without
@@
 
-Focussing our attention on the AUC reveals slightly different patterns. Here we
+In terms of AUC, we
 find that the correlation coefficient and the information gain are the feature
@@
 feature set from Dinh et al. \cite{Dinh2013a} nor the features suggested by
-the domain expert yielded better
+the expert yielded better
 results than automatic feature selection in all but \texttt{two?} cases. Neither
@@
 pre-processing step to improve the accuracy or AUC of a classifier, but also
-as a tool used separately that could help extract insights from medical data.
-
-Although we were able to achieve higher accuracy and AUC than the baseline,
-our best approaches also have a number of limitations which need to be pointed
-out. The first point to note is that using a wrapper method of feature
-selection is slower than using a filter approach (or manually picking the
-features before training), as the learning algorithm used as the wrapper
-needs to be invoked at each of the folds of cross-validation. Additionally,
-the $k$-NN algorithm becomes slower as the data set increases in the number
-of examples. We did not face such a problem during our work, but this will
-need to be addressed for larger data sets. It is also common to use more
-than one nearest neighbour, so care needs to be taken to find a suitable number
-of neighbours for a given problem. Finally, we achieved an improvement
-in AUC of 0.034 but with 29 features, which is still a fairly large
-number. The importance of using a fewer number of features is twofold: not
-only does this speed up the training time, it also means that we are able to
-make predictions without requiring a lot of information about a patient,
-which is a crucial consideration if this is to be implemented for use by
-physicians.
+as a tool used independently to extract insights from medical data.
 
 \subsection{Comparison of Classifiers}
-Figures \ref{fig:tr-nothreshold} and \ref{fig:tr-threshold} compare the
-accuracy and AUC of all classifiers under each feature selection method.
-The methods that do not select features based upon a specified threshold
-(manual and baseline as well as C4.5 wrapper and no feature selection)
-are graphed in Figure \ref{fig:tr-nothreshold}, and the methods that select
-based upon thresholds (correlation and information gain) are graphed
-in Figure \ref{fig:tr-threshold}, showing varying accuracy and AUC as the
-threshold is changed.
-
 In terms of classification accuracy, we did not find a lot of variation
@@
 \ref{fig:tr-nothreshold-cfs-acc} showing classifier performance with features
-selected by C4.5 decision trees and nearest
+selected by C4.5 and nearest
 neighbour methods have higher accuracy than logistic regression. This is also
@@
 examples. Even if the MLP had managed to achieve superior accuracy or AUC, the
-``black box'' nature of its prediction decisions are a barrier for their
-widespread use in medical decision-making.
+``black box'' nature of its prediction decisions is a barrier for their
+widespread adoption in medical decision-making.
 
@@
 be explored further. Unlike MLPs and SVMs, prediction decision of nearest
-neighbours are more transparent and understandable. However, the major drawback
-of NN algorithms is their storage requirement and the computational time
-required to search for the nearest neighbours to make the classification.
+neighbours are more transparent and understandable.
 
-There is a limitation to our findings that needs to be noted: we did not
+Although we were able to achieve higher accuracy and AUC than the baseline,
+our best approaches also have a number of limitations which need to be pointed
+out. The first point to note is that using a wrapper method of feature
+selection is slower than using a filter approach (or manually picking the
+features before training), as the learning algorithm used as the wrapper
+needs to be invoked at each of the folds of cross-validation. Additionally,
+the $k$-NN algorithm becomes slower as the data set increases in the number
+of examples. We did not face such a problem during our work, but this will
+need to be addressed for larger data sets. It is also common to use more
+than one nearest neighbour, so care needs to be taken to find a suitable number
+of neighbours for a given problem. Finally, we did not
 attempt to tune the parameters of the classifiers in order to find the ones

### 2015-02-14 papers/paper.tex
Small changes to headings

@@
 
+\subsection{Normalisation}
 $k$-NN allows all features to be taken
@@
 \subsubsection{Ranked Distance Nearest Neighbour}
-\paragraph{Intuition}
+\paragraph{Intuition.}
 As we mentioned earlier, all features are considered equally in a $k$-NN
@@
 
-\paragraph{Mathematical formulation}
+\paragraph{Mathematical formulation.}
 Given two instances $\mathbf{x} = (x_1,x_2,\ldots,x_n)$ and
@@
 
-\paragraph{Assignment of weights}
+\paragraph{Assignment of weights.}
 The weights $w_i$ can be tuned to match the particular problem that is being

### 2015-02-16 papers/paper.tex
Add graphs for no-threshold feature selectors with reduced classifier set

@@
 
-\subsection{Statistical Significance}
-
 \subsection{Experimental Setup and Procedure}
@@
 classifiers, with both the discretised and non-discretised data.
+\item Paired $t$-tests for statistical significance were carried out to
+determine whether or not discrepancies in performance were significant.
 \end{enumerate}
@@
 \begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{../thesis/images/results/tr-nofs-acc.eps}
+\includegraphics[width=\textwidth]{results/tr-nofs-acc.eps}
 \caption{}
@@
 \begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{../thesis/images/results/tr-nofs-auc.eps}
+\includegraphics[width=\textwidth]{results/tr-nofs-auc.eps}
 \caption{}
@@
 \begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{../thesis/images/results/tr-dinh-acc.eps}
+\includegraphics[width=\textwidth]{results/tr-dinh-acc.eps}
 \caption{}
@@
 \begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{../thesis/images/results/tr-dinh-auc.eps}
+\includegraphics[width=\textwidth]{results/tr-dinh-auc.eps}
 \caption{}
@@
 \begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{../thesis/images/results/tr-expert-acc.eps}
+\includegraphics[width=\textwidth]{results/tr-expert-acc.eps}
 \caption{}
@@
 \begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{../thesis/images/results/tr-expert-auc.eps}
+\includegraphics[width=\textwidth]{results/tr-expert-auc.eps}
 \caption{}
@@
 \begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{../thesis/images/results/tr-wrapper-acc.eps}
+\includegraphics[width=\textwidth]{results/tr-wrapper-acc.eps}
 \caption{}
@@
 \begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{../thesis/images/results/tr-wrapper-auc.eps}
+\includegraphics[width=\textwidth]{results/tr-wrapper-auc.eps}
 \caption{}

### 2015-03-07 papers/paper.tex
Add graphs for correlation and information gain feature selectors
Add overall tables and move plots into a separate file

@@
 
-\begin{figure*}[htbp]
-\begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{results/tr-nofs-acc.eps}
-\caption{}
-\label{}
-\end{subfigure}%
-\begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{results/tr-nofs-auc.eps}
-\caption{}
-\label{}
-\end{subfigure}
-
-\begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{results/tr-dinh-acc.eps}
-\caption{}
-\label{}
-\end{subfigure}%
-\begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{results/tr-dinh-auc.eps}
-\caption{}
-\label{}
-\end{subfigure}
-
-\begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{results/tr-expert-acc.eps}
-\caption{}
-\label{}
-\end{subfigure}%
-\begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{results/tr-expert-auc.eps}
-\caption{}
-\label{}
-\end{subfigure}
-
-\begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{results/tr-wrapper-acc.eps}
-\caption{}
-\label{}
-\end{subfigure}%
-\begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{results/tr-wrapper-auc.eps}
-\caption{}
-\label{}
-\end{subfigure}
-\caption[]{Comparison of accuracy and AUC between classifiers for each non-threshold feature selection method, grouped by whether or not discretisation was applied.}
-\label{}
-\end{figure*}
-
 Recall that in order for our work to be comparable to that of previous work on
@@
 train and widely applied in medicine.
+
+\input{results/overall-trauma-acc.tex}
+\input{results/overall-trauma-auc.tex}
 
@@
 as a tool used independently to extract insights from medical data.
+
+\input{results/tr-nothreshold-plots.tex}
+\input{results/tr-corr-ig-plots.tex}
 

### 2015-03-07 papers/results/overall-trauma-acc.tex
Add overall tables and move plots into a separate file

\begin{table}[htbp]
\resizebox{\linewidth}{!}{%
\begin{tabular}{r|cccccc}
Classifier & None & Baseline & Expert & C4.5 Wrapper & Correlation & IG \\ \hline
LR & \begin{tabular}[c]{@{}c@{}}77.62 (0.09)\\ D\end{tabular} & 75.06 (0.06)* & \begin{tabular}[c]{@{}c@{}}76.45 (0.06)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.3 (0.06)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.7 (0.09)\\ D, t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}77.65 (0.08)\\ D, t=0\end{tabular} \\
SVM & \begin{tabular}[c]{@{}c@{}}77.39 (0.08)\\ D\end{tabular} & 74.47 (0.07)* & \begin{tabular}[c]{@{}c@{}}75.67 (0.02)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}75.26 (0.07)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}77.42 (0.06)\\ D, t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.53 (0.07)\\ D, t=0.005\end{tabular}} \\
1NN & \begin{tabular}[c]{@{}c@{}}71.22 (0.1)\\ D\end{tabular} & 70.28 (0.08)* & \begin{tabular}[c]{@{}c@{}}72.56 (0.11)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.81 (0.07)\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.75 (0.13)\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.26 (0.13)\\ D, t=0.01\end{tabular} \\
20NN & \begin{tabular}[c]{@{}c@{}}75.9 (0.13)\\ D\end{tabular} & 73.59 (0.15)* & \begin{tabular}[c]{@{}c@{}}75.51 (0.19)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.48 (0.13)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.26 (0.2)\\ D, t=0.125\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.56 (0.18)\\ D, t=0.005\end{tabular}} \\
RD & \begin{tabular}[c]{@{}c@{}}76.92 (0.11)\\ D\end{tabular} & 69.68 (0.3)* & \begin{tabular}[c]{@{}c@{}}76.19 (0.16)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.22 (0.14)\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}76.87 (0.25)\\ D, t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.82 (0.16)\\ D, t=0.001\end{tabular} \\
K* & \begin{tabular}[c]{@{}c@{}}71.4 (0.1)\\ D\end{tabular} & 72.20 (0.11)* & \begin{tabular}[c]{@{}c@{}}75.08 (0.09)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.27 (0.07)\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}75.07 (0.14)\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.08 (0.12)\\ D, t=0.01\end{tabular} \\
MLP & \begin{tabular}[c]{@{}c@{}}73.79 (0.16)\\ D\end{tabular} & 73.03 (0.15)* & \begin{tabular}[c]{@{}c@{}}73.45 (0.16)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.66 (0.12)\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.65 (0.17)\\ T=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.03 (0.18)\\ D, t=0.001\end{tabular}
\end{tabular}
}
\caption{Overall best accuracy of each classifier with a particular feature selection method on the trauma data set. * indicates that results were identical regardless of discretisation. D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied. The absence of * or D means that not discretising produced a better result. The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
\label{tab:overall-tr-acc}
\end{table}

### 2015-03-07 papers/results/overall-trauma-auc.tex
Add overall tables and move plots into a separate file

\begin{table}[htbp]
\resizebox{\linewidth}{!}{%
\begin{tabular}{r|cccccc}
Classifier & None & Baseline & Expert & C4.5 Wrapper & Correlation & IG \\ \hline
LR & \begin{tabular}[c]{@{}c@{}}0.844 (0.0004)\\ D\end{tabular} & 0.812 (0.0003)* & \begin{tabular}[c]{@{}c@{}}0.825 (0.0003)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.831 (0.0002)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.846 (0.0003)\\ D, t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.844 (0.0005)\\ D, t=0\end{tabular} \\
SVM & \begin{tabular}[c]{@{}c@{}}0.843 (0.0004)\\ D\end{tabular} & 0.787 (0.0011)* & 0.807 (0.0003) & 0.806 (0.0003) & \textbf{\begin{tabular}[c]{@{}c@{}}0.844 (0.0004)\\ D, t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.842 (0.0004)\\ D, t=0\end{tabular} \\
1NN & \begin{tabular}[c]{@{}c@{}}0.726 (0.0013)\\ D\end{tabular} & 0.719 (0.001)* & \begin{tabular}[c]{@{}c@{}}0.760 (0.0011)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.816 (0.0007)\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.798 (0.0005)\\ D, t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.768 (0.0006)\\ T=0.1\end{tabular} \\
20NN & \begin{tabular}[c]{@{}c@{}}0.827 (0.0008)\\ D\end{tabular} & 0.792 (0.0007)* & \begin{tabular}[c]{@{}c@{}}0.808 (0.001)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.828 (0.0007)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.831 (0.0009)\\ D, t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.832 (0.0013)\\ D, t=0.01\end{tabular}} \\
RD & \begin{tabular}[c]{@{}c@{}}0.834 (0.0009)\\ D\end{tabular} & 0.742 (0.0016)* & \begin{tabular}[c]{@{}c@{}}0.807 (0.0012)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.825 (0.0009)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.836 (0.0011)\\ D, t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.834 (0.0011)\\ D, t=0.005\end{tabular} \\
K* & \begin{tabular}[c]{@{}c@{}}0.768 (0.0008)\\ D\end{tabular} & 0.774 (0.0008)* & \begin{tabular}[c]{@{}c@{}}0.803 (0.0006)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.838 (0.0003)\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.82 (0.0012)\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802 (0.0008)\\ D, t=0.01\end{tabular} \\
MLP & \begin{tabular}[c]{@{}c@{}}0.803 (0.0015)\\ D\end{tabular} & 0.781 (0.0014)* & \begin{tabular}[c]{@{}c@{}}0.784 (0.0016)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.825 (0.0008)\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.818 (0.0011)\\ T=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.803 (0.0017)\\ D, t=0\end{tabular}
\end{tabular}
}
\caption{Overall best AUC of each classifier with a particular feature selection method on the trauma data set. * indicates that results were identical regardless of discretisation. D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied. The absence of * or D means that not discretising produced a better result. The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
\label{tab:overall-tr-auc}
\end{table}

### 2015-03-07 papers/results/tr-corr-ig-plots.tex
Add overall tables and move plots into a separate file

--- 2014-11-02 thesis/results/tr-plots.tex
@@
 \begin{figure}[htbp]
-\begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-nofs-acc.eps}
-\caption{}
-\label{}
-\end{subfigure}%
-\begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-nofs-auc.eps}
-\caption{}
-\label{}
-\end{subfigure}
-
-\begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-dinh-acc.eps}
-\caption{}
-\label{}
-\end{subfigure}%
-\begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-dinh-auc.eps}
-\caption{}
-\label{}
-\end{subfigure}
-
-\begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-expert-acc.eps}
-\caption{}
-\label{}
-\end{subfigure}%
-\begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-expert-auc.eps}
-\caption{}
-\label{}
-\end{subfigure}
-
-\begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-cfs-acc.eps}
-\caption{}
-\label{fig:tr-nothreshold-cfs-acc}
-\end{subfigure}%
-\begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-cfs-auc.eps}
-\caption{}
-\label{}
-\end{subfigure}
-\caption{Comparison of accuracy and AUC between classifiers for each non-threshold feature selection method, grouped by whether or not discretisation was applied.}
-\label{fig:tr-nothreshold}
-\end{figure}
-
-\begin{figure}
-\ContinuedFloat
-
-\begin{subfigure}{.48\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-wrapper-acc.eps}
-\caption{}
-\label{}
-\end{subfigure}%
-\begin{subfigure}{.55\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-wrapper-auc.eps}
-\caption{}
-\label{}
-\end{subfigure}
-\caption{Comparison of accuracy and AUC between classifiers for each non-threshold feature selection method, grouped by whether or not discretisation was applied.}
-\label{}
-\end{figure}
-
-\begin{figure}[htbp]
-\begin{subfigure}{\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-corr.eps}
+\begin{subfigure}{1.1\textwidth}
+\includegraphics[width=\textwidth]{results/tr-corr.eps}
 \caption{}
@@
 \end{subfigure}
-\caption{Comparison of accuracy and AUC between classifiers for feature selectors that discard features based on a threshold.
+
+\begin{subfigure}{1.1\textwidth}
+\includegraphics[width=\textwidth]{results/tr-ig.eps}
+\caption{}
+\label{fig:tr-threshold-ig}
+\end{subfigure}
+\caption[]{Comparison of accuracy and AUC between classifiers for feature selectors that discard features based on a threshold.
 Solid lines correspond to a result without discretisation, and dashed lines correspond to a result with discretisation.}
@@
 \end{figure}
-
-\begin{figure}[htbp]
-\ContinuedFloat
-\begin{subfigure}{\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-ig.eps}
-\caption{}
-\label{fig:tr-threshold-ig}
-\end{subfigure}
-
-\begin{subfigure}{\textwidth}
-\includegraphics[width=\textwidth]{images/results/tr-oner.eps}
-\caption{}
-\label{fig:tr-threshold-oner}
-\end{subfigure}
-\caption{Comparison of accuracy and AUC between classifiers for feature selectors that discard features based on a threshold.
-Solid lines correspond to a result without discretisation, and dashed lines correspond to a result with discretisation.}
-\end{figure}

### 2015-03-07 papers/results/tr-nothreshold-plots.tex
Add overall tables and move plots into a separate file

\begin{figure*}[htbp]
\begin{subfigure}{.50\textwidth}
\includegraphics[width=\textwidth]{results/tr-nofs-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.57\textwidth}
\includegraphics[width=\textwidth]{results/tr-nofs-auc.eps}
\caption{}
\label{}
\end{subfigure}

\begin{subfigure}{.50\textwidth}
\includegraphics[width=\textwidth]{results/tr-dinh-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.57\textwidth}
\includegraphics[width=\textwidth]{results/tr-dinh-auc.eps}
\caption{}
\label{}
\end{subfigure}

\begin{subfigure}{.50\textwidth}
\includegraphics[width=\textwidth]{results/tr-expert-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.57\textwidth}
\includegraphics[width=\textwidth]{results/tr-expert-auc.eps}
\caption{}
\label{}
\end{subfigure}

\begin{subfigure}{.50\textwidth}
\includegraphics[width=\textwidth]{results/tr-wrapper-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.57\textwidth}
\includegraphics[width=\textwidth]{results/tr-wrapper-auc.eps}
\caption{}
\label{}
\end{subfigure}
\caption[]{Comparison of accuracy and AUC between classifiers for each non-threshold feature selection method, grouped by whether or not discretisation was applied.}
\label{}
\end{figure*}

### 2015-03-08 papers/paper.tex
Add correct references to tables and figures to the discussion

@@
 each classifier resulted from a reduced feature set obtained through a
-particular feature selection method. Out of 20 combinations of classifiers
-and discretisation (the 10 classifiers with and 10 without
+particular feature selection method. Out of 14 combinations of classifiers
+and discretisation (the 7 classifiers with and 7 without
 discretisation), the C4.5 wrapper method resulted in the best
-accuracy for 7 of these combinations, followed by 5 using information gain
-and 4 using the correlation coefficient and very few for the other methods.
-\texttt{TODO: recount the numbers in this paragraph.}
+accuracy for 5 of these combinations, followed by 5 using information gain,
+2 using the correlation coefficient and very few for the other methods.
 
@@
 equally well when used in conjunction with logistic regression and SVMs.
-Interestingly, although the C4.5
-wrapper method selects features which allow classifiers to achieve good
-accuracy, only in 2 out of the 20 combinations of classifiers did it give the
-best AUC result out of all the feature selectors.
+Interestingly, correlation coefficient as a feature selector produced the
+best AUC results for 6 out the 14 combinations, compared to only 2 when we
+considered accuracy instead.
 
@@
 the expert yielded better
-results than automatic feature selection in all but \texttt{two?} cases. Neither
+results than automatic feature selection in all cases. Neither
 feature selection method produced the best AUC result for any classifier.
@@
 the feature selector: in these cases, such as the results from Figure
-\ref{fig:tr-nothreshold-cfs-acc} showing classifier performance with features
+\ref{fig:tr-nothreshold-wrapper-acc} and \ref{fig:tr-nothreshold-wrapper-auc}
+showing classifier performance with features
 selected by C4.5 and nearest
